### 2004-06-28 doc/Simulation_and_Control_using_Supremica/functionblocks.tex
*** empty log message ***

\chapter{IEC 61499 Function Block Standard}
\label{functionblocks}

\begin{bf}
{This chapter is copied from Sachin Jain MSc Thesis}
\end{bf}

New technologies and standards are emerging that are set to have a
dramatic effect on the design and implementation of industrial
control systems. A new standard IEC-61131-3~\cite{IEC-61131-3} was
introduced in 1995 to standardize the representation of control
designs so as to enable cross-vendor integration. The standard
offered significant advantages over earlier vendor proprietary
control design representations~\cite{Book-61131}.
\begin{itemize}
\item It encouraged well-structured procedural
programs that allowed the development of different functional
elements into separate procedures or program units.
\item It provided standardized languages and methods of program execution
so that wide range of technological problems can be programmed as
vendor-independent software elements.
\item It provided support for various languages like structured
text, function block diagrams, ladder diagrams, instruction list
and sequential function charts.
\item The standard enforced strong data typing and supported user defined
data structures.
\end{itemize}

The standard however was based on a scanning mechanism where the
program execution was monitored by a continuous loop. Although
different parts of the program could be scanned at different
frequencies (based on fidelity requirements), it was tedious to
describe such behavior using sequential function charts. The
standard was based on polling mechanism rather than an event
driven system, and hence its execution order is not clearly
defined.

Further the standard was designed to support a software model and
languages for PLCs whose software was typically running on a
single processing resource. The communication between different
resources was supported through global variables and communication
blocks, both of which did not provide a clear and concise method
for defining the connections between distributed systems.

To overcome these problems, a function block model was sought that
would support distributed control systems in a scalable,
extensible and flexible manner. IEC-61499~\cite{IEC-61499}
standard has now been developed for distributed industrial
process, measurement and control systems.

In industrial systems, function blocks are now an established
concept for defining robust, re-usable software components. A {\it
Function Block} instance is defined to a software function unit
comprising an individual, named copy of a data structure and
associated operations specified by a corresponding function type.

Function blocks are analogous to objects in object-oriented
programming. Control designers can now take standard proven
encapsulated functionality in the form of blocks and link it
together in the quickest and most intuitive way possible. They
reduce the complexity since system integrators can now build large
systems using {\it black-box composition} i.e. without worrying
about how the internals of a block work.

The following sections are adapted largely from the various
discussions on the features of the standard in ~\cite{IEC-61499,
IEC-61131-3, Book-61131, Book-61499, christensen1, christensen2}

\section{Reference Models}
IEC-61499 has provided a set of models for describing distributed
systems that have been programmed using function blocks.
\begin{itemize}
\item {\bf System Model:} At the physical level, a distributed
system consists of a set of co-operating {\it applications}. An
application can exist on a single system such as application C in
figure \ref{f:System_Model} or have functionality distributed over
a number of devices such as applications A and B in Figure
\ref{f:System_Model}.

\item {\bf Device Model:} A resource provides independent
execution and control of networks of function blocks. It supports
zero or more {\it resources} and has at least one interface. The
interface could either be a 'process interface' that allows the
resources to exchange data with the input and output points on the
physical devices, or a communication interface that allows the
resources to communicate with external resources in remote
devices. The device model is illustrated in Figure
\ref{f:Device_Model}.

\item {\bf Resource Model:} A resource is an independent functional unit,
contained in a device which has independent control of its
operation. The resource can be created, configured, parameterized,
started up, deleted etc. without effecting other resources. Each
resource provides an interface to the communication systems and
tot the device specific process. The resource is therefore,
concerned with the mapping of data and event flows which pass
between function blocks in the local resource to remove resource
function blocks via the device communication interfaces.

Figure \ref{f:Resource_Model} depicts the main features of an IEC
61499 resource. Within the resource, it shows a network of
interconnected function blocks linked by data and event flows. A
scheduling function ensures that the events are executed in the
correct serial order. {\it Service Interface} function blocks are
a special form of function blocks that provide a link between
function blocks and the interface of the resource.

\item {\bf Application Model:} An application is defined as a
network of interconnected function blocks, linked by data and
event flows. An application can be fragmented and distributed over
many resources. The resources ensure that the events are executed
in the correct serial order based on the causal relationships
specified by the applications.

An application could be understood as an entire set of function
blocks and interconnections to solve a particular automation
control problem. Figure \ref{f:Application_Model} illustrates an
application, where each individual block is either a function
block or a subapplication.

\item {\bf Function Block Model:} A function block is a functional
unit of software that its own data structure which can be
manipulated by one or more algorithms. The formal description of
the data structure is provided by the function block type
definition.

The main features of function blocks are:
\begin{itemize}
\item Each function block has a type name and an instance name.
\item Each block ha a set of event inputs, which can receive
inputs from other blocks through event connections.
\item Each block has one or more outputs, through which it can
propagate its results to other blocks.
\item There is a set of data inputs that allow values to be passed
in from other blocks.
\item There is a set of data outputs, through which the block can
pass its results to other blocks.
\item Each block may have a set of internal variables that hold
its state between algorithm invocations.
\item The behavior of the block is defined in terms of algorithms
and state information. This behavior is defined in the function
block type definition.
\end{itemize}

Figure \ref{f:Function_Block_Char} shows the main characteristics
of a function block. The top part of the block, called the
'execution control' portion contains in some cases, a state
machine that maps events to the algorithms that should be
triggered as a consequence of the event. The lower portion of
block contains both the algorithms and the internal data, both of
which are encapsulated within the block.

\end{itemize}

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/System_Model.eps, scale=.8} \caption[System
Model]{System Model {\protect ~\cite{IEC-61499}}}
\label{f:System_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Device_Model.eps, scale=.8} \caption[Device
Model]{Device Model{\protect ~\cite{IEC-61499}}}
\label{f:Device_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Resource_Model.eps, scale=.8} \caption[Resource
Model]{Resource Model{\protect ~\cite{IEC-61499}}}
\label{f:Resource_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Application_Model.eps, scale=.8}
\caption[Application Model]{Application Model{\protect
~\cite{IEC-61499}}} \label{f:Application_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Function_Block_Char.eps, scale=.8}
\caption[Function Block Characteristics]{Function Block
Characteristics{\protect ~\cite{IEC-61499}}}
\label{f:Function_Block_Char}
\end{center}
\end{figure}
%%\doublespace
%

\section{Function Block Types}

A function block type is defined by a type name, formal
definitions for the input events and data and the output events
and data. The type definition also includes the internal behavior
of the block. The notion of a function block type is analogous to
the class definition in object-oriented programming while function
block instances correspond to the object instances of the classes.

\begin{itemize}
\item{\bf Basic Function Block Types:} The behavior of basic
function blocks is defined in terms of algorithms that are invoked
in response to input events. As algorithms execute, they trigger
output events to signal their state to other blocks. The basic
blocks hold their state in terms of internal variables. Basic
function blocks also can communicate their state using input and
output data variables.

The mapping of events onto algorithms is expressed using a state
transition notation known as execution control chart and is
described in the next section. Figure \ref{f:Basic_Block} shows an
example of a basic function block.

\item{\bf Composite Function Blocks:} The behavior of a composite
function block is described in terms of the event and data
connections between its component function blocks. A composite
function block can be used to hide the internal interactions
between the component blocks and presenting a simplified interface
for use by the outside world. A composite function block is
illustrated in Figure \ref{f:Composite_Block}.

\item {\bf Service Interface Function Blocks:} Service interface
function blocks provide an interface between the function block
domain and the external services. Since a service interface
function block is primarily concerned with data exchange, it is
defined using time sequence diagrams. Examples of service
interface function blocks are communication blocks for publishing
and subscribing over a network and management blocks for
dynamically creating component blocks and event connections.

\item {\bf Event Function Blocks:} Event function blocks are some
special basic function blocks defined in the standard to ease the
management of events. These blocks are used for control,
generation and detection of events and can be used by other
composite blocks. These blocks are defined using an execution
control chart or a time sequence diagram. These blocks can be
used, for instance, for the following purposes.
\begin{enumerate}
\item To split events to produce new events
\item To allow events to be merged.
\item To permit the propagation of certain events.
\item To select between two or more events.
\item To delay an event by a given period.
\item To generate a stream of events at a regular interval.
\end{enumerate}

\end{itemize}

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Basic_Function_Block.eps, scale=.8}
\caption[Basic Function Block]{Basic Function Block{\protect
~\cite{IEC-61499}}} \label{f:Basic_Block}
\end{center}
\end{figure}
%\doublespace
%

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Composite_Function_Block.eps, scale=.8}
\caption[Composite Function Block]{Composite Function
Block{\protect ~\cite{IEC-61499}}} \label{f:Composite_Block}
\end{center}
\end{figure}
%\doublespace
%

\section{Execution Model}

The execution of algorithms for basic function blocks is invoked
by the execution control portion of a function block instance in
response to events at event inputs. This invocation takes the form
of a request to the underlying scheduling function of the
associated resource to schedule the execution of algorithm's
operations. The model assumes that the resource in which a
function block exists provides a scheduling function that ensures
each phase of function block execution occurs in the correct order
and with the correct priority.

The execution model specifies eight discrete timing points as
shown in Figure \ref{f:Execution_Model} that must occur
sequentially for the basic function block to operate. These are:

\begin{description}
\item[$t_1$]: Values coming from external function blocks
as input data are stable at the current block.
\item[$t_2$]: The event associated with the input values occurs.
\item[$t_3$]: The execution control function notifies the resource
scheduling function that it has the input values and is ready to
execute the algorithm.
\item[$t_4$]: The scheduling function begins the algorithm execution.
\item[$t_5$]: The algorithm processes input values, (and in some
cases, internal variables) to create new output values that are
written to the block's outputs.
\item[$t_6$]: The algorithm notifies the scheduling function that it
has completed its execution.
\item[$t_7$]: The scheduling function invokes the execution control
to generate the output events.
\item[$t_8$]: The execution control generates the appropriate output
events at its event output ports. The event may be used by
downstream blocks to signal that they may use the output values
generated by this block.
\end{description}

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Execution_Control_Chart.eps, scale=.8}
\caption[Execution Model]{Execution Model{\protect
~\cite{IEC-61499}}} \label{f:Execution_Model}
\end{center}
\end{figure}
%\doublespace
%

To ensure consistency, the algorithms must use the snapshot of
input variables that exists when the input event arrived. Further
the scheduling function must be able to detect the arrival of
input events at a faster rate than the rate at which they can be
handled by the block.

\section{FBEditor}
Function Block Development Kit (FBDK) ~\cite{FBEditor} is a
software toolkit that enables users to build and test data types,
function block types, resource types, device types and system
configurations according to the IEC 61499 standard. FBEditor is a
part of the FBDK toolkit that provides a development and runtime
environment for applications based on the IEC 61499 standard.
Figure \ref{f:FBEditor} shows the editor being used to load a
function block used for controlling a motor.

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/FBEditor.eps, scale=.6}
\caption[FBEditor]{FBEditor{\protect ~\cite{FBEditor}}}
\label{f:FBEditor}
\end{center}
\end{figure}
%\doublespace
%

FBEditor enables the users to edit the block definitions through
both a text editor and through a graphical user interface. New
data types, basic function blocks, composite function blocks,
execution control charts and other higher level entities may be
created through the editor. The attributes of the blocks (data and
events) may also be examined through a graphical user interface.
The function blocks can be examined at the device level as shown
in Figure \ref{f:Editor-Devices} or at the level of individual
attributes like the connections, as shown in figure
\ref{f:Editor-Mapping}.

The editor supports the languages supported by the standard
(Structured text, execution control chart etc.). It uses XML as
the exchange format to ensure portability across various domains.
Figure \ref{f:Editor-Versions} shows how the editor can also be
used to manage meta properties like versions, owner entities,
authors etc.

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Devices.eps, scale=.6} \caption[FBEditor
- Editing system devices]{FBEditor - Editing system
devices{\protect ~\cite{FBEditor}}} \label{f:Editor-Devices}
\end{center}
\end{figure}
%\doublespace
%

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Mapping.eps, scale=.6} \caption[FBEditor
- Editing device connections]{FBEditor - Editing device
connections{\protect ~\cite{FBEditor}}} \label{f:Editor-Mapping}
\end{center}
\end{figure}
%\doublespace
%

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Versions.eps, scale=.6}
\caption[FBEditor - Editing meta properties]{FBEditor - Editing
meta properties{\protect ~\cite{FBEditor}}}
\label{f:Editor-Versions}
\end{center}
\end{figure}
%\doublespace
%
FBEditor has a navigation pane that allows the user to choose the
display on the worksheet area, including the external interface;
the Execution Control Chart (ECC) for basic function block type;
service primitive sequences for service interface function block
types; and function block networks for composite function block
types, device and resource types and instances, and system
configurations. The worksheet / viewport area contains the
function block diagram for the currently selected instance.
Different components of the diagram can be dragged around for
better readability.

FBEditor has been developed in Java language. Hence in order to
run simulations on applications written using the editor, one
needs to generate the Java classes that correspond to the source
files for the function blocks written in structured text (for
instance). The editor has a code generation mechanism through
which it parses through the source files and generates a Java
class corresponding to each block.

On occurrence of certain input events, the user needs to specify
the algorithms that should be scheduled to be executed and the
output events that should signal the completion of algorithms.
This may be done either by editing the generated Java files or by
enclosing the algorithm inside especially marked tags in the
source files. The files are then compiled and simulations can be
run from within the editor environment. Figure
\ref{f:Editor-Transitions} shows the assignment of output events
on the transitions through the editing of the execution control
chart.

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Transitions.eps, scale=.6}
\caption[FBEditor Editing the execution control chart]{FBEditor
Editing the execution control chart{\protect ~\cite{FBEditor}}}
\label{f:Editor-Transitions}
\end{center}
\end{figure}
%\doublespace
%

\section{EMBench and Function Blocks}

Since the FBEditor provided the aforementioned functionality, it
was decided that the EMBench should leverage it. In particular,
the bench relies on the editor for two basic features.
\begin{enumerate}
\item To generate the Java files from the IEC 61499 source files.
\item To manage the Java thread that initializes the blocks during
simulation.
\end{enumerate}

It has been visualized that the bench will eventually have the
editor as one of its components. The bench explicitly manages the
design activities of the user and generates the function block
definitions for the user design. It then requests the FBEditor
component to load the specified source files, generate the Java
classes and run the simulation. In addition, the user may make
behavioral changes to the blocks by editing the events, if he/she
so desires. The editor conforms to the published service
specifications of the bench as explained in chapter \ref{chap3}
and hence would present the same interface to the client.

Currently, however, the FBEditor is not a part of the bench
because it does not expose an API. Since it is external to the
bench, the bench has to circumvent this problem by allowing the
editor to be executed as a separate application and delegating the
simulation task completely to the editor. The data exchange
necessary between the two applications is handled via network
communication. This issue is explained in a greater detail in
chapter~\ref{chap4} and chapter~\ref{chap5}.

### 2004-06-28 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

\chapter{Automation Objects}
\label{automationobjects}

Introduce Automation Objects
Define the different views (Model-View-Controller)
	Configuration
	Human-Machine Interface
	Control
	Simulation

### 2004-06-28 doc/Simulation_and_Control_using_Supremica/introduction.tex
*** empty log message ***

\chapter{Introduction}
\label{introduction}

Define the problem.
Define other approaches, state of the art.
Introduce Function Blocks
Introduce Automation Objects
(Configuration, Simulation, Control, HMI)

### 2004-06-28 src/org/supremica/automata/IO/AutomataToIEC61499.java
GC

@@ -52,8 +52,11 @@ import java.io.*;
 import java.util.*;
 import java.text.DateFormat;
 
+
 import org.supremica.log.*;
 import org.supremica.automata.*;
+import org.supremica.util.SupremicaException;
+import org.supremica.automata.algorithms.SynchronizationType;
 
 /**
  * @author cengic
@@ -69,51 +72,38 @@ import org.supremica.automata.*;
  * 
  */
 
+// TODO: Get rid of the internal variables
+// TODO: Make the generated function behave according to agreement
+// TODO: Make state variables int
+// TODO: Make comments optional
+
+
 public class AutomataToIEC61499
 
 {
 	
 	private static Logger logger = LoggerFactory.createLogger(AutomataToIEC61499.class);
-	
 	private Project theProject;
-	
-	private AutomataToIEC1131 theAutomataToIEC1131;
-	
+	private IEC61131Helper theHelper;
 	private Alphabet allEvents;
-	
+	private SynchronizationType syncType = SynchronizationType.Prioritized;
 	
 	
 	
 	// Constructor
-	
 	public AutomataToIEC61499(Project theProject)
 	{
-		
 		this.theProject = theProject;
+		this.theHelper = IEC61131Helper.getInstance();
 		this.initialize();
-		
-		try 
-		{
-			theAutomataToIEC1131 = new AutomataToIEC1131(theProject);
-		} 
-		catch (Exception e) 
-		{
-			logger.error("Something wrong with instantiation of AutomataToIEC1131");
-		}
-	
 	}
 
-	
-	
-	
 	
 	
 	private void initialize()
 	{
 		allEvents = theProject.setIndicies();
-		
 	}
-
 
 
 
@@ -221,16 +211,35 @@ public class AutomataToIEC61499
 	{
 		pw.println("VAR");
 
+		// Not needed acctually
 		// Internal event variables
-		for (Iterator alphIt = allEvents.iterator(); alphIt.hasNext();) 
-		{
-			LabeledEvent currEvent = (LabeledEvent) alphIt.next();
-			pw.println("\te_" + currEvent.getSynchIndex() + " : BOOL;");
-		}
+		//for (Iterator alphIt = allEvents.iterator(); alphIt.hasNext();) 
+		//{
+		//	LabeledEvent currEvent = (LabeledEvent) alphIt.next();
+		//	pw.println("\te_" + currEvent.getSynchIndex() + " : BOOL;");
+		//}
 
-		// Variables needed for the AutomataToIEC1131
[…]

### 2004-06-28 src/org/supremica/gui/ActionMan.java

@@ -3817,7 +3817,7 @@ public class ActionMan
 						PrintWriter theWriter =
 							new PrintWriter(new FileWriter(currFile));
 
-						exporter.serializeStructuredText(theWriter);
+						exporter.printSource(theWriter);
 						theWriter.close();
 					} catch (Exception ex) {
 						logger.error(

### 2004-06-28 doc/Simulation_and_Control_using_Supremica/functionblocks.tex
*** empty log message ***

@@
 \begin{bf}
-{This chapter is copied from Sachin Jain MSc Thesis}
+{This chapter is with minor revisions copied from chapter 2 in \cite{j:des:2002}.}
 \end{bf}
@@
 
-\section{EMBench and Function Blocks}
+%\section{EMBench and Function Blocks}
 
-Since the FBEditor provided the aforementioned functionality, it
-was decided that the EMBench should leverage it. In particular,
-the bench relies on the editor for two basic features.
-\begin{enumerate}
-\item To generate the Java files from the IEC 61499 source files.
-\item To manage the Java thread that initializes the blocks during
-simulation.
-\end{enumerate}
+%Since the FBEditor provided the aforementioned functionality, it
+%was decided that the EMBench should leverage it. In particular,
+%the bench relies on the editor for two basic features.
+%\begin{enumerate}
+%\item To generate the Java files from the IEC 61499 source files.
+%\item To manage the Java thread that initializes the blocks during
+%simulation.
+%\end{enumerate}
 
-It has been visualized that the bench will eventually have the
-editor as one of its components. The bench explicitly manages the
-design activities of the user and generates the function block
-definitions for the user design. It then requests the FBEditor
-component to load the specified source files, generate the Java
-classes and run the simulation. In addition, the user may make
-behavioral changes to the blocks by editing the events, if he/she
-so desires. The editor conforms to the published service
-specifications of the bench as explained in chapter \ref{chap3}
-and hence would present the same interface to the client.
+%It has been visualized that the bench will eventually have the
+%editor as one of its components. The bench explicitly manages the
+%design activities of the user and generates the function block
+%definitions for the user design. It then requests the FBEditor
+%component to load the specified source files, generate the Java
+%classes and run the simulation. In addition, the user may make
+%behavioral changes to the blocks by editing the events, if he/she
+%so desires. The editor conforms to the published service
+%specifications of the bench as explained in chapter \ref{chap3}
+%and hence would present the same interface to the client.
 
-Currently, however, the FBEditor is not a part of the bench
-because it does not expose an API. Since it is external to the
-bench, the bench has to circumvent this problem by allowing the
-editor to be executed as a separate application and delegating the
-simulation task completely to the editor. The data exchange
-necessary between the two applications is handled via network
-communication. This issue is explained in a greater detail in
-chapter~\ref{chap4} and chapter~\ref{chap5}.
+%Currently, however, the FBEditor is not a part of the bench
+%because it does not expose an API. Since it is external to the
+%bench, the bench has to circumvent this problem by allowing the
+%editor to be executed as a separate application and delegating the
+%simulation task completely to the editor. The data exchange
+%necessary between the two applications is handled via network
+%communication. This issue is explained in a greater detail in
+%chapter~\ref{chap4} and chapter~\ref{chap5}.

### 2004-06-28 doc/Simulation_and_Control_using_Supremica/functionblocks.tex
*** empty log message ***

@@
 dramatic effect on the design and implementation of industrial
-control systems. A new standard IEC-61131-3~\cite{IEC-61131-3} was
+control systems. A new standard IEC-61131-3~\cite{iec:1131:1993} was
 introduced in 1995 to standardize the representation of control
@@
 offered significant advantages over earlier vendor proprietary
-control design representations~\cite{Book-61131}.
+control design representations~\cite{l:pro:1995}.
 \begin{itemize}
@@
 would support distributed control systems in a scalable,
-extensible and flexible manner. IEC-61499~\cite{IEC-61499}
+extensible and flexible manner. IEC-61499~\cite{iec:61499:2000}
 standard has now been developed for distributed industrial
@@
 The following sections are adapted largely from the various
-discussions on the features of the standard in ~\cite{IEC-61499,
-IEC-61131-3, Book-61131, Book-61499, christensen1, christensen2}
+discussions on the features of the standard in ~\cite{iec:61499:2000,
+iec:1131:1993, l:pro:1995, l:mod:2001, c:des:2002, c:ope:2002}
 
@@
 \epsfig{file=images/functionblocks/System_Model.eps, scale=.8} \caption[System
-Model]{System Model {\protect ~\cite{IEC-61499}}}
+Model]{System Model {\protect ~\cite{iec:61499:2000}}}
 \label{f:System_Model}
@@
 \epsfig{file=images/functionblocks/Device_Model.eps, scale=.8} \caption[Device
-Model]{Device Model{\protect ~\cite{IEC-61499}}}
+Model]{Device Model{\protect ~\cite{iec:61499:2000}}}
 \label{f:Device_Model}
@@
 \epsfig{file=images/functionblocks/Resource_Model.eps, scale=.8} \caption[Resource
-Model]{Resource Model{\protect ~\cite{IEC-61499}}}
+Model]{Resource Model{\protect ~\cite{iec:61499:2000}}}
 \label{f:Resource_Model}
@@
 \caption[Application Model]{Application Model{\protect
-~\cite{IEC-61499}}} \label{f:Application_Model}
+~\cite{iec:61499:2000}}} \label{f:Application_Model}
 \end{center}
@@
 \caption[Function Block Characteristics]{Function Block
-Characteristics{\protect ~\cite{IEC-61499}}}
+Characteristics{\protect ~\cite{iec:61499:2000}}}
 \label{f:Function_Block_Char}
@@
 \caption[Basic Function Block]{Basic Function Block{\protect
-~\cite{IEC-61499}}} \label{f:Basic_Block}
+~\cite{iec:61499:2000}}} \label{f:Basic_Block}
 \end{center}
@@
 \caption[Composite Function Block]{Composite Function
-Block{\protect ~\cite{IEC-61499}}} \label{f:Composite_Block}
+Block{\protect ~\cite{iec:61499:2000}}} \label{f:Composite_Block}
 \end{center}
@@
 \caption[Execution Model]{Execution Model{\protect
-~\cite{IEC-61499}}} \label{f:Execution_Model}
+~\cite{iec:61499:2000}}} \label{f:Execution_Model}
 \end{center}
@@
 \section{FBEditor}
-Function Block Development Kit (FBDK) ~\cite{FBEditor} is a
+Function Block Development Kit (FBDK) ~\cite{c:fun:2002} is a
 software toolkit that enables users to build and test data types,
@@
 \epsfig{file=images/functionblocks/FBEditor.eps, scale=.6}
-\caption[FBEditor]{FBEditor{\protect ~\cite{FBEditor}}}
+\caption[FBEditor]{FBEditor{\protect ~\cite{c:fun:2002}}}
 \label{f:FBEditor}
@@
 - Editing system devices]{FBEditor - Editing system
-devices{\protect ~\cite{FBEditor}}} \label{f:Editor-Devices}
+devices{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Devices}
 \end{center}
@@
 - Editing device connections]{FBEditor - Editing device
-connections{\protect ~\cite{FBEditor}}} \label{f:Editor-Mapping}
+connections{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Mapping}
 \end{center}
@@
 \caption[FBEditor - Editing meta properties]{FBEditor - Editing
-meta properties{\protect ~\cite{FBEditor}}}
+meta properties{\protect ~\cite{c:fun:2002}}}
 \label{f:Editor-Versions}
@@
 \caption[FBEditor Editing the execution control chart]{FBEditor
-Editing the execution control chart{\protect ~\cite{FBEditor}}}
+Editing the execution control chart{\protect ~\cite{c:fun:2002}}}
 \label{f:Editor-Transitions}

### 2004-06-28 src/org/supremica/automata/IO/AutomataToIEC61499.java
GC

@@ -233,7 +233,7 @@ public class AutomataToIEC61499
 				State currState = (State) stateIt.next();
 				int currStateIndex = currState.getSynchIndex();
 
-				theHelper.printBooleanVariableDeclaration(pw, "q_" + currAutomatonIndex + "_" + currStateIndex, currState.getName() + " in " + currAutomaton.getName(),2);
+				theHelper.printBooleanVariableDeclaration(pw, "Q_" + currAutomatonIndex + "_" + currStateIndex, currState.getName() + " in " + currAutomaton.getName(),1);
 			}
 		}
 		
@@ -296,29 +296,43 @@ public class AutomataToIEC61499
 		// RESET algorithm resets the automata in the function block. In other words it
 		// makes automata enter the initial state.
 		pw.println("ALGORITHM RESET IN ST :");
-		pw.println("\tinitialized := FALSE;");
+		// Not needed anymore
+		//pw.println("\tinitialized := FALSE;");
 		// theAutomataToIEC1131.printInitializationStructureAsST(pw);
 		//pw.println("\n\t(* Set the initial state *)");
 		//pw.println("\tIF (NOT initialized)");
 		//pw.println("\tTHEN");
 
+		// Set all state variables to FALSE
 		for (Iterator autIt = theProject.iterator(); autIt.hasNext(); )
 		{
 			Automaton currAutomaton = (Automaton) autIt.next();
-			
+			int currAutomatonIndex = currAutomaton.getSynchIndex();
+
+			for (Iterator stateIt = currAutomaton.stateIterator(); stateIt.hasNext(); )
+			{
+				State currState = (State) stateIt.next();
+				int currStateIndex = currState.getSynchIndex();
+
+				pw.println("\tQ_" + currAutomatonIndex + "_" + currStateIndex + " := FALSE;");
+			}
+		}
+
+		// Then set the initital states to TRUE
+		for (Iterator autIt = theProject.iterator(); autIt.hasNext();)
+		{
+			Automaton currAutomaton = (Automaton) autIt.next();
+
 			if (currAutomaton.getInitialState() == null)
 			{
-				String errMessage = "AutomataToIEC61499.printAlgorithmDeclarations: "
-									+ "all automata must have an initial state but automaton "
-									+ currAutomaton.getName()
-									+ "doesn't";
[…]

### 2004-06-28 src/org/supremica/automata/IO/AutomataToIEC61499.java
GC

@@ -72,8 +72,6 @@ import org.supremica.automata.algorithms.SynchronizationType;
  * 
  */
 
-// TODO: Get rid of the internal variables
-// TODO: Make the generated function behave according to agreement
 // TODO: Make state variables int
 // TODO: Make comments optional
 
@@ -115,16 +113,7 @@ public class AutomataToIEC61499
 		//pw.println("(* Supremica version: " + org.supremica.Version.version() + "*)");
 		//pw.println("(* Time of generation: + DateFormat.getDateTimeInstance().format(new Date()) +" *)");
 				
-		pw.println("FUNCTION_BLOCK AUTOGEN_AUTOMATA_FUNCTION_BLOCK"
-					+ "\t"
-					+ "(* " 
-					+ "Supremica version: " 
-					+ org.supremica.Version.version()
-					+ "\t"
-					+ "Time of generation: " 
-					+ DateFormat.getDateTimeInstance().format(new Date()) 
-					+ " *)"
-					);
+		pw.println("FUNCTION_BLOCK AUTOGEN_AUTOMATA_FUNCTION_BLOCK"); // + "\t" + "(* " + "Supremica version: " + org.supremica.Version.version() + "\t" + "Time of generation: " + DateFormat.getDateTimeInstance().format(new Date()) + " *)");
 
 	}
 
@@ -181,12 +170,13 @@ public class AutomataToIEC61499
 		
 		
 		// The input_variable_list. Input variables represent the automaton events and are of the
-		// bool type. Only one can be true when the OCCURRED event happens.
+		// bool type. Only one should be TRUE when the OCCURRED event happens but if several are TRUE
+		// all of the transitions will take place. The user have to make shure that this is possible!
 		pw.println("VAR_INPUT");		
 		for (Iterator alphIt = allEvents.iterator(); alphIt.hasNext();)
 		{
 			LabeledEvent currEvent = (LabeledEvent) alphIt.next();
-			pw.println("\tEI_" + currEvent.getSynchIndex() + " : BOOL;\t(* " + currEvent.getLabel() + " *)");
+			pw.println("\tEI_" + currEvent.getSynchIndex() + " : BOOL;"); //\t(* " + currEvent.getLabel() + " *)");
 		}
 		pw.println("END_VAR");
 
@@ -210,17 +200,6 @@ public class AutomataToIEC61499
 	private void printInternalVariableList(PrintWriter pw) 
 	{
 		pw.println("VAR");
-
-		// Not needed acctually
[…]

### 2004-07-01 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

@@
 
-Introduce Automation Objects
-Define the different views (Model-View-Controller)
-	Configuration
-	Human-Machine Interface
-	Control
-	Simulation
+%Introduce Automation Objects
+
+%% Why automation objects? Background.
+Using only function blocks efficient hierarchical control
+system design is very time consumig. To facilitate easier
+and more rapid design process a new kind of design object
+needs to be introduced. Before that the purpose of that kind
+of object has to be defined.
+
+%% Definition. Multi view objects where the view is the
+%% interface exposed for different purposes.
+
+% Define the different views (Model-View-Controller) if relevant.
+%% Configuration
+%% Human-Machine Interface
+%% Control
+%% Simulation
+
+% Introduce the example

### 2004-07-01 doc/Simulation_and_Control_using_Supremica/functionblocks.tex
*** empty log message ***

@@
 
-New technologies and standards are emerging that are set to have a
-dramatic effect on the design and implementation of industrial
-control systems. A new standard IEC-61131-3~\cite{iec:1131:1993} was
-introduced in 1995 to standardize the representation of control
-designs so as to enable cross-vendor integration. The standard
-offered significant advantages over earlier vendor proprietary
+New technologies and standards are emerging that are set to
+have a dramatic effect on the design and implementation of
+industrial control systems. A new standard
+IEC-61131-3~\cite{iec:1131:1993} was introduced in 1995 to
+standardize the representation of control designs so as to
+enable cross-vendor integration. The standard offered
+significant advantages over earlier vendor proprietary
 control design representations~\cite{l:pro:1995}.
 \begin{itemize}
-\item It encouraged well-structured procedural
-programs that allowed the development of different functional
-elements into separate procedures or program units.
-\item It provided standardized languages and methods of program execution
-so that wide range of technological problems can be programmed as
-vendor-independent software elements.
-\item It provided support for various languages like structured
-text, function block diagrams, ladder diagrams, instruction list
-and sequential function charts.
-\item The standard enforced strong data typing and supported user defined
-data structures.
+\item It encouraged well-structured procedural programs that
+  allowed the development of different functional elements
+  into separate procedures or program units.
+\item It provided standardized languages and methods of
+  program execution so that wide range of technological
+  problems can be programmed as vendor-independent software
+  elements.
+\item It provided support for various languages like
+  structured text, function block diagrams, ladder diagrams,
+  instruction list and sequential function charts.
+\item The standard enforced strong data typing and supported
+  user defined data structures.
 \end{itemize}
@@
 
-\section{FBEditor}
-Function Block Development Kit (FBDK) ~\cite{c:fun:2002} is a
-software toolkit that enables users to build and test data types,
-function block types, resource types, device types and system
-configurations according to the IEC 61499 standard. FBEditor is a
-part of the FBDK toolkit that provides a development and runtime
-environment for applications based on the IEC 61499 standard.
-Figure \ref{f:FBEditor} shows the editor being used to load a
-function block used for controlling a motor.
+% \section{FBEditor}
+% Function Block Development Kit (FBDK) ~\cite{c:fun:2002} is a
+% software toolkit that enables users to build and test data types,
+% function block types, resource types, device types and system
+% configurations according to the IEC 61499 standard. FBEditor is a
+% part of the FBDK toolkit that provides a development and runtime
+% environment for applications based on the IEC 61499 standard.
+% Figure \ref{f:FBEditor} shows the editor being used to load a
+% function block used for controlling a motor.
 
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/FBEditor.eps, scale=.6}
-\caption[FBEditor]{FBEditor{\protect ~\cite{c:fun:2002}}}
-\label{f:FBEditor}
-\end{center}
-\end{figure}
-%\doublespace
-%
+% %
+% %\singlespace
+% \begin{figure}
+% \begin{center}
+% \epsfig{file=images/functionblocks/FBEditor.eps, scale=.6}
+% \caption[FBEditor]{FBEditor{\protect ~\cite{c:fun:2002}}}
+% \label{f:FBEditor}
+% \end{center}
+% \end{figure}
+% %\doublespace
+% %
 
-FBEditor enables the users to edit the block definitions through
-both a text editor and through a graphical user interface. New
-data types, basic function blocks, composite function blocks,
-execution control charts and other higher level entities may be
-created through the editor. The attributes of the blocks (data and
-events) may also be examined through a graphical user interface.
-The function blocks can be examined at the device level as shown
-in Figure \ref{f:Editor-Devices} or at the level of individual
-attributes like the connections, as shown in figure
-\ref{f:Editor-Mapping}.
+% FBEditor enables the users to edit the block definitions through
+% both a text editor and through a graphical user interface. New
+% data types, basic function blocks, composite function blocks,
+% execution control charts and other higher level entities may be
+% created through the editor. The attributes of the blocks (data and
+% events) may also be examined through a graphical user interface.
+% The function blocks can be examined at the device level as shown
+% in Figure \ref{f:Editor-Devices} or at the level of individual
+% attributes like the connections, as shown in figure
+% \ref{f:Editor-Mapping}.
 
-The editor supports the languages supported by the standard
-(Structured text, execution control chart etc.). It uses XML as
-the exchange format to ensure portability across various domains.
-Figure \ref{f:Editor-Versions} shows how the editor can also be
-used to manage meta properties like versions, owner entities,
-authors etc.
+% The editor supports the languages supported by the standard
+% (Structured text, execution control chart etc.). It uses XML as
+% the exchange format to ensure portability across various domains.
+% Figure \ref{f:Editor-Versions} shows how the editor can also be
+% used to manage meta properties like versions, owner entities,
+% authors etc.
 
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Editor-Devices.eps, scale=.6} \caption[FBEditor
-- Editing system devices]{FBEditor - Editing system
-devices{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Devices}
-\end{center}
-\end{figure}
-%\doublespace
-%
+% %
+% %\singlespace
+% \begin{figure}
+% \begin{center}
+% \epsfig{file=images/functionblocks/Editor-Devices.eps, scale=.6} \caption[FBEditor
+% - Editing system devices]{FBEditor - Editing system
+% devices{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Devices}
+% \end{center}
+% \end{figure}
+% %\doublespace
+% %
 
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Editor-Mapping.eps, scale=.6} \caption[FBEditor
-- Editing device connections]{FBEditor - Editing device
-connections{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Mapping}
-\end{center}
-\end{figure}
-%\doublespace
-%
+% %
+% %\singlespace
+% \begin{figure}
+% \begin{center}
+% \epsfig{file=images/functionblocks/Editor-Mapping.eps, scale=.6} \caption[FBEditor
+% - Editing device connections]{FBEditor - Editing device
+% connections{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Mapping}
+% \end{center}
+% \end{figure}
+% %\doublespace
+% %
 
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Editor-Versions.eps, scale=.6}
-\caption[FBEditor - Editing meta properties]{FBEditor - Editing
-meta properties{\protect ~\cite{c:fun:2002}}}
-\label{f:Editor-Versions}
-\end{center}
-\end{figure}
-%\doublespace
-%
-FBEditor has a navigation pane that allows the user to choose the
-display on the worksheet area, including the external interface;
-the Execution Control Chart (ECC) for basic function block type;
-service primitive sequences for service interface function block
-types; and function block networks for composite function block
-types, device and resource types and instances, and system
-configurations. The worksheet / viewport area contains the
-function block diagram for the currently selected instance.
-Different components of the diagram can be dragged around for
-better readability.
+% %
+% %\singlespace
+% \begin{figure}
+% \begin{center}
+% \epsfig{file=images/functionblocks/Editor-Versions.eps, scale=.6}
+% \caption[FBEditor - Editing meta properties]{FBEditor - Editing
+% meta properties{\protect ~\cite{c:fun:2002}}}
+% \label{f:Editor-Versions}
+% \end{center}
+% \end{figure}
+% %\doublespace
+% %
+% FBEditor has a navigation pane that allows the user to choose the
+% display on the worksheet area, including the external interface;
+% the Execution Control Chart (ECC) for basic function block type;
+% service primitive sequences for service interface function block
+% types; and function block networks for composite function block
+% types, device and resource types and instances, and system
+% configurations. The worksheet / viewport area contains the
+% function block diagram for the currently selected instance.
+% Different components of the diagram can be dragged around for
+% better readability.
 
-FBEditor has been developed in Java language. Hence in order to
-run simulations on applications written using the editor, one
-needs to generate the Java classes that correspond to the source
-files for the function blocks written in structured text (for
-instance). The editor has a code generation mechanism through
-which it parses through the source files and generates a Java
-class corresponding to each block.
+% FBEditor has been developed in Java language. Hence in order to
+% run simulations on applications written using the editor, one
+% needs to generate the Java classes that correspond to the source
+% files for the function blocks written in structured text (for
+% instance). The editor has a code generation mechanism through
+% which it parses through the source files and generates a Java
+% class corresponding to each block.
 
-On occurrence of certain input events, the user needs to specify
-the algorithms that should be scheduled to be executed and the
-output events that should signal the completion of algorithms.
-This may be done either by editing the generated Java files or by
-enclosing the algorithm inside especially marked tags in the
-source files. The files are then compiled and simulations can be
-run from within the editor environment. Figure
-\ref{f:Editor-Transitions} shows the assignment of output events
-on the transitions through the editing of the execution control
-chart.
+% On occurrence of certain input events, the user needs to specify
+% the algorithms that should be scheduled to be executed and the
+% output events that should signal the completion of algorithms.
+% This may be done either by editing the generated Java files or by
+% enclosing the algorithm inside especially marked tags in the
+% source files. The files are then compiled and simulations can be
+% run from within the editor environment. Figure
+% \ref{f:Editor-Transitions} shows the assignment of output events
+% on the transitions through the editing of the execution control
+% chart.
 
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Editor-Transitions.eps, scale=.6}
-\caption[FBEditor Editing the execution control chart]{FBEditor
-Editing the execution control chart{\protect ~\cite{c:fun:2002}}}
-\label{f:Editor-Transitions}
-\end{center}
-\end{figure}
+% %
+% %\singlespace
+% \begin{figure}
+% \begin{center}
+% \epsfig{file=images/functionblocks/Editor-Transitions.eps, scale=.6}
+% \caption[FBEditor Editing the execution control chart]{FBEditor
+% Editing the execution control chart{\protect ~\cite{c:fun:2002}}}
+% \label{f:Editor-Transitions}
+% \end{center}
+% \end{figure}
+
 %\doublespace

### 2004-07-01 src/org/supremica/automata/IO/AutomataToIEC61499.java

@@ -52,30 +52,24 @@ import java.io.*;
 import java.util.*;
 import java.text.DateFormat;
 
-
 import org.supremica.log.*;
 import org.supremica.automata.*;
 import org.supremica.util.SupremicaException;
 import org.supremica.automata.algorithms.SynchronizationType;
 
-/**
- * @author cengic
- *
- * This class generates an IEC 61499 function block implementing 
- * the automata in the current project.
+//TODO: Make state variables int
+
+/** This class generates an IEC-61499 function block 
+ * implementing the automata in the current project.
  *  
- * An important notice is that the event in the IEC-61499 is not the 
- * same as the automaton event, that is why the
- * implementation of the interface is done this way
+ * <p>Note that the event in the IEC-61499 is not the 
+ * same as an automaton event. That is why the
+ * implementation of the interface is done this way. KA 
+ * and I are planing several different interfaces so this 
+ * is only the first one.
  * 
- * KA and I are planing several different interfaces so this is only the first one.
- * 
- */
-
-// TODO: Make state variables int
-// TODO: Make comments optional
-
-
+ * @author Goran Cengic
+ */ 
 public class AutomataToIEC61499
 
 {
@@ -85,48 +79,54 @@ public class AutomataToIEC61499
 	private IEC61131Helper theHelper;
 	private Alphabet allEvents;
 	private SynchronizationType syncType = SynchronizationType.Prioritized;
+	private boolean comments = false;	
 	
+	public void commentsOn()
+	{
+		comments = true;
+	}
 	
+	public void commentsOff()
+	{
+		comments = false;
+	}
 	
 	// Constructor
 	public AutomataToIEC61499(Project theProject)
 	{
 		this.theProject = theProject;
 		this.theHelper = IEC61131Helper.getInstance();
-		this.initialize();
-	}
-
-	
-	
-	private void initialize()
-	{
-		allEvents = theProject.setIndicies();
+		allEvents = this.theProject.setIndicies();
+		commentsOn();
 	}
 
 
-
-	// Makes the beginning of the function block type declaration
+	/** Makes the beginning of the function block type declaration. */
 	private void printBeginProgram(PrintWriter pw)
 	{
 
[…]

### 2004-07-07 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

--- 2004-06-28 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
@@
 
-Introduce Automation Objects
-Define the different views (Model-View-Controller)
-	Configuration
-	Human-Machine Interface
-	Control
-	Simulation
+%% Background
+With function blocks the main technological problem of
+distributed automation has been solved. A general framework
+for development and deployment of distributed automation
+system has been developed and established. What is still
+needed is a way of encapsulating function blocks together
+with the data about their properties and, more importantly,
+their functionality in a way that will enable diminishing of
+the time to market and facilitate effective rapid object
+oriented development of automation systems.
+%% Why automation objects? What do we want to do?
+After considering the following arguments it will become
+clear why such encapsulation is needed and how it can help
+in achieving the afore mentioned goal.
+
+%%     effective reuse of software
+Object oriented system development as presented in
+\cite{iec:614991:2000} is not enough for achieving
+\textit{effective reuse of software} in automation systems
+development. This is the case because effective software
+reuse demands efficient exchange of information about how
+the software that is supposed to be reused is used and what
+it does. In other words, interfaces of the software objects
+have to be well specified and well documented. The
+\cite{iec:614991:2000} makes sure that the interfaces are
+well specified but it does not provide anything that will
+assure that they are well documented.
+
+%%     effective decentralization of software development
+Effective software reuse facilitates \textit{effective
+  decentralization of software development}. If anyone could
+use the software that somebody else developed without
+wasting time on resolving the difficulties that arise from
+poorly documented interfaces, or resorting to
+reimplementation because of the same problem, than everyone
+could concentrate on providing the best implementation for
+their part of the automation system software.
+
+%%     easier integration and deployment of software
+Leveraging effective software reuse and decentralized
+software development \textit{makes integration and
+  deployment much easier}. System integrator's task is
+hereby reduced to picking the right software objects from
+the library and connecting them in correct patterns for the
+specific application.
+
+%%     development of libraries
+Using the encapsulation of software objects with well
+specified information about their usage makes it possible to
+\textit{build libraries} of elements that can easily be
+shared over the Internet since everything that is needed for
+the usage of the elements comes with the elements
+themselves. This is unlike the Java system where the objects
+are distributed in binary only format and the interface
+documentation has to be downloaded or reviewed on-line prior
+to usage.
+
+%%     all of these result in rapid development of automation systems
+The arguments presented are enough to support the idea that
+the software objects encapsulated with the information about
+their usage will increase the speed of the development of
+the automation systems.
+
+After a formal definition of the encapsulation the benefits
+beyond the ones discussed above will be presented followed
+by the presentation about the implementation issues and the
+support in Supremica that is planned for such data objects.
+
+%% Definition. 
+%%     multi view objects
+%%     view is the interface exposed for different purposes.
+An \textbf{automation object} is defined as a set of
+applications, subapplications and function blocks (as
+defined in \cite{iec:614991:2000}) encapsulated with a
+declaration of \textit{attributes} and \textit{views} of
+these.
+
+An \textbf{attribute} is user defined and specified data that
+describes some significant property of the automation
+object.
+
+A \textbf{view} is a description of a certain aspect of the
+automation object and information about how that aspect is
+used in the automation system.
+% view is functional description and its name should only
+% contain what is needed to understand its function
+% eg elevator, pump, oven, sensor
+% all the information about what hardware this automation
+% block is for should be in the attributes
+% eg attribute hardware='abb robot N25'
+
+%% How this definition supports the arguments above
+
+%% Further benefits of the the automation objects that could
+%% not be described prior to definition
+%%     views facilitate several different implementations of
+%%     the same functionality for different hardware and it
+%%     is the hardware vendors that should supply these
+
+% Support in Supremica
+
+%% implementation of automation objects
+%%     archive of XML files
+%%     attribute and view declarations in MetaInformation.xml
+
+%% libraries 
+%%     non-hierarchical pool of automation objects
+%%     libraries  may be Internet repositories
+%%     searching and sorting by multiple attributes
+%%     searching and sorting by multiple views
+
+% Example of usage in Supremica. Three scenarios with the
+% lab equipment (kulbanan).
+% 1. Student does everything.
+% 2. Library is provided and the logic is done by the student.
+% 3. Library is proviede and the logic is synthesised by the
+%    Supremica.

### 2004-07-07 src/org/supremica/automata/IO/AutomataToIEC61499.java

@@ -69,6 +69,7 @@ import org.supremica.automata.algorithms.SynchronizationType;
  * is only the first one.
  * 
  * @author Goran Cengic
+ * @author cengic@s2.chalmers.se
  */ 
 public class AutomataToIEC61499
 

### 2004-07-16 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

@@
 
-%% Background
-With function blocks the main technological problem of
+%% Intro
+With the function blocks the main technological problem of
 distributed automation has been solved. A general framework
@@
 oriented development of automation systems.
-%% Why automation objects? What do we want to do?
-After considering the following arguments it will become
-clear why such encapsulation is needed and how it can help
-in achieving the afore mentioned goal.
+
+\section{Background}
+%% Background. Why automation objects? What do we want to do?
+Considering the following arguments it is clear why such
+encapsulation is needed and how it can help in achieving the
+afore mentioned goal.
 
 %%     effective reuse of software
-Object oriented system development as presented in
-\cite{iec:614991:2000} is not enough for achieving
-\textit{effective reuse of software} in automation systems
-development. This is the case because effective software
-reuse demands efficient exchange of information about how
-the software that is supposed to be reused is used and what
-it does. In other words, interfaces of the software objects
-have to be well specified and well documented. The
-\cite{iec:614991:2000} makes sure that the interfaces are
-well specified but it does not provide anything that will
-assure that they are well documented.
+Object oriented system development as presented in the
+IEC~61499 standard is ideal for the purpose that is is
+designed for but not enough for achieving {\it effective
+  software reuse} in automation systems development.
+
+% grouping software objects in functional units (encapsulation)
+% abstraction of functionality (views)
+% properties of the functional unit (attributes)
 
@@
 
+\section{Definition}
 %% Definition. 
 %%     multi view objects
-%%     view is the interface exposed for different purposes.
-An \textbf{automation object} is defined as a set of
-applications, subapplications and function blocks (as
-defined in \cite{iec:614991:2000}) encapsulated with a
-declaration of \textit{attributes} and \textit{views} of
-these.
+An \textbf{automation object} is defined as \textit{a set of
+  software objects} encapsulated with a declaration of
+\textit{services}, \textit{attributes} and \textit{views} of
+the set.
 
-An \textbf{attribute} is user defined and specified data that
+The \textbf{set} specifies the applications, subapplications
+and function blocks, as defined in \cite{iec:614991:2000},
+contained within the automation object. Only one set per
+automation object is defined and it contains \textit{all} of
+the software objects required for the functionality of the
+automation object.
+% less dependency problems at the cost of storage, more on
+% this after the definition
+
+An \textbf{attribute} is a user defined and specified data that
 describes some significant property of the automation
@@
 
-A \textbf{view} is a description of a certain aspect of the
-automation object and information about how that aspect is
-used in the automation system.
-% view is functional description and its name should only
-% contain what is needed to understand its function
-% eg elevator, pump, oven, sensor
-% all the information about what hardware this automation
-% block is for should be in the attributes
-% eg attribute hardware='abb robot N25'
+A \textbf{view} is an interface to the object. Every
+interface has at least one implementation.
+
+A \textbf{service}
 
@@
 
+\section{Implementation and Support in Supremica}
 % Support in Supremica

### 2004-07-16 doc/Simulation_and_Control_using_Supremica/functionblocks.tex
*** empty log message ***

--- 2004-06-28 doc/Simulation_and_Control_using_Supremica/functionblocks.tex
@@
-\chapter{IEC 61499 Function Block Standard}
+\chapter{IEC~61499 Standard}
 \label{functionblocks}
 
-\begin{bf}
-{This chapter is with minor revisions copied from chapter 2 in \cite{j:des:2002}.}
-\end{bf}
+\textbf{This chapter,with minor revisions, is copied from
+  chapter 2 in \cite{j:des:2002}.}
 
-New technologies and standards are emerging that are set to have a
-dramatic effect on the design and implementation of industrial
-control systems. A new standard IEC-61131-3~\cite{iec:1131:1993} was
-introduced in 1995 to standardize the representation of control
-designs so as to enable cross-vendor integration. The standard
-offered significant advantages over earlier vendor proprietary
+New technologies and standards are emerging that are set to
+have a dramatic effect on the design and implementation of
+industrial control systems. A new standard
+IEC~61131-3~\cite{iec:1131:1993} was introduced in 1995 to
+standardize the representation of control designs so as to
+enable cross-vendor integration. The standard offered
+significant advantages over earlier vendor proprietary
 control design representations~\cite{l:pro:1995}.
 \begin{itemize}
-\item It encouraged well-structured procedural
-programs that allowed the development of different functional
-elements into separate procedures or program units.
-\item It provided standardized languages and methods of program execution
-so that wide range of technological problems can be programmed as
-vendor-independent software elements.
-\item It provided support for various languages like structured
-text, function block diagrams, ladder diagrams, instruction list
-and sequential function charts.
-\item The standard enforced strong data typing and supported user defined
-data structures.
+\item It encouraged well-structured procedural programs that
+  allowed the development of different functional elements
+  into separate procedures or program units.
+\item It provided standardized languages and methods of
+  program execution so that wide range of technological
+  problems can be programmed as vendor-independent software
+  elements.
+\item It provided support for various languages like
+  structured text, function block diagrams, ladder diagrams,
+  instruction list and sequential function charts.
+\item The standard enforced strong data typing and supported
+  user defined data structures.
 \end{itemize}
 
-The standard however was based on a scanning mechanism where the
-program execution was monitored by a continuous loop. Although
-different parts of the program could be scanned at different
-frequencies (based on fidelity requirements), it was tedious to
-describe such behavior using sequential function charts. The
-standard was based on polling mechanism rather than an event
-driven system, and hence its execution order is not clearly
-defined.
+The standard however was based on a scanning mechanism where
+the program execution was monitored by a continuous loop.
+Although different parts of the program could be scanned at
+different frequencies (based on fidelity requirements), it
+was tedious to describe such behavior using sequential
+function charts. The standard was based on polling mechanism
+rather than an event driven system, and hence its execution
+order is not clearly defined.
 
-Further the standard was designed to support a software model and
-languages for PLCs whose software was typically running on a
-single processing resource. The communication between different
-resources was supported through global variables and communication
-blocks, both of which did not provide a clear and concise method
-for defining the connections between distributed systems.
+Further the standard was designed to support a software
+model and languages for PLCs whose software was typically
+running on a single processing resource. The communication
+between different resources was supported through global
+variables and communication blocks, both of which did not
+provide a clear and concise method for defining the
+connections between distributed systems.
 
-To overcome these problems, a function block model was sought that
-would support distributed control systems in a scalable,
-extensible and flexible manner. IEC-61499~\cite{iec:61499:2000}
-standard has now been developed for distributed industrial
-process, measurement and control systems.
+To overcome these problems, a function block model was
+sought that would support distributed control systems in a
+scalable, extensible and flexible manner.
+IEC~61499-1~\cite{iec:614991:2000} standard has now been
+developed for distributed industrial process, measurement
+and control systems.
 
-In industrial systems, function blocks are now an established
-concept for defining robust, re-usable software components. A {\it
-Function Block} instance is defined to a software function unit
-comprising an individual, named copy of a data structure and
-associated operations specified by a corresponding function type.
+In industrial systems, function blocks are now an
+established concept for defining robust, re-usable software
+components. A {\it Function Block} instance is defined to a
+software function unit comprising an individual, named copy
+of a data structure and associated operations specified by a
+corresponding function type.
 
@@
 encapsulated functionality in the form of blocks and link it
-together in the quickest and most intuitive way possible. They
-reduce the complexity since system integrators can now build large
-systems using {\it black-box composition} i.e. without worrying
-about how the internals of a block work.
+together in the quickest and most intuitive way possible.
+They reduce the complexity since system integrators can now
+build large systems using {\it black-box composition} i.e.
+without worrying about how the internals of a block work.
 
 The following sections are adapted largely from the various
-discussions on the features of the standard in ~\cite{iec:61499:2000,
-iec:1131:1993, l:pro:1995, l:mod:2001, c:des:2002, c:ope:2002}
+discussions on the features of the standard in
+~\cite{iec:614991:2000, iec:1131:1993, l:pro:1995,
+  l:mod:2001, c:des:2002, c:ope:2002}
 
 \section{Reference Models}
-IEC-61499 has provided a set of models for describing distributed
-systems that have been programmed using function blocks.
-\begin{itemize}
-\item {\bf System Model:} At the physical level, a distributed
-system consists of a set of co-operating {\it applications}. An
-application can exist on a single system such as application C in
-figure \ref{f:System_Model} or have functionality distributed over
+IEC~61499-1 has provided a set of models for describing
+distributed systems that have been programmed using function
+blocks.
+
+\subsection{System Model} 
+At the physical level, a distributed system consists of a
+set of co-operating {\it applications}. An application can
+exist on a single system such as application C in figure
+\ref{f:System_Model} or have functionality distributed over
 a number of devices such as applications A and B in Figure
@@
 
-\item {\bf Device Model:} A resource provides independent
-execution and control of networks of function blocks. It supports
-zero or more {\it resources} and has at least one interface. The
-interface could either be a 'process interface' that allows the
-resources to exchange data with the input and output points on the
-physical devices, or a communication interface that allows the
-resources to communicate with external resources in remote
-devices. The device model is illustrated in Figure
+\begin{figure}
+  \begin{center}
+    \epsfig{file=images/functionblocks/System_Model.eps,scale=.8} 
+    \caption[System Model]
+            {System Model{\protect ~\cite{iec:614991:2000}}}
+    \label{f:System_Model}
+  \end{center}
+\end{figure}
+
+\subsection{Device Model} 
+A resource provides independent execution and control of
+networks of function blocks. It supports zero or more {\it
+  resources} and has at least one interface. The interface
+could either be a 'process interface' that allows the
+resources to exchange data with the input and output points
+on the physical devices, or a communication interface that
+allows the resources to communicate with external resources
+in remote devices. The device model is illustrated in Figure
 \ref{f:Device_Model}.
 
-\item {\bf Resource Model:} A resource is an independent functional unit,
-contained in a device which has independent control of its
-operation. The resource can be created, configured, parameterized,
-started up, deleted etc. without effecting other resources. Each
-resource provides an interface to the communication systems and
-tot the device specific process. The resource is therefore,
-concerned with the mapping of data and event flows which pass
-between function blocks in the local resource to remove resource
-function blocks via the device communication interfaces.
+\begin{figure}
+  \begin{center}
+    \epsfig{file=images/functionblocks/Device_Model.eps,scale=.8} 
+    \caption[Device Model]
+            {Device Model{\protect ~\cite{iec:614991:2000}}}
+    \label{f:Device_Model}
+  \end{center}
+\end{figure}
 
-Figure \ref{f:Resource_Model} depicts the main features of an IEC
-61499 resource. Within the resource, it shows a network of
-interconnected function blocks linked by data and event flows. A
-scheduling function ensures that the events are executed in the
-correct serial order. {\it Service Interface} function blocks are
-a special form of function blocks that provide a link between
-function blocks and the interface of the resource.
+\subsection{Resource Model} 
+A resource is an independent functional unit, contained in a
+device which has independent control of its operation. The
+resource can be created, configured, parameterized, started
+up, deleted etc. without effecting other resources. Each
+resource provides an interface to the communication systems
+and tot the device specific process. The resource is
+therefore, concerned with the mapping of data and event
+flows which pass between function blocks in the local
+resource to remove resource function blocks via the device
+communication interfaces.
 
-\item {\bf Application Model:} An application is defined as a
-network of interconnected function blocks, linked by data and
-event flows. An application can be fragmented and distributed over
-many resources. The resources ensure that the events are executed
-in the correct serial order based on the causal relationships
-specified by the applications.
+Figure \ref{f:Resource_Model} depicts the main features of
+an IEC 61499 resource. Within the resource, it shows a
+network of interconnected function blocks linked by data and
+event flows. A scheduling function ensures that the events
+are executed in the correct serial order. {\it Service
+  Interface} function blocks are a special form of function
+blocks that provide a link between function blocks and the
+interface of the resource.
 
-An application could be understood as an entire set of function
-blocks and interconnections to solve a particular automation
-control problem. Figure \ref{f:Application_Model} illustrates an
-application, where each individual block is either a function
-block or a subapplication.
+\begin{figure}
+  \begin{center}
+    \epsfig{file=images/functionblocks/Resource_Model.eps,scale=.8} 
+    \caption[Resource Model]
+            {Resource Model{\protect ~\cite{iec:614991:2000}}}
+    \label{f:Resource_Model}
+  \end{center}
+\end{figure}
 
-\item {\bf Function Block Model:} A function block is a functional
-unit of software that its own data structure which can be
-manipulated by one or more algorithms. The formal description of
-the data structure is provided by the function block type
-definition.
+\subsection{Application Model} 
+An application is defined as a network of interconnected
+function blocks, linked by data and event flows. An
+application can be fragmented and distributed over many
+resources. The resources ensure that the events are executed
+in the correct serial order based on the causal
+relationships specified by the applications.
+
+An application could be understood as an entire set of
+function blocks and interconnections to solve a particular
+automation control problem. Figure \ref{f:Application_Model}
+illustrates an application, where each individual block is
+either a function block or a subapplication.
+
+\begin{figure}
+  \begin{center}
+    \epsfig{file=images/functionblocks/Application_Model.eps, scale=.8}
+    \caption[Application Model]
+            {Application Model{\protect ~\cite{iec:614991:2000}}} 
+    \label{f:Application_Model}
+  \end{center}
+\end{figure}
+
+\subsection{Function Block Model} 
+A function block is a functional unit of software that its
+own data structure which can be manipulated by one or more
+algorithms. The formal description of the data structure is
+provided by the function block type definition.
 
@@
 \begin{itemize}
-\item Each function block has a type name and an instance name.
-\item Each block ha a set of event inputs, which can receive
-inputs from other blocks through event connections.
-\item Each block has one or more outputs, through which it can
-propagate its results to other blocks.
-\item There is a set of data inputs that allow values to be passed
-in from other blocks.
-\item There is a set of data outputs, through which the block can
-pass its results to other blocks.
-\item Each block may have a set of internal variables that hold
-its state between algorithm invocations.
-\item The behavior of the block is defined in terms of algorithms
-and state information. This behavior is defined in the function
-block type definition.
+\item Each function block has a type name and an instance
+  name.
+\item Each block ha a set of event inputs, which can
+  receive inputs from other blocks through event
+  connections.
+\item Each block has one or more outputs, through which it
+  can propagate its results to other blocks.
+\item There is a set of data inputs that allow values to
+  be passed in from other blocks.
+\item There is a set of data outputs, through which the
+  block can pass its results to other blocks.
+\item Each block may have a set of internal variables that
+  hold its state between algorithm invocations.
+\item The behavior of the block is defined in terms of
+  algorithms and state information. This behavior is
+  defined in the function block type definition.
 \end{itemize}
 
-Figure \ref{f:Function_Block_Char} shows the main characteristics
-of a function block. The top part of the block, called the
-'execution control' portion contains in some cases, a state
-machine that maps events to the algorithms that should be
-triggered as a consequence of the event. The lower portion of
-block contains both the algorithms and the internal data, both of
-which are encapsulated within the block.
+Figure \ref{f:Function_Block_Char} shows the main
+characteristics of a function block. The top part of the
+block, called the 'execution control' portion contains in
+some cases, a state machine that maps events to the
+algorithms that should be triggered as a consequence of the
+event. The lower portion of block contains both the
+algorithms and the internal data, both of which are
+encapsulated within the block.
 
-\end{itemize}
-
-%
-%%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/System_Model.eps, scale=.8} \caption[System
-Model]{System Model {\protect ~\cite{iec:61499:2000}}}
-\label{f:System_Model}
-\end{center}
-\end{figure}
-%%\doublespace
-%
-
-%
-%%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Device_Model.eps, scale=.8} \caption[Device
-Model]{Device Model{\protect ~\cite{iec:61499:2000}}}
-\label{f:Device_Model}
-\end{center}
-\end{figure}
-%%\doublespace
-%
-
-%
-%%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Resource_Model.eps, scale=.8} \caption[Resource
-Model]{Resource Model{\protect ~\cite{iec:61499:2000}}}
-\label{f:Resource_Model}
-\end{center}
-\end{figure}
-%%\doublespace
-%
-
-%
-%%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Application_Model.eps, scale=.8}
-\caption[Application Model]{Application Model{\protect
-~\cite{iec:61499:2000}}} \label{f:Application_Model}
-\end{center}
-\end{figure}
-%%\doublespace
-%
-
-%
-%%\singlespace
 \begin{figure}
@@
     \epsfig{file=images/functionblocks/Function_Block_Char.eps, scale=.8}
-\caption[Function Block Characteristics]{Function Block
-Characteristics{\protect ~\cite{iec:61499:2000}}}
+    \caption[Function Block Characteristics]
+            {Function Block Characteristics{\protect ~\cite{iec:614991:2000}}}
     \label{f:Function_Block_Char}
@@
 \end{figure}
-%%\doublespace
-%
 
 \section{Function Block Types}
+A function block type is defined by a type name, formal
+definitions for the input events and data and the output
+events and data. The type definition also includes the
+internal behavior of the block. The notion of a function
+block type is analogous to the class definition in
+object-oriented programming while function block instances
+correspond to the object instances of the classes.
 
-A function block type is defined by a type name, formal
-definitions for the input events and data and the output events
-and data. The type definition also includes the internal behavior
-of the block. The notion of a function block type is analogous to
-the class definition in object-oriented programming while function
-block instances correspond to the object instances of the classes.
+\subsection{Basic Function Block Types} 
+The behavior of basic function blocks is defined in terms of
+algorithms that are invoked in response to input events. As
+algorithms execute, they trigger output events to signal
+their state to other blocks. The basic blocks hold their
+state in terms of internal variables. Basic function blocks
+also can communicate their state using input and output data
+variables.
 
-\begin{itemize}
-\item{\bf Basic Function Block Types:} The behavior of basic
-function blocks is defined in terms of algorithms that are invoked
-in response to input events. As algorithms execute, they trigger
-output events to signal their state to other blocks. The basic
-blocks hold their state in terms of internal variables. Basic
-function blocks also can communicate their state using input and
-output data variables.
+The mapping of events onto algorithms is expressed using a
+state transition notation known as execution control chart
+and is described in the next section. Figure
+\ref{f:Basic_Block} shows an example of a basic function
+block.
 
-The mapping of events onto algorithms is expressed using a state
-transition notation known as execution control chart and is
-described in the next section. Figure \ref{f:Basic_Block} shows an
-example of a basic function block.
+\begin{figure}
+  \begin{center}
+    \epsfig{file=images/functionblocks/Basic_Function_Block.eps,scale=.8}
+    \caption[Basic Function Block]
+            {Basic Function Block{\protect ~\cite{iec:614991:2000}}} 
+    \label{f:Basic_Block}
+  \end{center}
+\end{figure}
 
-\item{\bf Composite Function Blocks:} The behavior of a composite
-function block is described in terms of the event and data
-connections between its component function blocks. A composite
-function block can be used to hide the internal interactions
-between the component blocks and presenting a simplified interface
-for use by the outside world. A composite function block is
-illustrated in Figure \ref{f:Composite_Block}.
+\subsection{Composite Function Blocks} 
+The behavior of a composite function block is described in
+terms of the event and data connections between its
+component function blocks. A composite function block can be
+used to hide the internal interactions between the component
+blocks and presenting a simplified interface for use by the
+outside world. A composite function block is illustrated in
+Figure \ref{f:Composite_Block}.
 
-\item {\bf Service Interface Function Blocks:} Service interface
-function blocks provide an interface between the function block
-domain and the external services. Since a service interface
-function block is primarily concerned with data exchange, it is
-defined using time sequence diagrams. Examples of service
-interface function blocks are communication blocks for publishing
-and subscribing over a network and management blocks for
+\begin{figure}
+  \begin{center}
+    \epsfig{file=images/functionblocks/Composite_Function_Block.eps, scale=.8}
+    \caption[Composite Function Block]
+            {Composite Function Block{\protect ~\cite{iec:614991:2000}}} 
+        \label{f:Composite_Block}
+  \end{center}
+\end{figure}
+
+\subsection{Service Interface Function Blocks} 
+Service interface function blocks provide an interface
+between the function block domain and the external services.
+Since a service interface function block is primarily
+concerned with data exchange, it is defined using time
+sequence diagrams. Examples of service interface function
+blocks are communication blocks for publishing and
+subscribing over a network and management blocks for
 dynamically creating component blocks and event connections.
 
-\item {\bf Event Function Blocks:} Event function blocks are some
-special basic function blocks defined in the standard to ease the
-management of events. These blocks are used for control,
-generation and detection of events and can be used by other
-composite blocks. These blocks are defined using an execution
-control chart or a time sequence diagram. These blocks can be
-used, for instance, for the following purposes.
+\subsection{Event Function Blocks} 
+Event function blocks are some special basic function blocks
+defined in the standard to ease the management of events.
+These blocks are used for control, generation and detection
+of events and can be used by other composite blocks. These
+blocks are defined using an execution control chart or a
+time sequence diagram. These blocks can be used for the
+following purposes:
 \begin{enumerate}
-\item To split events to produce new events
+\item To split events to produce new events.
 \item To allow events to be merged.
@@
 
-\end{itemize}
+\section{Execution Model}
+The execution of algorithms for basic function blocks is
+invoked by the execution control portion of a function block
+instance in response to events at event inputs. This
+invocation takes the form of a request to the underlying
+scheduling function of the associated resource to schedule
+the execution of algorithm's operations. The model assumes
+that the resource in which a function block exists provides
+a scheduling function that ensures each phase of function
+block execution occurs in the correct order and with the
+correct priority.
 
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Basic_Function_Block.eps, scale=.8}
-\caption[Basic Function Block]{Basic Function Block{\protect
-~\cite{iec:61499:2000}}} \label{f:Basic_Block}
-\end{center}
-\end{figure}
-%\doublespace
-%
-
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Composite_Function_Block.eps, scale=.8}
-\caption[Composite Function Block]{Composite Function
-Block{\protect ~\cite{iec:61499:2000}}} \label{f:Composite_Block}
-\end{center}
-\end{figure}
-%\doublespace
-%
-
-\section{Execution Model}
-
-The execution of algorithms for basic function blocks is invoked
-by the execution control portion of a function block instance in
-response to events at event inputs. This invocation takes the form
-of a request to the underlying scheduling function of the
-associated resource to schedule the execution of algorithm's
-operations. The model assumes that the resource in which a
-function block exists provides a scheduling function that ensures
-each phase of function block execution occurs in the correct order
-and with the correct priority.
-
-The execution model specifies eight discrete timing points as
-shown in Figure \ref{f:Execution_Model} that must occur
-sequentially for the basic function block to operate. These are:
+The execution model specifies eight discrete timing points
+as shown in Figure \ref{f:Execution_Model} that must occur
+sequentially for the basic function block to operate. These
+are:
 
 \begin{description}
-\item[$t_1$]: Values coming from external function blocks
-as input data are stable at the current block.
-\item[$t_2$]: The event associated with the input values occurs.
-\item[$t_3$]: The execution control function notifies the resource
-scheduling function that it has the input values and is ready to
-execute the algorithm.
-\item[$t_4$]: The scheduling function begins the algorithm execution.
-\item[$t_5$]: The algorithm processes input values, (and in some
-cases, internal variables) to create new output values that are
-written to the block's outputs.
-\item[$t_6$]: The algorithm notifies the scheduling function that it
-has completed its execution.
-\item[$t_7$]: The scheduling function invokes the execution control
-to generate the output events.
-\item[$t_8$]: The execution control generates the appropriate output
-events at its event output ports. The event may be used by
-downstream blocks to signal that they may use the output values
-generated by this block.
+\item[$t_1$]: Values coming from external function blocks as
+  input data are stable at the current block.
+\item[$t_2$]: The event associated with the input values
+  occurs.
+\item[$t_3$]: The execution control function notifies the
+  resource scheduling function that it has the input values
+  and is ready to execute the algorithm.
+\item[$t_4$]: The scheduling function begins the algorithm
+  execution.
+\item[$t_5$]: The algorithm processes input values, (and in
+  some cases, internal variables) to create new output
+  values that are written to the block's outputs.
+\item[$t_6$]: The algorithm notifies the scheduling function
+  that it has completed its execution.
+\item[$t_7$]: The scheduling function invokes the execution
+  control to generate the output events.
+\item[$t_8$]: The execution control generates the
+  appropriate output events at its event output ports. The
+  event may be used by downstream blocks to signal that they
+  may use the output values generated by this block.
 \end{description}
 
-%
-%\singlespace
 \begin{figure}
@@
     \epsfig{file=images/functionblocks/Execution_Control_Chart.eps, scale=.8}
-\caption[Execution Model]{Execution Model{\protect
-~\cite{iec:61499:2000}}} \label{f:Execution_Model}
+    \caption[Execution Model]
+            {Execution Model{\protect ~\cite{iec:614991:2000}}}
+    \label{f:Execution_Model}
   \end{center}
 \end{figure}
-%\doublespace
-%
 
-To ensure consistency, the algorithms must use the snapshot of
-input variables that exists when the input event arrived. Further
-the scheduling function must be able to detect the arrival of
-input events at a faster rate than the rate at which they can be
-handled by the block.
-
-\section{FBEditor}
-Function Block Development Kit (FBDK) ~\cite{c:fun:2002} is a
-software toolkit that enables users to build and test data types,
-function block types, resource types, device types and system
-configurations according to the IEC 61499 standard. FBEditor is a
-part of the FBDK toolkit that provides a development and runtime
-environment for applications based on the IEC 61499 standard.
-Figure \ref{f:FBEditor} shows the editor being used to load a
-function block used for controlling a motor.
-
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/FBEditor.eps, scale=.6}
-\caption[FBEditor]{FBEditor{\protect ~\cite{c:fun:2002}}}
-\label{f:FBEditor}
-\end{center}
-\end{figure}
-%\doublespace
-%
-
-FBEditor enables the users to edit the block definitions through
-both a text editor and through a graphical user interface. New
-data types, basic function blocks, composite function blocks,
-execution control charts and other higher level entities may be
-created through the editor. The attributes of the blocks (data and
-events) may also be examined through a graphical user interface.
-The function blocks can be examined at the device level as shown
-in Figure \ref{f:Editor-Devices} or at the level of individual
-attributes like the connections, as shown in figure
-\ref{f:Editor-Mapping}.
-
-The editor supports the languages supported by the standard
-(Structured text, execution control chart etc.). It uses XML as
-the exchange format to ensure portability across various domains.
-Figure \ref{f:Editor-Versions} shows how the editor can also be
-used to manage meta properties like versions, owner entities,
-authors etc.
-
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Editor-Devices.eps, scale=.6} \caption[FBEditor
-- Editing system devices]{FBEditor - Editing system
-devices{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Devices}
-\end{center}
-\end{figure}
-%\doublespace
-%
-
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Editor-Mapping.eps, scale=.6} \caption[FBEditor
-- Editing device connections]{FBEditor - Editing device
-connections{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Mapping}
-\end{center}
-\end{figure}
-%\doublespace
-%
-
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Editor-Versions.eps, scale=.6}
-\caption[FBEditor - Editing meta properties]{FBEditor - Editing
-meta properties{\protect ~\cite{c:fun:2002}}}
-\label{f:Editor-Versions}
-\end{center}
-\end{figure}
-%\doublespace
-%
-FBEditor has a navigation pane that allows the user to choose the
-display on the worksheet area, including the external interface;
-the Execution Control Chart (ECC) for basic function block type;
-service primitive sequences for service interface function block
-types; and function block networks for composite function block
-types, device and resource types and instances, and system
-configurations. The worksheet / viewport area contains the
-function block diagram for the currently selected instance.
-Different components of the diagram can be dragged around for
-better readability.
-
-FBEditor has been developed in Java language. Hence in order to
-run simulations on applications written using the editor, one
-needs to generate the Java classes that correspond to the source
-files for the function blocks written in structured text (for
-instance). The editor has a code generation mechanism through
-which it parses through the source files and generates a Java
-class corresponding to each block.
-
-On occurrence of certain input events, the user needs to specify
-the algorithms that should be scheduled to be executed and the
-output events that should signal the completion of algorithms.
-This may be done either by editing the generated Java files or by
-enclosing the algorithm inside especially marked tags in the
-source files. The files are then compiled and simulations can be
-run from within the editor environment. Figure
-\ref{f:Editor-Transitions} shows the assignment of output events
-on the transitions through the editing of the execution control
-chart.
-
-%
-%\singlespace
-\begin{figure}
-\begin{center}
-\epsfig{file=images/functionblocks/Editor-Transitions.eps, scale=.6}
-\caption[FBEditor Editing the execution control chart]{FBEditor
-Editing the execution control chart{\protect ~\cite{c:fun:2002}}}
-\label{f:Editor-Transitions}
-\end{center}
-\end{figure}
-%\doublespace
-%
-
-%\section{EMBench and Function Blocks}
-
-%Since the FBEditor provided the aforementioned functionality, it
-%was decided that the EMBench should leverage it. In particular,
-%the bench relies on the editor for two basic features.
-%\begin{enumerate}
-%\item To generate the Java files from the IEC 61499 source files.
-%\item To manage the Java thread that initializes the blocks during
-%simulation.
-%\end{enumerate}
-
-%It has been visualized that the bench will eventually have the
-%editor as one of its components. The bench explicitly manages the
-%design activities of the user and generates the function block
-%definitions for the user design. It then requests the FBEditor
-%component to load the specified source files, generate the Java
-%classes and run the simulation. In addition, the user may make
-%behavioral changes to the blocks by editing the events, if he/she
-%so desires. The editor conforms to the published service
-%specifications of the bench as explained in chapter \ref{chap3}
-%and hence would present the same interface to the client.
-
-%Currently, however, the FBEditor is not a part of the bench
-%because it does not expose an API. Since it is external to the
-%bench, the bench has to circumvent this problem by allowing the
-%editor to be executed as a separate application and delegating the
-%simulation task completely to the editor. The data exchange
-%necessary between the two applications is handled via network
-%communication. This issue is explained in a greater detail in
-%chapter~\ref{chap4} and chapter~\ref{chap5}.
+To ensure consistency, the algorithms must use the snapshot
+of input variables that exists when the input event arrived.
+Further the scheduling function must be able to detect the
+arrival of input events at a faster rate than the rate at
+which they can be handled by the block.

### 2004-11-09 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

@@
 for development and deployment of distributed automation
-system has been developed and established. What is still
-needed is a way of encapsulating function blocks together
-with the data about their properties and, more importantly,
-their functionality in a way that will enable diminishing of
-the time to market and facilitate effective rapid object
-oriented development of automation systems.
+systems has been developed. What is still needed is a way of
+encapsulating function blocks together with the data about
+their properties and, more importantly, their functionality
+in a way that will enable diminishing the time to market
+and facilitate effective rapid object oriented development
+of automation systems.
 
@@
 Object oriented system development as presented in the
-IEC~61499 standard is ideal for the purpose that is is
-designed for but not enough for achieving {\it effective
+IEC~61499 standard is ideal for the purpose that it is
+designed for but not enough for achieving \emph{effective
   software reuse} in automation systems development.
-
-% grouping software objects in functional units (encapsulation)
-% abstraction of functionality (views)
-% properties of the functional unit (attributes)
 
@@
 hereby reduced to picking the right software objects from
-the library and connecting them in correct patterns for the
-specific application.
+the library and connecting them for the specific
+application.
 
@@
 
-%%     all of these result in rapid development of automation systems
+%%     all of these result in rapid development of
+%%     automation systems
 The arguments presented are enough to support the idea that
@@
 
-After a formal definition of the encapsulation the benefits
-beyond the ones discussed above will be presented followed
-by the presentation about the implementation issues and the
-support in Supremica that is planned for such data objects.
+%After a specification and a formal definition of the
+%encapsulation the benefits beyond the ones discussed above
+%will be presented followed by the presentation about the
+%implementation issues and the support in Supremica that is
+%planned for such data objects.
 
-\section{Definition}
-%% Definition. 
-%%     multi view objects
-An \textbf{automation object} is defined as \textit{a set of
-  software objects} encapsulated with a declaration of
-\textit{services}, \textit{attributes} and \textit{views} of
-the set.
+\section{Specification}
+To realize all of the above mentioned arguments a new kind
+of software objects are needed, preliminary called
+\emph{automation objects}. Even though this kind of objects
+are still very much a research subject the following is the
+specification of the objects so far.
 
-The \textbf{set} specifies the applications, subapplications
-and function blocks, as defined in \cite{iec:614991:2000},
-contained within the automation object. Only one set per
-automation object is defined and it contains \textit{all} of
-the software objects required for the functionality of the
-automation object.
-% less dependency problems at the cost of storage, more on
-% this after the definition
+%Automation Object:
+%* consists of models, controllers, views and
+%  configuration registry
+Automation objects consists of models, views, controllers
+and configuration registry implemented using IEC-61499
+function blocks and their inteerfaces. Models are
+interchangable simulation models with varaying level of
+abstraction. One model might be used for the simulation on a
+high level of abstraction, typicaly behavioural model, and
+another for the simulation on a lower level. This is then
+used to implement dynamic zooming during the simulation. One
+of the models will implement the communication with the
+controlled equpiment thus allowing effortless switching
+between simulation and acutal control. This will enable
+gradual verification and on-line tuning of the big
+automation systems by allowing the simulation of the system
+components that are being configured or tuned while running
+the components that are already configured.
 
-An \textbf{attribute} is a user defined and specified data that
-describes some significant property of the automation
-object.
+Controllers are function blocks that control the model. The
+best examples of controllers are regulators as PI- or
+PID-regulator. On the higher level of the automation system
+these can be configurable implementations of different
+automation algorithms. This allows interchange of different
+controlling strategies for easy verification with simulation
+and the equipment.
 
-A \textbf{view} is an interface to the object. Every
-interface has at least one implementation.
+Views are the function blocks allowing user interaction with
+the automation object. These implement all the alowable ways
+of interaction with the automation object. At the lower
+levels of the automation system they will typically provide
+different event and data interfaces depending on the user's
+needs while at the higher leve they will implement full
+graphical user interfaces for the operators.
 
-A \textbf{service}
-
-%% How this definition supports the arguments above
-
-%% Further benefits of the the automation objects that could
-%% not be described prior to definition
-%%     views facilitate several different implementations of
-%%     the same functionality for different hardware and it
-%%     is the hardware vendors that should supply these
+%* service is a set of views that support some relevant task
+%  for the connected controller  
+%* is extensible with user defined services
+%* supports replacement of services defined within the
+%  object by user defined services
+%* supports extension of services
+%* supports distribution of services and model among
+%  different computing resources
+%* has a single topology involving model and services
+%* has an access protection mechanism
 
@@
 %%   archive of XML files
-%%     attribute and view declarations in MetaInformation.xml
+%%   models, views, controllers and registry are declared in
+%%   some kind of MetaInformation.xml
 
@@
 %%   libraries  may be Internet repositories
-%%     searching and sorting by multiple attributes
-%%     searching and sorting by multiple views
+%%   searching and sorting by multiple attributes/keys
+%%   searching and sorting by services
 

### 2004-11-11 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

@@
+% -*- TeX-master:"main" -*-
+
 \chapter{Automation Objects}
 \label{automationobjects}
-
 %% Intro
@@
 for development and deployment of distributed automation
-systems has been developed. What is still needed is a way of
-encapsulating function blocks together with the data about
-their properties and, more importantly, their functionality
-in a way that will enable diminishing the time to market
-and facilitate effective rapid object oriented development
-of automation systems.
+systems has been established. What is still needed is a way
+of encapsulating function blocks together with the data
+about their properties and, more importantly, about their
+functionality in a way that will enable diminishing of the
+time-to-market and facilitate effective rapid object
+oriented development of automation systems. In this chapter
+the concept of automation objects is presented as far as the
+research has come. There are still plenty of details to be
+worked out but this chapter should clarify the role and the
+benefits of using the automation objects in the development
+of the industrial, as well as other, automated control
+systems.
 
@@
 IEC~61499 standard is ideal for the purpose that it is
-designed for but not enough for achieving \emph{effective
-  software reuse} in automation systems development.
+designed for but not enough for achieving effective software
+reuse in automation systems development.
 
 %%     effective decentralization of software development 
-Effective software reuse facilitates \textit{effective
-  decentralization of software development}. If anyone could
+Effective software reuse facilitates effective
+decentralization of software development. If anyone could
 use the software that somebody else developed without
@@
 Leveraging effective software reuse and decentralized
-software development \textit{makes integration and
-  deployment much easier}. System integrator's task is
-hereby reduced to picking the right software objects from
-the library and connecting them for the specific
-application.
+software development makes integration and deployment much
+easier. System integrator's task is hereby reduced to
+picking the right software objects from the library and
+connecting them for the specific application.
 
@@
 specified information about their usage makes it possible to
-\textit{build libraries} of elements that can easily be
-shared over the Internet since everything that is needed for
-the usage of the elements comes with the elements
-themselves. This is unlike the Java system where the objects
-are distributed in binary only format and the interface
-documentation has to be downloaded or reviewed on-line prior
-to usage.
+build libraries of elements that can easily be shared over
+the Internet since everything that is needed for the usage
+of the elements comes with the elements themselves. This is
+unlike the Java system where the objects are distributed in
+binary only format and the interface documentation has to be
+downloaded or reviewed on-line prior to usage.
 
@@
 
-%After a specification and a formal definition of the
-%encapsulation the benefits beyond the ones discussed above
-%will be presented followed by the presentation about the
-%implementation issues and the support in Supremica that is
-%planned for such data objects.
-
 \section{Specification}
@@
 
-%Automation Object:
-%* consists of models, controllers, views and
+%* AO consists of models, controllers, views and
 %  configuration registry
-Automation objects consists of models, views, controllers
-and configuration registry implemented using IEC-61499
-function blocks and their inteerfaces. Models are
-interchangable simulation models with varaying level of
-abstraction. One model might be used for the simulation on a
-high level of abstraction, typicaly behavioural model, and
-another for the simulation on a lower level. This is then
-used to implement dynamic zooming during the simulation. One
-of the models will implement the communication with the
-controlled equpiment thus allowing effortless switching
-between simulation and acutal control. This will enable
-gradual verification and on-line tuning of the big
-automation systems by allowing the simulation of the system
-components that are being configured or tuned while running
-the components that are already configured.
+Automation objects consists of models, controllers, views
+and configuration registry implemented using IEC~61499
+function blocks and their interfaces, see
+figure~\ref{fig:AO}. Models are interchangeable simulation
+models with varying level of abstraction. One model might
+be used for the simulation on a high level of abstraction,
+typically behavioral model, and another for the simulation
+on a lower level. This is then used to implement dynamic
+zooming during the simulation. One of the models will
+implement the communication with the controlled equipment
+thus allowing effortless switching between simulation and
+actual control. This will enable gradual verification and
+on-line tuning of the big automation systems by allowing the
+simulation of the system components that are being
+configured or tuned while running the components that are
+already configured.
+\begin{figure}[!ht]
+  \centering
+  \includegraphics{images/automationobjects/AO}
+  \caption{The architecture of the automation object.}
+  \label{fig:AO}
+\end{figure}
 
@@
 controlling strategies for easy verification with simulation
-and the equipment.
+and the equipment. The automation object can have as many
+controllers as it is needed but only one controller may be
+active at any time.
 
 Views are the function blocks allowing user interaction with
-the automation object. These implement all the alowable ways
+the automation object. These implement all the allowable ways
 of interaction with the automation object. At the lower
 levels of the automation system they will typically provide
-different event and data interfaces depending on the user's
-needs while at the higher leve they will implement full
-graphical user interfaces for the operators.
+various interfaces to the object depending on the user's
+needs while at the higher levels they will implement the
+full graphical user interfaces for the operators. As with
+the controllers there may be as many views as needed for the
+automation object.
 
-%* service is a set of views that support some relevant task
+Configuration registry is a function block implementing the
+configuration database and interface to it. This
+function block holds all relevant attributes for the
+automation object and provides an interface for accessing
+and changing them. Configuration registry is used by models,
+controllers and the views for storing and retrieving the
+configuration data.
+
+%* Service is a set of views that support some relevant task
 %  for the connected controller  
-%* is extensible with user defined services
+For every automation object a number of services may be
+defined. A number of views that are needed for a certain
+task during the development of the control system are
+specified in the definition of the service thus the service
+operates on a certain controller. The service can than be
+invoked during the development and ease the task that it is
+designed for. One service may only include views that
+operate on the same controller since only one controller is
+active at any time. Several services that operate on the
+same controller may be activated at the same time. On the
+other hand, services that operate on different controllers
+are mutually exclusive. During the development of the
+control system at the highest level of abstraction user of
+the automation objects is interacting only with services.
+For example, think of a company that is producing conveyor
+belts. The company buys motors and produces the belts. With
+the motor comes the automation object for it from the motor
+manufacturer. When the motor and the belt are assembled the
+tuning service of the motor's automation object can be
+invoked and the motor parameters can be tuned for the
+conveyor belt. The company also develops an accompanying
+automation object for the conveyor belt that encapsulates
+the automation object for the motor and provides a service
+for the tuning of the conveyor belt when it is installed in
+some manufacturing process.
+
 %* supports replacement of services defined within the
 %  object by user defined services
+%* is extensible with user defined services
 %* supports extension of services
-%* supports distribution of services and model among
+All services defined in an automation object are replaceable
+with user defined services. In case a user is not satisfied
+with how a service works she can develop new views and
+define a new service that will replace the service defined
+in the automation object. The user can also customize
+existing services by adding views to or replacing views in
+the defined services. It is also possible to define new
+services for the automation object by combining existing
+views in different groups, adding new views or both.
+
+%* supports distribution of services, controller and model among
 %  different computing resources
-%* has a single topology involving model and services
+Since the automation objects are built using IEC~61499
+function blocks the model, controller and services are
+distributable among different computing resources. Using
+this property tuning of the conveyor belt in the earlier
+example may be done from any computer on the company's
+network.
+
 %* has an access protection mechanism
+Last but not the least, automation objects have a
+protection mechanism that will hide the implementation of
+the models, views and controllers from the users of the
+automation objects.
 
@@
 % Support in Supremica
+First implementation of the automation object concepts will
+be done in the Supremica. The architecture for the
+automation objects will be concurrently implemented with the
+development tools and functions. In this section
+implementation considerations are discussed and the
+development work flow is presented using some mockup
+screenshots.
 
@@
 %%   some kind of MetaInformation.xml 
+An automation object is distributed as compressed archive of
+XML-files. The archive contains XML definitions of models,
+views, controllers, registry and a file containing the meta
+information for the automation object. In the meta
+information file all information about the automation object
+(ie name, creator, version), all of the possible
+configuration attributes, service definitions and interface
+explanations are stored.
 
@@
 
-% Example of usage in Supremica. Three scenarios with the
-% lab equipment (kulbanan).
-% 1. Student does everything.
-% 2. Library is provided and the logic is done by the student.
-% 3. Library is proviede and the logic is synthesised by the
-%    Supremica.
+%% development tool is the control system
+%% how will it work (mockups)

### 2004-11-16 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

@@
 Configuration registry is a function block implementing the
-configuration database and interface to it. This
-function block holds all relevant attributes for the
-automation object and provides an interface for accessing
-and changing them. Configuration registry is used by models,
-controllers and the views for storing and retrieving the
-configuration data.
+configuration database and interface to it. This function
+block holds all relevant attributes and configuration
+variables for the automation object and provides an
+interface for accessing and changing them. Configuration
+registry is used by models, controllers and the views for
+storing and retrieving the configuration data. For example,
+controllers will use this function block to access their
+regulator parameters and other configurable parameters.
 
@@
 implementation considerations are discussed and the
-development work flow is presented using some mockup
-screenshots.
+development work flow is presented.
 
-%% implementation of automation objects
+%% format of automation objects
 %%   archive of XML files
@@
 information for the automation object. In the meta
-information file all information about the automation object
+information file all the automation object attributes
 (ie name, creator, version), all of the possible
-configuration attributes, service definitions and interface
-explanations are stored.
+configuration variables, service definitions and interface
+explanations are stored. In other words everything needed
+for successful reusage in the development, configuration and
+execution of the automation object.
 
-%% libraries 
+%% repositories
 %%   non-hierarchical pool of automation objects
-%%   libraries  may be Internet repositories
+%%   may be share using the internet
 %%   searching and sorting by multiple attributes/keys
 %%   searching and sorting by services
+It is envisioned that after the initial acceptance period
+significant amount of automation objects are developed.
+These are stored in the repositories for easy acess. The
+repositories are non-hierarchical pools of automation
+objects, probably implemented using database technology.
+These repositories are prefferably shared over the internet.
+The most important feature of the repositories is the
+ability for extensinve search and sorting by multiple
+attributes defined in the registry as well as the defined
+services.
 
 %% development tool is the control system
-%% how will it work (mockups)
+%% how will it work?
+The Supremica tool provides the user interface for the
+development and the deployment of the automation object
+applications. The implementation that supports the user
+interface consists of two cooperating parts. The development
+environment for the automation objects and the run-time
+environment for the function blocks.
+
+The DE provides the tools and functions for the repository
+handlig and automation object development. The user is able
+to choose among the available automation objects from the
+repository and include them in the application. While
+working with the objects the user can activate different
+services or bring up specific views and exchange the models
+and controllers. User can also connect the views of the
+different objects in the application. Working in this way
+the user can build and tune the application to it's liking.
+
+All the time during the development process the run-time
+environment is active and taking commands from the
+development environment to add function block types, add
+instances and connections to composite function block types,
+to add instances and connections to the application, to
+configure the resources, to start and stop the execution of
+the application and so, thus creating the application in the
+run-time environment as it is developed. These commands are
+all at the function block level. It is the development
+environment that transforms the automation object operations
+into the function blocks running in the run-time environment
+and thus making all of the automation objects alive
+throughout the development.
+
+While the user have one of the simulation models activated
+the run-time is executing the code that simulates the
+behaviour of the hardware that the object is supposed to
+control. Switching to the model that talks to the process
+interface issues a command to exchange the function blocks
+in the run-time and results in the control of hardware
+instead of the simulation of the control. This gives user
+the freedom to incrementally develop large systems.
+
+For example, consider a production plant that is connected
+to the network of control computers. Engineers responsible
+for each part can develop, tune and test their part on any
+computer on the network. As each one switches to the model
+that communcates with the hardware the plant becomes more
+and more functional until it is fully operational.
+
+To accomplish this the development environment uses the
+database technology to hold the application data that is
+shared. While working on the application the engineers can
+acess this data from any computer on the network. The
+application data includes the distribution configuration
+governing which automation objects are run by which
+computers on the network. By issuing the commands to the
+run-time environment of the computer running the automation
+object that is currently worked on, the development
+environment allows the user to work on the whole application
+from any computer on the network. This means that if
+something goes wrong or software/equipment needs upgrading
+anyone that can acess the application data can, from any
+computer on the network, switch the model of the automation
+object in question to simulation model so that the parts of
+the plant that don't need the resource being upgraded in the
+production can continue to function, resolv the problem or
+perform the upgrade and finally switch back. This feature
+can also be used for active monitoring, tuning and control
+of the system by a human operator given that the views
+implementing the human machine interface are developed.

### 2004-12-06 doc/Simulation_and_Control_using_Supremica/functionblocks.tex
*** empty log message ***

@@
+% -*- TeX-master:"main" -*-
+
 \chapter{IEC~61499 Standard}
@@
 
-\subsection{System Model} 
+\subsection*{System Model} 
 At the physical level, a distributed system consists of a
@@
 
-\subsection{Device Model} 
+\subsection*{Device Model} 
 A resource provides independent execution and control of
@@
 
-\subsection{Resource Model} 
+\subsection*{Resource Model} 
 A resource is an independent functional unit, contained in a
@@
 
-\subsection{Application Model} 
+\subsection*{Application Model} 
 An application is defined as a network of interconnected
@@
 
-\subsection{Function Block Model} 
+\subsection*{Function Block Model} 
 A function block is a functional unit of software that its
@@
 
-\subsection{Basic Function Block Types} 
+\subsection*{Basic Function Blocks} 
 The behavior of basic function blocks is defined in terms of
@@
 
-\subsection{Composite Function Blocks} 
+\subsection*{Composite Function Blocks} 
 The behavior of a composite function block is described in
@@
 
-\subsection{Service Interface Function Blocks} 
+\subsection*{Service Interface Function Blocks} 
 Service interface function blocks provide an interface
@@
 
-\subsection{Event Function Blocks} 
+\subsection*{Event Function Blocks} 
 Event function blocks are some special basic function blocks

### 2004-12-06 doc/Simulation_and_Control_using_Supremica/introduction.tex
*** empty log message ***

@@
+% -*- TeX-master:"main" -*-
+
 \chapter{Introduction}
@@
 
-Define the problem.
-Define other approaches, state of the art.
-Introduce Function Blocks
-Introduce Automation Objects
-(Configuration, Simulation, Control, HMI)
+%Define the problem.
+Ever increasing competition forces companies to diversify
+their product lines. To achieve customizable products the
+companies require constantly changing and evolving
+production plants. Such plants are called flexible
+production plants. The plant is constantly adopted to the
+products with goal of producing the highest quality products
+in as little time as posibble. To achieve this the software
+for the automatic control of such a plant has to be changed
+along with the plant. The problem is that using today
+established development methods this is impossible without
+stoping of the production.
+
+%Define other approaches, state of the art.
+
+%Introduce Function Blocks
+
+%Introduce Automation Objects

### 2004-12-06 src/org/supremica/automata/algorithms/AutomataSynthesizer.java
*** empty log message ***

@@ -544,7 +544,7 @@ public class AutomataSynthesizer
 
 		// We need to synthesize even if the result above is controllable
 		// Nonblocking may ruin controllability
-		// ARASH: choose between tripple and the single fixpoint algorithms:
+		// ARASH: choose between triple and the single fixpoint algorithms:
 		AutomatonSynthesizer synthesizer = singleFixpoint
 										   ? new AutomatonSynthesizerSingleFixpoint(retval.automaton, synthesizerOptions)
 										   : new AutomatonSynthesizer(retval.automaton, synthesizerOptions);

### 2004-12-06 doc/Simulation_and_Control_using_Supremica/introduction.tex
*** empty log message ***

@@
 production plants. The plant is constantly adopted to the
-products with goal of producing the highest quality products
-in as little time as posibble. To achieve this the software
-for the automatic control of such a plant has to be changed
-along with the plant. The problem is that using today
-established development methods this is impossible without
-stoping of the production.
+products with the goal of production of highest quality
+products in as little time as posibble. To achieve this the
+software for the automatic control of such a plant has to be
+developed and upgraded along with the plant. The problem is
+that when using today established development methods this
+upgrade is impossible without an interrupt of the production
+that can last for several weeks in some cases.
+
+To avoid the down time associated with the upgrade of the
+control software in the production plants a more agile
+architecture for the control software is needed. To be more
+specific an object oriented architecutre and development
+method are needed as opposed to the loop based architecture
+and structured development method implemented with PLCs. The
+advantages of object oriented software development are
+reaped across the IT industry for quite some time now. Why
+it has not been widely adopted in the automatic control
+software development is for the most part because of the
+high investment already made in the PLCs making the
+companies reluctant to invest in yet another expensive
+thechnology driven upgrade. In other words, as long as the
+automation control problems in the existing plants are
+possible to be solved with PLCs there will be no adoption of
+object orented architectures. Another big reason for slow
+adoption of object oriented development methods is the lack
+of standard object oreinted platform suitable for the
+automation software applications.
 

### 2004-12-07 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

@@
 significant amount of automation objects are developed.
-These are stored in the repositories for easy acess. The
+These are stored in the repositories for easy access. The
 repositories are non-hierarchical pools of automation
 objects, probably implemented using database technology.
-These repositories are prefferably shared over the internet.
+These repositories are preferably shared over the internet.
 The most important feature of the repositories is the
-ability for extensinve search and sorting by multiple
+ability for extensive search and sorting by multiple
 attributes defined in the registry as well as the defined
@@
 The DE provides the tools and functions for the repository
-handlig and automation object development. The user is able
+handling and automation object development. The user is able
 to choose among the available automation objects from the
@@
 the run-time is executing the code that simulates the
-behaviour of the hardware that the object is supposed to
+behavior of the hardware that the object is supposed to
 control. Switching to the model that talks to the process
@@
 computer on the network. As each one switches to the model
-that communcates with the hardware the plant becomes more
+that communicates with the hardware the plant becomes more
 and more functional until it is fully operational.
@@
 shared. While working on the application the engineers can
-acess this data from any computer on the network. The
+access this data from any computer on the network. The
 application data includes the distribution configuration
@@
 something goes wrong or software/equipment needs upgrading
-anyone that can acess the application data can, from any
+anyone that can access the application data can, from any
 computer on the network, switch the model of the automation
@@
 the plant that don't need the resource being upgraded in the
-production can continue to function, resolv the problem or
+production can continue to function, resolve the problem or
 perform the upgrade and finally switch back. This feature

### 2004-12-07 doc/Simulation_and_Control_using_Supremica/introduction.tex
*** empty log message ***

@@
 products with the goal of production of highest quality
-products in as little time as posibble. To achieve this the
+products in as little time as possible. To achieve this the
 software for the automatic control of such a plant has to be
@@
 architecture for the control software is needed. To be more
-specific an object oriented architecutre and development
+specific an object oriented architecture and development
 method are needed as opposed to the loop based architecture
@@
 companies reluctant to invest in yet another expensive
-thechnology driven upgrade. In other words, as long as the
+technology driven upgrade. In other words, as long as the
 automation control problems in the existing plants are
 possible to be solved with PLCs there will be no adoption of
-object orented architectures. Another big reason for slow
+object oriented architectures. Another big reason for slow
 adoption of object oriented development methods is the lack
-of standard object oreinted platform suitable for the
-automation software applications.
+of standard object oriented platform suitable for the
+automation software applications. The latter reason is being
+addressed by several vendors but their offerings are
+proprietary, incompatible and only add-ons to the IEC 61131
+standard which defines a non object oriented development
+platform.
 
@@
 %Introduce Function Blocks
+The International Electrotechnical Commission (IEC) is making
+an effort to standardize a platform for the software of next
+generation of systems for industrial measurement and
+control. The standard is named IEC 61499 and it defines the
+architecture, software tool requirements and application
+guidelines. The architecture defines object oriented
+platform for distributable software components and is
+described in more detail in chapter~\ref{functionblocks}.
 
 %Introduce Automation Objects
+The standard defines the common ground for the distributed
+industrial control applications but it is not enough. To
+ease the effective and rapid control application development
+a more user friendly software objects are needed. The kind
+of objects that will run on any IEC 61499 compatible system
+but provide a greater support for the task of application
+developing. That kind of objects are defined in
+chapter~\ref{automationobjects}.

### 2005-01-24 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

@@
 and configuration registry implemented using IEC~61499
-function blocks and their interfaces, see
+function blocks, and their interfaces, see
 figure~\ref{fig:AO}. Models are interchangeable simulation

### 2005-01-24 src/org/supremica/automata/IO/AutomataToIEC61499.java

@@ -74,659 +74,675 @@ import org.supremica.automata.algorithms.SynchronizationType;
 public class AutomataToIEC61499
 
 {
+    
+	private static Logger logger = LoggerFactory.createLogger(AutomataToIEC61499.class);
+	private Project theProject;
+	private IEC61131Helper theHelper;
+	private Alphabet allEvents;
+	private SynchronizationType syncType = SynchronizationType.Prioritized;
+	private boolean comments = false;	
 	
-    private static Logger logger = LoggerFactory.createLogger(AutomataToIEC61499.class);
-    private Project theProject;
-    private IEC61131Helper theHelper;
-    private Alphabet allEvents;
-    private SynchronizationType syncType = SynchronizationType.Prioritized;
-    private boolean comments = false;	
-	
-    public void commentsOn()
-    {
-	comments = true;
-    }
-	
-    public void commentsOff()
-    {
-	comments = false;
-    }
-	
-    // Constructor
-    public AutomataToIEC61499(Project theProject)
-    {
-	this.theProject = theProject;
-	this.theHelper = IEC61131Helper.getInstance();
-	allEvents = this.theProject.setIndicies();
-    }
-
-
-    /** Makes the beginning of the function block type declaration. */
-    private void printBeginProgram(PrintWriter pw)
-    {
-	if (comments)
+	public void commentsOn()
 	{
-	    pw.println("(* This function block was automatically generated from Supremica *)");
-	    pw.println("(* Supremica version: " + org.supremica.Version.version() + "*)");
-	    pw.println("(* Time of generation: " + DateFormat.getDateTimeInstance().format(new Date()) + " *)");
+		comments = true;
 	}
 	
-	if (theProject.getNbrOfAutomata() >= 2)
+	public void commentsOff()
 	{
-	    pw.println("FUNCTION_BLOCK AUTOGEN_FB");
+		comments = false;
 	}
-	else
-	{
-	    pw.println("FUNCTION_BLOCK AUTOGEN_" + theProject.getAutomatonAt(0).getName());	    
-	}
-    }
-
-
-
-
-    /** Makes the fb_interface_list production rule of the standard. */
-    private void printInterfaceList(PrintWriter pw) 
-    {
-
-
-	// The event_input_list. This is the same for all FBs of automata.
[…]

### 2005-01-24 src/org/supremica/gui/ActionMan.java

@@ -3694,57 +3694,54 @@ public class ActionMan
 			}
 		}
 	}
-
+    
 	// Generate IEC-61499 Function Block
 	public static void ProjectToIEC61499(Gui gui) 
 	{
 		// Automata selectedProject = gui.getselectedProject();
 		Project selectedProject = gui.getSelectedProject();
-
+		
 		if (selectedProject.size() < 1) {
 			JOptionPane.showMessageDialog(
-				gui.getComponent(),
-				"At least one automaton must be selected!",
-				"Alert",
-				JOptionPane.ERROR_MESSAGE);
-
+						      gui.getComponent(),
+						      "At least one automaton must be selected!",
+						      "Alert",
+						      JOptionPane.ERROR_MESSAGE);
+			
 			return;
 		}
-
+		
 		JFileChooser fileExporter = FileDialogs.getExportFileChooser(FileFormats.FBT);
-
-		if (fileExporter.showSaveDialog(gui.getComponent()) == JFileChooser.APPROVE_OPTION) {
+		
+		if (fileExporter.showSaveDialog(gui.getComponent()) == JFileChooser.APPROVE_OPTION) 
+		{
 			File currFile = fileExporter.getSelectedFile();
-
-			if (currFile != null) {
-				if (!currFile.isDirectory()) {
-					try {
-						AutomataToIEC61499 exporter =
-							new AutomataToIEC61499(selectedProject);
-						PrintWriter theWriter =
-							new PrintWriter(new FileWriter(currFile));
-
-						exporter.printSources(theWriter);
-						theWriter.close();
-					} catch (Exception ex) {
-						logger.error(
-							"Exception while generating IEC-61499 Function Block code to file "
-								+ currFile.getAbsolutePath());
-						logger.debug(ex.getMessage());
-						logger.debug(ex.getStackTrace());
-
-						return;
-					}
-
-					logger.info(
-						"IEC-61499 Function Block file successfully generated at "
-							+ currFile.getAbsolutePath());
+			
+			if (currFile != null) 
+			{
+				try 
+				{
+					AutomataToIEC61499 exporter = new AutomataToIEC61499(selectedProject);
+					exporter.printSources(currFile);
+				} 
+				catch (Exception ex) 
+				{
+					logger.error("Exception while generating IEC-61499 Function Block code to file "
+						     + currFile.getAbsolutePath());
+					logger.debug(ex.getMessage());
+					logger.debug(ex.getStackTrace());
+					
[…]

### 2005-01-31 doc/Sync_Execution_with_IEC61499/paper.tex
*** empty log message ***

\documentclass[conference]{IEEEtran}
\bibliographystyle{IEEEtran}

\usepackage{cite}
\usepackage[latin1]{inputenc}
\usepackage[dvips]{graphicx}
\usepackage{theorem}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{latexsym}
\usepackage{amsmath}

\newlength{\graphwidth}

\newcommand{\fig}[3][scale=0.5]
{
  \vspace{0.3cm}
  \begin{minipage}[t]{0in}
    \vspace{0pt}
    \makebox{\small #2}
  \end{minipage}
  \settowidth{\graphwidth}{\includegraphics[#1]{#3}}
  \nopagebreak\kern-1.5cm
  \begin{minipage}[t]{\graphwidth}
    \vspace{0.2cm} \includegraphics[#1]{#3}
  \end{minipage}
}

\newtheorem{definition}{Definition}

\begin{document}
\title{Implementation of Full Synchronous Composition of
  Discrete Event Models Using IEC 61499 Function Blocks}

\author{\authorblockN{Goran Cengic and Knut Åkesson}
\authorblockA{
  Automation Laboratory\\
  Department of Signals and Systems\\
  Chalmers University of Technology\\
  Sweden
}
\and
\authorblockN{Chengyin Yuan and Placid Ferreira}
\authorblockA{
  Department of Mechanical and Industrial Engineering\\
  University of Illinois at Urbana-Champaign\\
  USA
}
}
\maketitle
\begin{abstract}
%   Two methods for the implementation of synchronized
%   execution of the discrete event systems using the IEC
%   61499 function blocks are presented. The first method is
%   suitable for distributed control applications. Each
%   discrete event model is implemented in a single function
%   block type. These models are executed synchronously using
%   a special synchronization function block type. The second
%   method is suitable for non--distributed, i.e. local,
%   control applications and implements all discrete event
%   models that need to be synchronized in one function block
%   type. The methods are presented using a simple
%   manufacturing system. A supervisor for the example has been
%   synthesized using simpler resource models resulting in
%   more efficient synthesis. On the other hand complex models
%   have been used for the execution giving more accurate
%   control.
\end{abstract}

\section{Introduction}
% distributed control systems
In many man--made systems there is a need for distributed
control systems. The need is greatest in cases where the
system is distributed among different geographical locations
or dispersed over large area. Examples of such systems are
manufacturing systems and chemical processing systems.

Developing distributed control systems is costly and error-prone
due to that the existing communication standards are either
vendor dependent or at a low level of abstraction. An
emerging open standard for developing distributed control
system is IEC~61499~\cite{iec:614991:2000},
\cite{rl:mod:2001}. IEC 61499  extends the standard for
Programmable Logic Controller, IEC~61131~\cite{iec:61131},
with functionality for distributed execution of control objects.

Shorter product life-cycles increases the demands on the
manufacturing system. Since the product, to be produced, in
the manufacturing plant is changing frequently the
production equipment needs to be modified too. The frequent
changes of the system make it time consuming and expensive
to update the control system application to reflect the
changes. The same problem applies to all of the system types
mentioned above and it becomes even larger in mission
critical control applications since the strict requirenment
of guaranteeing the correct operation is imposed. Thus,
there is a need to introduce algorithms and tools to make is
a simple and safe as possible to modify the control system.

% formal methods and suprevisory control
Many man made systems can, at some abstraction level, be
modeled by discrete event models. A useful class of discrete
event models is the deterministic finite automata. Finite
automata can be used in a formal method,
called suprevisory control theory~\cite{rw:con:1989}, to
automatically generate correct control functions so that the
controlled system exhibits behavior given by some
specification.

% synchronization
In suprevisory control theory the interaction between models
is modeled by synchronization operator. This operator has to
be implemented in control application in order to transfer
theoretical results into operating control system. The
implementation has to be distributable for it to be useful
in control systems for the applications distributed over
geographical area, like those mentioned in the beginning of
the introduction.

There are existing methods for implementation of
synchronization operator using PLC programming techniques
\cite{hfl:syn:1999,hfl:mod:2001,fh:plc:1998} but the
distribution is hard to achieve in that programming
environment. A new method presented in this paper uses the
emerging standard for distributed control systems to remedy
that.

% paper organisation
Mathematical definitions that are used in the paper are
presented in the next section. After that, the most
important definitions and results of suprevisory control
theory are introduced followed by an exampel showing how
supervisory control theory can be applied in implementation
of control application for a small system. After the example
a method for implementation of the synchronization operator
for any number of synchronized models is presented together
with the argument for the correctness of the implementation.
In Section~\ref{sec:app_SCT} some considerations for
implementation of supervisory control theory in applications
are presented before the conclusion of the paper.

\section{Mathematical Preliminaries}
In~\cite{hmu:int:2003} many important results for automata
are presented. Below, the terminology in this paper is
introduced.\\
\textbf{Definition:} Deterministic Finite Automaton (DFA)\\
A deterministic finite automaton is a 5-tuple defined as
$${A} = \langle Q, \Sigma, \delta, q_i, Q_m\rangle$$
where
$Q$ is a nonempty finite set of \emph{states}, $\Sigma$ is
the \emph{alphabet} which is a finite set of symbols that
denote the occurrence of an event, $\delta:Q\times \Sigma
\rightarrow Q$ is a \emph{transition function}, $q_i \in Q$
is the \emph{initial state} and $Q_m \subseteq Q$ is a set
of \emph{marked states}. $\Box$\\
\textbf{Definition:} Enabled Event Function\\
The enabled event function $\Gamma$ is defined as
$$\Gamma:Q\rightarrow 2^{\Sigma}$$
where $\Gamma(q)$ is the
set of all events $\sigma$ for which $\delta(q,\sigma)$ is
defined. $\Box$\\
Automata are usually visualized by directed graphs with
states represented by nodes, transitions by arrows and
events upon which the transitions occur by labels on the
arrows.

Synchronous composition, \cite{h:com:1985}, is a useful
approach to model the interaction between automata and is
used in this paper to model the interaction between the
system models.\\
\textbf{Definition}: Full Synchronous Composition of two DFAs\\
The full synchronous composition of two DFAs, $A^1$ and
$A^2$, is denoted ${A}^1||{A}^2=\langle
Q,\Sigma,\delta,q_i,Q_m\rangle$ where $Q=Q^1 \times Q^2$,
$\Sigma=\Sigma^1 \cup \Sigma^2$, $q_i=\langle {q^1_{i}},
{q^2_{i}} \rangle$, $Q_m={Q^1_{m}} \times {Q^2_{m}}$ and the
transition function is
\begin{displaymath}
  \begin{array}{l}
    \delta(\langle{q^1,q^2}\rangle,\sigma)=\\
    =\left\{ {\begin{array}{*{10}l}
          {{ \langle \delta ^1 (q^1 ,\sigma ), \delta ^2 (q^2 ,\sigma ) 
              \rangle}} &  \sigma \in \Gamma^1(q^1) \cap \Gamma^2(q^2) \\
          {{ \langle \delta ^1 (q^1 ,\sigma ), q^2 \rangle} } &  \sigma \in 
          \Gamma^1(q^1) \setminus \Sigma^2 \\
          { { \langle q^1, \delta ^2 (q^2 ,\sigma )\rangle}} & \sigma \in 
          \Gamma^2(q^2) \setminus \Sigma^1 \\
          { \rm undefined} & {{\rm otherwise.}}
        \end{array}} \right.
  \end{array}
\end{displaymath} 
where $q^1 \in Q^1$, $q^2 \in Q^2$ and $\sigma \in
\Sigma$. $\Box$
In other words, if an event is in both alphabets and it is
enabled in both automatons, the transitions associated with
it are made simultaneously in both automatons. Otherwise,
only the automaton containing the event in its alphabet
makes the transition. This definition can be extended for
several automata.\\
\textbf{Definition}: Full Synchronous Composition of $n$ DFAs\\
The full synchronous composition of DFAs,
$A^1$, $A^2$ ... $A^n$, is denoted ${A}^1||...||{A}^n=\langle
Q,\Sigma,\delta,q_i,Q_m\rangle$ where 
\begin{eqnarray}
  Q & = & Q^1 \times Q^2 \times ... \times Q^n\nonumber\\
  \Sigma & = & \Sigma^1 \cup \Sigma^2 \cup ... \cup \Sigma^n\nonumber\\
  q_i & = & \langle q^1_i,q^2_i,...,q^n_i \rangle\nonumber\\
  Q_m & = & Q^1_m \times Q^2_m \times ... \times Q^n_m\nonumber
\end{eqnarray}
and the transition function is
\begin{multline}\label{eq:delta}
  \delta(\langle{q^1,...,q^n}\rangle,\sigma)=\\
  =
  \begin{cases}
    {\rm undefined} &, {\rm if\ } \exists A^k:\sigma\in\Sigma^k\land\sigma\notin\Gamma^k(q^k)\\
    \langle q^{1+},\dots,q^{n+}\rangle & {\rm , otherwise}
  \end{cases} 
\end{multline} 
where 
\begin{equation}\label{eq:qi}
  q^{k+}=
  \begin{cases}
    \delta^k(q^k,\sigma) & {\rm if\ }\sigma\in\Gamma^k(q^k)\\
    q^k                  & {\rm if\ }\sigma\notin\Sigma^k
  \end{cases}
\end{equation}
$\Box$\\
What this definiton states is that in a group of automatons
that share one event if there is one automaton that does not
have that event enabled in its current state, than the event
is not allowed to occur in any of the other automatons in
the group. That is, one automaton is able to prevent all the
others from executing an event if it does not enable it at
the same time as all the other automatons. This property of
the definiton is used to construct a supervisor which
prevents events from occurring. The concept of supervisor is
explained further in the next section which introduces the
supervisory control theory.

\section{Supervisory Control Theory}
\label{sec:sct}
Suprevisory control theory provides definitions and results
needed for caluation of allowed behavior of the system from
discrete event models of possible behavior and desired
behavior. The model of the possible behavior is called the
plant, the model of the desired behavior is called the
specification, and the result of the computation is called
the suprevisor.

In the supervisory control theory, the plant is assumed to
generate all events and the purpose of the supervisor is to
prevent the plant from generating events that will violate
the specfication. The assumption that the plant generates
all events might seem like a restriction. However, it is
often possible to introduce a controller between the
supervisor and the resources that initiate the events.
Treating the controller as part of the plant will make the
assumption that all events is generated by the plant a
natural one. In Figure~\ref{fig:sup_plant} such configuration
is shown.
\begin{figure}
  \centering \includegraphics[scale=0.5]{figures/sup_plant}
  \caption{Supervisory control configuration.}
  \label{fig:sup_plant}
\end{figure}

The supervisor $S$ in Figure~\ref{fig:sup_plant} has to
prevent the plant from breaking the specification and from
getting blocked in a state from where it can not reach a
marked state by using as few restrictions of the plant as
possible. By observing the observable events of the plant
and disabling the controlable events of the plant, the
supervisor decides when, and if, it has to limit the
activity of the plant in order to keep the plant's behavior
within the specification. An important result of the
supervisory control theory is that such a supervisor can be
calculated from the models of the plant and specification so
that the system behavior is given by
$$S||P'.$$

In cases when there are several plant models $P_1,..,P_n$
with appropriate specification models $Sp_1,...,Sp_n$ the
supervisor can also be generated and the system behavior
within specification is given by
$$S||P_1||...||P_n.$$
Thus to control the system so that it
satisfies the specification, the synchronous composition of
the plant and supervisor models has to be implemented in the
control software. Using the conventional implementation
methods, i.e. implementing the synchronization using PLC
programming techniques, the solution becomes complex which
makes it difficult, if not impossible, to distribute across
multiple control computers. Next section presents an example
illustrating how the syncrounous composition can be
implemented using the new standard.

\section{Example}
In previous section it is shown that the synchronization
operator is a central concept for the results of supervisory
control theory. Now an example is presented showing how that
theoretical concept can be implemented using IEC 61499
function blocks for a small system before a general method
is given.

Consider a system containing two plants, \texttt{P1} and
\texttt{P2}. The discrete event models for the plants are
shown in the figure~\ref{fig:plants}. All events in the
models are controllable and observable.The desired behavior
of the system is that event \texttt{b} must always precede
event \texttt{c}. The supervisor for this specification can
be generated and is shown in figure~\ref{fig:sup}.
\begin{figure}
  \centering 
  \fig[scale=0.55]{\texttt{P1}}{figures/P1_aut} 
  \fig[scale=0.55]{\texttt{P2}}{figures/P2_aut}
  \caption{Discrete event models of the plants.}
  \label{fig:plants}
\end{figure}
\begin{figure}
  \centering 
  \fig[scale=0.55]{\texttt{S}}{figures/Sp_aut}
  \caption{The model of the supervisor.}
  \label{fig:sup}
\end{figure}
\begin{figure}[!ht]
  \centering 
  \fig[scale=0.5]{\texttt{SYS}}{figures/Sys_aut}
  \caption{The model of possible system behavior while satisfying
    the desired system behavior, $S||P_1||P_2$.}
  \label{fig:sys}
\end{figure}
Full synchronous execution of these models gives the system
that behaves as freely as possible while fulfilling the
specification. The resulting model of the system is shown in
figure~\ref{fig:sys}.

% IEC 61499 intro
The emergin standard IEC 61499 specifies a generic
architecture for measurement and control of industrial
processes. The architecture is based on functional software
units called \textit{function blocks}. The most common type
of function block in an application is basic function block.
It executes algorithms based on the arriving events and
generate new events that are passed on upon the algorithms
execution. The algorithms use the internal variables and the
data associated with the incoming events to update the
internal variables and produce the output data associated
with the outgoing events. What algorithms are invoked is
controlled by the execution control chart (ECC) which is a
automaton with algorithms associated with different states.
The events are boolean expressions involving event inputs,
internal, input, and output variables. In
Figure~\ref{fig:61499_intro} a basic function block type is
shown using the graphical representation.
\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures/basic_block}\\(a)
  \vskip 5mm
  %\includegraphics[scale=0.5]{figures/basic_ecc}\\(b)
  \caption{(a) The basic function block type and (b) its
    execution control chart.}
  \label{fig:61499_intro}
\end{figure}

The event driven interface enables the distribution of the
function blocks among the multiple control computers.

Function block types that generate the behavior of the plant
and specification models for the example are shown in
figure~\ref{fig:systemFBs}. The event input of each block
type is \texttt{OCCURRED}. It signals that an discrete event
has occurred and that the model should update its state
according to the data inputs. The data inputs are boolean
variables representing the discrete events in the model's
alphabet. For well defined behavior only one of the data
inputs associated with the \texttt{OCCURRED} event must be
\texttt{true} when the event is received. The model's state
is then updated according to the discrete event that this
data input represents and the event output \texttt{DONE} is
generated. It signals that the model has updated its state
and that the data outputs are ready. The data outputs are
boolean variables representing the enabled events after the
model has updated its state, i.e. the events labeling the
transitions that are possible from the new state.

The function block type that is generated according to the
new method presented is shown in figure~\ref{fig:syncFB} and
is called synchronization function block type. The data
inputs and outputs of the block type are application
specific, since they are defined by the models that will be
synchronized. The event input and output, on the other hand,
are the same for all applications which also makes the ECC
of the block type the same for all applications. The ECC for
the block type is shown in figure~\ref{fig:ecc_sync}. The
\texttt{DISABLE\_EVENTS} algorithm associated with the
\texttt{SYNC} state is where the synchronized execution is
implemented. The transition labeled with the ``1'' character
is taken as soon as the \texttt{DISABLE\_EVENTS} algorithm
has finished executing and the \texttt{UPDATE} event is
sent.

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures/P1}\\(a)
  \vskip 5mm
  \includegraphics[scale=0.5]{figures/P2}\\(b)
  \vskip 5mm
  \includegraphics[scale=0.5]{figures/Sp}\\(c)
  \caption{The function block types implementing (a)
    resource 1, (b) resource 2, and (c) specification.}
  \label{fig:systemFBs}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures/SYNC}
  \caption{The function block type for the synchronization.}
  \label{fig:syncFB}
\end{figure}

The synchronization block reads the enabled events of the
synchronized models and disables the events that are not
enabled in all the models that have that event in their
alphabets. Each boolean data input signals if the event in
the model represented by that input is enabled
(\texttt{true}) or disabled (\texttt{false}). By setting the
corresponding data output to \texttt{false} the
synchronization block signals that the transition labeled by
that event is not allowed to occur in the model during the
next update cycle, if so needed. To calculate each data
output a logical expression consisting of all data inputs
that represent the same event, connected by logical
\texttt{and} operation, is used. This means that if the
event is enabled in all of the models that have the event in
their alphabet it is allowed to occur in the next update
cycle. Otherwise the expression is false and the the event
is disabled. Following implementation of the
\texttt{DISABLE\_EVENTS} algorithm in IEC 61131 structured
text does the computation:

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{figures/sync_ecc}
  \caption{The execution control chart for the
    synchronization function block type.}
  \label{fig:ecc_sync}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[scale=0.35]{figures/app}
  \caption{The control application for the example.}
  \label{fig:app}
\end{figure}

\begin{verbatim}
EO_R1_a := EI_R1_a;
EO_R1_b := EI_R1_b AND EI_Sp_b;
EO_R2_c := EI_R2_c AND EI_Sp_c;
EO_R2_d := EI_R2_d;
EO_Sp_b := EI_R1_b AND EI_Sp_b;
EO_Sp_c := EI_R2_c AND EI_Sp_c;
\end{verbatim}
Since the event \texttt{a} is in the alphabet of the
\texttt{R1} model only it is always allowed to occur if it
is enabled. The event \texttt{b} is allowed to occur in
\texttt{R1} model only if it is enabled in \texttt{R1} and
\texttt{Sp} models at the same time since it is contained in
both alphabets. The same applies to the event \texttt{b} in
\texttt{Sp} model, hence the same logical expression. Events
\texttt{d} and \texttt{c} are treated in the same manner.

\section{General Method}
The method for implementation of the full synchronous
composition for the general system containing any number of
discrete event models is made up of two major steps. The
first step is creation of the function block types that
generate the behavior of the discrete event models. The
interface that these blocks have to adhere to is shown in
figure~\ref{fig:model_interface}. These blocks are
application specific but they can be automatically generated
from discrete event models~\cite{aff:sup:2003},
\cite{www:sup:2004}.

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures/model_interface}
  \caption{The interface definition for the function block
    implementation of the discrete event models.}
  \label{fig:model_interface}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures/sync_interface}
  \caption{The interface definition for the synchronization
    function block type.}
  \label{fig:sync_interface}
\end{figure}

The second step is creation of the synchronization function
block type. The interface of this block is shown in the
figure~\ref{fig:sync_interface}. The execution control chart
for the function block type is the same as shown in
figure~\ref{fig:ecc_sync} since it is application generic.

The \texttt{DISABLE\_EVENTS} algorithm is also application
specific and it can be generated like this: Let $M$ be the
set of all models for synchronization. For all $m,e$ such
that $m\in M$ and $e\in\Sigma^m$:
\begin{equation}\label{eq:lexp}
  {\rm EO}\_m\_e:=\bigwedge_{\{n:n\in
    M\land e\in\Sigma^n\}}{\rm EI}\_n\_e
\end{equation}
The iteration over all events of all models is for the
generation of the logical expression assignments to all data
outputs, which are represented by EO\_$m$\_$e$. The right
hand side of~(\ref{eq:lexp}) generates the logical expression
that covers all the cases of the synchronization definition.
The first case in~(\ref{eq:delta}) is covered by
$e\in\Sigma^n$ and the data input being \texttt{false} which
gives that EO\_$m$\_$e$ is \texttt{false} and thus the event
is disabled. The first case in equation~\ref{eq:qi} is
covered by $f$ being in the $\Sigma^n\cap\Sigma^m$ and the
data input being \texttt{true}. Finally, the second case in
equation~\ref{eq:qi} is covered by $f$ being in the
$\Sigma^n\cap\Sigma^m$. What is left now is to connect
instances of the model function blocks with the
synchronization function block, which concludes the method.

\section{Application of supervisory control theory}
\label{sec:app_SCT}
Blah blah.

\section{Conclusion}
In many distributed control applications the need for
synchronous execution of discrete event models is present.
The two methods presented in this paper may be used to
implement such synchronous execution by using IEC 61499
function blocks. The first method is useful in case the
function blocks implementing the models that need to be
synchronized have to be distributed across different
computing resources in the control system. It also results
in a slightly more complex function block program. The
second method on the other hand results in a single function
block type but it is only useful when the synchronized
models can fit into single computing resource.

The structure of the function block types in both methods is
obvious. This makes implementation of computer programs for
the generation of such blocks a straight forward task. In
fact, the function block type generation for the second
method is already implemented in the Supremica tool,
\cite{aff:sup:2003} and \cite{www:sup:2004}. Work is under
way to implement automatic generation for the first method
too. Future considerations should be made to accommodate for
the controllable and uncontrollable events in the automatic
generation of the synchronization function block types.

The method and function block programming in general can be
tested using the Function Block Development Kit (FBDK),
\cite{c:fun:2002}. For further information about IEC 61499
function blocks see \cite{c:ope:2002}.

\bibliography{../../../Bibliographies/desbib.bib}

\end{document}

### 2005-01-31 doc/Sync_Execution_with_IEC61499/paper.tex
*** empty log message ***

@@
 
-\author{\authorblockN{Goran Cengic and Knut Åkesson}
+\author{\authorblockN{Goran Cengic, Knut Åkesson and
+    Bengt Lennartson}
 \authorblockA{
@@
 \begin{abstract}
-%   Two methods for the implementation of synchronized
-%   execution of the discrete event systems using the IEC
-%   61499 function blocks are presented. The first method is
-%   suitable for distributed control applications. Each
-%   discrete event model is implemented in a single function
-%   block type. These models are executed synchronously using
-%   a special synchronization function block type. The second
-%   method is suitable for non--distributed, i.e. local,
-%   control applications and implements all discrete event
-%   models that need to be synchronized in one function block
-%   type. The methods are presented using a simple
-%   manufacturing system. A supervisor for the example has been
-%   synthesized using simpler resource models resulting in
-%   more efficient synthesis. On the other hand complex models
-%   have been used for the execution giving more accurate
-%   control.
+  A new method for implementation of synchronous composition
+  is presented. Using an emerging standard, IEC 61499, the
+  method enables efficient development of distributed
+  control applications. It also brings together supervisory
+  control theory with distributed control application
+  development since synchronous composition of discrete
+  event models is used to model interaction between systems
+  in supervisory control theory. Method is demonstrated
+  using a simple example before the formulation for the
+  general case is presented. A new structure of control
+  applications suitable for application of supervisory
+  control theory to control application development is
+  presented also.
 \end{abstract}
@@
 % distributed control systems
-In many man--made systems there is a need for distributed
+In many man-made systems there is a need for distributed
 control systems. The need is greatest in cases where the
@@
 
-Developing distributed control systems is costly and error-prone
-due to that the existing communication standards are either
-vendor dependent or at a low level of abstraction. An
-emerging open standard for developing distributed control
-system is IEC~61499~\cite{iec:614991:2000},
-\cite{rl:mod:2001}. IEC 61499  extends the standard for
-Programmable Logic Controller, IEC~61131~\cite{iec:61131},
-with functionality for distributed execution of control objects.
+Developing distributed control systems is costly and
+error-prone due to the fact that existing communication
+standards are either vendor dependent or at a low level of
+abstraction. An emerging open standard for developing
+distributed control system is
+IEC~61499~\cite{iec:614991:2000,rl:mod:2001}. IEC 61499
+extends the programming standard for Programmable Logic
+Controller, IEC~61131~\cite{iec:1131:1993}, with
+functionality for distributed execution of control objects.
 
-Shorter product life-cycles increases the demands on the
+Shorter product life-cycles increase the demands on the
 manufacturing system. Since the product, to be produced, in
@@
 to update the control system application to reflect the
-changes. The same problem applies to all of the system types
+changes. The same problem applies to both system types
 mentioned above and it becomes even larger in mission
@@
 of guaranteeing the correct operation is imposed. Thus,
-there is a need to introduce algorithms and tools to make is
-a simple and safe as possible to modify the control system.
+there is a need to introduce algorithms and tools to make it
+simple and as safe as possible to modify the control system.
+The supervisory control theory,\cite{rw:con:1989}, is a
+promising framework that might be useful for this purpose.
 
-% formal methods and suprevisory control
-Many man made systems can, at some abstraction level, be
+% suprevisory control
+Many man-made systems can, at some abstraction level, be
 modeled by discrete event models. A useful class of discrete
-event models is the deterministic finite automata. Finite
-automata can be used in a formal method,
-called suprevisory control theory~\cite{rw:con:1989}, to
-automatically generate correct control functions so that the
-controlled system exhibits behavior given by some
-specification.
+event models is the deterministic finite automata which can
+be used in suprevisory control theory. In the theory these
+models are used, together with a formal specification of the
+desired behavior, to automatically generate a control
+function which makes the system behave within specification.
 
 % synchronization
-In suprevisory control theory the interaction between models
-is modeled by synchronization operator. This operator has to
-be implemented in control application in order to transfer
-theoretical results into operating control system. The
-implementation has to be distributable for it to be useful
-in control systems for the applications distributed over
-geographical area, like those mentioned in the beginning of
-the introduction.
+The interaction between models is modeled by synchronization
+operator. This operator has to be implemented in control
+application in order to transfer theoretical results into
+operating control system. The implementation has to be
+distributable for it to be useful in control systems for the
+applications distributed over geographical area, like the
+ones mentioned in the beginning of the introduction.
 
@@
 \end{displaymath} 
-where $q^1 \in Q^1$, $q^2 \in Q^2$ and $\sigma \in
-\Sigma$. $\Box$
-In other words, if an event is in both alphabets and it is
-enabled in both automatons, the transitions associated with
-it are made simultaneously in both automatons. Otherwise,
-only the automaton containing the event in its alphabet
-makes the transition. This definition can be extended for
-several automata.\\
+where $q^1 \in Q^1$, $q^2 \in Q^2$ and $\sigma \in \Sigma$.
+$\Box$ In other words, if an event is in both alphabets and
+it is enabled in both automatons, the transitions associated
+with it are made simultaneously in both automatons.
+Otherwise, only the automaton containing the event in its
+alphabet makes the transition. This definition can be
+extended for several automata.\\
 \textbf{Definition}: Full Synchronous Composition of $n$ DFAs\\
-The full synchronous composition of DFAs,
-$A^1$, $A^2$ ... $A^n$, is denoted ${A}^1||...||{A}^n=\langle
+The full synchronous composition of DFAs, $A^1$, $A^2$ ...
+$A^n$, is denoted ${A}^1||...||{A}^n=\langle
 Q,\Sigma,\delta,q_i,Q_m\rangle$ where
@@
   \begin{cases}
-    {\rm undefined} &, {\rm if\ } \exists A^k:\sigma\in\Sigma^k\land\sigma\notin\Gamma^k(q^k)\\
-    \langle q^{1+},\dots,q^{n+}\rangle & {\rm , otherwise}
+    {\rm undefined} &{\rm if\ } \exists A^k:\sigma\in\Sigma^k\land\sigma\notin\Gamma^k(q^k)\\
+    \langle q^{1+},\dots,q^{n+}\rangle & {\rm otherwise}
   \end{cases} 
@@
 \label{sec:sct}
-Suprevisory control theory provides definitions and results
-needed for caluation of allowed behavior of the system from
-discrete event models of possible behavior and desired
-behavior. The model of the possible behavior is called the
-plant, the model of the desired behavior is called the
-specification, and the result of the computation is called
-the suprevisor.
+Suprevisory control theory provides algorithms for
+calcuation of allowed behavior of the system from discrete
+event models of possible behavior and desired behavior. The
+model of the possible behavior is called the plant, the
+model of the desired behavior is called the specification,
+and the result of the computation is called the suprevisor.
 
@@
 all events might seem like a restriction. However, it is
-often possible to introduce a controller between the
-supervisor and the resources that initiate the events.
-Treating the controller as part of the plant will make the
-assumption that all events is generated by the plant a
-natural one. In Figure~\ref{fig:sup_plant} such configuration
-is shown.
+often possible to introduce a controller which will initiate
+the events between the supervisor and the resource. Treating
+the controller as part of the plant will make the assumption
+that all events are generated by the plant a natural one. In
+Fig.~\ref{fig:sup_plant} such configuration is shown.
 \begin{figure}
-  \centering \includegraphics[scale=0.5]{figures/sup_plant}
+  \centering \includegraphics[scale=0.6]{figures/sup_plant}
   \caption{Supervisory control configuration.}
@@
 
-The supervisor $S$ in Figure~\ref{fig:sup_plant} has to
+The supervisor, S in Fig.~\ref{fig:sup_plant}, has to
 prevent the plant from breaking the specification and from
 getting blocked in a state from where it can not reach a
-marked state by using as few restrictions of the plant as
+marked state by using as few restrictions on the plant as
 possible. By observing the observable events of the plant
-and disabling the controlable events of the plant, the
+and disabling the controllable events of the plant, the
 supervisor decides when, and if, it has to limit the
@@
 calculated from the models of the plant and specification so
-that the system behavior is given by
-$$S||P'.$$
+that
+$$S||P'$$
+gives the system behavior.
 
 In cases when there are several plant models $P_1,..,P_n$
-with appropriate specification models $Sp_1,...,Sp_n$ the
-supervisor can also be generated and the system behavior
-within specification is given by
-$$S||P_1||...||P_n.$$
-Thus to control the system so that it
-satisfies the specification, the synchronous composition of
-the plant and supervisor models has to be implemented in the
-control software. Using the conventional implementation
-methods, i.e. implementing the synchronization using PLC
-programming techniques, the solution becomes complex which
-makes it difficult, if not impossible, to distribute across
-multiple control computers. Next section presents an example
-illustrating how the syncrounous composition can be
-implemented using the new standard.
+with given specification models $Sp_1,...,Sp_n$ the
+supervisor can also be generated and
+$$S||P_1||...||P_n$$
+gives the system behavior within specification.
 
-\section{Example}
+Thus to control the system so that it satisfies the
+specification, the synchronous composition of the plant and
+supervisor models has to be implemented in the control
+software. Using the conventional implementation methods,
+i.e. implementing the synchronization using PLC programming
+techniques, the solution becomes complex which makes it
+difficult to distribute across multiple control computers.
+Next section presents an example illustrating how the
+synchronous composition can be implemented using the new
+standard.
+
+\section{Synchronization: An Example}
+\label{sec:example}
 In previous section it is shown that the synchronization
@@
 Consider a system containing two plants, \texttt{P1} and
-\texttt{P2}. The discrete event models for the plants are
-shown in the figure~\ref{fig:plants}. All events in the
-models are controllable and observable.The desired behavior
-of the system is that event \texttt{b} must always precede
-event \texttt{c}. The supervisor for this specification can
-be generated and is shown in figure~\ref{fig:sup}.
-\begin{figure}
+\texttt{P2}, shown in Fig.~\ref{fig:plants}.These plant
+models may represent two manufacturing resources with a
+1-place buffer in-between. All events in the models are
+controllable and observable. The desired behavior of the
+system is that event \texttt{b} must always precede event
+\texttt{c}, which means that a product must leave buffer
+before another one can enter. The supervisor for this
+specification can be generated and is shown in
+Fig.~\ref{fig:sup}. In cases where the specification is
+more complex it needs to be modeled using the automaton and
+provided for the supervisor generation.
+\begin{figure}[b]
   \centering 
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[b]
   \centering 
@@
 \end{figure}
-\begin{figure}[!ht]
+\begin{figure}[t]
   \centering 
-  \fig[scale=0.5]{\texttt{SYS}}{figures/Sys_aut}
+  \fig[scale=0.45]{\texttt{SYS}}{figures/Sys_aut}
   \caption{The model of possible system behavior while satisfying
@@
 specification. The resulting model of the system is shown in
-figure~\ref{fig:sys}.
+Fig.~\ref{fig:sys}.
 
 % IEC 61499 intro
-The emergin standard IEC 61499 specifies a generic
+The emerging standard IEC 61499 specifies a generic
 architecture for measurement and control of industrial
 processes. The architecture is based on functional software
-units called \textit{function blocks}. The most common type
-of function block in an application is basic function block.
-It executes algorithms based on the arriving events and
-generate new events that are passed on upon the algorithms
-execution. The algorithms use the internal variables and the
-data associated with the incoming events to update the
-internal variables and produce the output data associated
-with the outgoing events. What algorithms are invoked is
-controlled by the execution control chart (ECC) which is a
-automaton with algorithms associated with different states.
-The events are boolean expressions involving event inputs,
-internal, input, and output variables. In
-Figure~\ref{fig:61499_intro} a basic function block type is
-shown using the graphical representation.
+units called function blocks. The most common type of
+function block in an application is basic function block and
+it is shown in Fig.~\ref{fig:61499_intro} using the
+graphical representation. This type of block executes
+algorithms based on the arriving events and generates new
+events that are passed on when the algorithms finish
+execution. The algorithms use data associated with incoming
+events to update the internal variables and produce the
+output data associated with the outgoing events. Which
+algorithms are invoked is controlled by the execution
+control chart (ECC) which is an automaton with algorithms
+associated with different states. When a state is entered
+the algorithms associated with it are executed once and the
+ECC stays in the state until a condition for leaving is
+fulfilled. The conditions upon which transitions occur are
+boolean expressions involving event input variables,
+internal variables, input variables, and output variables.
+When the expression evaluates to \texttt{true} the
+transition occur. Special case is transition labled with
+\texttt{1}, it is taken as soon as the algorithms finish
+execution.
 \begin{figure}
   \centering
-  \includegraphics[scale=0.5]{figures/basic_block}\\(a)
-  \vskip 5mm
-  %\includegraphics[scale=0.5]{figures/basic_ecc}\\(b)
-  \caption{(a) The basic function block type and (b) its
-    execution control chart.}
+  \includegraphics[scale=0.4]{figures/basic_block}
+  \caption{Anatomy of a basic function block type.}
   \label{fig:61499_intro}
@@
 
-The event driven interface enables the distribution of the
-function blocks among the multiple control computers.
-
 Function block types that generate the behavior of the plant
-and specification models for the example are shown in
-figure~\ref{fig:systemFBs}. The event input of each block
+and supervisor models for the example are shown in
+Fig.~\ref{fig:systemFBs}. The event input of each block
 type is \texttt{OCCURRED}. It signals that an discrete event
@@
 The function block type that is generated according to the
-new method presented is shown in figure~\ref{fig:syncFB} and
+new method presented is shown in Fig.~\ref{fig:syncFB} and
 is called synchronization function block type. The data
@@
 of the block type the same for all applications. The ECC for
-the block type is shown in figure~\ref{fig:ecc_sync}. The
+the block type is shown in Fig.~\ref{fig:ecc_sync}. The
 \texttt{DISABLE\_EVENTS} algorithm associated with the
 \texttt{SYNC} state is where the synchronized execution is
-implemented. The transition labeled with the ``1'' character
-is taken as soon as the \texttt{DISABLE\_EVENTS} algorithm
-has finished executing and the \texttt{UPDATE} event is
-sent.
+implemented. The \texttt{UPDATE} event is sent as soon as
+the \texttt{DISABLE\_EVENTS} algorithm has finished
+execution.
 
@@
   \caption{The function block types implementing (a)
-    resource 1, (b) resource 2, and (c) specification.}
+    plant 1, (b) plant 2, and (c) supervisor.}
   \label{fig:systemFBs}
@@
   \label{fig:syncFB}
+\end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.6]{figures/sync_ecc}
+  \caption{The execution control chart for the
+    synchronization function block type.}
+  \label{fig:ecc_sync}
 \end{figure}
@@
 that event is not allowed to occur in the model during the
-next update cycle, if so needed. To calculate each data
-output a logical expression consisting of all data inputs
-that represent the same event, connected by logical
-\texttt{and} operation, is used. This means that if the
-event is enabled in all of the models that have the event in
-their alphabet it is allowed to occur in the next update
-cycle. Otherwise the expression is false and the the event
-is disabled. Following implementation of the
-\texttt{DISABLE\_EVENTS} algorithm in IEC 61131 structured
-text does the computation:
+next update cycle. To calculate each data output a logical
+expression consisting of all data inputs that represent the
+same event, connected by logical \texttt{and} operation, is
+used. This means that if the event is enabled in all of the
+models that have the event in their alphabet it is allowed
+to occur in the next update cycle. Otherwise the expression
+is false and the event is disabled. Following implementation
+of the \texttt{DISABLE\_EVENTS} algorithm in IEC 61131
+structured text does the computation:
+\begin{verbatim}
+EO_P1_a := EI_P1_a;
+EO_P1_b := EI_P1_b AND EI_S_b;
+EO_P2_c := EI_P2_c AND EI_S_c;
+EO_P2_d := EI_P2_d;
+EO_S_b := EI_P1_b AND EI_S_b;
+EO_S_c := EI_P2_c AND EI_S_c;
+\end{verbatim}
+Since the event \texttt{a} is in the alphabet of the
+\texttt{P1} model only, it is always allowed to occur if it
+is enabled. The event \texttt{b} is allowed to occur in
+\texttt{P1} model only if it is enabled in \texttt{P1} and
+\texttt{S} models at the same time since it is contained in
+both alphabets. The same applies to the event \texttt{b} in
+\texttt{S} model, hence the same logical expression. Events
+\texttt{d} and \texttt{c} are treated in the same manner.
 
+The control application connecting all of the function
+blocks together is shown in Fig.~\ref{fig:app}. Small
+circles in the figure are event multiplexer and
+demultiplexer. Event multiplexer generates an event on the
+output every time an event arrives at any one of the inputs.
+Event demultiplexer does the opossite and generates an event
+on all of the outputs whenever an event arrives at the
+input.
 \begin{figure}
   \centering
-  \includegraphics[scale=0.6]{figures/sync_ecc}
-  \caption{The execution control chart for the
-    synchronization function block type.}
-  \label{fig:ecc_sync}
-\end{figure}
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.35]{figures/app}
+  \includegraphics[scale=0.45]{figures/app}
   \caption{The control application for the example.}
@@
 
-\begin{verbatim}
-EO_R1_a := EI_R1_a;
-EO_R1_b := EI_R1_b AND EI_Sp_b;
-EO_R2_c := EI_R2_c AND EI_Sp_c;
-EO_R2_d := EI_R2_d;
-EO_Sp_b := EI_R1_b AND EI_Sp_b;
-EO_Sp_c := EI_R2_c AND EI_Sp_c;
-\end{verbatim}
-Since the event \texttt{a} is in the alphabet of the
-\texttt{R1} model only it is always allowed to occur if it
-is enabled. The event \texttt{b} is allowed to occur in
-\texttt{R1} model only if it is enabled in \texttt{R1} and
-\texttt{Sp} models at the same time since it is contained in
-both alphabets. The same applies to the event \texttt{b} in
-\texttt{Sp} model, hence the same logical expression. Events
-\texttt{d} and \texttt{c} are treated in the same manner.
+The method for generation of the \texttt{DISABLE\_EVENTS}
+algorithm can be generalized for any number of models. This
+general method is presented in the next section.
 
-\section{General Method}
+\section{Synchronization: General Approach}
 The method for implementation of the full synchronous
@@
 interface that these blocks have to adhere to is shown in
-figure~\ref{fig:model_interface}. These blocks are
+Fig.~\ref{fig:model_interface}. These blocks are
 application specific but they can be automatically generated
-from discrete event models~\cite{aff:sup:2003},
-\cite{www:sup:2004}.
+from discrete event models.
 
-\begin{figure}
+\begin{figure}[b]
   \centering
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[t]
   \centering
@@
 block type. The interface of this block is shown in the
-figure~\ref{fig:sync_interface}. The execution control chart
+Fig.~\ref{fig:sync_interface}. The execution control chart
 for the function block type is the same as shown in
-figure~\ref{fig:ecc_sync} since it is application generic.
+Fig.~\ref{fig:ecc_sync} since it is application generic.
 
-The \texttt{DISABLE\_EVENTS} algorithm is also application
-specific and it can be generated like this: Let $M$ be the
-set of all models for synchronization. For all $m,e$ such
-that $m\in M$ and $e\in\Sigma^m$:
+The \texttt{DISABLE\_EVENTS} algorithm can be generated as
+follows: Let $M$ be the set of all models for
+synchronization. For all $m,e$ such that $m\in M$ and
+$e\in\Sigma^m$:
 \begin{equation}\label{eq:lexp}
@@
 The iteration over all events of all models is for the
-generation of the logical expression assignments to all data
-outputs, which are represented by EO\_$m$\_$e$. The right
-hand side of~(\ref{eq:lexp}) generates the logical expression
-that covers all the cases of the synchronization definition.
-The first case in~(\ref{eq:delta}) is covered by
+generation of the logical expression assignments for all
+data outputs, which are represented by EO\_$m$\_$e$. The
+right hand side of~(\ref{eq:lexp}) generates the logical
+expression that covers all the cases of the synchronization
+definition. The first case in~(\ref{eq:delta}) is covered by
 $e\in\Sigma^n$ and the data input being \texttt{false} which
 gives that EO\_$m$\_$e$ is \texttt{false} and thus the event
-is disabled. The first case in equation~\ref{eq:qi} is
-covered by $f$ being in the $\Sigma^n\cap\Sigma^m$ and the
-data input being \texttt{true}. Finally, the second case in
-equation~\ref{eq:qi} is covered by $f$ being in the
-$\Sigma^n\cap\Sigma^m$. What is left now is to connect
+is disabled. The first case in (\ref{eq:qi}) is covered by
+$e\in\Sigma^n$ and the data input being \texttt{true}, and
+the second case in (\ref{eq:qi}) is covered by
+$e\notin\Sigma^n$.
+
+To further illustrate how the general method is applied
+refer to the \texttt{DISABLE\_EVENTS} algorithm of the
+example in Section~\ref{sec:example}. First two lines are
+generated for the \texttt{P1} model since it has two events
+in its alphabet. The first event gives the left hand side of
+the first line, \texttt{EO\_P1\_a}. The right hand side is
+generated by iterating over all models and looking for the
+event \texttt{a} in the alphabets. If the event is found the
+boolean variable for that model's event is added to the
+logical expression. In example, event \texttt{a} is found
+only in model \texttt{P1} and therefore the right hand side
+contains only variable \texttt{EI\_P1\_a}. The second event
+of the model \texttt{P1} gives the left hand side of the
+second line to be \texttt{EO\_P1\_b}. Looking for the event
+\texttt{b} in the models it is found in \texttt{P1} and
+\texttt{S} so the right hand side of (\ref{eq:lexp}) gives
+\texttt{EI\_P1\_b AND EI\_S\_b}. Continuing this procedure
+produces the whole \texttt{DISABLE\_EVENTS} algorithm in the
+example.
+
+When algorithm is generated what is left is to connect
 instances of the model function blocks with the
-synchronization function block, which concludes the method.
+synchronization function block to generate the control
+application. Some further considerations for the
+implementaion of supervisory control theory are presented in
+the following section.
 
-\section{Application of supervisory control theory}
+\section{Applications in supervisory control theory}
 \label{sec:app_SCT}
-Blah blah.
+Since the general method presented is suitable for computer
+implementation it is interesting to investigate how this
+fact can ease the implementation of control applications
+using supervisory control theory.
+
+Today implementation of a distributed control application is
+done using application structure similar to the structure in
+Fig.~\ref{fig:old_app_struct}. Supervisor \texttt{S} in
+the application is manualy crafted using informal
+specificatons of the system behavior, and transfered to the
+PLC-code. For each application the communication between the
+suprevisor and the plants has to be implemented in a
+application specific way since the existing standards for
+communication between PLCs are at a low level of
+abstraction. Using IEC 61499 the communication part of the
+application is abstracted far away from the hardware level
+and the control engineer is only considering the event and
+data passing between the functional parts of the
+application, i.e. function block instances. When the system
+plants are upgraded or chaged the procedure of supervisor
+crafting has to be done once again and very small parts of
+the code from the application already implemented are
+reused, i.e. the communication parts.
+\begin{figure}
+  \centering 
+  \includegraphics[scale=0.6]{figures/old_app_struct}
+  \caption{The structure of distributed control applications today.}
+  \label{fig:old_app_struct}
+\end{figure}
+
+In software tool Supremica \cite{aff:sup:2003,www:sup:2004},
+algorithms from supervisory control theory are implemented
+so that a supervisor can be synthesised from automata models
+of the plants and specifications. The algorithm that
+implements the general method for synchronization according
+to this paper is also implemented in the same tool. What
+this means is that the control engineer can focus on the
+correct modeling of the plants and transfering of the
+informal specifications to the specification models for use
+in the Supremica tool. Then the tool automatically generates
+the supervisor in form of an automaton model, function block
+types that implement the skeleton for the plants, the
+supervisor automaton, synchronous composition function block
+and finally the control application that connects the
+instances of the generated function block types as seen in
+Fig.~\ref{fig:app} for the example. The sekelton in this
+case refers to the fact that the communication with the
+hardware has to be implemented by hand but all the behavior
+of the models is automatically generated. Using the tool to
+generate the control application gives new application
+structure as shown in Fig.~\ref{fig:new_app_struct}. In
+this structure, only the models for the generation of the
+plant blocks and the hardware control portion of those
+blocks have to be upgraded whenever the plants in the system
+change. All other parts of the application are regenerated
+automatically.
+\begin{figure}[b]
+  \centering 
+  \includegraphics[scale=0.6]{figures/new_app_struct}
+  \caption{A new structure for distributed control applications.}
+  \label{fig:new_app_struct}
+\end{figure}
+
+When trying to implement the control application for complex
+systems with many plants and specifications the problem of
+state space explosion is encountered when trying to
+synthesise the supervisor model. There are methods that
+represent the plants, specifications and the supervisor in
+different ways to try to cope with the problem. For example,
+a supervisor may be expressed as a singel automaton,
+multiple automata, and a binary decision diagram for
+off-line synthesis. Or the supervisor may be synthesized
+on-line while the control application is executing. The
+method presented in this paper can be used for
+synchronization of all types of representations since the
+model interface (see Fig.~\ref{fig:model_interface}) is
+based on events which are a general characteristic of the
+discrete event models.
 
 \section{Conclusion}
-In many distributed control applications the need for
-synchronous execution of discrete event models is present.
-The two methods presented in this paper may be used to
-implement such synchronous execution by using IEC 61499
-function blocks. The first method is useful in case the
-function blocks implementing the models that need to be
-synchronized have to be distributed across different
-computing resources in the control system. It also results
-in a slightly more complex function block program. The
-second method on the other hand results in a single function
-block type but it is only useful when the synchronized
-models can fit into single computing resource.
+Using supervisory control theory in construction of control
+applications can significantly reduce development time but
+until now it was dificult to apply the theory to
+implementation of distributed control applications. Using
+the emerging standard IEC 61499 it becomes significantly
+easier. The method presented in this paper brings
+supervisory control theory together with distributed control
+application development in an efficient way.
 
-The structure of the function block types in both methods is
-obvious. This makes implementation of computer programs for
-the generation of such blocks a straight forward task. In
-fact, the function block type generation for the second
-method is already implemented in the Supremica tool,
-\cite{aff:sup:2003} and \cite{www:sup:2004}. Work is under
-way to implement automatic generation for the first method
-too. Future considerations should be made to accommodate for
-the controllable and uncontrollable events in the automatic
-generation of the synchronization function block types.
+Since the method is implemented in the Supremica tool it
+significantly contributes to the shorter control application
+development time since it is a part in automatic generation
+of control applications. The method is also future proof by
+using an inherent characteristic of discrete event models
+for the interface upon which synchronization function block
+is generated.
 
@@
 
-\bibliography{../../../Bibliographies/desbib.bib}
+\bibliography{/home/cengic/research/Bibliographies/desbib.bib}
 

### 2005-01-31 doc/Sync_Execution_with_IEC61499/paper.tex
*** empty log message ***

@@
 mentioned above and it becomes even larger in mission
-critical control applications since the strict requirenment
+critical control applications since the strict requirement
 of guaranteeing the correct operation is imposed. Thus,
@@
 
-% suprevisory control
+% supervisory control
 Many man-made systems can, at some abstraction level, be
@@
 event models is the deterministic finite automata which can
-be used in suprevisory control theory. In the theory these
+be used in supervisory control theory. In the theory these
 models are used, together with a formal specification of the
@@
 
-% paper organisation
+% paper organization
 Mathematical definitions that are used in the paper are
 presented in the next section. After that, the most
-important definitions and results of suprevisory control
-theory are introduced followed by an exampel showing how
+important definitions and results of supervisory control
+theory are introduced followed by an example showing how
 supervisory control theory can be applied in implementation
@@
 $\Box$\\
-What this definiton states is that in a group of automatons
+What this definition states is that in a group of automatons
 that share one event if there is one automaton that does not
@@
 the same time as all the other automatons. This property of
-the definiton is used to construct a supervisor which
+the definition is used to construct a supervisor which
 prevents events from occurring. The concept of supervisor is
@@
 \label{sec:sct}
-Suprevisory control theory provides algorithms for
-calcuation of allowed behavior of the system from discrete
+Supervisory control theory provides algorithms for
+calculation of allowed behavior of the system from discrete
 event models of possible behavior and desired behavior. The
@@
 model of the desired behavior is called the specification,
-and the result of the computation is called the suprevisor.
+and the result of the computation is called the supervisor.
 
@@
 prevent the plant from generating events that will violate
-the specfication. The assumption that the plant generates
+the specification. The assumption that the plant generates
 all events might seem like a restriction. However, it is
@@
 When the expression evaluates to \texttt{true} the
-transition occur. Special case is transition labled with
+transition occur. Special case is transition labeled with
 \texttt{1}, it is taken as soon as the algorithms finish
@@
 output every time an event arrives at any one of the inputs.
-Event demultiplexer does the opossite and generates an event
+Event demultiplexer does the opposite and generates an event
 on all of the outputs whenever an event arrives at the
@@
 application. Some further considerations for the
-implementaion of supervisory control theory are presented in
+implementation of supervisory control theory are presented in
 the following section.
@@
 Fig.~\ref{fig:old_app_struct}. Supervisor \texttt{S} in
-the application is manualy crafted using informal
-specificatons of the system behavior, and transfered to the
+the application is manually crafted using informal
+specifications of the system behavior, and transfered to the
 PLC-code. For each application the communication between the
-suprevisor and the plants has to be implemented in a
+supervisor and the plants has to be implemented in a
 application specific way since the existing standards for
@@
 application, i.e. function block instances. When the system
-plants are upgraded or chaged the procedure of supervisor
+plants are upgraded or changed the procedure of supervisor
 crafting has to be done once again and very small parts of
@@
 algorithms from supervisory control theory are implemented
-so that a supervisor can be synthesised from automata models
+so that a supervisor can be synthesized from automata models
 of the plants and specifications. The algorithm that
@@
 this means is that the control engineer can focus on the
-correct modeling of the plants and transfering of the
+correct modeling of the plants and transferring of the
 informal specifications to the specification models for use
@@
 instances of the generated function block types as seen in
-Fig.~\ref{fig:app} for the example. The sekelton in this
+Fig.~\ref{fig:app} for the example. The skeleton in this
 case refers to the fact that the communication with the
@@
 state space explosion is encountered when trying to
-synthesise the supervisor model. There are methods that
+synthesize the supervisor model. There are methods that
 represent the plants, specifications and the supervisor in
 different ways to try to cope with the problem. For example,
-a supervisor may be expressed as a singel automaton,
+a supervisor may be expressed as a single automaton,
 multiple automata, and a binary decision diagram for
@@
 applications can significantly reduce development time but
-until now it was dificult to apply the theory to
+until now it was difficult to apply the theory to
 implementation of distributed control applications. Using

### 2005-01-31 doc/Sync_Execution_with_IEC61499/paper.tex
*** empty log message ***

@@
   event models is used to model interaction between systems
-  in supervisory control theory. Method is demonstrated
+  in supervisory control theory. The method is demonstrated
   using a simple example before the formulation for the
@@
 critical control applications since the strict requirement
-of guaranteeing the correct operation is imposed. Thus,
+of guaranteeing, the correct operation is imposed. Thus,
 there is a need to introduce algorithms and tools to make it
@@
 others from executing an event if it does not enable it at
-the same time as all the other automatons. This property of
+the same time as all the other automata. This property of
 the definition is used to construct a supervisor which
@@
 i.e. implementing the synchronization using PLC programming
-techniques, the solution becomes complex which makes it
+techniques, the solution becomes complex, which makes it
 difficult to distribute across multiple control computers.
 Next section presents an example illustrating how the
-synchronous composition can be implemented using the new
+synchronous composition might be implemented using the new
 standard.
@@
 boolean variables representing the enabled events after the
-model has updated its state, i.e. the events labeling the
-transitions that are possible from the new state.
+model has updated its state, i.e. the events that are possible from the new state.
 
@@
 The method for implementation of the full synchronous
-composition for the general system containing any number of
-discrete event models is made up of two major steps. The
+composition for the general system, containing any number of
+discrete event models, is made up of two major steps. The
 first step is creation of the function block types that

### 2005-01-31 doc/Sync_Execution_with_IEC61499/paper.tex
*** empty log message ***

@@
 \begin{abstract}
-  A new method for implementation of synchronous composition
-  is presented. Using an emerging standard, IEC 61499, the
-  method enables efficient development of distributed
-  control applications. It also brings together supervisory
-  control theory with distributed control application
-  development since synchronous composition of discrete
-  event models is used to model interaction between systems
-  in supervisory control theory. The method is demonstrated
-  using a simple example before the formulation for the
-  general case is presented. A new structure of control
-  applications suitable for application of supervisory
-  control theory to control application development is
-  presented also.
+  One reason that the development of distributed control
+  system is hard is that the communication primitives are at
+  a low abstraction level. To put less burden on the
+  developer it would be an advantage to have support for
+  high-level communication primitives. One such primitive is
+  the synchronous composition operator. This operator can be
+  used to model the interaction between an arbitrary number
+  of concurrently executing applications. This paper shows
+  how to implement the synchronous composition operator
+  using an emerging communication standard for distributed
+  control systems, IEC 61499. The operator is used in the
+  supervisory control theory, a framework for automatic
+  generation of provable correct supervisors. In
+  manufacturing systems, for example, the supervisor might
+  have to interact with a physically distributed plant. This
+  paper concludes therefore by showing how to use the
+  proposed implementation of the synchronous composition
+  operator to execute a supervisor against a physically
+  distributed plant.
+
 \end{abstract}
+
+% supervisory control theory, discrete event systems,
+% manufacturing systems, synchronous composition, IEC 61499,
+% Distributed Control Systems
 
@@
 synchronization operator using PLC programming techniques
-\cite{hfl:syn:1999,hfl:mod:2001,fh:plc:1998} but the
-distribution is hard to achieve in that programming
+\cite{hfl:syn:1999,hfl:mod:2001,fh:plc:1998,a:met:2002} but
+the distribution is hard to achieve in that programming
 environment. A new method presented in this paper uses the
@@
 \begin{figure}
-  \centering \includegraphics[scale=0.6]{figures/sup_plant}
-  \caption{Supervisory control configuration.}
+  \centering 
+  \includegraphics[scale=0.6]{figures/sup_plant}
+  \caption{Supervisory control configuration with a resource
+  \texttt{R}, controller \texttt{C}, and a supervisor \texttt{S}.}
   \label{fig:sup_plant}
@@
 
-The supervisor, S in Fig.~\ref{fig:sup_plant}, has to
-prevent the plant from breaking the specification and from
-getting blocked in a state from where it can not reach a
-marked state by using as few restrictions on the plant as
-possible. By observing the observable events of the plant
-and disabling the controllable events of the plant, the
-supervisor decides when, and if, it has to limit the
-activity of the plant in order to keep the plant's behavior
-within the specification. An important result of the
-supervisory control theory is that such a supervisor can be
-calculated from the models of the plant and specification so
-that
-$$S||P'$$
+The supervisor, \texttt{S} (see Fig.~\ref{fig:sup_plant}),
+has to prevent the plant, \texttt{P}, from breaking the
+specification and from getting blocked in a state from where
+it can not reach a marked state by using as few restrictions
+on the plant as possible. By observing the observable events
+of the plant and disabling the controllable events of the
+plant, the supervisor decides when, and if, it has to limit
+the activity of the plant in order to keep the plant's
+behavior within the specification. An important result of
+the supervisory control theory is that such a supervisor can
+be calculated from the models of the plant and specification
+so that
+$$S||P$$
 gives the system behavior.
@@
   \fig[scale=0.55]{\texttt{P2}}{figures/P2_aut}
-  \caption{Discrete event models of the plants.}
+  \caption{Discrete event models of the plants. These may
+    represent two manufacturing resources with a 1-place
+    buffer in-between.}
   \label{fig:plants}
@@
   \fig[scale=0.55]{\texttt{S}}{figures/Sp_aut}
-  \caption{The model of the supervisor.}
+  \caption{The model of the supervisor that make shure that
+    event \texttt{b} always occurs before event \texttt{c}.}
   \label{fig:sup}
@@
   \centering
-  \includegraphics[scale=0.4]{figures/basic_block}
+  \includegraphics[scale=0.65]{figures/basic_block}
   \caption{Anatomy of a basic function block type.}
@@
 and supervisor models for the example are shown in
-Fig.~\ref{fig:systemFBs}. The event input of each block
-type is \texttt{OCCURRED}. It signals that an discrete event
-has occurred and that the model should update its state
+Fig.~\ref{fig:systemFBs}. The event input of each block type
+is \texttt{OCCURRED}. It signals that an discrete event has
+occurred and that the model should update its state
 according to the data inputs. The data inputs are boolean
@@
 boolean variables representing the enabled events after the
-model has updated its state, i.e. the events that are possible from the new state.
+model has updated its state, i.e. the events that are
+possible from the new state.
 
@@
   \centering
-  \includegraphics[scale=0.5]{figures/P1}\\(a)
+  \includegraphics[scale=0.45]{figures/P1}\\(a)
   \vskip 5mm
-  \includegraphics[scale=0.5]{figures/P2}\\(b)
+  \includegraphics[scale=0.45]{figures/P2}\\(b)
   \vskip 5mm
-  \includegraphics[scale=0.5]{figures/Sp}\\(c)
+  \includegraphics[scale=0.45]{figures/Sp}\\(c)
   \caption{The function block types implementing (a)
@@
   \includegraphics[scale=0.5]{figures/SYNC}
-  \caption{The function block type for the synchronization.}
+  \caption{The function block type for synchronization.}
   \label{fig:syncFB}
@@
   \centering
-  \includegraphics[scale=0.6]{figures/sync_ecc}
+  \includegraphics[scale=0.45]{figures/sync_ecc}
   \caption{The execution control chart for the
@@
   \includegraphics[scale=0.45]{figures/app}
-  \caption{The control application for the example.}
+  \caption{The control application for the example with
+  function block instances for plants, \texttt{P1} and
+  \texttt{P2}, supervisor, \texttt{S}, and synchronization
+  block, \texttt{SYNC}.}
   \label{fig:app}
@@
 
-Today implementation of a distributed control application is
-done using application structure similar to the structure in
-Fig.~\ref{fig:old_app_struct}. Supervisor \texttt{S} in
-the application is manually crafted using informal
-specifications of the system behavior, and transfered to the
-PLC-code. For each application the communication between the
-supervisor and the plants has to be implemented in a
-application specific way since the existing standards for
-communication between PLCs are at a low level of
-abstraction. Using IEC 61499 the communication part of the
-application is abstracted far away from the hardware level
-and the control engineer is only considering the event and
-data passing between the functional parts of the
-application, i.e. function block instances. When the system
-plants are upgraded or changed the procedure of supervisor
-crafting has to be done once again and very small parts of
-the code from the application already implemented are
-reused, i.e. the communication parts.
+Today most implementations of distributed control
+applications are done using application structure similar to
+the structure in Fig.~\ref{fig:old_app_struct}. The
+controller, \texttt{C}, in application is manually crafted
+using informal specifications of the system behavior, and
+transfered to the PLC-code. For each application the
+communication between the supervisor and the plants has to
+be implemented in an application and vendor specific way
+since the existing standards for communication between PLCs
+are at a low level of abstraction. Using IEC 61499 the
+communication part of the application is abstracted away
+from the hardware level and the control engineer is only
+considering the event and data passing between the
+functional parts of the application, i.e. function block
+instances. When the system plants are upgraded or changed
+the procedure of supervisor crafting has to be done once
+again and very small parts of the code from the application
+already implemented are reused, i.e. the communication
+parts.
 \begin{figure}
@@
   \includegraphics[scale=0.6]{figures/old_app_struct}
-  \caption{The structure of distributed control applications today.}
+  \caption{The structure of distributed control applications
+  today with plants, \texttt{P1} and \texttt{P2}, and
+  controller, \texttt{C}.}
   \label{fig:old_app_struct}

### 2005-02-01 doc/Sync_Execution_with_IEC61499/paper.tex
*** empty log message ***

@@
 \begin{document}
-\title{Implementation of Full Synchronous Composition of
-  Discrete Event Models Using IEC 61499 Function Blocks}
+\title{Implementation of Full Synchronous Composition Using
+  IEC 61499 Function Blocks}
 
@@
 Shorter product life-cycles increase the demands on the
-manufacturing system. Since the product, to be produced, in
-the manufacturing plant is changing frequently the
-production equipment needs to be modified too. The frequent
-changes of the system make it time consuming and expensive
-to update the control system application to reflect the
-changes. The same problem applies to both system types
-mentioned above and it becomes even larger in mission
-critical control applications since the strict requirement
-of guaranteeing, the correct operation is imposed. Thus,
-there is a need to introduce algorithms and tools to make it
-simple and as safe as possible to modify the control system.
-The supervisory control theory,\cite{rw:con:1989}, is a
-promising framework that might be useful for this purpose.
+manufacturing system. Since products to be produced in
+manufacturing plants are changing frequently, the production
+equipment also needs to be modified. Frequent changes of the
+system make it time consuming and expensive to update the
+control system application to reflect the such changes. The
+problem becomes even larger in mission critical control
+applications since strict requirement of guaranteeing
+correct operation is then imposed. Thus, there is a need to
+introduce algorithms and tools to make it simple and as safe
+as possible to modify the control system. The supervisory
+control theory,\cite{rw:con:1989}, is a promising framework
+that might be useful for this purpose.
 
@@
 modeled by discrete event models. A useful class of discrete
-event models is the deterministic finite automata which can
+event models is the deterministic finite automata, which can
 be used in supervisory control theory. In the theory these
-models are used, together with a formal specification of the
-desired behavior, to automatically generate a control
-function which makes the system behave within specification.
+models are used together with a formal specification of the
+desired behavior to automatically generate a control
+function which makes the system behave within the given
+specification.
 
 % synchronization
-The interaction between models is modeled by synchronization
-operator. This operator has to be implemented in control
-application in order to transfer theoretical results into
-operating control system. The implementation has to be
-distributable for it to be useful in control systems for the
-applications distributed over geographical area, like the
-ones mentioned in the beginning of the introduction.
+The interaction between subsystems is modeled by Hoare's
+full synchronous composition (FSC)~\cite{h:com:1985}. This
+operator has to be implemented in control applications in
+order to transfer theoretical results into operating control
+systems. The implementation has to be distributable to be
+applicable in control systems distributed over geographical
+areas, like the ones mentioned above.
 
-There are existing methods for implementation of
-synchronization operator using PLC programming techniques
+There are existing methods for implementation of FSC using
+PLC programming techniques
 \cite{hfl:syn:1999,hfl:mod:2001,fh:plc:1998,a:met:2002} but
@@
 Mathematical definitions that are used in the paper are
-presented in the next section. After that, the most
-important definitions and results of supervisory control
-theory are introduced followed by an example showing how
-supervisory control theory can be applied in implementation
-of control application for a small system. After the example
-a method for implementation of the synchronization operator
-for any number of synchronized models is presented together
-with the argument for the correctness of the implementation.
-In Section~\ref{sec:app_SCT} some considerations for
-implementation of supervisory control theory in applications
-are presented before the conclusion of the paper.
+presented in the next section. Important definitions and
+results of supervisory control theory are introduced,
+followed by an example showing how supervisory control
+theory can be applied in implementation of control
+application for a small system. After the example a method
+for implementation of FSC for any number of synchronized
+models is presented together with the argument for the
+correctness of the implementation. Finally some
+considerations for implementation of supervisory control
+theory in applications are presented before the conclusion
+of the paper.
 
@@
 are presented. Below, the terminology in this paper is
-introduced.\\
-\textbf{Definition:} Deterministic Finite Automaton (DFA)\\
+introduced.
+
+\textbf{Definition 1:} Deterministic Finite Automaton (DFA)\\
 A deterministic finite automaton is a 5-tuple defined as
@@
 is the \emph{initial state} and $Q_m \subseteq Q$ is a set
-of \emph{marked states}. $\Box$\\
-\textbf{Definition:} Enabled Event Function\\
+of \emph{marked states}.\hfill$\Box$
+
+\textbf{Definition 2:} Enabled Event Function\\
 The enabled event function $\Gamma$ is defined as
@@
 set of all events $\sigma$ for which $\delta(q,\sigma)$ is
-defined. $\Box$\\
+defined.\hfill$\Box$\\
 Automata are usually visualized by directed graphs with
@@
 used in this paper to model the interaction between the
-system models.\\
-\textbf{Definition}: Full Synchronous Composition of two DFAs\\
+system models.
+
+\textbf{Definition 3}: Full Synchronous Composition of two DFAs\\
 The full synchronous composition of two DFAs, $A^1$ and
@@
 where $q^1 \in Q^1$, $q^2 \in Q^2$ and $\sigma \in \Sigma$.
-$\Box$ In other words, if an event is in both alphabets and
-it is enabled in both automatons, the transitions associated
-with it are made simultaneously in both automatons.
+\hfill$\Box$\\
+In other words, if an event is in both alphabets and
+it is enabled in both automata, the transitions associated
+with it are made simultaneously in both automata.
 Otherwise, only the automaton containing the event in its
 alphabet makes the transition. This definition can be
-extended for several automata.\\
-\textbf{Definition}: Full Synchronous Composition of $n$ DFAs\\
+extended for several automata.
+
+\textbf{Definition 4}: Full Synchronous Composition of $n$ DFAs\\
 The full synchronous composition of DFAs, $A^1$, $A^2$ ...
@@
 \end{equation}
-$\Box$\\
-What this definition states is that in a group of automatons
-that share one event if there is one automaton that does not
+\hfill$\Box$\\
+What this definition states is that, in a group of automata
+that share one event, if there is one automaton that does not
 have that event enabled in its current state, than the event
-is not allowed to occur in any of the other automatons in
+is not allowed to occur in any of the other automata in
 the group. That is, one automaton is able to prevent all the
@@
 has to prevent the plant, \texttt{P}, from breaking the
-specification and from getting blocked in a state from where
-it can not reach a marked state by using as few restrictions
+specification and from getting blocked in a state, from where
+it can not reach a marked state, by using as few restrictions
 on the plant as possible. By observing the observable events
-of the plant and disabling the controllable events of the
-plant, the supervisor decides when, and if, it has to limit
+of the plant and disabling controllable events of the
+plant, the supervisor decides when and if it has to limit
 the activity of the plant in order to keep the plant's
@@
 $$S||P$$
-gives the system behavior.
+gives the desired system behavior with as few
+restriction of the plant as possible.
 
@@
 $$S||P_1||...||P_n$$
-gives the system behavior within specification.
+gives the system behavior within the specification.
 
-Thus to control the system so that it satisfies the
-specification, the synchronous composition of the plant and
-supervisor models has to be implemented in the control
-software. Using the conventional implementation methods,
-i.e. implementing the synchronization using PLC programming
-techniques, the solution becomes complex, which makes it
-difficult to distribute across multiple control computers.
-Next section presents an example illustrating how the
-synchronous composition might be implemented using the new
-standard.
+Hence, to control the system so that it satisfies the
+specification, the FSC of the plants and supervisor models
+has to be implemented in the control software. Using the
+conventional implementation methods, i.e. implementing the
+synchronization using PLC programming techniques makes it
+difficult to distribute the application across multiple
+control computers, which is needed when the plants are
+distributed over geographical areas. Next section presents
+an example illustrating how the synchronous composition
+might be implemented using the new standard.
 
@@
 Consider a system containing two plants, \texttt{P1} and
-\texttt{P2}, shown in Fig.~\ref{fig:plants}.These plant
-models may represent two manufacturing resources with a
-1-place buffer in-between. All events in the models are
-controllable and observable. The desired behavior of the
-system is that event \texttt{b} must always precede event
-\texttt{c}, which means that a product must leave buffer
-before another one can enter. The supervisor for this
-specification can be generated and is shown in
-Fig.~\ref{fig:sup}. In cases where the specification is
-more complex it needs to be modeled using the automaton and
-provided for the supervisor generation.
+\texttt{P2}, shown in Fig.~\ref{fig:plants}. All events in
+the models are controllable and observable. The desired
+behavior of the system is that event \texttt{b} must always
+precede event \texttt{c}. These plant models together with
+the specification may represent two manufacturing resources
+with a 1-place buffer in-between. The specification means
+then that a product must leave buffer before another one can
+enter. The supervisor for this specification can be
+generated and is shown in Fig.~\ref{fig:sup}.
 \begin{figure}[b]
@@
   \fig[scale=0.55]{\texttt{S}}{figures/Sp_aut}
-  \caption{The model of the supervisor that make shure that
+  \caption{The model of the supervisor that makes sure 
     event \texttt{b} always occurs before event \texttt{c}.}
@@
 \begin{figure}[t]
-  \centering 
-  \fig[scale=0.45]{\texttt{SYS}}{figures/Sys_aut}
+  \centering \fig[scale=0.4]{\texttt{SYS}}{figures/Sys_aut}
   \caption{The model of possible system behavior while satisfying
@@
 \end{figure}
-Full synchronous execution of these models gives the system
-that behaves as freely as possible while fulfilling the
-specification. The resulting model of the system is shown in
-Fig.~\ref{fig:sys}.
+Full synchronous execution of \texttt{P1}, \texttt{P2} and
+\texttt{S} gives the system that behaves as freely as
+possible while fulfilling the specification. The resulting
+model of the system is shown in Fig.~\ref{fig:sys}.
 
-% IEC 61499 intro
+\subsection{IEC 61499 Basics}
 The emerging standard IEC 61499 specifies a generic
@@
 units called function blocks. The most common type of
-function block in an application is basic function block and
-it is shown in Fig.~\ref{fig:61499_intro} using the
-graphical representation. This type of block executes
-algorithms based on the arriving events and generates new
-events that are passed on when the algorithms finish
-execution. The algorithms use data associated with incoming
-events to update the internal variables and produce the
-output data associated with the outgoing events. Which
-algorithms are invoked is controlled by the execution
-control chart (ECC) which is an automaton with algorithms
-associated with different states. When a state is entered
-the algorithms associated with it are executed once and the
-ECC stays in the state until a condition for leaving is
-fulfilled. The conditions upon which transitions occur are
-boolean expressions involving event input variables,
-internal variables, input variables, and output variables.
-When the expression evaluates to \texttt{true} the
-transition occur. Special case is transition labeled with
-\texttt{1}, it is taken as soon as the algorithms finish
-execution.
+function block in an application is the basic function block
+shown in Fig.~\ref{fig:61499_intro} using the graphical
+representation. This type of block executes algorithms based
+on the arriving events and generates new events that are
+passed on when the algorithms finish execution. The
+algorithms use data associated with incoming events to
+update the internal variables and produce the output data
+associated with the outgoing events. Which algorithms are
+invoked is controlled by the execution control chart (ECC),
+which is an automaton with algorithms associated with
+different states. When a state is entered the algorithms
+associated with it are executed once and the ECC stays in
+the state until a condition for leaving is fulfilled. The
+conditions upon which transitions occur are boolean
+expressions involving event input variables, internal
+variables, input variables, and output variables. When the
+expression evaluates to \texttt{true} the transition occur.
+Special case is transition labeled with \texttt{1}; it is
+taken as soon as the algorithms finish execution.
 \begin{figure}
   \centering
-  \includegraphics[scale=0.65]{figures/basic_block}
+  \includegraphics[scale=0.8]{figures/basic_block}
   \caption{Anatomy of a basic function block type.}
@@
 
+\subsection{Example Implementation}
 Function block types that generate the behavior of the plant
@@
 Fig.~\ref{fig:systemFBs}. The event input of each block type
-is \texttt{OCCURRED}. It signals that an discrete event has
+is \texttt{OCCURRED}. It signals that a discrete event has
 occurred and that the model should update its state
@@
 are the same for all applications which also makes the ECC
-of the block type the same for all applications. The ECC for
-the block type is shown in Fig.~\ref{fig:ecc_sync}. The
-\texttt{DISABLE\_EVENTS} algorithm associated with the
-\texttt{SYNC} state is where the synchronized execution is
-implemented. The \texttt{UPDATE} event is sent as soon as
-the \texttt{DISABLE\_EVENTS} algorithm has finished
-execution.
+of the block type the same for all applications.
+
+The synchronization block reads the enabled events of the
+synchronized models and disables the events that are not
+enabled in all the models that have that event in their
+alphabets. Each boolean data input signals if the event in
+the model represented by that input is enabled
+(\texttt{true}) or disabled (\texttt{false}). By setting the
+corresponding data output to \texttt{false} the
+synchronization block signals that the transition labeled by
+that event is not allowed to occur in the model during the
+next update cycle.
+
+The ECC for the block type is shown in
+Fig.~\ref{fig:ecc_sync}. The \texttt{DISABLE\_EVENTS}
+algorithm associated with the \texttt{SYNC} state is where
+the synchronized execution is implemented. The
+\texttt{UPDATE} event is sent as soon as the
+\texttt{DISABLE\_EVENTS} algorithm has finished execution.
+The following implementation of the algorithm in IEC 61131
+structured text does the computation:
+\begin{verbatim}
+EO_P1_a := EI_P1_a;
+EO_P1_b := EI_P1_b AND EI_S_b;
+EO_P2_c := EI_P2_c AND EI_S_c;
+EO_P2_d := EI_P2_d;
+EO_S_b := EI_P1_b AND EI_S_b;
+EO_S_c := EI_P2_c AND EI_S_c;
+\end{verbatim}
+To calculate each data output a logical expression
+consisting of all data inputs that represent the same event,
+connected by logical \texttt{and} operation, is used. This
+means that if the event is enabled in all of the models that
+have the event in their alphabet it is allowed to occur in
+the next update cycle. Otherwise the expression is false and
+the event is disabled.
+
+Since the event \texttt{a} is in the alphabet of the
+\texttt{P1} model only, it is always allowed to occur if it
+is enabled. The event \texttt{b} is allowed to occur in
+\texttt{P1} model only if it is enabled in \texttt{P1} and
+\texttt{S} models at the same time since it is contained in
+both alphabets. The same applies to the event \texttt{b} in
+\texttt{S} model, hence the same logical expression. Events
+\texttt{d} and \texttt{c} are treated in the same manner.
 
@@
   \centering
-  \includegraphics[scale=0.45]{figures/P1}\\(a)
+  \includegraphics[scale=0.5]{figures/P1}\\(a)
   \vskip 5mm
-  \includegraphics[scale=0.45]{figures/P2}\\(b)
+  \includegraphics[scale=0.5]{figures/P2}\\(b)
   \vskip 5mm
-  \includegraphics[scale=0.45]{figures/Sp}\\(c)
+  \includegraphics[scale=0.5]{figures/Sp}\\(c)
   \caption{The function block types implementing (a)
@@
   \centering
-  \includegraphics[scale=0.45]{figures/sync_ecc}
+  \includegraphics[scale=0.5]{figures/sync_ecc}
   \caption{The execution control chart for the
@@
 \end{figure}
-
-The synchronization block reads the enabled events of the
-synchronized models and disables the events that are not
-enabled in all the models that have that event in their
-alphabets. Each boolean data input signals if the event in
-the model represented by that input is enabled
-(\texttt{true}) or disabled (\texttt{false}). By setting the
-corresponding data output to \texttt{false} the
-synchronization block signals that the transition labeled by
-that event is not allowed to occur in the model during the
-next update cycle. To calculate each data output a logical
-expression consisting of all data inputs that represent the
-same event, connected by logical \texttt{and} operation, is
-used. This means that if the event is enabled in all of the
-models that have the event in their alphabet it is allowed
-to occur in the next update cycle. Otherwise the expression
-is false and the event is disabled. Following implementation
-of the \texttt{DISABLE\_EVENTS} algorithm in IEC 61131
-structured text does the computation:
-\begin{verbatim}
-EO_P1_a := EI_P1_a;
-EO_P1_b := EI_P1_b AND EI_S_b;
-EO_P2_c := EI_P2_c AND EI_S_c;
-EO_P2_d := EI_P2_d;
-EO_S_b := EI_P1_b AND EI_S_b;
-EO_S_c := EI_P2_c AND EI_S_c;
-\end{verbatim}
-Since the event \texttt{a} is in the alphabet of the
-\texttt{P1} model only, it is always allowed to occur if it
-is enabled. The event \texttt{b} is allowed to occur in
-\texttt{P1} model only if it is enabled in \texttt{P1} and
-\texttt{S} models at the same time since it is contained in
-both alphabets. The same applies to the event \texttt{b} in
-\texttt{S} model, hence the same logical expression. Events
-\texttt{d} and \texttt{c} are treated in the same manner.
 
@@
 
-When algorithm is generated what is left is to connect
+When the algorithm is generated what is left is to connect
 instances of the model function blocks with the
@@
 application. Some further considerations for the
-implementation of supervisory control theory are presented in
-the following section.
+implementation of supervisory control theory are presented
+in the following section.
 
 \section{Applications in supervisory control theory}
-\label{sec:app_SCT}
+\label{sec:app_sct}
 Since the general method presented is suitable for computer
@@
 
-In software tool Supremica \cite{aff:sup:2003,www:sup:2004},
-algorithms from supervisory control theory are implemented
-so that a supervisor can be synthesized from automata models
-of the plants and specifications. The algorithm that
-implements the general method for synchronization according
-to this paper is also implemented in the same tool. What
-this means is that the control engineer can focus on the
-correct modeling of the plants and transferring of the
-informal specifications to the specification models for use
-in the Supremica tool. Then the tool automatically generates
-the supervisor in form of an automaton model, function block
+In the software tool Supremica
+\cite{aff:sup:2003,www:sup:2004}, algorithms from
+supervisory control theory are implemented so that a
+supervisor can be synthesized from automata models of the
+plants and specifications. The algorithm that implements the
+general method for synchronization according to this paper
+is also implemented in the same tool. What this means is
+that the control engineer can focus on the correct modeling
+of the plants and transferring of the informal
+specifications to the specification models for use in the
+Supremica tool. Then the tool automatically generates the
+supervisor in form of an automaton model, function block
 types that implement the skeleton for the plants, the
@@
 generate the control application gives new application
-structure as shown in Fig.~\ref{fig:new_app_struct}. In
-this structure, only the models for the generation of the
-plant blocks and the hardware control portion of those
-blocks have to be upgraded whenever the plants in the system
-change. All other parts of the application are regenerated
+structure as shown in Fig.~\ref{fig:new_app_struct}. In this
+structure, only the models for the generation of the plant
+blocks and the hardware control portion of those blocks have
+to be upgraded whenever the plants in the system change. All
+other parts of the application are regenerated
 automatically.
-\begin{figure}[b]
-  \centering 
-  \includegraphics[scale=0.6]{figures/new_app_struct}
-  \caption{A new structure for distributed control applications.}
-  \label{fig:new_app_struct}
-\end{figure}
 
@@
 discrete event models.
+\begin{figure}[!hb]
+  \centering 
+  \includegraphics[scale=0.6]{figures/new_app_struct}
+  \caption{A new structure for distributed control applications.}
+  \label{fig:new_app_struct}
+\end{figure}
+\vfil
 
@@
 Using supervisory control theory in construction of control
-applications can significantly reduce development time but
+applications can significantly reduce development time, but
 until now it was difficult to apply the theory to
@@
 significantly contributes to the shorter control application
-development time since it is a part in automatic generation
+development time, since it is a part in automatic generation
 of control applications. The method is also future proof by
@@
 \cite{c:fun:2002}. For further information about IEC 61499
-function blocks see \cite{c:ope:2002}.
+function blocks, see \cite{c:ope:2002}.
 

### 2005-03-07 src/org/supremica/functionblocks/model/ECC.java
*** empty log message ***

@@ -51,6 +51,8 @@ public class ECC
 
 	private ECState initialState = null;
 
+	
+
 	ECC()
 	{
 		System.out.println("ECC(): Creating new empty ECC");
@@ -83,7 +85,49 @@ public class ECC
 		
 		System.out.println("ECC.execute()");
 		
-		// return the new state
-		return null;
-	}
+		// find all transitions that have currentECState in their source
+		List ecTransitionsWithSameSource = new LinkedList();
+
+		for(Iterator iter = ecTransitions.iterator();iter.hasNext();)
+		{
+			ECTransition temp = (ECTransition) iter.next();
+			if (temp.getSource() == currentECState)
+			{
+				ecTransitionsWithSameSource.add(temp);
+			}
+		}
+		
+		// evalute all of their conditions and remove the ones that can not be taken
+		if(ecTransitionsWithSameSource.size()>0)
+		{
+			for(Iterator iter = ecTransitionsWithSameSource.iterator();iter.hasNext();)
+			{
+				ECTransition temp = (ECTransition) iter.next();
+				if(!temp.getCondition().evaluate(vars)) iter.remove();
+			}
+		}
+		else
+		{
+			System.out.println("ECC.execute(): no possible transitions from state" + currentECState.getName());
+			System.out.println("\t Check your ECC for deadlocks!");
+		}
+		
+		// if more than one condition is true report it and don't do anything,
+		// otherwise take the transition
+		if(ecTransitionsWithSameSource.size()>0)
+		{
+			if(ecTransitionsWithSameSource.size()==1)
+			{
+				return ((ECTransition) ecTransitionsWithSameSource.get(0)).getDestination();
+			}
+			else
+			{
+				System.out.println("ECC.execute(): more than one possible transitions from state" + currentECState.getName());
+				System.out.println("\t Check your ECC for determinism!");
+			}		
+		}
+		
+		// didn't find any transitions that could be taken
+		return currentECState;
+	}	
 }

### 2005-03-07 src/org/supremica/functionblocks/model/ECCondition.java

@@ -51,6 +51,11 @@
  */
 package org.supremica.functionblocks.model;
 
+import java.lang.Exception;
+import java.io.StringReader;
+import java.io.Reader;
+import java_cup.runtime.Scanner;
+import java_cup.runtime.Symbol;
 import org.supremica.functionblocks.model.interpreters.st.Lexer;
 import org.supremica.functionblocks.model.interpreters.st.Parser;
 
@@ -58,23 +63,39 @@ import org.supremica.functionblocks.model.interpreters.st.Parser;
 public class ECCondition
 {
 	private String condition = null;
+	private StringReader reader = null;
+	private Lexer lexer = null;
+	private Parser parser = null;
+	
+	private ECCondition(){}
 
 	public ECCondition(String cond)
 	{
-		condition = cond;
+		setCondition(cond);
 	}
 
 	public void setCondition(String cond)
 	{
 		condition = cond;
-		// parse condition
+
+		reader = new StringReader(condition);
+		
+		lexer = new Lexer((Reader) reader);
+
+		parser = new Parser((Scanner) lexer);
+
+
 	}
 
-	public boolean evaluateCondition(Variables vars)
+	public boolean evaluate(Variables vars)
 	{
-		// interpret the condition string and evaluate
-		// for now all conditions are true, ie labeled by '1'
+		
+		try{
+			System.out.println("ECCondition.evaluate(): parser returned symbol:" + parser.parse().toString());
+		} 
+		catch(Exception e)
+		{}
 		return true;
 	}
-
+	
 }

### 2005-03-07 src/org/supremica/functionblocks/model/FBType.java

@@ -44,10 +44,9 @@ package org.supremica.functionblocks.model;
 
 import java.util.*;
 
-public abstract class FBType
+public abstract class FBType extends NamedObject
 {
 
-    String name;
     Resource resource;
     String comment;
     List versionInfo = new LinkedList();

### 2005-03-07 src/org/supremica/functionblocks/model/Resource.java

@@ -75,6 +75,7 @@ public class Resource
 
 	// create a dummy FB type and instance
 	fbType = new BasicFBType("DummyFBType", this);
+	fbType.getECC().
 	fbType.addAlgorithm(new TestAlgorithm("DummyAlgorithm"));
 	fbInstance = fbType.createInstance("DummyFBInstance");
 	fbInstance.addEventInputQueue(new EventQueue());

### 2005-03-07 src/org/supremica/functionblocks/model/Resource.java
*** empty log message ***

@@ -75,7 +75,6 @@ public class Resource
 
 	// create a dummy FB type and instance
 	fbType = new BasicFBType("DummyFBType", this);
-	fbType.getECC().
 	fbType.addAlgorithm(new TestAlgorithm("DummyAlgorithm"));
 	fbInstance = fbType.createInstance("DummyFBInstance");
 	fbInstance.addEventInputQueue(new EventQueue());

### 2005-11-09 doc/EFA/efa.tex
*** empty log message ***

\documentclass{article}

\usepackage{amsthm,amsmath,amssymb}

\newtheorem{theorem}{Theorem}

\begin{document}
\tableofcontents
\section{Introduction}

\subsection{Background}

\section{Extended Finite Automaton (EFA)}
An Extended Deterministic Finite Automaton (EFA) is an
augmentation of a regular finite automaton. The regular automaton
$A$ is defined as $A=<Q,\Sigma,\delta, q_i,Q_m>$. Let
$T=\{(q,\sigma)\in Q \times \Sigma: \delta(q,\sigma)!\}$ denote
the set of transitions in the deterministic automaton $A$ and let
$\tau_k$, $k=1,...,|T|$ denote the elements (transitions) of $T$.

The automaton $A$ is extended by a set $X=(x_1,...,x_n)$ of $n$
integer variables $x_{k}$ and by a $|T|$-tuple of function pairs
$G=\big((g_{\tau_1},a_{\tau_1}),...,
(g_{\tau_{|T|}},a_{\tau_{|T|}})\big)$, where each pair
$(g_{\tau_k},a_{\tau_k})$ is associated to the corresponding
transition $\tau_k$ in $T$. $g_{\tau_k}$ is an evaluation of a
boolean formula (a guard formula) over $X$ i.e.
$g_{\tau_k}:{\mathbb{N}}^{n}\rightarrow \{0,1\}$ saying if the
corresponding transition $\tau_k$ is enabled or not. $a_{\tau_k}$
is an updating (action) function from ${\mathbb{N}}^{n}$ to
${\mathbb{N}}^{n}$ of the integer variables $X$.

With the above notations we will formally define an EFA as the
$8$-tuple $A_{e}=<Q, \Sigma, \delta_{e}, q_i, Q_m, X, G>$, where
$\delta_{e}: Q \times \Sigma \times X \rightarrow Q\times X $ is
an extention of the transition function $\delta:Q \times \Sigma
\rightarrow Q$. $\delta_e$ is defined by means of $\delta$ and the
elements of $G$ as
\begin{eqnarray}
\delta_{e}(q,\sigma,X)=\left\{
\begin{array}{ll}
(\delta(q,\sigma),a_{q \sigma}(X)\big)\quad\quad g_{q \sigma}(X)=1\\
\big(q,X \big)\quad\quad\quad\quad g_{q \sigma}(X)=0
\end{array}\right.
\end{eqnarray}
If the guard formula $g_{q \sigma}$ is undefined then
$\delta_{e}(q,\sigma,X)$ is undefined.

\subsection{Synchronous composition of EFA's}
The synchronous composition of two EFA's $A=<Q^A, \Sigma^A,
\delta^A, q^A_i, Q^A_m, X, G^A>$ and $B=<Q^B, \Sigma^B, \delta^B,
q^B_i, Q^B_m, Y, G^B>$ exists if and only if the automata are
consistent. The consistency condition requires that the action
functions triggered by a common enabled  event $\sigma\in
\Sigma_{A}\cap \Sigma_{B}$ in the two automata update common
variables to the same values, i.e. if the $i^{ith}$ variable in
$X$ and the $j^{ith}$ variable in $Y$ are the same variable and
$g_{\alpha \sigma}(X)\wedge g_{\beta \sigma}(Y)=1$ where
$(\alpha,\beta)\in Q^A\times Q^B$ then $(a_{\alpha \sigma}(X))_i =
(a_{\beta \sigma}(Y))_j$.

The consistency condition can be verified by inspection. However,
if the vectors $X$ or $Y$ disable the corresponding guard formula
($g_{\alpha \sigma}(X)\wedge g_{\beta \sigma}(Y)$) it may happen
that $a_{\alpha \sigma}(X))_i \neq (a_{\beta \sigma}(Y))_j$ even
though the two automata are consistent. Therefore it may be a
time-consuming task to verify the consistency condition by
inspection. In section \ref{Implementation} we will present
algorithms that includes a consistency check of the synchronous
composition of two EFA's. \\

Formally we define the synchronous product $A||B$ of two EFA's
$A=<Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X, G^A>$ and $B=<Q^B,
\Sigma^B, \delta^B, q^B_i, Q^B_m, Y, G^B>$ as
\begin{eqnarray}
A||B=<Q^A\times Q^B, \Sigma^A \cup \Sigma^B , \delta, q_i,
Q^A_m\times Q^B_m, X \cup Y, G>.
\end{eqnarray}
To define the guard functions, the action functions and the
transition function for the product we make use of the
corresponding functions in $A$ and $B$.

The guard functions $g_{q \sigma}(Z)=g_{q \sigma}(X\cup Y)$ are
defined as
\begin{eqnarray}
g_{q \sigma}(Z)=\left\{
\begin{array}{ll}
g_{q^A\sigma}(X)\wedge g_{q^B\sigma}(Y) \quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
g_{q^A\sigma}(X) \quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
g_{q^B\sigma}(Y) \quad\quad\sigma\in \Sigma^B \setminus \Sigma^A\\
\textrm{undefined       otherwise}
\end{array}\right.
\end{eqnarray}
If $X\cap Y\neq \emptyset$ we introduce $Y'=Y\setminus X$ and
$X'=X\setminus Y$ and look at the restrictions of the updating
functions $a^A_\delta$ and $a^B_\delta$ to the sets $X'$ and $Y'$.
The updating functions are defined as (if $X\cap Y=\emptyset$ then
$Y'=Y$ and $X'=X$)
\begin{eqnarray}
a_{q \sigma}(Z)=\left\{
\begin{array}{ll}
 \big(a^A_{q^A \sigma}(X), a^B_{q^B \sigma}(Y')\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
\big(a^A_{q^A \sigma}(X), Y'\big)\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
\big(X', a^B_{q^B \sigma}(Y)\big)\quad\quad\sigma\in \Sigma^B \setminus \Sigma^A\\
\textrm{undefined       otherwise}
\end{array}\right.
\end{eqnarray}

 The transition function
$\delta(q,\sigma,Z)=\delta(q^A,q^B,\sigma,X\cup Y)$ is defined
from $\delta^A$ and $\delta^B$ as
\begin{eqnarray}
\delta(q,\sigma,Z)=\left\{
\begin{array}{ll}
\big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(Z)\big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(Z)=1\\
\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(Z)\big) \quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(Z)=1\\
\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(Z)\big) \quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(Z)=1\\
\textrm{undefined       otherwise}
\end{array}\right.
\end{eqnarray}

\noindent By the above definition of synchronization we now obtain
the following property

\begin{theorem}
   Full synchronous composition of EFAs is an associative
   operation.
\end{theorem}

\noindent \textbf{Proof}: We show this by considering three
automata $A, A'$ and $A''$ and showing that $(A\|A')\|A'' =
A\|(A'\|A'')$.

The 8-tuple representing $(A\|A')\|A''$ is

\begin{center}
\begin{tabular}{ll}
  $(A\|A')\|A'' =$ & $\Big \langle (Q \times Q') \times Q'',$\\
  & $ \big(\Sigma \bigcup \Sigma' \big) \bigcup \Sigma'',$\\
  & $\delta^{(A\|A')\|A''},$\\
  & $\big((q_i, q'_i) , q''_i\big),$\\
  & $(Q_m \times Q_m') \times Q_m'',$\\
  & $\big(X \bigcup X') \bigcup X'',$\\
  & $G^{(A\|A')\|A''},$\\
  & $\big(X_m \bigcup X'_m) \bigcup X''_m \Big \rangle$
\end{tabular}
\end{center}

\noindent In order to establish associativity we now have to show
this for each of the elements in the 8-tuple. For the states, the
alphabet, the initial state, the marked states and the variables,
associativity is clear by the laws of set operations. We will now
in turn deal with the guard/action functions and the transition
function.

For the guard function $g^{(A\|A')\|A''} : X^c \rightarrow
\{\mathrm{true}, \mathrm{false}\}$ of the composite automaton we
have by definition

\begin{equation}
  g^{(A\|A')\|A''}(X^c) = \big[g(X) \wedge g'(X') \big] \wedge g''(X'') = g(X) \wedge \big[g'(X') \wedge g''(X'')] = g^{A\|(A'\|A'')}(X^c)
\end{equation}

\noindent Now consider the action function $a^{(A\|A')\|A''} : X^c
\rightarrow X^c$. Here things get a bit more messy, since we need
to make sure that each variable is updated at most one time even
though we might have $X\bigcap X'\bigcap X'' \neq \emptyset$

\section{Implementation of EFA's with hierarchy in Supremica}

\subsection{Supremica}

\subsection{Implementation}\label{Implementation}
\subsection{Examples}
\end{document}

### 2005-11-09 doc/EFA/efa.tex
*** empty log message ***

@@
 augmentation of a regular finite automaton. The regular automaton
-$A$ is defined as $A=<Q,\Sigma,\delta, q_i,Q_m>$. Let
-$T=\{(q,\sigma)\in Q \times \Sigma: \delta(q,\sigma)!\}$ denote
-the set of transitions in the deterministic automaton $A$ and let
-$\tau_k$, $k=1,...,|T|$ denote the elements (transitions) of $T$.
+$A$ is defined as $A=\langle Q,\Sigma,\delta, q_i,Q_m \rangle$.
+Let $T=\{(q,\sigma)\in Q \times \Sigma: \delta(q,\sigma)!\}$
+denote the set of transitions in the deterministic automaton $A$
+and let $\tau_k$, $k=1,...,|T|$ denote the elements (transitions)
+of $T$.
 
-The automaton $A$ is extended by a set $X=(x_1,...,x_n)$ of $n$
-integer variables $x_{k}$ and by a $|T|$-tuple of function pairs
-$G=\big((g_{\tau_1},a_{\tau_1}),...,
+The automaton $A$ is extended by a vector $X=(x_1,...,x_n)$ of $n$
+finite integer variables $x_{k}$ and by a $|T|$-tuple of function
+pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
 (g_{\tau_{|T|}},a_{\tau_{|T|}})\big)$, where each pair
@@
 transition $\tau_k$ in $T$. $g_{\tau_k}$ is an evaluation of a
-boolean formula (a guard formula) over $X$ i.e.
+boolean formula (a guard formula) of $X$ i.e.
 $g_{\tau_k}:{\mathbb{N}}^{n}\rightarrow \{0,1\}$ saying if the
@@
 is an updating (action) function from ${\mathbb{N}}^{n}$ to
-${\mathbb{N}}^{n}$ of the integer variables $X$.
+${\mathbb{N}}^{n}$ of the integer variables $X=(x_1,...,x_n)$.
 
 With the above notations we will formally define an EFA as the
-$8$-tuple $A_{e}=<Q, \Sigma, \delta_{e}, q_i, Q_m, X, G>$, where
-$\delta_{e}: Q \times \Sigma \times X \rightarrow Q\times X $ is
-an extention of the transition function $\delta:Q \times \Sigma
-\rightarrow Q$. $\delta_e$ is defined by means of $\delta$ and the
-elements of $G$ as
+$8$-tuple $A_{e}=\langle Q, \Sigma, \delta_{e}, q_i, Q_m, X, G
+,X_m \rangle$, where $X_m$ are the marked integer values of the
+variables in $X$ and $\delta_{e}: Q \times \Sigma \times
+{\mathbb{N}}^{n} \rightarrow Q\times {\mathbb{N}}^{n} $ is a new
+extended version of the transition function $\delta:Q \times
+\Sigma \rightarrow Q$. The extended transition function $\delta_e$
+is defined by means of $\delta$ and the elements of $G$ as
 \begin{eqnarray}
@@
 (\delta(q,\sigma),a_{q \sigma}(X)\big)\quad\quad g_{q \sigma}(X)=1\\
-\big(q,X \big)\quad\quad\quad\quad g_{q \sigma}(X)=0
+\textrm{undefined} \quad\quad\quad\quad\quad g_{q \sigma}(X)=0
 \end{array}\right.
 \end{eqnarray}
-If the guard formula $g_{q \sigma}$ is undefined then
-$\delta_{e}(q,\sigma,X)$ is undefined.
+If $(q,\sigma)\notin T$ then $g_{q\sigma}$ is undefined and
+therefore $\delta_{e}(q,\sigma,X)$ is undefined. Notice that for
+the set of transitions $T_e$ in the extended automaton $A_e$ we
+have $T_e=\{(q,\sigma)\in Q \times \Sigma: g_{q\sigma}(X)=1
+\textrm{ for some } X\}\subset \{(q,\sigma)\in Q \times \Sigma:
+g_{q\sigma} \textrm{ is defined } \}\ =T$.
 
 \subsection{Synchronous composition of EFA's}
-The synchronous composition of two EFA's $A=<Q^A, \Sigma^A,
-\delta^A, q^A_i, Q^A_m, X, G^A>$ and $B=<Q^B, \Sigma^B, \delta^B,
-q^B_i, Q^B_m, Y, G^B>$ exists if and only if the automata are
-consistent. The consistency condition requires that the action
-functions triggered by a common enabled  event $\sigma\in
-\Sigma_{A}\cap \Sigma_{B}$ in the two automata update common
-variables to the same values, i.e. if the $i^{ith}$ variable in
-$X$ and the $j^{ith}$ variable in $Y$ are the same variable and
-$g_{\alpha \sigma}(X)\wedge g_{\beta \sigma}(Y)=1$ where
-$(\alpha,\beta)\in Q^A\times Q^B$ then $(a_{\alpha \sigma}(X))_i =
-(a_{\beta \sigma}(Y))_j$.
+The synchronous composition of two EFA's $A=\langle Q^A, \Sigma^A,
+\delta^A, q^A_i, Q^A_m, X^{A}, G^A, X^A_m \rangle$ and $B=\langle
+Q^B, \Sigma^B, \delta^B, q^B_i, Q^B_m, X^{B}, G^B, X^B_m \rangle$
+exists if and only if the automata are consistent. The consistency
+condition requires that the action functions triggered by a common
+enabled event $\sigma\in \Sigma_{A}\cap \Sigma_{B}$ in the two
+automata update common variables to the same values, i.e. if the
+$i^{ith}$ variable in $X^{A}$ and the $j^{ith}$ variable in
+$X^{B}$ are the same variable and $g_{q^A \sigma}(X^{A})\wedge
+g_{q^B \sigma}(X^{B})=1$ where $(q^A,q^B)\in Q^A\times Q^B$ then
+$(a_{q^A \sigma}(X^{A}))_i = (a_{q^B \sigma}(X^{B}))_j$.
 
 The consistency condition can be verified by inspection. However,
-if the vectors $X$ or $Y$ disable the corresponding guard formula
-($g_{\alpha \sigma}(X)\wedge g_{\beta \sigma}(Y)$) it may happen
-that $a_{\alpha \sigma}(X))_i \neq (a_{\beta \sigma}(Y))_j$ even
-though the two automata are consistent. Therefore it may be a
-time-consuming task to verify the consistency condition by
-inspection. In section \ref{Implementation} we will present
-algorithms that includes a consistency check of the synchronous
+if the vectors $X^{A}$ or $X^{B}$ disable the corresponding guard
+formula ($g_{q^A \sigma}(X^{A})\wedge g_{q^B \sigma}(X^{B})$) it
+may happen that $a_{q^A \sigma}(X^{A}))_i \neq (a_{q^B
+\sigma}(X^{B}))_j$ even though the two automata are consistent.
+Therefore it may be a time-consuming task to verify the
+consistency condition by inspection. In section
+\ref{Implementation} we will present algorithms that includes a
+consistency check of the synchronous
 composition of two EFA's. \\
@@
  Formally we define the synchronous product $A||B$ of two EFA's
-$A=<Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X, G^A>$ and $B=<Q^B,
-\Sigma^B, \delta^B, q^B_i, Q^B_m, Y, G^B>$ as
+$A=\langle Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X^{A}, G^A,X^A_m
+\rangle$ and $B=\langle Q^B, \Sigma^B, \delta^B, q^B_i, Q^B_m,
+X^{B}, G^B,X^B_m
+ \rangle$ as
 \begin{eqnarray}
-A||B=<Q^A\times Q^B, \Sigma^A \cup \Sigma^B , \delta, q_i,
-Q^A_m\times Q^B_m, X \cup Y, G>.
+A||B=\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B , \delta, q_i,
+Q^A_m\times Q^B_m, Z, G, Z_m \rangle.
 \end{eqnarray}
-To define the guard functions, the action functions and the
-transition function for the product we make use of the
-corresponding functions in $A$ and $B$.
+To define the integer variables, the guard functions, the action
+functions and the transition function for the product we make use
+of the corresponding elements in $A$ and $B$.
 
-The guard functions $g_{q \sigma}(Z)=g_{q \sigma}(X\cup Y)$ are
-defined as
-\begin{eqnarray}
+ Let $Z^{AB}$ denote the common variables of $X^{A}$ and $X^{B}$, and let $Z^A$
+and $Z^B$ be the distinct variables of $X^{A}$ and $X^{B}$,
+respectively. We define the integer variables $Z$ of the
+synchronous product of $A$ and $B$ as the vector $Z=(Z^A, Z^{AB},
+Z^B)$. The guard functions $g_{q \sigma}(Z)$ are defined as
+\begin{eqnarray}\label{guard_sync_1}
 g_{q \sigma}(Z)=\left\{
 \begin{array}{ll}
-g_{q^A\sigma}(X)\wedge g_{q^B\sigma}(Y) \quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
-g_{q^A\sigma}(X) \quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
-g_{q^B\sigma}(Y) \quad\quad\sigma\in \Sigma^B \setminus \Sigma^A\\
-\textrm{undefined       otherwise}
+g_{q^A\sigma}(X^{A})\wedge g_{q^B\sigma}(X^{B}) \textrm{ if }\sigma\in \Sigma^A \cap \Sigma^B\\
+g_{q^A\sigma}(X^{A}) \quad\quad \textrm{ if } \sigma\in \Sigma^A \setminus \Sigma^B\\
+g_{q^B\sigma}(X^{B}) \quad\quad \textrm{ if }
+\sigma\in\Sigma^B\setminus \Sigma^A
 \end{array}\right.
 \end{eqnarray}
-If $X\cap Y\neq \emptyset$ we introduce $Y'=Y\setminus X$ and
-$X'=X\setminus Y$ and look at the restrictions of the updating
-functions $a^A_\delta$ and $a^B_\delta$ to the sets $X'$ and $Y'$.
-The updating functions are defined as (if $X\cap Y=\emptyset$ then
-$Y'=Y$ and $X'=X$)
+If any one of the guards functions in the above definition are
+undefined the guard function for the product $g_{q \sigma}$ is
+also undefined. To simplify the notation further, we use "dummy"
+guard functions in the following way. If $\sigma\in \Sigma^A
+\setminus \Sigma^B$ we introduce fictional guards functions
+$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
+\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. With this convention we can
+write equation (\ref{guard_sync_1}) as
+\begin{eqnarray}
+g_{q \sigma}(Z)= g_{q^A\sigma}(X^{A})\wedge g_{q^B\sigma}(X^{B}).
+\end{eqnarray}
+To define the action functions for $A||B$ we look at the
+restrictions of the updating functions $a^A_\delta$ and
+$a^B_\delta$ to the sets $Z^A$ and $Z^B$. The updating functions
+are defined as (if $X^{A}$ and $X^{B}$ have no common variables
+then $Z^A=X^{A}$ and $Z^B=X^{B}$)
 \begin{eqnarray}
@@
 \begin{array}{ll}
- \big(a^A_{q^A \sigma}(X), a^B_{q^B \sigma}(Y')\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
-\big(a^A_{q^A \sigma}(X), Y'\big)\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
-\big(X', a^B_{q^B \sigma}(Y)\big)\quad\quad\sigma\in \Sigma^B \setminus \Sigma^A\\
-\textrm{undefined       otherwise}
+ \big(a^A_{q^A \sigma}(X^{A}), a^B_{q^B \sigma}(Z^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
+\big(a^A_{q^A \sigma}(X^{A}), Z^{B}\big)\quad\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
+\big(Z^{A}, a^B_{q^B \sigma}(X^{B})\big)\quad\quad\quad\sigma\in
+\Sigma^B \setminus \Sigma^A
 \end{array}\right.
+\end{eqnarray}
+As in the definition of the guard functions the notion of being
+undefined is inherited from the action functions of $A$ and $B$.
+Similar to the "dummy" guard functions we introduce fictional
+action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
+introduce fictional action functions $a_{q^B\sigma}(X^B)=X^B$ for
+$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
+$a_{q^A\sigma}(X^A)=X^A$. If $X^{A}$ and $X^{B}$ have no common
+variables then the formula for the action functions of the product
+simplifies to
+\begin{eqnarray}
+a_{q \sigma}(Z)=
+ \big(a^A_{q^A \sigma}(X^{A}), a^B_{q^B \sigma}(X^{B})\big).
 \end{eqnarray}
@@
  The transition function
-$\delta(q,\sigma,Z)=\delta(q^A,q^B,\sigma,X\cup Y)$ is defined
-from $\delta^A$ and $\delta^B$ as
+$\delta(q,\sigma,Z)$ is defined from $\delta^A$ and $\delta^B$ as
 \begin{eqnarray}
@@
 \big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(Z)\big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(Z)=1\\
-\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(Z)\big) \quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(Z)=1\\
-\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(Z)\big) \quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(Z)=1\\
-\textrm{undefined       otherwise}
+\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(Z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(Z)=1\\
+\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(Z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(Z)=1\\
+\textrm{
+undefined}\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+g_{q\sigma}(Z)=0
 \end{array}\right.
@@
 
-\noindent By the above definition of synchronization we now obtain
+The synchronous product $A||B$ of two FSA $A$ and $B$ can be
+extended to $(A||B)_e$ by taking the synchronous product of the
+extended components  $A_e$ and $B_e$. In other words, the
+operations extension and product commute, $(A||B)_e=A_e||B_e$.
+
+Moreover, by the above definition of synchronization we now obtain
 the following property
@@
 A\|(A'\|A'')$.
-
-The 8-tuple representing $(A\|A')\|A''$ is
-
+\newline\newline
+\noindent The 8-tuple representing $(A\|A')\|A''$ is
 \begin{center}
@@
   & $(Q_m \times Q_m') \times Q_m'',$\\
-  & $\big(X \bigcup X') \bigcup X'',$\\
+  & $\big[[X, X'], X'']$\\
   & $G^{(A\|A')\|A''},$\\
-  & $\big(X_m \bigcup X'_m) \bigcup X''_m \Big \rangle$
+  & $\big[[X_m, X'_m], X''_m \big] \Big \rangle$
 \end{tabular}
@@
 \noindent Now consider the action function $a^{(A\|A')\|A''} : X^c
-\rightarrow X^c$. Here things get a bit more messy, since we need
-to make sure that each variable is updated at most one time even
-though we might have $X\bigcap X'\bigcap X'' \neq \emptyset$
+\rightarrow X^c$.
 

### 2005-11-10 doc/EFA/efa.tex

@@
 
-\noindent \textbf{Proof}: We show this by considering three
-automata $A, A'$ and $A''$ and showing that $(A\|A')\|A'' =
-A\|(A'\|A'')$.
+\noindent \textbf{Proof}: To prove this associativity we
+considering three automata $A, A'$ and $A''$, and show that
+$(A\|A')\|A'' = A\|(A'\|A'')$. First, we assume that the three
+automata have no common variables.
 \newline\newline
@@
 
+klkljkll
+
 \section{Implementation of EFA's with hierarchy in Supremica}

### 2005-11-10 doc/EFA/efa.tex
*** empty log message ***

@@
 
-\noindent \textbf{Proof}: To prove this associativity we
+\noindent \textbf{Proof}: To prove the associativity we
 considering three automata $A, A'$ and $A''$, and show that
@@
 
-For the guard function $g^{(A\|A')\|A''} : X^c \rightarrow
-\{\mathrm{true}, \mathrm{false}\}$ of the composite automaton we
-have by definition
+The variables for the product are $X^c=(X_m, X'_m , X''_m)$ and
+the number of variables are $n=|X^c|$. For the guard functions
+$g^{(A\|A')\|A''} : {\mathbb{N}}^{n} \rightarrow \{\mathrm{true},
+\mathrm{false}\}$ of the composite automaton we have by definition
 

### 2005-11-10 doc/EFA/efa.tex
*** empty log message ***

@@
 are defined as (if $X^{A}$ and $X^{B}$ have no common variables
-then $Z^A=X^{A}$ and $Z^B=X^{B}$)
-\begin{eqnarray}
+then $Z=(X^{A},X^{B})$)
+\begin{eqnarray}\label{action_sync_1}
 a_{q \sigma}(Z)=\left\{
@@
 $a_{q^A\sigma}(X^A)=X^A$. If $X^{A}$ and $X^{B}$ have no common
-variables then the formula for the action functions of the product
-simplifies to
+variables then equation (\ref{action_sync_1}), that defines the
+action functions of the product, simplifies to
 \begin{eqnarray}

### 2005-11-10 doc/EFA/efa.tex
*** empty log message ***

--- 2005-11-10 doc/EFA/efa.tex
@@
 
-The automaton $A$ is extended by a vector $X=(x_1,...,x_n)$ of $n$
-finite integer variables $x_{k}$ and by a $|T|$-tuple of function
-pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
+The automaton $A$ is extended by a set $X=X_1\times ... \times
+X_n$, where $X_k$, $k=1,...,n$, are a finite nonempty subsets of
+$\mathbb{Z}$. The set $X$ is the domain of definition of an
+$n$-tuple of integer variables $x=(x_1,...,x_n)$ with initial
+value $x_0=(x_{10},...,x_{n0})$. We also add to $A$ an $|T|$-tuple
+of function pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
 (g_{\tau_{|T|}},a_{\tau_{|T|}})\big)$, where each pair
@@
 transition $\tau_k$ in $T$. $g_{\tau_k}$ is an evaluation of a
-boolean formula (a guard formula) of $X$ i.e.
-$g_{\tau_k}:{\mathbb{N}}^{n}\rightarrow \{0,1\}$ saying if the
-corresponding transition $\tau_k$ is enabled or not. $a_{\tau_k}$
-is an updating (action) function from ${\mathbb{N}}^{n}$ to
-${\mathbb{N}}^{n}$ of the integer variables $X=(x_1,...,x_n)$.
+boolean formula (a guard formula) over $X$ i.e. $g_{\tau_k}:X
+\rightarrow \{0,1\}$ saying if the corresponding transition
+$\tau_k$ is enabled or not. $a_{\tau_k}$ is an updating (action)
+function from $X$ to $X$ of the integer variables
+$x=(x_1,...,x_n)$.
 
 With the above notations we will formally define an EFA as the
-$8$-tuple $A_{e}=\langle Q, \Sigma, \delta_{e}, q_i, Q_m, X, G
-,X_m \rangle$, where $X_m$ are the marked integer values of the
-variables in $X$ and $\delta_{e}: Q \times \Sigma \times
-{\mathbb{N}}^{n} \rightarrow Q\times {\mathbb{N}}^{n} $ is a new
-extended version of the transition function $\delta:Q \times
-\Sigma \rightarrow Q$. The extended transition function $\delta_e$
-is defined by means of $\delta$ and the elements of $G$ as
+$9$-tuple $A_{e}=\langle Q, \Sigma, \delta_{e}, q_i, Q_m, X, G
+,x_0, X_m \rangle$, where $X_m$ are the marked integer values of
+the variables $x$ and $\delta_{e}: Q \times \Sigma \times X
+\rightarrow Q\times X $ is a new extended version of the
+transition function $\delta:Q \times \Sigma \rightarrow Q$. The
+extended transition function $\delta_e$ is defined by means of
+$\delta$ and the elements of $G$ as
 \begin{eqnarray}
-\delta_{e}(q,\sigma,X)=\left\{
+\delta_{e}(q,\sigma,x)=\left\{
 \begin{array}{ll}
-(\delta(q,\sigma),a_{q \sigma}(X)\big)\quad\quad g_{q \sigma}(X)=1\\
-\textrm{undefined} \quad\quad\quad\quad\quad g_{q \sigma}(X)=0
+(\delta(q,\sigma),a_{q \sigma}(x)\big)\quad\quad g_{q \sigma}(x)=1\\
+\textrm{undefined} \quad\quad\quad\quad\quad g_{q \sigma}(x)=0
 \end{array}\right.
@@
 If $(q,\sigma)\notin T$ then $g_{q\sigma}$ is undefined and
-therefore $\delta_{e}(q,\sigma,X)$ is undefined. Notice that for
+therefore $\delta_{e}(q,\sigma,x)$ is undefined. Notice that for
 the set of transitions $T_e$ in the extended automaton $A_e$ we
-have $T_e=\{(q,\sigma)\in Q \times \Sigma: g_{q\sigma}(X)=1
-\textrm{ for some } X\}\subset \{(q,\sigma)\in Q \times \Sigma:
-g_{q\sigma} \textrm{ is defined } \}\ =T$.
+have $T_e=\{(q,\sigma)\in Q \times \Sigma: g_{q\sigma}(x)=1
+\textrm{ for some } x\in X\}\subset \{(q,\sigma)\in Q \times
+\Sigma: g_{q\sigma} \textrm{ is defined } \}\ =T$.
 
@@
 automata update common variables to the same values, i.e. if the
-$i^{ith}$ variable in $X^{A}$ and the $j^{ith}$ variable in
-$X^{B}$ are the same variable and $g_{q^A \sigma}(X^{A})\wedge
-g_{q^B \sigma}(X^{B})=1$ where $(q^A,q^B)\in Q^A\times Q^B$ then
-$(a_{q^A \sigma}(X^{A}))_i = (a_{q^B \sigma}(X^{B}))_j$.
+$i^{ith}$ variable in $x^{A}$ and the $j^{ith}$ variable in
+$x^{B}$ are the same variable and $g_{q^A \sigma}(x^{A})\wedge
+g_{q^B \sigma}(x^{B})=1$ where $(q^A,q^B)\in Q^A\times Q^B$ then
+$(a_{q^A \sigma}(x^{A}))_i = (a_{q^B \sigma}(x^{B}))_j$.
 
 The consistency condition can be verified by inspection. However,
-if the vectors $X^{A}$ or $X^{B}$ disable the corresponding guard
-formula ($g_{q^A \sigma}(X^{A})\wedge g_{q^B \sigma}(X^{B})$) it
-may happen that $a_{q^A \sigma}(X^{A}))_i \neq (a_{q^B
-\sigma}(X^{B}))_j$ even though the two automata are consistent.
+if the vectors $x^{A}$ or $x^{B}$ disable the corresponding guard
+formula ($g_{q^A \sigma}(x^{A})\wedge g_{q^B \sigma}(x^{B})$) it
+may happen that $a_{q^A \sigma}(x^{A}))_i \neq (a_{q^B
+\sigma}(x^{B}))_j$ even though the two automata are consistent.
 Therefore it may be a time-consuming task to verify the
@@
 
- Let $Z^{AB}$ denote the common variables of $X^{A}$ and $X^{B}$, and let $Z^A$
-and $Z^B$ be the distinct variables of $X^{A}$ and $X^{B}$,
-respectively. We define the integer variables $Z$ of the
-synchronous product of $A$ and $B$ as the vector $Z=(Z^A, Z^{AB},
-Z^B)$. The guard functions $g_{q \sigma}(Z)$ are defined as
+ Let $z^{AB}$ denote the common variables of $x^{A}$ and $x^{B}$, and let $z^A$
+and $z^B$ be the distinct variables of $x^{A}$ and $x^{B}$,
+respectively. We define the integer variables $z$ of the
+synchronous product of $A$ and $B$ as the vector $z=(z^A, z^{AB},
+z^B)$. The guard functions $g_{q \sigma}(z)$ are defined as
 \begin{eqnarray}\label{guard_sync_1}
-g_{q \sigma}(Z)=\left\{
+g_{q \sigma}(z)=\left\{
 \begin{array}{ll}
-g_{q^A\sigma}(X^{A})\wedge g_{q^B\sigma}(X^{B}) \textrm{ if }\sigma\in \Sigma^A \cap \Sigma^B\\
-g_{q^A\sigma}(X^{A}) \quad\quad \textrm{ if } \sigma\in \Sigma^A \setminus \Sigma^B\\
-g_{q^B\sigma}(X^{B}) \quad\quad \textrm{ if }
+g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \textrm{ if }\sigma\in \Sigma^A \cap \Sigma^B\\
+g_{q^A\sigma}(x^{A}) \quad\quad \textrm{ if } \sigma\in \Sigma^A \setminus \Sigma^B\\
+g_{q^B\sigma}(x^{B}) \quad\quad \textrm{ if }
 \sigma\in\Sigma^B\setminus \Sigma^A
@@
 \begin{eqnarray}
-g_{q \sigma}(Z)= g_{q^A\sigma}(X^{A})\wedge g_{q^B\sigma}(X^{B}).
+g_{q \sigma}(z)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
 \end{eqnarray}
@@
 $a^B_\delta$ to the sets $Z^A$ and $Z^B$. The updating functions
-are defined as (if $X^{A}$ and $X^{B}$ have no common variables
-then $Z^A=X^{A}$ and $Z^B=X^{B}$)
-\begin{eqnarray}
-a_{q \sigma}(Z)=\left\{
+are defined as (if $x^{A}$ and $x^{B}$ have no common variables
+then $z=(x^{A},x^{B})$)
+\begin{eqnarray}\label{action_sync_1}
+a_{q \sigma}(z)=\left\{
 \begin{array}{ll}
- \big(a^A_{q^A \sigma}(X^{A}), a^B_{q^B \sigma}(Z^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
-\big(a^A_{q^A \sigma}(X^{A}), Z^{B}\big)\quad\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
-\big(Z^{A}, a^B_{q^B \sigma}(X^{B})\big)\quad\quad\quad\sigma\in
+ \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(z^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
+\big(a^A_{q^A \sigma}(x^{A}), z^{B}\big)\quad\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
+\big(z^{A}, a^B_{q^B \sigma}(x^{B})\big)\quad\quad\quad\sigma\in
 \Sigma^B \setminus \Sigma^A
@@
 action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
-introduce fictional action functions $a_{q^B\sigma}(X^B)=X^B$ for
+introduce fictional action functions $a_{q^B\sigma}(x^B)=x^B$ for
 $B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
-$a_{q^A\sigma}(X^A)=X^A$. If $X^{A}$ and $X^{B}$ have no common
-variables then the formula for the action functions of the product
-simplifies to
+$a_{q^A\sigma}(x^A)=x^A$. If $x^{A}$ and $x^{B}$ have no common
+variables then equation (\ref{action_sync_1}), that defines the
+action functions of the product, simplifies to
 \begin{eqnarray}
-a_{q \sigma}(Z)=
- \big(a^A_{q^A \sigma}(X^{A}), a^B_{q^B \sigma}(X^{B})\big).
+a_{q \sigma}(z)=
+ \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
 \end{eqnarray}
@@
  The transition function
-$\delta(q,\sigma,Z)$ is defined from $\delta^A$ and $\delta^B$ as
+$\delta(q,\sigma,z)$ is defined from $\delta^A$ and $\delta^B$ as
 \begin{eqnarray}
-\delta(q,\sigma,Z)=\left\{
+\delta(q,\sigma,z)=\left\{
 \begin{array}{ll}
-\big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(Z)\big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(Z)=1\\
-\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(Z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(Z)=1\\
-\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(Z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(Z)=1\\
+\big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(z)\big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(z)=1\\
+\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(z)=1\\
+\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(z)=1\\
 \textrm{
 undefined}\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
-g_{q\sigma}(Z)=0
+g_{q\sigma}(z)=0
 \end{array}\right.

### 2005-11-10 doc/EFA/efa.tex
*** empty log message ***

@@
 The automaton $A$ is extended by a set $X=X_1\times ... \times
-X_n$, where $X_k$, $k=1,...,n$, are a finite nonempty subsets of
+X_n$, where $X_k$, $k=1,...,n$, are finite nonempty subsets of
 $\mathbb{Z}$. The set $X$ is the domain of definition of an
@@
 \subsection{Synchronous composition of EFA's}
+
+Formally we define the synchronous product $A||B$ of two EFA's
+\begin{eqnarray}
+A=\langle Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X^{A},
+G^A,x^A_0,X^A_m \rangle
+\end{eqnarray} and \begin{eqnarray} B=\langle Q^B, \Sigma^B, \delta^B, q^B_i,
+Q^B_m, X^{B}, G^B,x^B_0,X^B_m \rangle
+\end{eqnarray} as
+\begin{eqnarray}
+A||B&=&\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B ,
+\delta^{A||B}, (q^A_i,q^B_i) ,\nonumber\\
+&& Q^A_m\times Q^B_m, X^A \times X^B, G^{A||B},(x^A_0,x^B_0),
+X^A_m \times X^B_m \rangle.
+\end{eqnarray}
+To define the guard functions, the action functions and the
+transition function of the product we make use of the
+corresponding functions in $A$ and $B$. To denote the elements of
+$A||B$ we remove unnecessary  superscripts and write $A||B=\langle
+Q, \Sigma, \delta, q_i, Q_m, X, G ,x_0, X_m \rangle$. The guard
+functions $g_{q \sigma}(x)$ are defined as
+\begin{eqnarray}\label{guard_sync_1}
+g_{q \sigma}(x)=\left\{
+\begin{array}{ll}
+g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
+g_{q^A\sigma}(x^{A})\quad\quad\quad\quad\quad\quad\quad\sigma\in \Sigma^A\setminus\Sigma^B\\
+g_{q^B\sigma}(x^{B})\quad\quad\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
+\end{array}\right.
+\end{eqnarray}
+If any one of the guards functions in the above definition are
+undefined the guard function for the product $g_{q \sigma}$ is
+also undefined. To simplify the notation further, we use "dummy"
+guard functions in the following way. If $\sigma\in \Sigma^A
+\setminus \Sigma^B$ we introduce fictional guards functions
+$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
+\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. Doing this, we can write
+equation (\ref{guard_sync_1}) as
+\begin{eqnarray}
+g_{q \sigma}(x)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
+\end{eqnarray}
+The updating functions $a_{q \sigma}(x)$ are defined as
+\begin{eqnarray}\label{action_sync_1}
+a_{q \sigma}(x)=\left\{
+\begin{array}{ll}
+ \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
+\big(a^A_{q^A\sigma}(x^{A}),x^{B}\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^A\setminus\Sigma^B\\
+\big(x^{A},a^B_{q^B\sigma}(x^{B})\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
+\end{array}\right.
+\end{eqnarray}
+As in the definition of the guard functions the notion of being
+undefined is inherited from the action functions of $A$ and $B$.
+Similar to the "dummy" guard functions we introduce fictional
+action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
+introduce fictional action functions $a_{q^B\sigma}(x^B)=x^B$ for
+$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
+$a_{q^A\sigma}(x^A)=x^A$. With this convention equation
+(\ref{action_sync_1}) simplifies to
+\begin{eqnarray}
+a_{q \sigma}(x)=
+ \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
+\end{eqnarray}
+
+ The transition function
+$\delta(q,\sigma,x)$ is defined from $\delta^A$ and $\delta^B$ as
+\begin{eqnarray}
+\delta(q,\sigma,x)=\left\{
+\begin{array}{ll}
+\big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(x)\big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(x)=1\\
+\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(x)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(x)=1\\
+\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(x)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(x)=1\\
+\textrm{
+undefined}\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+g_{q\sigma}(x)=0
+\end{array}\right.
+\end{eqnarray}
+
+The synchronous product $A||B$ of two FSA $A$ and $B$ can be
+extended to $(A||B)_e$ by taking the synchronous product of the
+extended components  $A_e$ and $B_e$. In other words, the
+operations extension and product commute, $(A||B)_e=A_e||B_e$.
+
+Moreover, by the above definition of synchronization we now obtain
+the following property
+
+\begin{theorem}
+   Full synchronous composition of EFA's is an associative
+   operation.
+\end{theorem}
+
+\noindent \textbf{Proof}: To prove the associativity we
+considering three automata $A, A'$ and $A''$, and show that
+$(A\|A')\|A'' = A\|(A'\|A'')$. Here, we assume that the three
+automata have no common variables.
+\newline\newline
+\noindent The 8-tuple representing $(A\|A')\|A''$ is
+\begin{center}
+\begin{tabular}{ll}
+  $(A\|A')\|A'' =$ & $\Big \langle (Q \times Q') \times Q'',$\\
+  & $ \big(\Sigma \bigcup \Sigma' \big) \bigcup \Sigma'',$\\
+  & $\delta^{(A\|A')\|A''},$\\
+  & $\big((q_i, q'_i) , q''_i\big),$\\
+  & $(Q_m \times Q_m') \times Q_m'',$\\
+  & $\big[[X, X'], X'']$\\
+  & $G^{(A\|A')\|A''},$\\
+  & $\big[[X_m, X'_m], X''_m \big] \Big \rangle$
+\end{tabular}
+\end{center}
+
+\noindent In order to establish associativity we now have to show
+this for each of the elements in the 8-tuple. For the states, the
+alphabet, the initial state, the marked states and the variables,
+associativity is clear by the laws of set operations. We will now
+in turn deal with the guard/action functions and the transition
+function.
+
+For the guard functions $g^{(A\|A')\|A''} : X^c \rightarrow
+\{\mathrm{true}, \mathrm{false}\}$ of the composite automaton we
+have by definition
+
+\begin{equation}
+  g^{(A\|A')\|A''}(X^c) = \big[g(X) \wedge g'(X') \big] \wedge g''(X'') = g(X) \wedge \big[g'(X') \wedge g''(X'')] = g^{A\|(A'\|A'')}(X^c)
+\end{equation}
+
+\noindent Now consider the action function $a^{(A\|A')\|A''} : X^c
+\rightarrow X^c$.
+
+\subsection{Synchronous composition of EFA's with common variables}
+
 The synchronous composition of two EFA's $A=\langle Q^A, \Sigma^A,
@@
 Q^B, \Sigma^B, \delta^B, q^B_i, Q^B_m, X^{B}, G^B, X^B_m \rangle$
-exists if and only if the automata are consistent. The consistency
-condition requires that the action functions triggered by a common
-enabled event $\sigma\in \Sigma_{A}\cap \Sigma_{B}$ in the two
-automata update common variables to the same values, i.e. if the
-$i^{ith}$ variable in $x^{A}$ and the $j^{ith}$ variable in
-$x^{B}$ are the same variable and $g_{q^A \sigma}(x^{A})\wedge
-g_{q^B \sigma}(x^{B})=1$ where $(q^A,q^B)\in Q^A\times Q^B$ then
-$(a_{q^A \sigma}(x^{A}))_i = (a_{q^B \sigma}(x^{B}))_j$.
+with common variables exists if and only if the automata are
+consistent. The consistency condition requires that the action
+functions triggered by a common enabled event $\sigma\in
+\Sigma_{A}\cap \Sigma_{B}$ in the two automata update common
+variables to the same values, i.e. if the $i^{ith}$ variable in
+$x^{A}$ and the $j^{ith}$ variable in $x^{B}$ are the same
+variable and $g_{q^A \sigma}(x^{A})\wedge g_{q^B \sigma}(x^{B})=1$
+where $(q^A,q^B)\in Q^A\times Q^B$ then $(a_{q^A \sigma}(x^{A}))_i
+= (a_{q^B \sigma}(x^{B}))_j$.
 
@@
 z^B)$. The guard functions $g_{q \sigma}(z)$ are defined as
-\begin{eqnarray}\label{guard_sync_1}
+\begin{eqnarray}\label{guard_sync_2}
 g_{q \sigma}(z)=\left\{
@@
 \Sigma^A$ we set $g_{q^A\sigma}(.)=1$. With this convention we can
-write equation (\ref{guard_sync_1}) as
+write equation (\ref{guard_sync_2}) as
 \begin{eqnarray}
@@
 then $z=(x^{A},x^{B})$)
-\begin{eqnarray}\label{action_sync_1}
+\begin{eqnarray}\label{action_sync_2}
 a_{q \sigma}(z)=\left\{
@@
 $a_{q^A\sigma}(x^A)=x^A$. If $x^{A}$ and $x^{B}$ have no common
-variables then equation (\ref{action_sync_1}), that defines the
+variables then equation (\ref{action_sync_2}), that defines the
 action functions of the product, simplifies to
@@
 
-The synchronous product $A||B$ of two FSA $A$ and $B$ can be
-extended to $(A||B)_e$ by taking the synchronous product of the
-extended components  $A_e$ and $B_e$. In other words, the
-operations extension and product commute, $(A||B)_e=A_e||B_e$.
-
-Moreover, by the above definition of synchronization we now obtain
-the following property
-
-\begin{theorem}
-   Full synchronous composition of EFAs is an associative
-   operation.
-\end{theorem}
-
-\noindent \textbf{Proof}: To prove the associativity we
-considering three automata $A, A'$ and $A''$, and show that
-$(A\|A')\|A'' = A\|(A'\|A'')$. First, we assume that the three
-automata have no common variables.
-\newline\newline
-\noindent The 8-tuple representing $(A\|A')\|A''$ is
-\begin{center}
-\begin{tabular}{ll}
-  $(A\|A')\|A'' =$ & $\Big \langle (Q \times Q') \times Q'',$\\
-  & $ \big(\Sigma \bigcup \Sigma' \big) \bigcup \Sigma'',$\\
-  & $\delta^{(A\|A')\|A''},$\\
-  & $\big((q_i, q'_i) , q''_i\big),$\\
-  & $(Q_m \times Q_m') \times Q_m'',$\\
-  & $\big[[X, X'], X'']$\\
-  & $G^{(A\|A')\|A''},$\\
-  & $\big[[X_m, X'_m], X''_m \big] \Big \rangle$
-\end{tabular}
-\end{center}
-
-\noindent In order to establish associativity we now have to show
-this for each of the elements in the 8-tuple. For the states, the
-alphabet, the initial state, the marked states and the variables,
-associativity is clear by the laws of set operations. We will now
-in turn deal with the guard/action functions and the transition
-function.
-
-The variables for the product are $X^c=(X_m, X'_m , X''_m)$ and
-the number of variables are $n=|X^c|$. For the guard functions
-$g^{(A\|A')\|A''} : {\mathbb{N}}^{n} \rightarrow \{\mathrm{true},
-\mathrm{false}\}$ of the composite automaton we have by definition
-
-\begin{equation}
-  g^{(A\|A')\|A''}(X^c) = \big[g(X) \wedge g'(X') \big] \wedge g''(X'') = g(X) \wedge \big[g'(X') \wedge g''(X'')] = g^{A\|(A'\|A'')}(X^c)
-\end{equation}
-
-\noindent Now consider the action function $a^{(A\|A')\|A''} : X^c
-\rightarrow X^c$.
-
-klkljkll
-
 \section{Implementation of EFA's with hierarchy in Supremica}

### 2005-11-11 doc/EFA/efa.tex
*** empty log message ***

@@
 \rightarrow X^c$.
+\subsection{The translation of an EFA to  a FA}
+To simplify the realization we shall restrict us to quite simple
+action- and guard functions. For the updating functions we assume
+that $a_{q \sigma}(x)=(a^1_{q \sigma}(x),...,a^n_{q
+\sigma}(x))=(a^1_{q \sigma}(x_1),...,a^n_{q \sigma}(x_n))$. In
+other words, the updating of the variables are independent from
+each other. For the guard functions we assume that they have the
+following form $g_{q \sigma}(x)=g^1_{q \sigma}(x_1)\wedge
+...\wedge g^n_{q \sigma}(x_n)$. These assumptions makes straight
+forward to translate an EFA to a FA.
+
+-realize an automata for each variable.
 

### 2005-11-11 doc/EFA/efa.tex
*** empty log message ***

@@
 
--realize an automata for each variable.
+Given an extended finite automaton $A=\langle Q, \Sigma, \delta,
+q_i, Q_m, X, G ,x_0, X_m \rangle$ with simple guard and action
+functions, the steps are:\\
+- Form a regular FA, $A'$, from $A$ by removing $G$, $X$, $X_m$,
+$x_0$ and rename (associate with the corresponding state) all
+enabled events $\sigma$ with $q\sigma$.\\
+- Realize an automata $A_{x_k}$ for each variable $x_k$, with marked "states" $X_{m,k}$ and with as many events as elements in $G$ .\\
+- Synchronize all automata.
 

### 2005-11-11 doc/EFA/efa.tex
*** empty log message ***

--- 2005-11-10 doc/EFA/efa.tex
@@
 \subsection{Background}
-
+\newpage
 \section{Extended Finite Automaton (EFA)}
@@
 \end{equation}
-
 \noindent Now consider the action function $a^{(A\|A')\|A''} : X^c
 \rightarrow X^c$.
+\newpage
+\subsection{The translation of an EFA to  an FA}
+To simplify the realization we shall restrict us to quite simple
+action- and guard functions. For the updating functions we assume
+that $a_{q \sigma}(x)=(a^1_{q \sigma}(x),...,a^n_{q
+\sigma}(x))=(a^1_{q \sigma}(x_1),...,a^n_{q \sigma}(x_n))$. In
+other words, the updating of the variables are independent from
+each other. For the guard functions we assume that they are given
+in disjunctive normal form $g_{q \sigma}(x)=g^1_{q \sigma}(x)\vee
+...\vee g^j_{q \sigma}(x)$. We shall denote the number of
+components $j$ in such an expression by $|g_{q \sigma}|$. These
+assumptions makes it straight forward to translate an EFA to an
+FA.
 
+Given an extended finite automaton $A=\langle Q, \Sigma, \delta,
+q_i, Q_m, X, G ,x_0, X_m \rangle$ with simple guard and action
+functions, the algorithm can be divided into four steps:\\
+- Form a regular FA, $A'$, from $A$ by removing $G$, $X$, $X_m$,
+$x_0$ and change the names of all enabled events $\sigma$ to
+$q\sigma$ (associate with the corresponding
+state). Divide the event(transition) $q\sigma$ further into $|g_{q \sigma}|$ identical copies $\{(q\sigma)_k\}^{|g_{q \sigma}|}_{k=1}$.\\
+- Realize an automata $A_{x_k}$ for each variable $x_k$, with
+marked "states" $X_{mk}$,
+ initial state $x_{k 0}$ and with as many events the total number of events of type $q\sigma$ in $A'$.\\
+- Synchronize all automata.\\
+- Change back to original event names.
+\newpage
 \subsection{Synchronous composition of EFA's with common variables}

### 2005-11-14 doc/EFA/efa.tex
*** empty log message ***

@@
 in disjunctive normal form $g_{q \sigma}(x)=g^1_{q \sigma}(x)\vee
-...\vee g^j_{q \sigma}(x)$. We shall denote the number of
+...\vee g^j_{q \sigma}(x)$. We shall denote the number of "OR"
 components $j$ in such an expression by $|g_{q \sigma}|$. These
@@
  initial state $x_{k 0}$ and with as many events the total number of events of type $q\sigma$ in $A'$.\\
-- Synchronize all automata.\\
+- Synchronize all automata $A_{x_k}$, $k=1,...,n$ and $A'$.\\
 - Change back to original event names.

### 2005-11-15 doc/EFA/efa.tex
*** empty log message ***

@@
 \newpage
+\section{Finite Automaton (FA) with hierarchy}
+
 \section{Extended Finite Automaton (EFA)}
+The regular deterministic finite automaton $A$ is a $5$-tuple
+\begin{eqnarray}
+A=\langle Q,\Sigma,\delta, q_i,Q_m \rangle\nonumber
+\end{eqnarray} where $Q$ is a finite set of states; $\Sigma$ (the alphabet) is a nonempty
+set of events; $\delta$ is the transition function; $q_i$ is the
+initial state and $Q_m \subseteq Q$ is the set of marked states.
 An Extended Deterministic Finite Automaton (EFA) is an
-augmentation of a regular finite automaton. The regular automaton
-$A$ is defined as $A=\langle Q,\Sigma,\delta, q_i,Q_m \rangle$.
-Let $T=\{(q,\sigma)\in Q \times \Sigma: \delta(q,\sigma)!\}$
-denote the set of transitions in the deterministic automaton $A$
-and let $\tau_k$, $k=1,...,|T|$ denote the elements (transitions)
-of $T$.
+augmentation of the regular finite automaton with guard and action
+formulas associated to the transitions. The transitions in the EFA
+are enabled if and only if a guard formula is true. Moreover, when
+a transition in an EFA is taken, updating actions of a set of
+finite integer variables may follow. Formally this is done as
+follows.
 
-The automaton $A$ is extended by a set $X=X_1\times ... \times
-X_n$, where $X_k$, $k=1,...,n$, are finite nonempty subsets of
-$\mathbb{Z}$. The set $X$ is the domain of definition of an
-$n$-tuple of integer variables $x=(x_1,...,x_n)$ with initial
-value $x_0=(x_{10},...,x_{n0})$. We also add to $A$ an $|T|$-tuple
-of function pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
+Given a regular automaton $A=\langle Q,\Sigma,\delta, q_i,Q_m
+\rangle$, let $T=\{(q,\sigma)\in Q \times \Sigma:
+\delta(q,\sigma)!\}$ denote the set of transitions in $A$ and let
+$\tau_k$, $k=1,...,|T|$ denote the elements (transitions) of $T$.
+First, we add to $A$ a set $X=X_1\times ... \times X_n$, where
+$X_k$, $k=1,...,n$, are finite nonempty subsets of $\mathbb{Z}$.
+The set $X$ is the domain of definition of an $n$-tuple of integer
+variables $x=(x_1,...,x_n)$ with initial value
+$x_0=(x_{10},...,x_{n0})$. We also add to $A$ an $|T|$-tuple of
+function pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
 (g_{\tau_{|T|}},a_{\tau_{|T|}})\big)$, where each pair

### 2005-11-15 doc/EFA/efa2fa.tex
*** empty log message ***

\documentclass{article}

\usepackage{amsthm,amsmath,amssymb}

\newtheorem{theorem}{Theorem}

\pagestyle{empty}

\begin{document}

\section*{The translation of an EFA to  an FA}
The main body of theory and algorithms within supervisory control
theory has been developed for standard finite automata. For the
extended finite automata to be useful we need to fit them within
this framework. In other words, in order to do formal analysis
based on these new automata, what we need is a way to convert EFAs
into standard FAs.

Fortunately, as it turns out, there is a rather elegant
transformation of EFAs into FAs based on synchronization. In
essence, the recipe is as follows:

\begin{enumerate}
    \item Form a stripped EFA by removing the variables and all
    guard/action pairs from the EFA. This yields a regular FA.
    \item Introduce a new unique event for each guard/action pair and
     relabel the transitions of the stripped automaton correspondingly.
    \item Introduce \emph{one} automaton $A_{x_k}$ for each variable $x_k$ with
    the number of states equal to the range of $x_k$.

    \item
    For the states of the variable automata $A_{x_k}$ that fulfill the guard conditions,
    add transitions that realize the updating of the variables as specified by the action
    functions.

    \item Construct the FA-equivalent of the EFA by synchronizing
    the stripped EFA with all variable automata $A_{x_k}$.
    \item Change back to original event names.
\end{enumerate}

\noindent Below, an example of how this can be done is given.
\end{document}

### 2005-11-15 doc/EFA/efa.tex
*** empty log message ***

@@
 \newpage
-\section{Finite Automaton (FA) with hierarchy}
+\section{Finite Automata (FA) with hierarchy}
 
-\section{Extended Finite Automaton (EFA)}
+\section{Extended Finite Automata (EFA)}
 The regular deterministic finite automaton $A$ is a $5$-tuple
@@
 \end{eqnarray}
-As in the definition of the guard functions the notion of being
-undefined is inherited from the action functions of $A$ and $B$.
-Similar to the "dummy" guard functions we introduce fictional
-action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
-introduce fictional action functions $a_{q^B\sigma}(x^B)=x^B$ for
-$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
+As in the definition of the guard functions (\ref{guard_sync_1}),
+the notion of being undefined is inherited from the action
+functions of $A$ and $B$. Similar to the "dummy" guard functions
+we introduce fictional action functions. If $\sigma\in \Sigma^A
+\setminus \Sigma^B$ we introduce fictional action functions
+$a_{q^B\sigma}(x^B)=x^B$ for $B$ and if
+$\sigma\in\Sigma^B\setminus \Sigma^A$ we set
 $a_{q^A\sigma}(x^A)=x^A$. With this convention equation
@@
 $x_0$ and change the names of all enabled events $\sigma$ to
-$q\sigma$ (associate with the corresponding
-state). Divide the event(transition) $q\sigma$ further into $|g_{q \sigma}|$ identical copies $\{(q\sigma)_k\}^{|g_{q \sigma}|}_{k=1}$.\\
+$q\sigma$ i.e. the associate them with the corresponding
+transition.\\
+- Rename the events $q\sigma$ in $A'$ further into $|g_{q \sigma}|$ identical transitions with event names: $\{(q\sigma)_k\}^{|g_{q \sigma}|}_{k=1}$.\\
 - Realize an automata $A_{x_k}$ for each variable $x_k$, with
-marked "states" $X_{mk}$,
- initial state $x_{k 0}$ and with as many events the total number of events of type $q\sigma$ in $A'$.\\
+marked "states" $X_{mk}$, initial state $x_{k 0}$ and with as many events the total number of events in $A'$.
+The transitions correspond to the action functions and they are enabled if the guard formula is true.\\
 - Synchronize all automata $A_{x_k}$, $k=1,...,n$ and $A'$.\\

### 2005-11-21 src/org/supremica/functionblocks/model/JavaTextAlgorithm.java
*** empty log message ***

@@ -111,6 +111,11 @@ public class JavaTextAlgorithm extends Algorithm
 						BooleanVariable tmpVar = (BooleanVariable) vars.getVariable(curName); 
 						interpreter.set(curName,tmpVar.getValue().booleanValue());
 					}
+					else if (curVariable instanceof ObjectVariable)
+					{
+						ObjectVariable tmpVar = (ObjectVariable) vars.getVariable(curName); 
+						interpreter.set(curName,tmpVar.getValue());
+					}
 				}
 				catch (Exception e)
 				{
@@ -169,6 +174,12 @@ public class JavaTextAlgorithm extends Algorithm
 						BooleanVariable tmpVar = (BooleanVariable) vars.getVariable(curName); 
 						boolean tmpValue = ((Boolean) interpreter.get(curName)).booleanValue();
 						((BooleanVariable) vars.getVariable(curName)).setValue(tmpValue);
+					}
+					else if (curVariable instanceof ObjectVariable)
+					{
+						ObjectVariable tmpVar = (ObjectVariable) vars.getVariable(curName); 
+						Object tmpValue = (Object) interpreter.get(curName);
+						((ObjectVariable) vars.getVariable(curName)).setValue(tmpValue);
 					}
 				}
 				catch (Exception e)

### 2005-11-21 src/org/supremica/functionblocks/model/ObjectVariable.java

@@ -49,25 +49,24 @@
 
 package org.supremica.functionblocks.model;
 
-import java.lang.String;
 
-public class StringVariable extends Variable
+public class ObjectVariable extends Variable
 {
 	
-	private String value = "";
+	private Object value = null;
 
-	public StringVariable(String type, String s)
+	public ObjectVariable(String type, Object o)
 	{
 		super(type);
-		value=s;
+		value=o;
 	}
 
-	public synchronized void setValue(String v)
+	public synchronized void setValue(Object o)
 	{
-		value = v;
+		value = o;
 	}
 
-	public synchronized String getValue()
+	public synchronized Object getValue()
 	{
 		return value;
 		
@@ -75,12 +74,12 @@ public class StringVariable extends Variable
 
 	public Object clone()
 	{
-		return new StringVariable(getType(),getValue());
+		return new ObjectVariable(getType(),getValue());
 	}
-
+	
 	public String toString()
 	{
-		return getType() + " " + value;
+		return getType() + " " + value.toString();
 	}
 
 }

### 2005-11-21 src/org/supremica/functionblocks/model/FBRuntime.java
*** empty log message ***

@@ -49,9 +49,8 @@
 /*
  * @author cengic
  */
-package org.supremica.apps;
+package org.supremica.functionblocks.model;
 
-import org.supremica.functionblocks.model.*;
 import java.io.File;
 
 public class FBRuntime

### 2005-11-21 src/org/supremica/automata/algorithms/scheduling/AbstractAstar.java
*** empty log message ***

@@ -899,6 +899,10 @@ logger.info("in getParent()");
     public Automaton buildScheduleAutomaton(int[] currNode) 
 		throws Exception
 	{
+		logger.warn("closedKeySet...");
+		for (Iterator<Integer> keys = closedTree.keySet().iterator(); keys.hasNext(); )
+			logger.info("key = " + keys.next().intValue());
+
 		Automaton scheduleAuto = new Automaton();
 		scheduleAuto.setComment("Schedule");
 	

### 2005-11-21 doc/EFA/efa.tex
*** empty log message ***

@@
 \newpage
+\subsection{Translating hierarchical structures to EFA's}
+It is well known that hierarchical structures of states and
+transitions reduce the descriptive complexity of automata models.
+We shall in this section show how the hierarchy in state charts
+can be translated into EFA's.
+
+In what follows, states are represented by boxes and hierarchy is
+illustrated by being inside or outside the boxes. Arrows between
+states represents transitions and are labelled with the lower case
+Greek letters, $\alpha$-$\omega$. An arrow that leaves a box
+applies to all sub-states inside the box. The common way to enter
+a group of states its by the system's history in that group. The
+simplest way of doing this is to enter the most recently visited
+state in the group. Hierarchical structures with quite general
+history functions can easily be translated into EFA's.
+
 \subsection{Synchronous composition of EFA's with common variables}

### 2005-11-21 src/org/supremica/automata/algorithms/scheduling/AbstractAstar.java
*** empty log message ***

@@ -92,6 +92,8 @@ public abstract class AbstractAstar
 	 */
 	protected TreeMap<Integer, int[]> closedTree;
 
+	protected int maxOpenSize = 0;
+
 	///////////////////////////////////////////////////////////////////////////////////////
 	//                                 NEW_TRY_END (051117 - ...)                        //
 	///////////////////////////////////////////////////////////////////////////////////////
@@ -209,6 +211,9 @@ public abstract class AbstractAstar
 			// 		//				infoStr += prep2Info;
 			// 	    }
 	    
+			//Tillf
+			long millis = System.currentTimeMillis();
+
 			timer.start();
 			searchCounter = 0;
 
@@ -216,9 +221,14 @@ public abstract class AbstractAstar
 			if (accNode == null)
 				throw new RuntimeException("no marked state found, nr of iteration = " + searchCounter); 
 
-			infoStr += "\tA*-iterations: " + searchCounter + " in time: " + timer.elapsedTime() + " ms.\n";
+			infoStr += "\tA*-iterations (nr of search calls through the closed tree): " + searchCounter + "\n";
+			infoStr += "\tIn time: " + timer.elapsedTime() + " ms\n";
+			infoStr += "\tMillis-time: " + (System.currentTimeMillis() - millis) + "ms\n";
+			infoStr += "\tThe CLOSED tree contains (at the end) " + closedTree.size() + " elements\n";
+			infoStr += "\tMax{OPEN.size} = " + maxOpenSize + "\n";
 			// 	    infoStr += "\t\t"+ printNodeSignature(accNode);
 			infoStr += "\t\t" + "g = " + accNode[accCostIndex];
+			
 			logger.info(infoStr);
 	    
 			return accNode;
@@ -294,14 +304,22 @@ public abstract class AbstractAstar
 			// NEW_TRY_START
 		
 			openTree.clear();
+			closedTree.clear();
+
 			openTree.add(initNode);
 
 			while(! openTree.isEmpty())
 			{
 				searchCounter++;
 
+				// Tillf ...
+				int currOpenSize = openTree.size();
+				if (currOpenSize > maxOpenSize)
+					maxOpenSize = currOpenSize;
+				// ... fram hit
+
 				currNode = openTree.first();
-				
+
 				if (isAcceptingNode(currNode))
 					return currNode;
 
@@ -798,15 +816,11 @@ public abstract class AbstractAstar
 		throws Exception 
 	{
 		// NEW_TRY_START
[…]

### 2005-11-21 src/org/supremica/automata/algorithms/scheduling/ModifiedAstar.java

@@ -55,13 +55,20 @@ public class ModifiedAstar
 			// 					// TODO: Lägg currNode till closedTree på rätt ställe
 			// 				}
 						
-			updateClosedTree(currNode);
+			boolean currNodeIsAddedToClosed = updateClosedTree(currNode);
 			
-			Iterator childIter = expander.expandNode(currNode, activeAutomataIndex).iterator();
-			while (childIter.hasNext()) {
-				int[] nextNode = (int[])childIter.next();
-				
-				openTree.add(nextNode);
+			if (currNodeIsAddedToClosed)
+			{
+				Iterator childIter = expander.expandNode(currNode, activeAutomataIndex).iterator();
+				while (childIter.hasNext()) 
+				{
+					int[] nextNode = (int[])childIter.next();
+					
+					// Calculate the estimate function of the expanded node and store it at the appropriate position
+					nextNode[ESTIMATE_INDEX] = calcEstimatedCost(nextNode);
+					
+					openTree.add(nextNode);
+				}
 			}
 		}
 		catch (Exception e)
@@ -79,8 +86,12 @@ public class ModifiedAstar
 	 * closedTree. If there are ties (in some directions one node is better than the other, but in others
 	 * it is worse), this node, as well as all the discovered tie nodes are stored in the closedTree
 	 * (note that all examined nodes that are always worse than "node" are discarded).
+	 * 
+	 * @param int[] node - the new node that might be added to the CLOSED tree.
+	 * @return true, if the new node is added to the CLOSED tree
+	 *         false, otherwise.
 	 */
-	private void updateClosedTree(int[] node)
+	private boolean updateClosedTree(int[] node)
 	{
 		// The nodes corresponding to the same logical state (but different paths from the initial state)
 		// as the new node. They are stored as one int[]-variable in the closedTree.
@@ -124,7 +135,7 @@ public class ModifiedAstar
 				if (newNodeIsAlwaysWorse)
 				{
 					closedTree.put(new Integer(getKey(node)), correspondingClosedNodes);
-					return;
+					return false;
 				}
 				// else if the examined node is neither worse nor better, its index is added to the tieIndices
 				else if (! newNodeIsAlwaysWorse && ! newNodeIsAlwaysBetter)
@@ -152,7 +163,10 @@ public class ModifiedAstar
 
 			closedTree.put(new Integer(getKey(node)), newClosedNode);
 		}
+
+		return true;
 	}
+
 	// NEW_START_END
 		
 //     protected void branch(int[] currNode) 

### 2005-11-21 doc/EFA/efa.tex
*** empty log message ***

@@
 \end{eqnarray}
-If any one of the guards functions in the above definition are
+If in any of the three cases, the corresponding guard function is
 undefined the guard function for the product $g_{q \sigma}$ is
-also undefined. To simplify the notation further, we use "dummy"
-guard functions in the following way. If $\sigma\in \Sigma^A
-\setminus \Sigma^B$ we introduce fictional guards functions
-$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
-\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. Doing this, we can write
-equation (\ref{guard_sync_1}) as
+also undefined. To simplify the notation and increase readability,
+we use "dummy" guard functions in the following way. If $\sigma\in
+\Sigma^A \setminus \Sigma^B$ we introduce fictional guards
+functions $g_{q^B\sigma}(.)=1$ for $B$ and if
+$\sigma\in\Sigma^B\setminus \Sigma^A$ we set $g_{q^A\sigma}(.)=1$.
+Doing this, we can write equation (\ref{guard_sync_1}) as
 \begin{eqnarray}
@@
 \begin{array}{ll}
- \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
-\big(a^A_{q^A\sigma}(x^{A}),x^{B}\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^A\setminus\Sigma^B\\
-\big(x^{A},a^B_{q^B\sigma}(x^{B})\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
+ \big(a_{q^A \sigma}(x^{A}), a_{q^B \sigma}(x^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
+\big(a_{q^A\sigma}(x^{A}),x^{B}\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^A\setminus\Sigma^B\\
+\big(x^{A},a_{q^B\sigma}(x^{B})\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
 \end{array}\right.
@@
 a_{q \sigma}(x)=
- \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
+ \big(a_{q^A \sigma}(x^{A}), a_{q^B \sigma}(x^{B})\big).
 \end{eqnarray}
 
- The transition function
-$\delta(q,\sigma,x)$ is defined from $\delta^A$ and $\delta^B$ as
+ The transition function $\delta^{A||B}$ is defined from $\delta^A$ and $\delta^B$ as
 \begin{eqnarray}
-\delta(q,\sigma,x)=\left\{
+\delta^{A||B}(q,\sigma,x)=\left\{
 \begin{array}{ll}
@@
 \end{array}\right.
+\end{eqnarray}
+
+\subsubsection{Properties of the product}
+$(A||B)||C$:
+
+\begin{eqnarray}
+g^{(A||B)||C}_{q \sigma}(x)&=& g_{q^{AB}\sigma}(x^{AB})\wedge
+g_{q^C\sigma}(x^{C})\nonumber\\
+&=&g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B})\wedge
+g_{q^C\sigma}(x^{C})\nonumber\\
+&=& g_{q^{A}\sigma}(x^{A})\wedge g_{q^{BC}\sigma}(x^{BC})\nonumber\\
+&=& g^{A||(B||C)}_{q \sigma}(x)
+\end{eqnarray}
+
+\begin{eqnarray}
+a^{(A||B)||C}_{q \sigma}(x)&=&
+ \big(a_{q^{AB} \sigma}(x^{AB}), a_{q^{C}
+ \sigma}(x^{C})\big)\nonumber\\
+&=&\big(a_{q^{A} \sigma}(x^{A}), a_{q^{B}
+ \sigma}(x^{B}), a_{q^{C}
+ \sigma}(x^{C})\big)\nonumber\\
+&=&\big(a_{q^{A} \sigma}(x^{A}), a_{q^{BC}
+ \sigma}(x^{BC})\big)\nonumber\\
+&=&a^{A||(B||C)}_{q \sigma}(x)
 \end{eqnarray}

### 2005-11-21 doc/EFA/efa.tex
*** empty log message ***

@@
 
-Formally we define the synchronous product $A||B$ of two EFA's
+Formally we define the synchronous product $A\|B$ of two EFA's
 \begin{eqnarray}
@@
 \begin{eqnarray}
-A||B&=&\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B ,
-\delta^{A||B}, (q^A_i,q^B_i) ,\nonumber\\
-&& Q^A_m\times Q^B_m, X^A \times X^B, G^{A||B},(x^A_0,x^B_0),
+A\|B&=&\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B ,
+\delta^{A\|B}, (q^A_i,q^B_i) ,\nonumber\\
+&& Q^A_m\times Q^B_m, X^A \times X^B, G^{A\|B},(x^A_0,x^B_0),
 X^A_m \times X^B_m \rangle.
@@
 corresponding functions in $A$ and $B$. To denote the elements of
-$A||B$ we remove unnecessary  superscripts and write $A||B=\langle
+$A\|B$ we remove unnecessary  superscripts and write $A\|B=\langle
 Q, \Sigma, \delta, q_i, Q_m, X, G ,x_0, X_m \rangle$. The guard
@@
 
- The transition function $\delta^{A||B}$ is defined from $\delta^A$ and $\delta^B$ as
+ The transition function $\delta^{A\|B}$ is defined from $\delta^A$ and $\delta^B$ as
 \begin{eqnarray}
-\delta^{A||B}(q,\sigma,x)=\left\{
+\delta^{A\|B}(q,\sigma,x)=\left\{
 \begin{array}{ll}
@@
 \subsubsection{Properties of the product}
-$(A||B)||C$:
 
+Consider three EFA's $A$, $B$ and $C$. From the definition of the
+synchronization we shall now show that
+ $(A\|B)\|C=A\|(B\|C)$ i.e. we have:
+\begin{theorem}
+   Full synchronous composition of EFA's is an associative
+   operation.
+\end{theorem}
+
+\noindent \textbf{Proof}: We need to prove that $(A\|B)\|C =
+A\|(B\|C)$. Here, we assume that the three automata have no common
+variables.
+\newline\newline
+\noindent The 9-tuple representing $(A\|B)\|C$ is
+\begin{center}
+\begin{tabular}{ll}
+  $(A\|B)\|C =$ & $\Big \langle (Q^{A} \times Q^{B}) \times Q^{C},$\\
+  & $ \big(\Sigma^{A} \bigcup \Sigma^{B} \big) \bigcup \Sigma^{C},$\\
+  & $\delta^{(A\|B)\|C},$\\
+  & $\big((q^{A}_i, q^{B}_i) , q^{C}_i\big),$\\
+  & $(Q^{A}_m \times Q^{B}_m) \times Q^{C}_m,$\\
+  & $\big[[X^{A}, X^{B}], X^{C}]$\\
+  & $G^{(A\|B)\|C},$\\
+  & $\big((x^{A}_0, x^{B}_0) , x^{C}_0\big),$\\
+  & $\big[[X^{A}_m, X^{B}_m], X^{C}_m \big] \Big \rangle$
+\end{tabular}
+\end{center}
+
+\noindent In order to establish associativity we now have to show
+this for each of the elements in the tuple. For the states, the
+alphabet, the initial state, the marked states and the variables,
+associativity is clear by the laws of set operations. We will now
+in turn deal with the guard/action functions and the transition
+function.
+
+For the guard functions and the action functions of the composite
+automaton we have
 \begin{eqnarray}
-g^{(A||B)||C}_{q \sigma}(x)&=& g_{q^{AB}\sigma}(x^{AB})\wedge
+g^{(A\|B)\|C}_{q \sigma}(x)&=& g_{q^{AB}\sigma}(x^{AB})\wedge
 g_{q^C\sigma}(x^{C})\nonumber\\
@@
 &=& g_{q^{A}\sigma}(x^{A})\wedge g_{q^{BC}\sigma}(x^{BC})\nonumber\\
-&=& g^{A||(B||C)}_{q \sigma}(x)
+&=& g^{A\|(B\|C)}_{q \sigma}(x)
 \end{eqnarray}
-
+and
 \begin{eqnarray}
-a^{(A||B)||C}_{q \sigma}(x)&=&
+a^{(A\|B)\|C}_{q \sigma}(x)&=&
  \big(a_{q^{AB} \sigma}(x^{AB}), a_{q^{C}
@@
  \sigma}(x^{BC})\big)\nonumber\\
-&=&a^{A||(B||C)}_{q \sigma}(x)
+&=&a^{A\|(B\|C)}_{q \sigma}(x).
 \end{eqnarray}
+This shows that $G^{(A\|B)\|C}=G^{A\|(B\|C)}$, and since the
+transition function for the standard $FA$ is associative it
+follows that $\delta^{(A\|B)\|C}=\delta^{A\|(B\|C)}$.
+\begin{flushright}$\square$\end{flushright}
 
-The synchronous product $A||B$ of two FSA $A$ and $B$ can be
-extended to $(A||B)_e$ by taking the synchronous product of the
+The synchronous product $A\|B$ of two FSA $A$ and $B$ can be
+extended to $(A\|B)_e$ by taking the synchronous product of the
 extended components  $A_e$ and $B_e$. In other words, the
-operations extension and product commute, $(A||B)_e=A_e||B_e$.
+operations extension and product commute, $(A\|B)_e=A_e\|B_e$.
 
-Moreover, by the above definition of synchronization we now obtain
-the following property
-
-\begin{theorem}
-   Full synchronous composition of EFA's is an associative
-   operation.
-\end{theorem}
-
-\noindent \textbf{Proof}: To prove the associativity we
-considering three automata $A, A'$ and $A''$, and show that
-$(A\|A')\|A'' = A\|(A'\|A'')$. Here, we assume that the three
-automata have no common variables.
-\newline\newline
-\noindent The 8-tuple representing $(A\|A')\|A''$ is
-\begin{center}
-\begin{tabular}{ll}
-  $(A\|A')\|A'' =$ & $\Big \langle (Q \times Q') \times Q'',$\\
-  & $ \big(\Sigma \bigcup \Sigma' \big) \bigcup \Sigma'',$\\
-  & $\delta^{(A\|A')\|A''},$\\
-  & $\big((q_i, q'_i) , q''_i\big),$\\
-  & $(Q_m \times Q_m') \times Q_m'',$\\
-  & $\big[[X, X'], X'']$\\
-  & $G^{(A\|A')\|A''},$\\
-  & $\big[[X_m, X'_m], X''_m \big] \Big \rangle$
-\end{tabular}
-\end{center}
-
-\noindent In order to establish associativity we now have to show
-this for each of the elements in the 8-tuple. For the states, the
-alphabet, the initial state, the marked states and the variables,
-associativity is clear by the laws of set operations. We will now
-in turn deal with the guard/action functions and the transition
-function.
-
-For the guard functions $g^{(A\|A')\|A''} : X^c \rightarrow
-\{\mathrm{true}, \mathrm{false}\}$ of the composite automaton we
-have by definition
-
-\begin{equation}
-  g^{(A\|A')\|A''}(X^c) = \big[g(X) \wedge g'(X') \big] \wedge g''(X'') = g(X) \wedge \big[g'(X') \wedge g''(X'')] = g^{A\|(A'\|A'')}(X^c)
-\end{equation}
-\noindent Now consider the action function $a^{(A\|A')\|A''} : X^c
-\rightarrow X^c$.
-\newpage
 \subsection{The translation of an EFA to  an FA}
@@
 
-Formally we define the synchronous product $A||B$ of two EFA's
+Formally we define the synchronous product $A\|B$ of two EFA's
 $A=\langle Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X^{A}, G^A,X^A_m
@@
 \begin{eqnarray}
-A||B=\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B , \delta, q_i,
+A\|B=\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B , \delta, q_i,
 Q^A_m\times Q^B_m, Z, G, Z_m \rangle.
@@
 \end{eqnarray}
-To define the action functions for $A||B$ we look at the
+To define the action functions for $A\|B$ we look at the
 restrictions of the updating functions $a^A_\delta$ and

### 2006-03-02 doc/EFA/efa.tex
EFA_branch merged into HEAD

@@
 
-\usepackage{amsthm,amsmath,amssymb}
+\usepackage{amsthm,amsmath,amssymb, mathrsfs, graphicx,makeidx,multicol}
 
@@
 \begin{document}
+\title{Modular Verification Using Extended Finite Automata}
+\author{Markus Sk\"oldstam, Martin Byr\"od}
+%\institute{Signals and Systems}
+\maketitle
 \tableofcontents
 \section{Introduction}
+To formally verify and validate PLC programs one needs a model of the controller.
+There exists various methods to build models of
+controllers,
+\begin{enumerate}
+   \item [1] modelling controllers with Signal Interpreted Petri Nets,
+   \item [2] modelling controllers with SFC,
+   \item [3] building transition systems from existing controllers,
+   \item [4] using SOL and Function Blocks to describe controllers,
+   \item [5] applying algorithms based on modular finite state machines.
 
-\subsection{Background}
-\newpage
-\section{Finite Automata (FA) with hierarchy}
+   \end{enumerate}
+ These models are usually translated to finite state machines
+ or finite automata and validated with different supporting tools. The main difference
+ between them lies in the formalism and the way the behavior of the plant is included.
+ What one would like is a decision table that assists a user in finding
+ the best approach for the application at hand. Here, we present a general framework using
+ extended finite automata that models the plant and the specification.
+ The plant and the specification are then synthesized, resulting in a model of the controller:)
 
-\section{Extended Finite Automata (EFA)}
-The regular deterministic finite automaton $A$ is a $5$-tuple
+\section{The formal definition}
+An Extended Finite Automaton (EFA) is an
+augmentation of the regular automaton with guard and action
+formulas associated to the transitions. The transitions in the EFA
+are enabled if and only if a guard formula is true. Moreover, when
+a transition in an EFA is taken, updating actions of a set of
+variables may follow. The formal definition of an EFA is here
+explained by extending the definition of the regular finite automaton.
+Typically, system behaviours are modelled using deterministic finite-state automata.
+Nondeterminism is used here to support the possibility of having different guards on transitions
+that are triggered by common events. To avoid contradictions when synchronizing
+automata, we do not allow this for the updating actions of variables
+that are shared between automata.\\
+
+The regular finite nondeterministic automaton
+$A$ is a $5$-tuple
 \begin{eqnarray}
@@
 \end{eqnarray} where $Q$ is a finite set of states; $\Sigma$ (the alphabet) is a nonempty
-set of events; $\delta$ is the transition function; $q_i$ is the
+set of events; $\delta$ is the transition relation mapping elements of $Q\times\Sigma$ into subsets of $Q$; $q_i$ is the
 initial state and $Q_m \subseteq Q$ is the set of marked states.
-An Extended Deterministic Finite Automaton (EFA) is an
-augmentation of the regular finite automaton with guard and action
-formulas associated to the transitions. The transitions in the EFA
-are enabled if and only if a guard formula is true. Moreover, when
-a transition in an EFA is taken, updating actions of a set of
-finite integer variables may follow. Formally this is done as
-follows.
 
-Given a regular automaton $A=\langle Q,\Sigma,\delta, q_i,Q_m
-\rangle$, let $T=\{(q,\sigma)\in Q \times \Sigma:
-\delta(q,\sigma)!\}$ denote the set of transitions in $A$ and let
-$\tau_k$, $k=1,...,|T|$ denote the elements (transitions) of $T$.
 First, we add to $A$ a set $X=X_1\times ... \times X_n$, where
-$X_k$, $k=1,...,n$, are finite nonempty subsets of $\mathbb{Z}$.
-The set $X$ is the domain of definition of an $n$-tuple of integer
-variables $x=(x_1,...,x_n)$ with initial value
-$x_0=(x_{10},...,x_{n0})$. We also add to $A$ an $|T|$-tuple of
-function pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
+$X_k$, $k=1,...,n$, are finite nonempty sets. For instance, $X_k$
+could be any finite subset of $\mathbb{Z}$ or the set of states in  some
+other finite deterministic automaton. The finite set $X$ is the
+domain of definition of an $n$-tuple of variables
+$x=(x_1,...,x_n)$ with initial value $x_0=(x_{10},...,x_{n0})$. If
+a variable corresponds to the set of states in an automaton, we
+shall call it a state variable. The initial value of a state
+variable must equal the initial state of its automaton and we do
+not admit updating actions of a state variable. To tell apart the
+ordinary variables and the state variables we use the subscripts
+$v$ and $s$, respectively. Using this convention we divide the
+domain of definition and the variables into parts, $X=X_v \times
+X_s$ and $x=(x_v,x_s)$.
+
+ Let $T=\{(q,\sigma,p)\in Q \times \Sigma\times Q:
+\delta(q,\sigma)=p\}$ denote the set of transitions in $A$ and let
+$\tau_k$, $k=1,...,|T|$ denote the elements (transitions) of $T$. In the next step we
+ add a $|T|$-tuple of function pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
 (g_{\tau_{|T|}},a_{\tau_{|T|}})\big)$, where each pair
@@
 $\tau_k$ is enabled or not. $a_{\tau_k}$ is an updating (action)
-function from $X$ to $X$ of the integer variables
-$x=(x_1,...,x_n)$.
+function from $X_v$ to $X_v$ of the ordinary variables $x_v$. To
+indicate the source($q$), event($\sigma$) and target($p$) of a
+function pair we write $(g_{q\sigma p},a_{q\sigma p})$. The guards and
+action formulas are partitioned into subsets $(g_{q\sigma},a_{q\sigma})=\{(g_{q\sigma p},a_{q\sigma p})\in G:
+\delta(q,\sigma)=p\}$ whose elements have the same source and target.
 
-With the above notations we will formally define an EFA as the
+Using the above notations we define an EFA as the
 $9$-tuple $A_{e}=\langle Q, \Sigma, \delta_{e}, q_i, Q_m, X, G
-,x_0, X_m \rangle$, where $X_m$ are the marked integer values of
-the variables $x$ and $\delta_{e}: Q \times \Sigma \times X
+,x_0, X_m \rangle$, where $X_m$ are the marked values of the
+variables $x$ and $\delta_{e}: Q \times \Sigma \times X
 \rightarrow Q\times X $ is a new extended version of the
-transition function $\delta:Q \times \Sigma \rightarrow Q$. The
-extended transition function $\delta_e$ is defined by means of
-$\delta$ and the elements of $G$ as
+transition mapping $\delta:Q \times \Sigma \rightarrow Q$. The
+extended mapping $\delta_e$ is defined by means of
+$\delta$, the subsets $(g_{q\sigma},a_{q\sigma})$ and the variables $x=(x_v, x_s)$ as
 \begin{eqnarray}
@@
 \begin{array}{ll}
-(\delta(q,\sigma),a_{q \sigma}(x)\big)\quad\quad g_{q \sigma}(x)=1\\
-\textrm{undefined} \quad\quad\quad\quad\quad g_{q \sigma}(x)=0
+(\delta(q,\sigma),a_{q \sigma}(x_v),x_s \big)\quad\quad g_{q \sigma}(x)=1\\
+\textrm{undefined  } \quad\quad\quad\quad\quad\quad g_{q
+\sigma}(x)=0
+\end{array}\right..
+\end{eqnarray}
+The meaning of $\delta_{e}(q,\sigma,x)$ is that when scanning
+the symbol $\sigma$ the extended automaton moves to anyone of the
+states $p$ in the set $\delta(q,\sigma)$ whose guard $g_{q
+\sigma p}(x)$ is true. When the transition takes place the variables are
+updated by the action function.
+
+The set of transitions in the extended automaton is
+$T_e=\{(q,\sigma,x,p)\in Q \times \Sigma\times X\times Q:
+\big(\delta(q,\sigma)=p \big)\wedge \big(g_{q
+\sigma p}(x)=1\big)\}$. Notice that two transitions are
+identical if and only if the events, the states and the values
+of all variables are the same. Hence, a transition $(q,\sigma,p)$ in the
+regular automaton $A$ is divided into as many transitions as the number
+of elements in the set $\{x\in X:g_{q
+\sigma p}(x)=1\}$. However, the
+alphabets of the automata are the same and the language of the
+extended automaton is subset of the language of the regular automaton:
+\begin{eqnarray}
+L(A_e)=\{s\in \Sigma^*:
+\delta_e(q_i,s,x_0)!\}\subseteq \{s\in \Sigma^*:
+\delta(q_i,s)!\}=L(A).
+\end{eqnarray}
+In some sense EFA's have small languages and a few number of
+states to the expense of many transitions. This is an important observation
+since it is the main reason why EFA's enable efficient modular algorithms to perform
+verifications of supervisors.
+\subsection{How to use the variables}
+All logical systems can easily be represented by
+EFA models. When using EFA's as a modelling tool it important to understand how to choose the variables.
+A variable $x$ can represent many things, some examples  are: the states that an object can be in,
+ the number of customers in a queue system, the different types of customers in a queue or the time it
+takes for a process to execute different operations. In many applications the variables can be represented by
+ integers or enumeration objects.
+
+\subsubsection{Integer variables}
+
+\subsubsection{Enumeration variables}
+
+\section{Translating EFA to FA}
+
+An EFA can be translated into its corresponding FA.
+If $A$ is the translation of $A_e$ we shall use the notation
+$A=trans(A_e)$. When an EFA is "flatten out" the number of states
+increases to $Q \times X_v $. The reason for this is that all the
+actions are removed and translated into events in the regular FA. Notice
+that when $A_e$ only has ordinary variables $X=X_v$. If $A_e$ also has
+state variables the guard conditions on these variables are evaluated. When a variable
+corresponds to the set of states
+in an other automata this is not possible. Therefore its guard conditions are set to
+TRUE when the translation is performed. To simplify the
+algorithm we shall restrict us to quite simple action- and guard
+functions. We shall not compare variables with each other in the guard functions.
+The updating of the variables are assumed to be independent
+from each other. This implies that they can be written as
+tuples of functions of one variable i.e.
+$a_{q \sigma p}(x)=(a^1_{q \sigma p}(x),...,a^n_{q \sigma p}(x))=(a^1_{q
+\sigma p}(x_1),...,a^n_{q \sigma p}(x_n))$. We shall assume that the
+guards are parsed and written in disjunctive normal form $g_{q
+\sigma p}(x)=g_1(x)\vee ...\vee g_j(x)= (g^{1}_{1}(x_1)\wedge
+...\wedge g^{n}_{1}(x_n) )\vee ...\vee (g^{1}_{j}(x_1)\wedge
+...\wedge g^{n}_{j}(x_n) )$.
+
+These assumptions makes it
+straight forward to translate an EFA to an FA. Given an extended
+finite automaton $A_{e}=\langle Q, \Sigma, \delta, q_i, Q_m, X, G
+,x_0, X_m \rangle$ with simple guard and
+action functions, the steps are:\\
+
+\begin{enumerate}
+   \item [0] In the guard formulas (written in disjunctive normal form) the guards on
+             the state variables are evaluated. If an and-clause belonging to a state variable is
+             FALSE the "OR" component that it belongs to is removed, and if it is TRUE the and-clause
+             is removed. The number of "OR" components in the modified
+             guards will in what follows be denoted by $|g_{q \sigma p}|$.
+
+   \item [1] Form a regular FA, $A'$, from $A_{e}$ by:
+   \begin{enumerate}
+   \item [(i)] removing $G$, $X$, $X_m$, $x_0$ and the alphabet $\Sigma$,
+
+   \item [(ii)] form the alphabet of $A'$ by giving unique names to all events that are enabled in $A_{e}$,
+
+   \item [(iii)] update the transitions by changing the names of the events accordingly.
+
+   \item [(iv)] divide each event in $A'$ into $|g_{q \sigma p}|$ events with unique
+   names,
+   \item [(v)] split the corresponding transition into $|g_{q \sigma p}|$
+transitions using the new event names.
+
+   \end{enumerate}
+
+ \item [2]  Realize an automaton $A_{k}$ for each one of the ordinary variables $x_{v}$,
+ with as many events the total number of events in $A'$. The events in $A_k$ are associated
+ to the guard function belonging
+ to the right "OR" component of the guard. The transitions
+correspond to the action functions and they are enabled if the corresponding
+guard function is true.
+ \item [3]  The wanted finite automaton $A$ is formed by synchronizing $A_{k}$, $k=1,...,m$ with $A'$.
+ Here, the index $m$ is the total number of ordinary variables in $A_e$.
+ \item [4] Change the event names of $A$ to the original ones. Add to $A$ the possible events
+ in $A_{e}$ not present in any transitions, so that the two alphabets agree.
+\end{enumerate}
+The number of states in the variable automata equals the range of the corresponding variable.
+In particular, the marked states of a variable automaton are those that correspond to the set
+of marked values of that variable.
+
+In the last two steps  of the algorithm we obtain a monolithic description of $A_e$. To avoid the so called
+ state-space explosion problem, we must investigate the modular automata model obtained in step 2. However,
+ the modular structure in the algorithm does not have the correct alphabet and can therefore not be used to
+ verify specifications against a given plant model. For instance, suppose we have a model of the plant and
+ a specification in EFA form. To analyze the specification against the plant and at the same time
+use the modular structure in the translation algorithm we must translate them simultaneously.
+
+Note that if two EFA's has common variables, we cannot first translate them
+separately into standard FA and then synchronize them. Either this
+has to be done simultaneously or the automata must be synchronized
+on a higher level. The same problem occurs when translating EFA's with
+state variables. However, by synchronizing all automata with
+common variables an EFA with local ordinary variables can be constructed.
+
+\section{The synchronous product}
+
+The synchronous composition of two EFA's must consider that the automata
+may have both events and variables in common. When synchronizing two automata
+$A_e=\langle Q^A, \Sigma^A, \delta^A_e, q^A_i, Q^A_m,
+X^A, G^{A}, x^{A}_0, X^{A}_m \rangle$ and $B_e=\langle Q^B, \Sigma^B,
+\delta^B_e, q^B_i, Q^B_m, X^B, G^{B}, x^{B}_0, X^{B}_m \rangle$ that have
+common variables one must check that $A_e$ and $B_e$ are
+consistent. First of all, the common variables must have the same
+initial values. Secondly, we require that the action functions
+triggered by a common enabled event $\sigma\in \Sigma_{A}\cap
+\Sigma_{B}$ in the two automata update common variables to the
+same values, i.e. if the $i^{ith}$ variable in $x^{A}$ and the
+$j^{ith}$ variable in $x^{B}$ are the same then $(a_{q^A \sigma}(x^{A}))_i = (a_{q^B
+\sigma}(x^{B}))_j$. However, if the corresponding guard
+formula is false it may happen that $a_{q^A \sigma p^A}(x^{A}))_i \neq (a_{q^B
+\sigma p^B}(x^{B}))_j$ even though the two automata are consistent.
+To avoid such situations we require that action functions of common
+variables are identical.
+
+Before we define the synchronous product we introduce two more
+subscripts, $l$ and $c$. These are to be used for the variables.
+The subscript $l$ indicates local variables and the subscript $c$
+indicates common variables. In particular the subscripts can be
+used together with the subscripts for ordinary variables and state
+variables ($v$  and $s$, respectively). For instance, if we write
+$x_{cv}$, then $x_{cv}$ is a common ordinary variable.
+
+\subsection{Defining the product}\label{defining the product}
+
+Formally we can write the synchronous product $A_e \| B_e$ of two
+EFA's $A_e=\langle Q^A, \Sigma^A, \delta^A_e, q^A_i, Q^A_m, X^A,
+G^{A}, x^{A}_0, X^{A}_m \rangle$ and $B_e=\langle Q^B, \Sigma^B,
+\delta^B_e, q^B_i, Q^B_m, X^B, G^{B}, x^{B}_0, X^{B}_m \rangle$ as
+\begin{eqnarray}
+A_e \| B_e=\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B ,\nonumber\\
+\delta^{A\|B}_e, (q^{A}_i, q^{B}_i),\nonumber\\
+ Q^A_m\times Q^B_m, X^{A\|B},\nonumber\\
+G^{A\|B}, x^{A\|B}_0, X^{A\|B}_m\rangle.
+\end{eqnarray}
+We need to define the variables, the guard functions, the action
+functions and the transition function of the product. Let $x_{c}$
+denote the common variables of $x^{A}$ and $x^{B}$, and let
+$x^A_l$ and $x^B_l$ be the distinct variables of $x^{A}$ and
+$x^{B}$, respectively. The variables of the product of $A$ and $B$
+can be written as the vector $x^{A\|B}=(x^A_l, x_{c}, x^B_l)$. We
+define the guard functions $g_{q \sigma}(x^{A\|B})$ as
+\begin{eqnarray}\label{guard_sync_2}
+g_{q \sigma}(x^{A\|B})=\left\{
+\begin{array}{ll}
+g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \textrm{ if }\sigma\in \Sigma^A \cap \Sigma^B\\
+g_{q^A\sigma}(x^{A}) \quad\quad \textrm{ if } \sigma\in \Sigma^A \setminus \Sigma^B\\
+g_{q^B\sigma}(x^{B}) \quad\quad \textrm{ if }
+\sigma\in\Sigma^B\setminus \Sigma^A
 \end{array}\right.
 \end{eqnarray}
-If $(q,\sigma)\notin T$ then $g_{q\sigma}$ is undefined and
-therefore $\delta_{e}(q,\sigma,x)$ is undefined. Notice that for
-the set of transitions $T_e$ in the extended automaton $A_e$ we
-have $T_e=\{(q,\sigma)\in Q \times \Sigma: g_{q\sigma}(x)=1
-\textrm{ for some } x\in X\}\subset \{(q,\sigma)\in Q \times
-\Sigma: g_{q\sigma} \textrm{ is defined } \}\ =T$.
+If any one of the guards functions in the above definition are
+undefined the guard function for the product $g_{q \sigma}$ is
+also undefined. To simplify the notation further, we use "dummy"
+guard functions in the following way. If $\sigma\in \Sigma^A
+\setminus \Sigma^B$ we introduce fictional guards functions
+$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
+\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. With this convention we can
+write equation (\ref{guard_sync_2}) as
+\begin{eqnarray}
+g_{q \sigma}(x^{A\|B})= g_{q^A\sigma}(x^{A})\wedge
+g_{q^B\sigma}(x^{B}).
+\end{eqnarray}
+By $a^B_{q \sigma}(x^{B}_{vl})$, we mean the restriction of
+$a^B_{q^B \sigma}(x^{B}_{v})$ to the local ordinary variables of
+$B$. The updating functions are defined as (if $x^{A}$ and $x^{B}$
+have no common variables then $x^{A\|B}_v=(x^{A}_v,x^{B}_v)$)
+\begin{eqnarray}\label{action_sync_2}
+a_{q \sigma}(x^{A\|B}_v)=\left\{
+\begin{array}{ll}
+ \big(a^A_{q^A \sigma}(x^{A}_v), a^B_{q^B \sigma}(x^{B}_{vl})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
+\big(a^A_{q^A \sigma}(x^{A}_v), x^{B}_{vl}\big)\quad\quad\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
+\big(x^{A}_{vl}, a^B_{q^B
+\sigma}(x^{B}_v)\big)\quad\quad\quad\quad\sigma\in \Sigma^B
+\setminus \Sigma^A
+\end{array}\right.
+\end{eqnarray}
+As in the definition of the guard functions the notion of being
+undefined is inherited from the action functions of $A$ and $B$.
+Similar to the "dummy" guard functions we introduce fictional
+action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
+introduce fictional action functions $a_{q^B\sigma}(x^B_v)=x^B_v$
+for $B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
+$a_{q^A\sigma}(x^A_v)=x^A_v$. If $x^{A}_v$ and $x^{B}_v$ have no
+common variables then equation (\ref{action_sync_2}), that defines
+the action functions of the product, simplifies to
+\begin{eqnarray}
+a_{q \sigma}(x^{A\|B}_v)=
+ \big(a^A_{q^A \sigma}(x^{A}_v), a^B_{q^B \sigma}(x^{B}_v)\big).
+\end{eqnarray}
 
-\subsection{Synchronous composition of EFA's}
+ The transition function
+$\delta^{A\|B}_e$ is defined using the regular transitions
+functions $\delta^A$ and $\delta^B$ together with the guard and
+action functions of the product,
+\begin{eqnarray}
+\delta^{A\|B}_{e}(q,\sigma,x^{A\|B})=\left\{
+\begin{array}{ll}
+\big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(x^{A\|B}_v), x^{A\|B}_s \big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(x^{A\|B})=1\\
+\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(x^{A\|B}_v), x^{A\|B}_s\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(x^{A\|B})=1\\
+\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(x^{A\|B}_v), x^{A\|B}_s\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(x^{A\|B})=1\\
+\textrm{
+undefined}\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+g_{q\sigma}(x^{A\|B})=0
+\end{array}\right.
+\end{eqnarray}
 
-Formally we define the synchronous product $A\|B$ of two EFA's
+\subsection{Properties of the product}
+
+\subsubsection{When the product commutes with the translation algorithm}
+
+The synchronous product $A_e\|B_e$ of two EFA that only have local ordinary
+variables can be translated to $A\|B$ by translating
+$A_e$ and $B_e$ separately into finite automata and then
+synchronizing. In other words, the operations translating into
+$FA$ and product commute, $trans({A\|B}_e) =trans(A_e)\|trans(B_e)$.
+Before we prove this important property we shall look at the formulas
+ for the product when the components have local ordinary variables.
+ They become much easier to read.
+
+In section \ref{defining the product} we defined the synchronous product $A_e\|B_e$
+of two EFA's. Since here, we only consider automata with ordinary local variables,
+the variables for the product simplifies to $x=(x^A,x^B)$. Hence, can write the product as
 \begin{eqnarray}
-A=\langle Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X^{A},
-G^A,x^A_0,X^A_m \rangle
-\end{eqnarray} and \begin{eqnarray} B=\langle Q^B, \Sigma^B, \delta^B, q^B_i,
-Q^B_m, X^{B}, G^B,x^B_0,X^B_m \rangle
-\end{eqnarray} as
-\begin{eqnarray}
-A\|B&=&\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B ,
-\delta^{A\|B}, (q^A_i,q^B_i) ,\nonumber\\
+A_e\|B_e&=&\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B ,
+\delta^{A\|B}_e, (q^A_i,q^B_i) ,\nonumber\\
 && Q^A_m\times Q^B_m, X^A \times X^B, G^{A\|B},(x^A_0,x^B_0),
@@
 \end{eqnarray}
-To define the guard functions, the action functions and the
-transition function of the product we make use of the
-corresponding functions in $A$ and $B$. To denote the elements of
-$A\|B$ we remove unnecessary  superscripts and write $A\|B=\langle
-Q, \Sigma, \delta, q_i, Q_m, X, G ,x_0, X_m \rangle$. The guard
-functions $g_{q \sigma}(x)$ are defined as
-\begin{eqnarray}\label{guard_sync_1}
-g_{q \sigma}(x)=\left\{
-\begin{array}{ll}
-g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
-g_{q^A\sigma}(x^{A})\quad\quad\quad\quad\quad\quad\quad\sigma\in \Sigma^A\setminus\Sigma^B\\
-g_{q^B\sigma}(x^{B})\quad\quad\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
-\end{array}\right.
+Removing unnecessary  superscripts we can write $A_e\|B_e=\langle
+Q, \Sigma, \delta_e, q_i, Q_m, X, G ,x_0, X_m \rangle$. To simplify the
+notation and increase readability, we use "dummy" guard functions
+and fictional action function as in section \ref{defining the product}.
+Doing this, we can write the guard and action formulas for the product as
+\begin{eqnarray}
+g_{q \sigma}(x)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B})
 \end{eqnarray}
-If in any of the three cases, the corresponding guard function is
-undefined the guard function for the product $g_{q \sigma}$ is
-also undefined. To simplify the notation and increase readability,
-we use "dummy" guard functions in the following way. If $\sigma\in
-\Sigma^A \setminus \Sigma^B$ we introduce fictional guards
-functions $g_{q^B\sigma}(.)=1$ for $B$ and if
-$\sigma\in\Sigma^B\setminus \Sigma^A$ we set $g_{q^A\sigma}(.)=1$.
-Doing this, we can write equation (\ref{guard_sync_1}) as
-\begin{eqnarray}
-g_{q \sigma}(x)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
-\end{eqnarray}
-The updating functions $a_{q \sigma}(x)$ are defined as
-\begin{eqnarray}\label{action_sync_1}
-a_{q \sigma}(x)=\left\{
-\begin{array}{ll}
- \big(a_{q^A \sigma}(x^{A}), a_{q^B \sigma}(x^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
-\big(a_{q^A\sigma}(x^{A}),x^{B}\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^A\setminus\Sigma^B\\
-\big(x^{A},a_{q^B\sigma}(x^{B})\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
-\end{array}\right.
-\end{eqnarray}
-As in the definition of the guard functions (\ref{guard_sync_1}),
-the notion of being undefined is inherited from the action
-functions of $A$ and $B$. Similar to the "dummy" guard functions
-we introduce fictional action functions. If $\sigma\in \Sigma^A
-\setminus \Sigma^B$ we introduce fictional action functions
-$a_{q^B\sigma}(x^B)=x^B$ for $B$ and if
-$\sigma\in\Sigma^B\setminus \Sigma^A$ we set
-$a_{q^A\sigma}(x^A)=x^A$. With this convention equation
-(\ref{action_sync_1}) simplifies to
+and
 \begin{eqnarray}
@@
 \end{eqnarray}
-
- The transition function $\delta^{A\|B}$ is defined from $\delta^A$ and $\delta^B$ as
+The transition function $\delta^{A\|B}_e$ simplifies since we do not have any
+state variables ($\delta^A$ and $\delta^B$ are the transition functions
+for the corresponding regular automata)
 \begin{eqnarray}
-\delta^{A\|B}(q,\sigma,x)=\left\{
+\delta^{A\|B}_e(q,\sigma,x)=\left\{
 \begin{array}{ll}
@@
 
-\subsubsection{Properties of the product}
+\begin{theorem}
+Two automata with no common variables $A_e$ and $B_e$ fulfills
+$trans({A\|B}_e) =trans(A_e)\|trans(B_e)$.
+\end{theorem}
+
+\subsubsection{Associativity}
 
@@
 \noindent \textbf{Proof}: We need to prove that $(A\|B)\|C =
-A\|(B\|C)$. Here, we assume that the three automata have no common
-variables.
+A\|(B\|C)$. For simplicity, we assume that the three automata have
+no common variables.
 \newline\newline
@@
 
-The synchronous product $A\|B$ of two FSA $A$ and $B$ can be
-extended to $(A\|B)_e$ by taking the synchronous product of the
-extended components  $A_e$ and $B_e$. In other words, the
-operations extension and product commute, $(A\|B)_e=A_e\|B_e$.
-
-\subsection{The translation of an EFA to  an FA}
-To simplify the realization we shall restrict us to quite simple
-action- and guard functions. For the updating functions we assume
-that $a_{q \sigma}(x)=(a^1_{q \sigma}(x),...,a^n_{q
-\sigma}(x))=(a^1_{q \sigma}(x_1),...,a^n_{q \sigma}(x_n))$. In
-other words, the updating of the variables are independent from
-each other. For the guard functions we assume that they are given
-in disjunctive normal form $g_{q \sigma}(x)=g^1_{q \sigma}(x)\vee
-...\vee g^j_{q \sigma}(x)$. We shall denote the number of "OR"
-components $j$ in such an expression by $|g_{q \sigma}|$. These
-assumptions makes it straight forward to translate an EFA to an
-FA.
-
-Given an extended finite automaton $A=\langle Q, \Sigma, \delta,
-q_i, Q_m, X, G ,x_0, X_m \rangle$ with simple guard and action
-functions, the algorithm can be divided into four steps:\\
-- Form a regular FA, $A'$, from $A$ by removing $G$, $X$, $X_m$,
-$x_0$ and change the names of all enabled events $\sigma$ to
-$q\sigma$ i.e. the associate them with the corresponding
-transition.\\
-- Rename the events $q\sigma$ in $A'$ further into $|g_{q \sigma}|$ identical transitions with event names: $\{(q\sigma)_k\}^{|g_{q \sigma}|}_{k=1}$.\\
-- Realize an automata $A_{x_k}$ for each variable $x_k$, with
-marked "states" $X_{mk}$, initial state $x_{k 0}$ and with as many events the total number of events in $A'$.
-The transitions correspond to the action functions and they are enabled if the guard formula is true.\\
-- Synchronize all automata $A_{x_k}$, $k=1,...,n$ and $A'$.\\
-- Change back to original event names.
-\newpage
-\subsection{Translating hierarchical structures to EFA's}
-It is well known that hierarchical structures of states and
-transitions reduce the descriptive complexity of automata models.
-We shall in this section show how the hierarchy in state charts
-can be translated into EFA's.
+\section{Translating hierarchical structures to EFA's} It is well
+known that hierarchical structures of states and transitions
+reduce the descriptive complexity of automata models. We shall in
+this section show how the hierarchy in state charts can be
+translated into EFA's.
 
@@
 
-\subsection{Synchronous composition of EFA's with common variables}
+\section{Remarks}
+\begin{enumerate}
+   \item [1] Shall we remove marked integer values from our definition of
+EFA's? Will the user only want to mark the states?
 
-The synchronous composition of two EFA's $A=\langle Q^A, \Sigma^A,
-\delta^A, q^A_i, Q^A_m, X^{A}, G^A, X^A_m \rangle$ and $B=\langle
-Q^B, \Sigma^B, \delta^B, q^B_i, Q^B_m, X^{B}, G^B, X^B_m \rangle$
-with common variables exists if and only if the automata are
-consistent. The consistency condition requires that the action
-functions triggered by a common enabled event $\sigma\in
-\Sigma_{A}\cap \Sigma_{B}$ in the two automata update common
-variables to the same values, i.e. if the $i^{ith}$ variable in
-$x^{A}$ and the $j^{ith}$ variable in $x^{B}$ are the same
-variable and $g_{q^A \sigma}(x^{A})\wedge g_{q^B \sigma}(x^{B})=1$
-where $(q^A,q^B)\in Q^A\times Q^B$ then $(a_{q^A \sigma}(x^{A}))_i
-= (a_{q^B \sigma}(x^{B}))_j$.
+    \item [2] Waters can only load one file. This must be fixed
+    before synch. of EFA's.
 
-The consistency condition can be verified by inspection. However,
-if the vectors $x^{A}$ or $x^{B}$ disable the corresponding guard
-formula ($g_{q^A \sigma}(x^{A})\wedge g_{q^B \sigma}(x^{B})$) it
-may happen that $a_{q^A \sigma}(x^{A}))_i \neq (a_{q^B
-\sigma}(x^{B}))_j$ even though the two automata are consistent.
-Therefore it may be a time-consuming task to verify the
-consistency condition by inspection. In section
-\ref{Implementation} we will present algorithms that includes a
-consistency check of the synchronous
-composition of two EFA's. \\
+    \item [3] Can we translate EFA's to Tord's stuff?
 
-Formally we define the synchronous product $A\|B$ of two EFA's
-$A=\langle Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X^{A}, G^A,X^A_m
-\rangle$ and $B=\langle Q^B, \Sigma^B, \delta^B, q^B_i, Q^B_m,
-X^{B}, G^B,X^B_m
- \rangle$ as
-\begin{eqnarray}
-A\|B=\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B , \delta, q_i,
-Q^A_m\times Q^B_m, Z, G, Z_m \rangle.
-\end{eqnarray}
-To define the integer variables, the guard functions, the action
-functions and the transition function for the product we make use
-of the corresponding elements in $A$ and $B$.
+    \item [4] We must find examples where EFA's are easy to use.
 
- Let $z^{AB}$ denote the common variables of $x^{A}$ and $x^{B}$, and let $z^A$
-and $z^B$ be the distinct variables of $x^{A}$ and $x^{B}$,
-respectively. We define the integer variables $z$ of the
-synchronous product of $A$ and $B$ as the vector $z=(z^A, z^{AB},
-z^B)$. The guard functions $g_{q \sigma}(z)$ are defined as
-\begin{eqnarray}\label{guard_sync_2}
-g_{q \sigma}(z)=\left\{
-\begin{array}{ll}
-g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \textrm{ if }\sigma\in \Sigma^A \cap \Sigma^B\\
-g_{q^A\sigma}(x^{A}) \quad\quad \textrm{ if } \sigma\in \Sigma^A \setminus \Sigma^B\\
-g_{q^B\sigma}(x^{B}) \quad\quad \textrm{ if }
-\sigma\in\Sigma^B\setminus \Sigma^A
-\end{array}\right.
-\end{eqnarray}
-If any one of the guards functions in the above definition are
-undefined the guard function for the product $g_{q \sigma}$ is
-also undefined. To simplify the notation further, we use "dummy"
-guard functions in the following way. If $\sigma\in \Sigma^A
-\setminus \Sigma^B$ we introduce fictional guards functions
-$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
-\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. With this convention we can
-write equation (\ref{guard_sync_2}) as
-\begin{eqnarray}
-g_{q \sigma}(z)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
-\end{eqnarray}
-To define the action functions for $A\|B$ we look at the
-restrictions of the updating functions $a^A_\delta$ and
-$a^B_\delta$ to the sets $Z^A$ and $Z^B$. The updating functions
-are defined as (if $x^{A}$ and $x^{B}$ have no common variables
-then $z=(x^{A},x^{B})$)
-\begin{eqnarray}\label{action_sync_2}
-a_{q \sigma}(z)=\left\{
-\begin{array}{ll}
- \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(z^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
-\big(a^A_{q^A \sigma}(x^{A}), z^{B}\big)\quad\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
-\big(z^{A}, a^B_{q^B \sigma}(x^{B})\big)\quad\quad\quad\sigma\in
-\Sigma^B \setminus \Sigma^A
-\end{array}\right.
-\end{eqnarray}
-As in the definition of the guard functions the notion of being
-undefined is inherited from the action functions of $A$ and $B$.
-Similar to the "dummy" guard functions we introduce fictional
-action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
-introduce fictional action functions $a_{q^B\sigma}(x^B)=x^B$ for
-$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
-$a_{q^A\sigma}(x^A)=x^A$. If $x^{A}$ and $x^{B}$ have no common
-variables then equation (\ref{action_sync_2}), that defines the
-action functions of the product, simplifies to
-\begin{eqnarray}
-a_{q \sigma}(z)=
- \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
-\end{eqnarray}
+    \item [5] We want to find examples where it is better to
+    synch. two EFA's than to translate to FA's and then synch.
 
- The transition function
-$\delta(q,\sigma,z)$ is defined from $\delta^A$ and $\delta^B$ as
-\begin{eqnarray}
-\delta(q,\sigma,z)=\left\{
-\begin{array}{ll}
-\big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(z)\big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(z)=1\\
-\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(z)=1\\
-\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(z)=1\\
-\textrm{
-undefined}\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
-g_{q\sigma}(z)=0
-\end{array}\right.
-\end{eqnarray}
+    \item [6] Find smart algorithm's  than convert FA's to EFA's.
 
-\section{Implementation of EFA's with hierarchy in Supremica}
+     \item [7] Decide what to write in our article.
 
-\subsection{Supremica}
+    \item [8] Look at EFA's as a supervisory control tool? this is
+    done in all standard papers.
 
-\subsection{Implementation}\label{Implementation}
+    \end{enumerate}
+
+\section{Implementation}\label{Implementation}
 \subsection{Examples}
+
+\begin{thebibliography}{99}
+
+\bibitem{Chen} Y-L. Chen and F. Lin (2000). Modeling of Discrete Event Systems
+Using Finite State Machines with Parameters, emph{Proc. of the 2000 IEEE International Conference on Control Applications},
+Anchorag, Alaska, USA, pp. 941-946.
+
+\bibitem{Yang}Y. Yang and P. Gohari(2005). Embedded supervisory control of discrete-event systems,
+\emph{Proc. of the 2005 IEEE International Conference on Automation Science and Engineering},
+Edmonton, Canada, pp. 410-415.
+
+\end{thebibliography}
+
 \end{document}

### 2006-03-02 src/net/sourceforge/waters/gui/ControlledSurface.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.gui
 //# CLASS:   ControlledSurface
 //###########################################################################
-//# $Id: ControlledSurface.java,v 1.61 2006-01-23 23:32:34 siw4 Exp $
+//# $Id: ControlledSurface.java,v 1.62 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
  
 package net.sourceforge.waters.gui;
@@ -323,6 +323,14 @@ public class ControlledSurface
 					EditorEdgePopupMenu popup = new EditorEdgePopupMenu(this, object);
 					popup.show(this, e.getX(), e.getY());
 				}
+				if (o.getType() == EditorObject.LABELGROUP) {
+					EditorLabelGroup object =
+						(EditorLabelGroup) o;
+					
+					EditorLabelBlockPopupMenu popup = new
+						EditorLabelBlockPopupMenu(this, object);
+					popup.show(this, e.getX(), e.getY());
+				}
 			}
 			else
 			{
@@ -489,7 +497,8 @@ public class ControlledSurface
 						operation = DnDConstants.ACTION_COPY;
 					} else if (o instanceof EditorEdge) {
 						operation = DnDConstants.ACTION_COPY;
-					} else if (o instanceof EditorLabelGroup) {
+					} else if (o instanceof EditorLabelGroup && 
+							!(o instanceof EditorGuardActionBlock)) {
 						operation = DnDConstants.ACTION_COPY;
 					} else if (o instanceof EditorLabel) {
 						operation = DnDConstants.ACTION_COPY;
@@ -504,7 +513,8 @@ public class ControlledSurface
 						   ek.equals(EventKind.UNCONTROLLABLE)) {
 					if (o instanceof EditorEdge) {
 						operation = DnDConstants.ACTION_COPY;
-					} else if (o instanceof EditorLabelGroup) {
+					} else if (o instanceof EditorLabelGroup && 
+							!(o instanceof EditorGuardActionBlock)) {
 						operation = DnDConstants.ACTION_COPY;
 					}
 				}
@@ -545,7 +555,8 @@ public class ControlledSurface
 						addToEdge((EditorEdge) o, ip, e);
 						e.dropComplete(true);
 						return;
-					} else if (o instanceof EditorLabelGroup) {
+					} else if (o instanceof EditorLabelGroup && 
+							!(o instanceof EditorGuardActionBlock)) {
[…]

### 2006-03-02 src/net/sourceforge/waters/gui/EditorAction.java

@@ -1,0 +1,83 @@
+package net.sourceforge.waters.gui;
+
+import java.awt.event.FocusEvent;
+import java.awt.event.FocusListener;
+import java.awt.event.MouseEvent;
+import java.awt.event.MouseMotionListener;
+import javax.swing.*;
+
+import net.sourceforge.waters.model.compiler.CompilerOperatorTable;
+import net.sourceforge.waters.model.expr.ExpressionParser;
+import net.sourceforge.waters.model.expr.ParseException;
+import net.sourceforge.waters.model.module.BinaryExpressionProxy;
+import net.sourceforge.waters.model.module.SimpleExpressionProxy;
+import net.sourceforge.waters.subject.module.BinaryExpressionSubject;
+import net.sourceforge.waters.subject.module.GuardActionBlockSubject;
+import net.sourceforge.waters.subject.module.ModuleSubjectFactory;
+
+public class EditorAction extends JTextField 
+	implements FocusListener{
+
+	private SimpleExpressionProxy mActionExpression;
+	private ExpressionParser mParser;
+	private ModuleSubjectFactory mFactory;
+	private GuardActionBlockSubject mGuardActionBlock;
+	private EditorGuardActionBlock mParent;
+	
+	public EditorAction(BinaryExpressionProxy action,
+			EditorGuardActionBlock editorGuardActionBlock) {
+		super(" " + action.toString());
+		mParent = editorGuardActionBlock;
+		mGuardActionBlock = editorGuardActionBlock.getGuardActionBlock();
+		mFactory = ModuleSubjectFactory.getInstance();
+		mParser = new ExpressionParser(mFactory,
+				CompilerOperatorTable.getInstance());
+		mActionExpression = action;
+		this.addFocusListener(this);
+	}
+
+	private void update() {
+		SimpleExpressionProxy newExpression = null;
+		
+		if(this.getText().trim().equals("")) {
+			mParent.remove(this);
+			mParent.resizePanel();
+			mParent.getParent().getParent().repaint();
+			return;
+		}
+		
+		//check syntax
+		try {
+			newExpression = mParser.parse(this.getText());
+		} catch (ParseException e) {
+			// TODO Auto-generated catch block
+			e.printStackTrace();
+		}
+		
+		BinaryExpressionSubject action = new BinaryExpressionSubject(
+				((BinaryExpressionSubject) newExpression).getOperator(),
[…]

### 2006-03-02 src/net/sourceforge/waters/gui/EditorColor.java

@@ -13,6 +13,9 @@ class EditorColor
 
 	/** The default color of marked (accepting) nodes. */
 	public static Color DEFAULTMARKINGCOLOR = Color.GRAY; 
+	
+	/** The default color of marked (accepting) nodes. */
+	public static Color GUARDACTIONHEADER = Color.GRAY; 
 
 	/** The color of the drag-select area. */
 	public static Color DRAGSELECTCOLOR = new Color(0,0,255,32); 

### 2006-03-02 src/net/sourceforge/waters/gui/EditorEdge.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.gui
 //# CLASS:   EditorEdge
 //###########################################################################
-//# $Id: EditorEdge.java,v 1.34 2006-01-17 02:00:07 siw4 Exp $
+//# $Id: EditorEdge.java,v 1.35 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.gui;
@@ -58,7 +58,13 @@ public class EditorEdge
 	private EditorObject startNode; 
 	/** The end is a node. */
 	private EditorNode endNode;
-
+	
+	/** The label block (events) **/
+	private EditorLabelGroup mEditorLabelGroup;
+	
+	/** The guard+action block **/
+	private EditorGuardActionBlock mEditorGuardActionBlock;
+	
 	/** Boolean keeping track of whether the edge is a straight line or not. */
 	private boolean straight;
 
@@ -72,24 +78,24 @@ public class EditorEdge
 	private boolean dragT = false;	
 	private Rectangle2D.Double center = new Rectangle2D.Double();
 	private boolean dragC = false;
-
 	private EdgeSubject subject;
 	//private Point2D start;
 	private static double tearRatio = .8;
 	private static int WIDTHD = 2;
 	private static int WIDTHS = 5;
+	private EditorSurface mParent;
 
-	public EditorEdge(EditorObject iStartNode, EditorNode iEndNode, int x, int y, EdgeSubject e)
+	public EditorEdge(EditorObject iStartNode, EditorNode iEndNode,
+			int x, int y, EdgeSubject e, EditorSurface parent)
 	{
 		// This is an edge
 		type = EDGE;
-		
 		Point2D start = null;
-
 		startNode = iStartNode;
 		endNode = iEndNode;
 		subject = e;
 		endNode.attach(this);
+		mParent = parent;
 		if (startNode.getType() == NODE)
 		{
 			EditorNode s = (EditorNode) startNode;
@@ -1163,11 +1169,51 @@ public class EditorEdge
 		arrowAtEnd = set;
 	}
 
+	public EditorGuardActionBlock getEditorGuardActionBlock() {
+		return mEditorGuardActionBlock;
+	}
+
+	public void setEditorGuardActionBlock(
+			EditorGuardActionBlock editorGuardActionBlock) {
[…]

### 2006-03-02 src/net/sourceforge/waters/gui/EditorGuard.java

@@ -1,0 +1,72 @@
+package net.sourceforge.waters.gui;
+
+import java.awt.event.FocusEvent;
+import java.awt.event.FocusListener;
+
+import javax.swing.JTextField;
+
+import net.sourceforge.waters.model.compiler.CompilerOperatorTable;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable;
+import net.sourceforge.waters.model.expr.ExpressionParser;
+import net.sourceforge.waters.model.expr.ParseException;
+import net.sourceforge.waters.model.module.BinaryExpressionProxy;
+import net.sourceforge.waters.model.module.SimpleExpressionProxy;
+import net.sourceforge.waters.subject.module.BinaryExpressionSubject;
+import net.sourceforge.waters.subject.module.GuardActionBlockSubject;
+import net.sourceforge.waters.subject.module.ModuleSubjectFactory;
+
+public class EditorGuard extends JTextField
+	implements FocusListener{
+
+	private String mGuardExpression;
+	private ExpressionParser mParser;
+	private ModuleSubjectFactory mFactory;
+	private GuardActionBlockSubject mGuardActionBlock;
+	private EditorGuardActionBlock mParent;
+	
+	public EditorGuard(String guard, EditorGuardActionBlock editorGuardActionBlock) {
+		super(" " + guard.toString());
+		mParent = editorGuardActionBlock;
+		mGuardActionBlock = editorGuardActionBlock.getGuardActionBlock();;
+		mFactory = ModuleSubjectFactory.getInstance();
+		mParser = new ExpressionParser(mFactory,
+				GuardExpressionOperatorTable.getInstance());
+		mGuardExpression = guard;
+		this.addFocusListener(this);
+	}
+
+	private void update() {
+		SimpleExpressionProxy newExpression = null;
+		
+		if(this.getText().trim().equals("")) {
+			mParent.remove(this);
+			mParent.resizePanel();
+			mParent.getParent().getParent().repaint();
+			return;
+		}
+		
+		//check syntax
+		try {
+			newExpression = mParser.parse(this.getText());
+		} catch (ParseException e) {
+			// TODO Auto-generated catch block
+			e.printStackTrace();
+		}
+		
+		mGuardActionBlock.setGuard(this.getText());
+		mGuardExpression = this.getText();
+		mParent.resizePanel();
+		mParent.getParent().getParent().repaint();
+	}
+	
[…]

### 2006-03-02 src/net/sourceforge/waters/gui/EditorGuardActionBlock.java

@@ -1,0 +1,223 @@
+package net.sourceforge.waters.gui;
+
+import java.awt.Component;
+import java.awt.Container;
+import java.awt.Font;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.awt.event.MouseEvent;
+import java.awt.event.MouseListener;
+import java.util.List;
+import java.util.ArrayList;
+
+import javax.swing.BorderFactory;
+import javax.swing.JComponent;
+import javax.swing.JLabel;
+import javax.swing.JPanel;
+import javax.swing.JTextField;
+import javax.swing.border.EmptyBorder;
+import javax.swing.event.DocumentEvent;
+import javax.swing.event.DocumentListener;
+
+import net.sourceforge.waters.model.base.Proxy;
+import net.sourceforge.waters.model.module.BinaryExpressionProxy;
+import net.sourceforge.waters.subject.module.BinaryExpressionSubject;
+import net.sourceforge.waters.subject.base.AbstractSubject;
+import net.sourceforge.waters.subject.base.ModelChangeEvent;
+import net.sourceforge.waters.subject.module.GuardActionBlockSubject;
+import net.sourceforge.waters.subject.module.IdentifierSubject;
+import net.sourceforge.waters.subject.module.SimpleIdentifierSubject;
+
+public class EditorGuardActionBlock extends EditorLabelGroup
+	implements MouseListener{
+
+	private static final int LEFTMARGIN = 2;
+	private GuardActionBlockSubject mGuardActionBlock;
+	private String mGuardExpression;
+	private ArrayList<JComponent> mGuardActionExpressionLabels;
+	private int mGuardLabelIndex;
+	private int mActionLabelIndex;
+	private Component[] mPanelContent;
+	private JLabel mCollapsedContent;
+	private boolean isCollapsed;
+	private boolean mIsHighlighted = false;
+	private boolean mHasGuard = false;
+
+	public EditorGuardActionBlock(EditorEdge parentEdge, EditorSurface editorSurface) {
+		super(parentEdge, editorSurface);
+		// This is a GuardActionBlock
+		type = GUARDACTIONBLOCK;
+		Font defaultFont = new Font(null);
+		panel.removeAll();
+		isCollapsed = false;
+		mCollapsedContent = new JLabel(" G/A");
[…]

### 2006-03-02 src/net/sourceforge/waters/gui/EditorLabelBlockPopupMenu.java

@@ -1,0 +1,78 @@
+
+package net.sourceforge.waters.gui;
+
+import java.awt.*;
+import java.awt.event.*;
+import javax.swing.*;
+import org.supremica.util.VPopupMenu;
+import net.sourceforge.waters.model.module.SimpleNodeProxy;
+import net.sourceforge.waters.gui.command.Command;
+import net.sourceforge.waters.gui.command.DeleteEdgeCommand;
+
+/**
+ * Popup for editing attributes of a node.
+ */
+class EditorLabelBlockPopupMenu
+	extends VPopupMenu
+	implements ActionListener
+{
+	private EditorEdge edge;
+	private ControlledSurface parent;
+
+	private JMenuItem addGuardItem;
+	private JMenuItem addActionItem;
+	private EditorLabelGroup mBlock;	
+
+	public EditorLabelBlockPopupMenu(ControlledSurface parent, EditorLabelGroup block)
+	{
+		this.parent = parent;
+		this.mBlock = block;
+
+		init();
+	}
+
+	/**
+	 * Initialize the menu.
+	 */
+	private void init()
+	{
+		JMenuItem item;
+
+		item = new JMenuItem("Add guard");
+		item.addActionListener(this);
+		this.add(item);
+		addGuardItem = item;
+
+		item = new JMenuItem("Add action");
+		item.addActionListener(this);
+		this.add(item);
+		addActionItem = item;
+
+		// Disable "Add guard" if there is already a guard expression
+		if(((EditorEdge) mBlock.getParent()).getEditorGuardActionBlock() != null) {
+			if (((EditorEdge) mBlock.getParent()).getEditorGuardActionBlock().hasGuard())
+			{
+				addActionItem.setEnabled(false);
+				addActionItem.setToolTipText("Label is already in default position");
+			}
+		}
+	}
+
+	public void actionPerformed(ActionEvent e) 
+	{
+		if (e.getSource() == addGuardItem)
+		{
+			/*Command deleteEdge = new DeleteEdgeCommand(parent, edge);
+			parent.getEditorInterface().getUndoInterface().executeCommand(deleteEdge);
+			this.hide();*/
+			//TODO: add functionality
+		}
+
+		if (e.getSource() == addActionItem)
+		{
+			//TODO: add functionality
+		}
+
+		parent.repaint();
+	}
+}

### 2006-03-02 src/net/sourceforge/waters/gui/EditorLabelGroup.java

@@ -4,12 +4,13 @@
 //# PACKAGE: net.sourceforge.waters.gui
 //# CLASS:   EditorLabelGroup
 //###########################################################################
-//# $Id: EditorLabelGroup.java,v 1.24 2006-01-20 02:34:58 siw4 Exp $
+//# $Id: EditorLabelGroup.java,v 1.25 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 
 package net.sourceforge.waters.gui;
 
+import java.awt.BorderLayout;
 import java.awt.Color;
 import java.awt.Point;
 import java.awt.Rectangle;
@@ -49,6 +50,7 @@ public class EditorLabelGroup
 	implements ModelObserver
 {
 
+	protected boolean mSizeChanged;
 	public void removeFromSurface(EditorSurface e)
 	{
 		e.remove(panel);
@@ -409,11 +411,14 @@ public class EditorLabelGroup
 		shadowPanel.add(new JLabel(l.getText()));
 		*/
 	}
-
-	private void resizePanel()
+	protected void resizePanel() {
+		if(parent != null) {
+			parent.resizePanels();
+		}
+	}
+	protected int resizePanel(int width)
 	{
 		int height = 0;
-		int width = 0;
 
 		for (int i = 0; i < panel.getComponentCount(); i++)
 		{
@@ -421,26 +426,20 @@ public class EditorLabelGroup
 			c.setLocation(0, height);
 			c.setSize(c.getPreferredSize());
 
-			/*
-			c = (JComponent) shadowPanel.getComponent(i);
-			c.setLocation(0, height);
-			c.setSize(c.getPreferredSize());
-			*/
-
 			height += c.getHeight();
-
-			if (width < c.getWidth())
-			{
+			
+			if (width < c.getWidth()) {
 				width = c.getWidth();
 			}
 		}
 
-		panel.setSize(width, height);
+		panel.setSize(width + 2, height);
 		/*
 		shadowPanel.setSize(width, height);
 		*/
 
 		panel.repaint();
+		return width;
 	}
 
 	public int getX()
@@ -468,10 +467,9 @@ public class EditorLabelGroup
 		this(par.getSubject().getLabelBlock(), e);
 		parent = par;
 	}
-
+	
 	public EditorLabelGroup(LabelBlockSubject label, final EditorSurface surface)
 	{
-		
 		mUndo = surface.getEditorInterface().getUndoInterface();
 		// This is a labelgroup
 		type = LABELGROUP;
[…]

### 2006-03-02 src/net/sourceforge/waters/gui/EditorNewDialog.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.gui
 //# CLASS:   EditorNewDialog
 //###########################################################################
-//# $Id: EditorNewDialog.java,v 1.4 2005-11-03 01:24:15 robi Exp $
+//# $Id: EditorNewDialog.java,v 1.5 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 
@@ -124,6 +124,7 @@ public class EditorNewDialog
 			}
 			final LabelBlockSubject blocked = new LabelBlockSubject();
 			final GraphSubject graph = new GraphSubject(blocked);
+			
 			final SimpleComponentSubject comp =
 				new SimpleComponentSubject(ident, kind, graph);
 			//TODO: Make this create the component

### 2006-03-02 src/net/sourceforge/waters/gui/EditorObject.java

@@ -4,8 +4,8 @@
 //# PACKAGE: wnet.sourceforge.aters.gui
 //# CLASS:   EditorObject
 //###########################################################################
-//# $Id: EditorObject.java,v 1.23 2006-01-17 02:00:07 siw4 Exp $
-//# $Id: EditorObject.java,v 1.23 2006-01-17 02:00:07 siw4 Exp $
+//# $Id: EditorObject.java,v 1.24 2006-03-02 12:12:49 martin Exp $
+//# $Id: EditorObject.java,v 1.24 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 
@@ -33,6 +33,7 @@ public abstract class EditorObject
 	public static int NODEGROUP = 3;
 	public static int LABEL = 4;
 	public static int LABELGROUP = 5;
+	public static final int GUARDACTIONBLOCK = 6;		
 
     /** is not being draggedOver*/
     public static int NOTDRAG = 0;
@@ -57,7 +58,7 @@ public abstract class EditorObject
 	/** Thick line used for drawing shadows. */
 	public static final Stroke SHADOWSTROKE = new BasicStroke(10, BasicStroke.CAP_BUTT, BasicStroke.JOIN_ROUND);
 	/** Used as the basic stroke when printing - "hairline" width. */
-	public static final Stroke THINSTROKE = new BasicStroke(0.25f);		
+	public static final Stroke THINSTROKE = new BasicStroke(0.25f);
 	/** The default pen size. Is not {@code final} since it changes when printing. */
 	public static Stroke BASICSTROKE = SINGLESTROKE;
 

### 2006-03-02 src/net/sourceforge/waters/gui/EditorSurface.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.gui
 //# CLASS:   EditorSurface
 //###########################################################################
-//# $Id: EditorSurface.java,v 1.43 2006-01-23 23:47:46 siw4 Exp $
+//# $Id: EditorSurface.java,v 1.44 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 
@@ -32,6 +32,7 @@ import net.sourceforge.waters.subject.base.ModelChangeEvent;
 import net.sourceforge.waters.subject.base.ModelObserver;
 import net.sourceforge.waters.subject.module.*;
 import net.sourceforge.waters.xsd.module.AnchorPosition;
+import net.sourceforge.waters.xsd.module.GuardActionBlock;
 
 
 /**
@@ -51,6 +52,7 @@ public class EditorSurface
 	implements Printable,
 			   ModelObserver
 {
+	public static final int panelMarginAdjust = 2; //increase bounds for label & guardAction panels
 	protected boolean showGrid = true;
 	protected EditorWindowInterface root;
 	protected int gridSize = 16;
@@ -63,6 +65,7 @@ public class EditorSurface
 	protected ArrayList lines;
 	protected ArrayList shades;
 	protected ArrayList events;
+	protected ArrayList mGuardActionBlocks;
 	protected int dragStartX;
 	protected int dragStartY;
 	protected int dragNowX;
@@ -321,6 +324,23 @@ public class EditorSurface
 								20, 20);
 			}
 		}
+		
+		for (int i = 0; i < mGuardActionBlocks.size(); i++)
+		{
+
+			EditorGuardActionBlock GA = (EditorGuardActionBlock) mGuardActionBlocks.get(i);
+
+			GA.setPanelLocation();
+
+			if (GA.shadow && GA.isHighlighted())
+			{
+				Rectangle bounds = GA.getBounds();
+				g.setColor(GA.getShadowColor(isSelected(GA) || isSelected(GA.getParent())));
+				g.fillRoundRect((int) bounds.getX()-panelMarginAdjust, (int) bounds.getY()-panelMarginAdjust, 
[…]

### 2006-03-02 src/net/sourceforge/waters/gui/ModuleWindow.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.gui
 //# CLASS:   ModuleWindow
 //###########################################################################
-//# $Id: ModuleWindow.java,v 1.23 2006-02-20 22:20:21 robi Exp $
+//# $Id: ModuleWindow.java,v 1.24 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.gui;
@@ -31,6 +31,7 @@ import javax.swing.undo.UndoManager;
 import javax.swing.undo.UndoableEdit;
 import javax.xml.bind.JAXBException;
 
+
 import java.awt.datatransfer.Transferable;
 import java.awt.datatransfer.DataFlavor;
 import java.awt.datatransfer.UnsupportedFlavorException;
@@ -42,12 +43,10 @@ import java.awt.dnd.DragSourceAdapter;
 import java.awt.dnd.DragSourceDragEvent;
 import java.awt.dnd.DragSourceListener;
 import java.awt.dnd.InvalidDnDOperationException;
-
 import net.sourceforge.waters.gui.command.Command;
 import net.sourceforge.waters.gui.command.UndoableCommand;
 import net.sourceforge.waters.gui.command.UndoInterface;
 import net.sourceforge.waters.gui.observer.Observer;
-import net.sourceforge.waters.gui.observer.Subject;
 import net.sourceforge.waters.gui.observer.EditorChangedEvent;
 import net.sourceforge.waters.gui.observer.UndoRedoEvent;
 import net.sourceforge.waters.model.compiler.CompilerOperatorTable;
@@ -76,6 +75,10 @@ import net.sourceforge.waters.subject.module.ParameterBindingSubject;
 import net.sourceforge.waters.subject.module.ParameterSubject;
 import net.sourceforge.waters.subject.module.SimpleComponentSubject;
 import net.sourceforge.waters.subject.module.SimpleExpressionSubject;
+
+//EFA-----------------
+import net.sourceforge.waters.subject.module.*;
+
 import net.sourceforge.waters.xsd.base.ComponentKind;
 
 import net.sourceforge.waters.subject.module.IdentifierSubject;
@@ -220,6 +223,9 @@ public class ModuleWindow
 		int nbrOfAddedAutomata = supremica.addProject(supremicaProject);
 
 		System.err.println(nbrOfAddedAutomata);
+	//EFA question: 
[…]

### 2006-03-02 src/net/sourceforge/waters/gui/command/CreateEdgeCommand.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.gui.command
 //# CLASS:   CreateEdgeCommand
 //###########################################################################
-//# $Id: CreateEdgeCommand.java,v 1.7 2006-01-17 21:13:50 siw4 Exp $
+//# $Id: CreateEdgeCommand.java,v 1.8 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 
@@ -59,9 +59,8 @@ public class CreateEdgeCommand
 		final LabelBlockSubject labelBlock =
 			new LabelBlockSubject(empty, null);
 		final EdgeSubject edgeSubject =
-			new EdgeSubject(sourceSubject, targetSubject, labelBlock,
-							null, null, null);
-		mCreated = new EditorEdge(source, target, x, y, edgeSubject);
+			new EdgeSubject(sourceSubject, targetSubject, labelBlock);
+		mCreated = new EditorEdge(source, target, x, y, edgeSubject, surface);
     }
 
     /**

### 2006-03-02 src/net/sourceforge/waters/model/compiler/CompiledBooleanValue.java

@@ -1,0 +1,61 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.model.compiler
+//# CLASS:   CompiledBooleanValue
+//###########################################################################
+//# $Id$
+//###########################################################################
+
+package net.sourceforge.waters.model.compiler;
+
+import net.sourceforge.waters.model.expr.BooleanValue;
+
+
+class CompiledBooleanValue implements BooleanValue
+{
+
+  //#########################################################################
+  //# Constructors
+  CompiledBooleanValue(final boolean value)
+  {
+    mValue = value;
+  }
+
+
+  //#########################################################################
+  //# Interface net.sourceforge.waters.model.expr.BooleanValue
+  public boolean getValue()
+  {
+    return mValue;
+  }
+
+
+  //#########################################################################
+  //# Overrides for java.lang.Object
+  public String toString()
+  {
+    return Boolean.toString(mValue);
+  }
+
+  public int hashCode()
+  {
+    return mValue ? 1 : 0;
+  }
+
+  public boolean equals(final Object partner)
+  {
+    if (partner != null && getClass() == partner.getClass()) {
+      final CompiledBooleanValue value = (CompiledBooleanValue) partner;
+      return mValue == value.mValue;
+    } else {
+      return false;
+    }    
+  }
+
+
+  //#########################################################################
+  //# Data Members
+  private final boolean mValue;
+
+}

### 2006-03-02 src/net/sourceforge/waters/model/compiler/CompilerOperatorTable.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.compiler
 //# CLASS:   CompilerOperatorTable
 //###########################################################################
-//# $Id: CompilerOperatorTable.java,v 1.2 2005-11-03 01:24:15 robi Exp $
+//# $Id: CompilerOperatorTable.java,v 1.3 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.compiler;
@@ -45,6 +45,10 @@ public class CompilerOperatorTable extends AbstractOperatorTable {
     store(new BinaryEqualsOperator());
     store(new BinaryRangeOperator());
     store(new UnaryMinusOperator());
+
+    store(new BinaryIncreaseOperator());
+    store(new BinaryDecreaseOperator());
+    store(new BinaryAssignmentOperator());
   }
 
 
@@ -171,6 +175,105 @@ public class CompilerOperatorTable extends AbstractOperatorTable {
     int eval(final int lhs, final int rhs)
     {
       return lhs + rhs;
+    }
+
+  }
+  
+  private static class BinaryIncreaseOperator extends BinaryIntOperator
+  {
+
+    //#######################################################################
+    //# Interface net.sourceforge.waters.model.expr.Operator
+    public String getName()
+    {
+      return OPNAME_INCREASE;
+    }
+
+    public int getPriority()
+    {
+      return PRIORITY_ASSIGNMENT;
+    }
+
+
+    //#######################################################################
+    //# Interface net.sourceforge.waters.model.expr.BinaryOperator
+    public int getAssociativity()
+    {
+      return BinaryOperator.ASSOC_LEFT;
+    }
+
+
+    //#######################################################################
+    //# Overrides for Abstract Base Class BinaryIntOperator
+    int eval(final int lhs, final int rhs)
+    {
+      return lhs + rhs;
+    }
+
+  }
+  
+  private static class BinaryDecreaseOperator extends BinaryIntOperator
+  {
+
+    //#######################################################################
+    //# Interface net.sourceforge.waters.model.expr.Operator
+    public String getName()
+    {
+      return OPNAME_DECREASE;
+    }
+
+    public int getPriority()
+    {
+      return PRIORITY_ASSIGNMENT;
+    }
+
+
+    //#######################################################################
+    //# Interface net.sourceforge.waters.model.expr.BinaryOperator
[…]

### 2006-03-02 src/net/sourceforge/waters/model/compiler/GuardExpressionHandler.java

@@ -1,0 +1,1003 @@
+package net.sourceforge.waters.model.compiler;
+
+import java.lang.reflect.Array;
+import java.util.Arrays;
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.LinkedList;
+import java.util.List;
+import java.util.Map;
+import java.util.Set;
+
+import net.sourceforge.waters.model.base.ProxyVisitor;
+import net.sourceforge.waters.model.base.VisitorException;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.BinaryAndOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.BinaryBooleanOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.BinaryEqualsOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.BinaryGreaterEqualOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.BinaryGreaterThanOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.BinaryLessEqualOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.BinaryLessThanOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.BinaryNotEqualOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.BinaryOrOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.UnaryBooleanOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.UnaryNotOperator;
+import net.sourceforge.waters.model.compiler.GuardExpressionOperatorTable.IdentityOperator;
+import net.sourceforge.waters.model.expr.BinaryOperator;
+import net.sourceforge.waters.model.expr.BooleanOperator;
+import net.sourceforge.waters.model.expr.BooleanValue;
+import net.sourceforge.waters.model.expr.EvalException;
+import net.sourceforge.waters.model.expr.ExpressionParser;
+import net.sourceforge.waters.model.expr.IntValue;
+import net.sourceforge.waters.model.expr.Operator;
+import net.sourceforge.waters.model.expr.ParseException;
+import net.sourceforge.waters.model.expr.UnaryOperator;
+import net.sourceforge.waters.model.expr.Value;
[…]

### 2006-03-02 src/net/sourceforge/waters/model/compiler/GuardExpressionOperatorTable.java

@@ -1,0 +1,821 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.model.compiler
+//# CLASS:   CompilerOperatorTable
+//###########################################################################
+//# $Id$
+//###########################################################################
+
+package net.sourceforge.waters.model.compiler;
+
+import net.sourceforge.waters.model.expr.AbstractOperatorTable;
+import net.sourceforge.waters.model.expr.BinaryOperator;
+import net.sourceforge.waters.model.expr.BooleanOperator;
+import net.sourceforge.waters.model.expr.BooleanValue;
+import net.sourceforge.waters.model.expr.EvalException;
+import net.sourceforge.waters.model.expr.IntValue;
+import net.sourceforge.waters.model.expr.Operator;
+import net.sourceforge.waters.model.expr.OperatorTable;
+import net.sourceforge.waters.model.expr.UnaryOperator;
+import net.sourceforge.waters.model.expr.Value;
+
+
+/**
+ * <P>The guard+action expression operator table.</P>
+ *
+ * @author Martin Byröd
+ */
+
+public class GuardExpressionOperatorTable extends AbstractOperatorTable {
+
+  //#########################################################################
+  //# Static Class Methods
+  public static GuardExpressionOperatorTable getInstance()
+  {
+    return INSTANCE;
+  }
+
+
+  //#########################################################################
+  //# Constructors
+  private GuardExpressionOperatorTable()
+  {
+    super(16, OPCHAR_MIN, OPCHAR_MAX);
+    store(new BinaryNotEqualOperator());
+    store(new UnaryNotOperator());
+    store(new BinaryEqualsOperator());
+    store(new BinaryLessThanOperator());
+    store(new BinaryGreaterThanOperator());
+    store(new BinaryLessEqualOperator());
+    store(new BinaryGreaterEqualOperator());
+    store(new BinaryOrOperator());
+    store(new BinaryAndOperator());
+    store(new UnaryMinusOperator());
+  }
+
+
+//#########################################################################
+  //# Inner Class BinaryIntOperator
+  private static abstract class BinaryIntOperator implements BinaryOperator
+  {
+
+    //#######################################################################
+    //# Interface net.sourceforge.waters.model.expr.BinaryOperator
+    public int getLHSTypes()
+    {
[…]

### 2006-03-02 src/net/sourceforge/waters/model/compiler/ModuleCompiler.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.module
 //# CLASS:   ModuleCompiler
 //###########################################################################
-//# $Id: ModuleCompiler.java,v 1.4 2006-02-21 21:50:40 robi Exp $
+//# $Id: ModuleCompiler.java,v 1.5 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.compiler;
@@ -13,8 +13,11 @@ import java.io.IOException;
 import java.net.URI;
 import java.util.ArrayList;
 import java.util.Collection;
+import java.util.HashMap;
+import java.util.HashSet;
 import java.util.IdentityHashMap;
 import java.util.Iterator;
+import java.util.LinkedList;
 import java.util.List;
 import java.util.Map;
 import java.util.TreeMap;
@@ -34,8 +37,10 @@ import net.sourceforge.waters.model.des.TransitionProxy;
 import net.sourceforge.waters.model.expr.AtomValue;
 import net.sourceforge.waters.model.expr.BinaryOperator;
 import net.sourceforge.waters.model.expr.EvalException;
+import net.sourceforge.waters.model.expr.ExpressionParser;
 import net.sourceforge.waters.model.expr.IndexValue;
 import net.sourceforge.waters.model.expr.IntValue;
+import net.sourceforge.waters.model.expr.ParseException;
 import net.sourceforge.waters.model.expr.RangeValue;
 import net.sourceforge.waters.model.expr.UnaryOperator;
 import net.sourceforge.waters.model.expr.Value;
@@ -45,6 +50,7 @@ import net.sourceforge.waters.model.marshaller.WatersUnmarshalException;
 import net.sourceforge.waters.model.module.AbstractModuleProxyVisitor;
 import net.sourceforge.waters.model.module.AliasProxy;
 import net.sourceforge.waters.model.module.BinaryExpressionProxy;
+import net.sourceforge.waters.model.module.BooleanConstantProxy;
 import net.sourceforge.waters.model.module.EdgeProxy;
 import net.sourceforge.waters.model.module.EnumSetExpressionProxy;
 import net.sourceforge.waters.model.module.EventDeclProxy;
@@ -71,187 +77,182 @@ import net.sourceforge.waters.model.module.SimpleNodeProxy;
 import net.sourceforge.waters.model.module.SimpleParameterProxy;
[…]

### 2006-03-02 src/net/sourceforge/waters/model/expr/BooleanOperator.java

@@ -1,0 +1,25 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.model.expr
+//# CLASS:   BinaryOperator
+//###########################################################################
+//# $Id$
+//###########################################################################
+
+package net.sourceforge.waters.model.expr;
+
+
+/**
+ * <P>A boolean operator.</P>
+ *
+ * @see Operator
+ *
+ * @author Martin Byröd
+ */
+
+public interface BooleanOperator extends Operator {
+	//Returns the negated version of this operator
+	//i.e. "==" becomes "!=", "<" becomes ">=", etc.
+	public BooleanOperator getNegative();
+}

### 2006-03-02 src/net/sourceforge/waters/model/expr/BooleanValue.java

@@ -1,0 +1,16 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.model.expr
+//# CLASS:   IntValue
+//###########################################################################
+//# $Id$
+//###########################################################################
+
+package net.sourceforge.waters.model.expr;
+
+
+public interface BooleanValue extends IndexValue
+{
+  public boolean getValue();
+}

### 2006-03-02 src/net/sourceforge/waters/model/expr/IntValue.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.expr
 //# CLASS:   IntValue
 //###########################################################################
-//# $Id: IntValue.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: IntValue.java,v 1.3 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.expr;
@@ -12,7 +12,5 @@ package net.sourceforge.waters.model.expr;
 
 public interface IntValue extends IndexValue
 {
-
   public int getValue();
-
 }

### 2006-03-02 src/net/sourceforge/waters/model/expr/Operator.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.waters.model.expr
 //# CLASS:   Operator
 //###########################################################################
-//# $Id: Operator.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: Operator.java,v 1.3 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.expr;
@@ -31,14 +31,14 @@ public interface Operator {
 
   public int getPriority();
 
-
   //#########################################################################
   //# Class Constants
   public static final int TYPE_INT = 1;
   public static final int TYPE_ATOM = 2;
   public static final int TYPE_RANGE = 4;
   public static final int TYPE_NAME = 8;
+  public static final int TYPE_BOOLEAN = 16;
   public static final int TYPE_ANY =
-    TYPE_INT | TYPE_ATOM | TYPE_RANGE | TYPE_NAME;
+    TYPE_INT | TYPE_ATOM | TYPE_RANGE | TYPE_NAME | TYPE_BOOLEAN;
 
 }

### 2006-03-02 src/net/sourceforge/waters/model/marshaller/GuardActionBlockActionListHandler.java

@@ -1,0 +1,66 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters(EFA)
+//# PACKAGE: net.sourceforge.waters.model.marshaller
+//# CLASS:   GuardActionBlockActionListHandler
+//###########################################################################
+//# $Id: GuardActionBlockActionListHandler.java,v 1.2 2006-03-02 12:12:50 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.model.marshaller;
+
+import java.util.List;
+import javax.xml.bind.JAXBException;
+
+
+import net.sourceforge.waters.model.module.BinaryExpressionProxy;
+
+import net.sourceforge.waters.xsd.module.ActionListType;
+import net.sourceforge.waters.xsd.module.GuardActionBlockType;
+import net.sourceforge.waters.xsd.module.ObjectFactory;
+
+
+class GuardActionBlockActionListHandler
+  extends JAXBListHandler<GuardActionBlockType,ActionListType, BinaryExpressionProxy>
+{
+
+
+  //#########################################################################
+  //# Constructors
+	GuardActionBlockActionListHandler()
+  {
+    this(null);
+  }
+
+  GuardActionBlockActionListHandler(final ObjectFactory factory)
+  {
+    mFactory = factory;
+  }
+
+
+  //#########################################################################
+  //# Overrides for Abstract Base Class JAXBCheckedListHandler
+  ActionListType createListElement(final GuardActionBlockType container)
+    throws JAXBException
+  {
+    final ActionListType listelem = mFactory.createActionList();
+    container.setActionList(listelem);
+    return listelem;
+  }
+  
+  ActionListType getListElement(final GuardActionBlockType container)
+  {
+    return container.getActionList();
+  }
+
+  List getList(final ActionListType listelem)
+  {
+    return listelem.getList();
+  }
+
+
+  //#########################################################################
+  //# Data Members
+  private final ObjectFactory mFactory;
+
+}

### 2006-03-02 src/net/sourceforge/waters/model/marshaller/JAXBModuleExporter.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.marshaller
 //# CLASS:   JAXBModuleExporter
 //###########################################################################
-//# $Id: JAXBModuleExporter.java,v 1.4 2005-12-18 21:11:32 robi Exp $
+//# $Id: JAXBModuleExporter.java,v 1.5 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.marshaller;
@@ -26,6 +26,11 @@ import net.sourceforge.waters.model.base.VisitorException;
 import net.sourceforge.waters.model.expr.BinaryOperator;
 import net.sourceforge.waters.model.expr.UnaryOperator;
 import net.sourceforge.waters.model.module.AbstractModuleProxyVisitor;
+//EFA----------
+import net.sourceforge.waters.model.module.GuardActionBlockProxy;
+import net.sourceforge.waters.model.module.VariableProxy;
+import net.sourceforge.waters.model.module.BooleanConstantProxy;
+//-------------
 import net.sourceforge.waters.model.module.AliasProxy;
 import net.sourceforge.waters.model.module.BinaryExpressionProxy;
 import net.sourceforge.waters.model.module.BoxGeometryProxy;
@@ -114,8 +119,11 @@ import net.sourceforge.waters.xsd.module.SimpleNodeType;
 import net.sourceforge.waters.xsd.module.SimpleParameterType;
 import net.sourceforge.waters.xsd.module.SplineGeometryType;
 import net.sourceforge.waters.xsd.module.UnaryExpressionType;
-
-
+//EFA-----------------
+import net.sourceforge.waters.xsd.module.GuardActionBlockType;
+import net.sourceforge.waters.xsd.module.VariableType;
+import net.sourceforge.waters.xsd.module.BooleanConstantType;
+//-----------------------
 public class JAXBModuleExporter
   extends JAXBDocumentExporter<ModuleProxy,ModuleType>
   implements ModuleProxyVisitor
@@ -132,7 +140,44 @@ public class JAXBModuleExporter
 
   //#########################################################################
   //# Interface net.sourceforge.waters.model.module.ModuleProxyVisitor
-  public ElementType visitAliasProxy(final AliasProxy proxy)
+//EFA------------------- 
+  public GuardActionBlockType visitGuardActionBlockProxy(
+			final GuardActionBlockProxy proxy) throws VisitorException {
+		try {
+			final GuardActionBlockType element = mFactory
[…]

### 2006-03-02 src/net/sourceforge/waters/model/marshaller/JAXBModuleImporter.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.marshaller
 //# CLASS:   JAXBModuleImporter
 //###########################################################################
-//# $Id: JAXBModuleImporter.java,v 1.7 2006-02-23 01:52:34 robi Exp $
+//# $Id: JAXBModuleImporter.java,v 1.8 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.marshaller;
@@ -30,6 +30,11 @@ import net.sourceforge.waters.model.base.Proxy;
 import net.sourceforge.waters.model.expr.BinaryOperator;
 import net.sourceforge.waters.model.expr.OperatorTable;
 import net.sourceforge.waters.model.expr.UnaryOperator;
+//EFA-----------
+import net.sourceforge.waters.model.module.GuardActionBlockProxy;
+import net.sourceforge.waters.model.module.VariableProxy;
+import net.sourceforge.waters.model.module.BooleanConstantProxy;
+//-------------
 import net.sourceforge.waters.model.module.AliasProxy;
 import net.sourceforge.waters.model.module.BinaryExpressionProxy;
 import net.sourceforge.waters.model.module.BoxGeometryProxy;
@@ -117,12 +122,14 @@ import net.sourceforge.waters.xsd.module.SimpleParameterType;
 import net.sourceforge.waters.xsd.module.SplineGeometryType;
 import net.sourceforge.waters.xsd.module.SplineKind;
 import net.sourceforge.waters.xsd.module.UnaryExpressionType;
-
-
+//EFA----------------
+import net.sourceforge.waters.xsd.module.GuardActionBlockType;
+import net.sourceforge.waters.xsd.module.VariableType;
+import net.sourceforge.waters.xsd.module.BooleanConstantType;
+//-------------------
 public class JAXBModuleImporter
   extends JAXBDocumentImporter<ModuleProxy,ModuleType>
 {
-
   //#########################################################################
   //# Constructors
   public JAXBModuleImporter(final ModuleProxyFactory factory,
@@ -195,7 +202,41 @@ public class JAXBModuleImporter
       }
     };
     mHandlerMap.put
-      (net.sourceforge.waters.xsd.module.Edge.class, handler);
+    (net.sourceforge.waters.xsd.module.Edge.class, handler);
+    
+    //-------------------------- EFA
+    handler = new ImportHandler() {
[…]

### 2006-03-02 src/net/sourceforge/waters/model/marshaller/SimpleComponentVariableListHandler.java

@@ -1,0 +1,66 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters(EFA)
+//# PACKAGE: net.sourceforge.waters.model.marshaller
+//# CLASS:   SimpleComponentVariableListHandler
+//###########################################################################
+//# $Id: SimpleComponentVariableListHandler.java,v 1.2 2006-03-02 12:12:50 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.model.marshaller;
+
+import java.util.List;
+import javax.xml.bind.JAXBException;
+
+
+import net.sourceforge.waters.model.module.VariableProxy;
+
+import net.sourceforge.waters.xsd.module.VariableListType;
+import net.sourceforge.waters.xsd.module.SimpleComponentType;
+import net.sourceforge.waters.xsd.module.ObjectFactory;
+
+
+class SimpleComponentVariableListHandler
+  extends JAXBListHandler<SimpleComponentType,VariableListType,VariableProxy>
+{
+
+
+  //#########################################################################
+  //# Constructors
+	SimpleComponentVariableListHandler()
+  {
+    this(null);
+  }
+
+  SimpleComponentVariableListHandler(final ObjectFactory factory)
+  {
+    mFactory = factory;
+  }
+
+
+  //#########################################################################
+  //# Overrides for Abstract Base Class JAXBCheckedListHandler
+  VariableListType createListElement(final SimpleComponentType container)
+    throws JAXBException
+  {
+    final VariableListType listelem = mFactory.createVariableList();
+    container.setVariableList(listelem);
+    return listelem;
+  }
+  
+  VariableListType getListElement(final SimpleComponentType container)
+  {
+    return container.getVariableList();
+  }
+
+  List getList(final VariableListType listelem)
+  {
+    return listelem.getList();
+  }
+
+
+  //#########################################################################
+  //# Data Members
+  private final ObjectFactory mFactory;
+
+}

### 2006-03-02 src/net/sourceforge/waters/model/module/AbstractModuleProxyVisitor.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.module
 //# CLASS:   AbstractModuleProxyVisitor
 //###########################################################################
-//# $Id: AbstractModuleProxyVisitor.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: AbstractModuleProxyVisitor.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.module;
@@ -25,6 +25,12 @@ public class AbstractModuleProxyVisitor
   }
 
   public Object visitBinaryExpressionProxy(final BinaryExpressionProxy proxy)
+    throws VisitorException
+  {
+    return visitSimpleExpressionProxy(proxy);
+  }
+
+  public Object visitBooleanConstantProxy(final BooleanConstantProxy proxy)
     throws VisitorException
   {
     return visitSimpleExpressionProxy(proxy);
@@ -118,6 +124,12 @@ public class AbstractModuleProxyVisitor
     throws VisitorException
   {
     return visitNodeProxy(proxy);
+  }
+
+  public Object visitGuardActionBlockProxy(final GuardActionBlockProxy proxy)
+    throws VisitorException
+  {
+    return visitProxy(proxy);
   }
 
   public Object visitIdentifiedProxy(final IdentifiedProxy proxy)
@@ -250,6 +262,12 @@ public class AbstractModuleProxyVisitor
     throws VisitorException
   {
     return visitSimpleExpressionProxy(proxy);
+  }
+
+  public Object visitVariableProxy(final VariableProxy proxy)
+    throws VisitorException
+  {
+    return visitProxy(proxy);
   }
 
 }

### 2006-03-02 src/net/sourceforge/waters/model/module/BooleanConstantProxy.java

@@ -1,0 +1,29 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.model.module
+//# CLASS:   BooleanConstantProxy
+//###########################################################################
+//# $Id: BooleanConstantProxy.java,v 1.2 2006-03-02 12:12:49 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.model.module;
+
+
+/**
+ * An expression representing an integer constant.
+ *
+ * @author Markus Sköldstam
+ */
+
+
+public interface BooleanConstantProxy extends SimpleExpressionProxy {
+
+  //#########################################################################
+  //# Simple Access
+  /**
+   * Gets the boolean value of this constant.
+   */
+  public boolean isValue();
+
+}

### 2006-03-02 src/net/sourceforge/waters/model/module/EdgeProxy.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.module
 //# CLASS:   EdgeProxy
 //###########################################################################
-//# $Id: EdgeProxy.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: EdgeProxy.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.module;
@@ -46,8 +46,15 @@ public interface EdgeProxy extends Proxy {
    */
   // @ref
   public NodeProxy getTarget();
-
+  
+  
+  
   public LabelBlockProxy getLabelBlock();
+  
+  // @optional
+  public GuardActionBlockProxy getGuardActionBlock();
+  
+  
 
   /**
    * Gets the rendering information for this edge.

### 2006-03-02 src/net/sourceforge/waters/model/module/GuardActionBlockProxy.java

@@ -1,0 +1,33 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.model.module
+//# CLASS:   GuardActionBlockProxy
+//###########################################################################
+//# $Id: GuardActionBlockProxy.java,v 1.2 2006-03-02 12:12:49 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.model.module;
+import java.util.List;
+
+import net.sourceforge.waters.model.base.Proxy;
+
+/**
+ * Guard and Actions with associated geometry information. A GuardActionBlock 
+ * is found is on the edges of a graph. 
+ *
+ * @author Markus Sköldstam
+ */
+
+public interface GuardActionBlockProxy extends Proxy {
+
+  //#########################################################################
+  //# Getters and Setters
+  
+  // @optional 
+  public String getGuard();
+   
+  public List<BinaryExpressionProxy> getActionList();
+  
+  public LabelGeometryProxy getGeometry();
+}

### 2006-03-02 src/net/sourceforge/waters/model/module/ModuleProxyFactory.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.module
 //# CLASS:   ModuleProxyFactory
 //###########################################################################
-//# $Id: ModuleProxyFactory.java,v 1.5 2006-02-22 03:35:07 robi Exp $
+//# $Id: ModuleProxyFactory.java,v 1.6 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.module;
@@ -48,6 +48,13 @@ public interface ModuleProxyFactory
        SimpleExpressionProxy right);
 
   /**
+   * Creates a new boolean constant.
+   * @param value The boolean value of the new boolean constant.
+   */
+  public BooleanConstantProxy createBooleanConstantProxy
+      (boolean value);
+
+  /**
    * Creates a new box geometry.
    * @param rectangle The rectangle of the new box geometry.
    */
@@ -73,6 +80,7 @@ public interface ModuleProxyFactory
    * @param source The source node of the new edge.
    * @param target The target node of the new edge.
    * @param labelBlock The label block of the new edge.
+   * @param guardActionBlock The guard action block of the new edge, or <CODE>null</CODE>.
    * @param geometry The rendering information of the new edge, or <CODE>null</CODE>.
    * @param startPoint The rendering information for the start point of the new edge, or <CODE>null</CODE>.
    * @param endPoint The rendering information for the end point of the new edge, or <CODE>null</CODE>.
@@ -81,6 +89,7 @@ public interface ModuleProxyFactory
       (NodeProxy source,
        NodeProxy target,
        LabelBlockProxy labelBlock,
+       GuardActionBlockProxy guardActionBlock,
        SplineGeometryProxy geometry,
        PointGeometryProxy startPoint,
        PointGeometryProxy endPoint);
@@ -88,6 +97,7 @@ public interface ModuleProxyFactory
   /**
    * Creates a new edge using default values.
    * This method creates an edge with
+   * the guard action block set to <CODE>null</CODE>,
    * the rendering information set to <CODE>null</CODE>,
    * the rendering information for the start point set to <CODE>null</CODE>, and
    * the rendering information for the end point set to <CODE>null</CODE>.
@@ -289,6 +299,26 @@ public interface ModuleProxyFactory
        EventListExpressionProxy propositions);
 
   /**
+   * Creates a new guard action block.
+   * @param guard The guard of the new guard action block, or <CODE>null</CODE>.
[…]

### 2006-03-02 src/net/sourceforge/waters/model/module/ModuleProxyVisitor.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.module
 //# CLASS:   ModuleProxyVisitor
 //###########################################################################
-//# $Id: ModuleProxyVisitor.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: ModuleProxyVisitor.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.module;
@@ -21,6 +21,9 @@ public interface ModuleProxyVisitor
     throws VisitorException;
 
   public Object visitBinaryExpressionProxy(BinaryExpressionProxy proxy)
+    throws VisitorException;
+
+  public Object visitBooleanConstantProxy(BooleanConstantProxy proxy)
     throws VisitorException;
 
   public Object visitBoxGeometryProxy(BoxGeometryProxy proxy)
@@ -66,6 +69,9 @@ public interface ModuleProxyVisitor
     throws VisitorException;
 
   public Object visitGroupNodeProxy(GroupNodeProxy proxy)
+    throws VisitorException;
+
+  public Object visitGuardActionBlockProxy(GuardActionBlockProxy proxy)
     throws VisitorException;
 
   public Object visitIdentifiedProxy(IdentifiedProxy proxy)
@@ -132,6 +138,9 @@ public interface ModuleProxyVisitor
     throws VisitorException;
 
   public Object visitUnaryExpressionProxy(UnaryExpressionProxy proxy)
+    throws VisitorException;
+
+  public Object visitVariableProxy(VariableProxy proxy)
     throws VisitorException;
 
 }

### 2006-03-02 src/net/sourceforge/waters/model/module/SimpleComponentProxy.java

@@ -4,13 +4,13 @@
 //# PACKAGE: net.sourceforge.waters.model.decl
 //# CLASS:   SimpleComponentProxy
 //###########################################################################
-//# $Id: SimpleComponentProxy.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: SimpleComponentProxy.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.module;
 
 import net.sourceforge.waters.xsd.base.ComponentKind;
-
+import java.util.List;
 
 /**
  * <P>A component representing a single finite-state machine.</P>
@@ -44,5 +44,7 @@ public interface SimpleComponentProxy extends ComponentProxy {
   public ComponentKind getKind();
 
   public GraphProxy getGraph();
+  
+  public List<VariableProxy> getVariables();
 
 }

### 2006-03-02 src/net/sourceforge/waters/model/module/VariableProxy.java

@@ -1,0 +1,52 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.model.module
+//# CLASS:   VariableProxy
+//###########################################################################
+//# $Id: VariableProxy.java,v 1.2 2006-03-02 12:12:49 martin Exp $
+//###########################################################################
+
+
+package net.sourceforge.waters.model.module;
+
+import net.sourceforge.waters.model.base.Proxy;
+
+
+/**
+ * <P>A variable is used to define the guards
+ *  and action on transitions.</P>
+ *
+ * 
+ *
+ * @author Markus S.
+ */
+
+public interface VariableProxy extends Proxy{
+
+  //#########################################################################
+  //# Getters
+  
+  /**
+   * Gets the Name of this variable.
+   */
+  
+ 
+  public String getName(); 
+  
+  /**
+   * Gets the type of this variable. Possible types:
+   * EnumSetExpressionProxy (enum), BinaryExpression (integer) 
+   * or SimpleIdentifierProxy (boolean).
+   */
+ 
+  public SimpleExpressionProxy getType(); 
+  
+  /**
+   * Gets the initial value of this variable.
+   */
+ 
+  public SimpleExpressionProxy getInitialValue();
+
+  
+}

### 2006-03-02 src/net/sourceforge/waters/model/printer/ModuleProxyPrinter.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.printer
 //# CLASS:   ModuleProxyPrinter
 //###########################################################################
-//# $Id: ModuleProxyPrinter.java,v 1.3 2005-11-10 21:54:42 robi Exp $
+//# $Id: ModuleProxyPrinter.java,v 1.4 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.printer;
@@ -60,6 +60,10 @@ import net.sourceforge.waters.model.module.SimpleParameterProxy;
 import net.sourceforge.waters.model.module.SplineGeometryProxy;
 import net.sourceforge.waters.model.module.UnaryExpressionProxy;
 
+import net.sourceforge.waters.model.module.VariableProxy;
+import net.sourceforge.waters.model.module.GuardActionBlockProxy;
+import net.sourceforge.waters.model.module.BooleanConstantProxy;
+
 import net.sourceforge.waters.xsd.base.ComponentKind;
 import net.sourceforge.waters.xsd.base.EventKind;
 
@@ -89,6 +93,33 @@ public class ModuleProxyPrinter
 
   //#########################################################################
   //# Interface net.sourceforge.waters.model.module.ModuleProxyVisitor
+  
+//EFA-------------------
+  public Object visitGuardActionBlockProxy
+  (final GuardActionBlockProxy proxy)
+ throws VisitorException
+  {
+	  //Temporary changes
+	     return null;
+  }
+  
+  public Object visitVariableProxy
+  (final VariableProxy proxy)
+ throws VisitorException
+  {
+	  //Temporary changes
+	    return null;
+  }
+  public Object visitBooleanConstantProxy 
+  (final BooleanConstantProxy proxy)
+  throws VisitorException
+	  {
+	    print(Boolean.toString(proxy.isValue()));
+	    return null;
+   }
+  
+  //------------------
+  
   public Object visitAliasProxy
       (final AliasProxy proxy)
     throws VisitorException
@@ -544,6 +575,7 @@ public class ModuleProxyPrinter
         print('(');
       }
       mAssocBraces = false;
+      print(operator.getName());
       final SimpleExpressionProxy subTerm = proxy.getSubTerm();
       subTerm.acceptVisitor(this);
       if (needBraces) {

### 2006-03-02 src/net/sourceforge/waters/plain/module/AliasElement.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.plain.module
 //# CLASS:   AliasElement
 //###########################################################################
-//# $Id: AliasElement.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: AliasElement.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.plain.module;

### 2006-03-02 src/net/sourceforge/waters/plain/module/BinaryExpressionElement.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.plain.module
 //# CLASS:   BinaryExpressionElement
 //###########################################################################
-//# $Id: BinaryExpressionElement.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: BinaryExpressionElement.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.plain.module;

### 2006-03-02 src/net/sourceforge/waters/plain/module/BooleanConstantElement.java

@@ -1,0 +1,85 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.plain.module
+//# CLASS:   BooleanConstantElement
+//###########################################################################
+//# $Id: BooleanConstantElement.java,v 1.2 2006-03-02 12:12:49 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.plain.module;
+
+import net.sourceforge.waters.model.base.ProxyVisitor;
+import net.sourceforge.waters.model.base.VisitorException;
+import net.sourceforge.waters.model.module.BooleanConstantProxy;
+import net.sourceforge.waters.model.module.ModuleProxyVisitor;
+
+
+/**
+ * An immutable implementation of the {@link BooleanConstantProxy} interface.
+ *
+ * @author Robi Malik
+ */
+
+public final class BooleanConstantElement
+  extends SimpleExpressionElement
+  implements BooleanConstantProxy
+{
+
+  //#########################################################################
+  //# Constructors
+  /**
+   * Creates a new boolean constant.
+   * @param value The boolean value of the new boolean constant.
+   */
+  public BooleanConstantElement(final boolean value)
+  {
+    mIsValue = value;
+  }
+
+
+  //#########################################################################
+  //# Cloning
+  public BooleanConstantElement clone()
+  {
+    return (BooleanConstantElement) super.clone();
+  }
+
+
+  //#########################################################################
+  //# Equality
+  public boolean equals(final Object partner)
+  {
+    if (super.equals(partner)) {
+      final BooleanConstantElement downcast = (BooleanConstantElement) partner;
+      return
+        (mIsValue == downcast.mIsValue);
+    } else {
+      return false;
+    }
+  }
+
+
+  //#########################################################################
+  //# Interface net.sourceforge.waters.model.base.Proxy
+  public Object acceptVisitor(final ProxyVisitor visitor)
+    throws VisitorException
+  {
+    final ModuleProxyVisitor downcast = (ModuleProxyVisitor) visitor;
+    return downcast.visitBooleanConstantProxy(this);
+  }
+
+
+  //#########################################################################
+  //# Interface net.sourceforge.waters.model.module.BooleanConstantProxy
+  public boolean isValue()
+  {
+    return mIsValue;
+  }
+
+
+  //#########################################################################
+  //# Data Members
+  private final boolean mIsValue;
+
+}

### 2006-03-02 src/net/sourceforge/waters/plain/module/ComponentElement.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.plain.module
 //# CLASS:   ComponentElement
 //###########################################################################
-//# $Id: ComponentElement.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: ComponentElement.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.plain.module;

### 2006-03-02 src/net/sourceforge/waters/plain/module/EdgeElement.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.plain.module
 //# CLASS:   EdgeElement
 //###########################################################################
-//# $Id: EdgeElement.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: EdgeElement.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.plain.module;
@@ -13,6 +13,7 @@ import net.sourceforge.waters.model.base.Geometry;
 import net.sourceforge.waters.model.base.ProxyVisitor;
 import net.sourceforge.waters.model.base.VisitorException;
 import net.sourceforge.waters.model.module.EdgeProxy;
+import net.sourceforge.waters.model.module.GuardActionBlockProxy;
 import net.sourceforge.waters.model.module.LabelBlockProxy;
 import net.sourceforge.waters.model.module.ModuleProxyVisitor;
 import net.sourceforge.waters.model.module.NodeProxy;
@@ -39,6 +40,7 @@ public final class EdgeElement
    * @param source The source node of the new edge.
    * @param target The target node of the new edge.
    * @param labelBlock The label block of the new edge.
+   * @param guardActionBlock The guard action block of the new edge, or <CODE>null</CODE>.
    * @param geometry The rendering information of the new edge, or <CODE>null</CODE>.
    * @param startPoint The rendering information for the start point of the new edge, or <CODE>null</CODE>.
    * @param endPoint The rendering information for the end point of the new edge, or <CODE>null</CODE>.
@@ -46,6 +48,7 @@ public final class EdgeElement
   public EdgeElement(final NodeProxy source,
                      final NodeProxy target,
                      final LabelBlockProxy labelBlock,
+                     final GuardActionBlockProxy guardActionBlock,
                      final SplineGeometryProxy geometry,
                      final PointGeometryProxy startPoint,
                      final PointGeometryProxy endPoint)
@@ -53,6 +56,7 @@ public final class EdgeElement
     mSource = source;
     mTarget = target;
     mLabelBlock = labelBlock;
+    mGuardActionBlock = guardActionBlock;
     mGeometry = geometry;
     mStartPoint = startPoint;
     mEndPoint = endPoint;
@@ -61,6 +65,7 @@ public final class EdgeElement
   /**
[…]

### 2006-03-02 src/net/sourceforge/waters/plain/module/GuardActionBlockElement.java

@@ -1,0 +1,153 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.plain.module
+//# CLASS:   GuardActionBlockElement
+//###########################################################################
+//# $Id: GuardActionBlockElement.java,v 1.2 2006-03-02 12:12:49 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.plain.module;
+
+import java.util.ArrayList;
+import java.util.Collection;
+import java.util.Collections;
+import java.util.List;
+
+import net.sourceforge.waters.model.base.Geometry;
+import net.sourceforge.waters.model.base.ProxyVisitor;
+import net.sourceforge.waters.model.base.VisitorException;
+import net.sourceforge.waters.model.module.BinaryExpressionProxy;
+import net.sourceforge.waters.model.module.GuardActionBlockProxy;
+import net.sourceforge.waters.model.module.LabelGeometryProxy;
+import net.sourceforge.waters.model.module.ModuleProxyVisitor;
+import net.sourceforge.waters.plain.base.Element;
+
+
+/**
+ * An immutable implementation of the {@link GuardActionBlockProxy} interface.
+ *
+ * @author Robi Malik
+ */
+
+public final class GuardActionBlockElement
+  extends Element
+  implements GuardActionBlockProxy
+{
+
+  //#########################################################################
+  //# Constructors
+  /**
+   * Creates a new guard action block.
+   * @param guard The guard of the new guard action block, or <CODE>null</CODE>.
+   * @param actionList The action list of the new guard action block.
+   * @param geometry The geometry of the new guard action block, or <CODE>null</CODE>.
+   */
+  public GuardActionBlockElement(final String guard,
+                                 final Collection<? extends BinaryExpressionProxy> actionList,
+                                 final LabelGeometryProxy geometry)
+  {
+    mGuard = guard;
+    final List<BinaryExpressionProxy> actionListModifiable =
+      new ArrayList<BinaryExpressionProxy>(actionList);
+    mActionList =
+      Collections.unmodifiableList(actionListModifiable);
+    mGeometry = geometry;
+  }
+
+  /**
+   * Creates a new guard action block using default values.
+   * This constructor creates a guard action block with
+   * the guard set to <CODE>null</CODE>,
+   * an empty action list, and
[…]

### 2006-03-02 src/net/sourceforge/waters/plain/module/ModuleElementFactory.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.plain.module
 //# CLASS:   ModuleElementFactory
 //###########################################################################
-//# $Id: ModuleElementFactory.java,v 1.5 2006-02-22 03:35:07 robi Exp $
+//# $Id: ModuleElementFactory.java,v 1.6 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.plain.module;
@@ -19,6 +19,7 @@ import net.sourceforge.waters.model.base.Proxy;
 import net.sourceforge.waters.model.expr.BinaryOperator;
 import net.sourceforge.waters.model.expr.UnaryOperator;
 import net.sourceforge.waters.model.module.AliasProxy;
+import net.sourceforge.waters.model.module.BinaryExpressionProxy;
 import net.sourceforge.waters.model.module.BoxGeometryProxy;
 import net.sourceforge.waters.model.module.ColorGeometryProxy;
 import net.sourceforge.waters.model.module.EdgeProxy;
@@ -26,6 +27,7 @@ import net.sourceforge.waters.model.module.EventDeclProxy;
 import net.sourceforge.waters.model.module.EventListExpressionProxy;
 import net.sourceforge.waters.model.module.ExpressionProxy;
 import net.sourceforge.waters.model.module.GraphProxy;
+import net.sourceforge.waters.model.module.GuardActionBlockProxy;
 import net.sourceforge.waters.model.module.IdentifierProxy;
 import net.sourceforge.waters.model.module.LabelBlockProxy;
 import net.sourceforge.waters.model.module.LabelGeometryProxy;
@@ -37,6 +39,7 @@ import net.sourceforge.waters.model.module.PointGeometryProxy;
 import net.sourceforge.waters.model.module.SimpleExpressionProxy;
 import net.sourceforge.waters.model.module.SimpleIdentifierProxy;
 import net.sourceforge.waters.model.module.SplineGeometryProxy;
+import net.sourceforge.waters.model.module.VariableProxy;
 
 import net.sourceforge.waters.xsd.base.ComponentKind;
 import net.sourceforge.waters.xsd.base.EventKind;
@@ -88,6 +91,16 @@ public class ModuleElementFactory
   }
 
   /**
+   * Creates a new boolean constant.
+   * @param value The boolean value of the new boolean constant.
+   */
+  public BooleanConstantElement createBooleanConstantProxy
+      (final boolean value)
+  {
[…]

### 2006-03-02 src/net/sourceforge/waters/plain/module/SimpleComponentElement.java

@@ -4,10 +4,15 @@
 //# PACKAGE: net.sourceforge.waters.plain.module
 //# CLASS:   SimpleComponentElement
 //###########################################################################
-//# $Id: SimpleComponentElement.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: SimpleComponentElement.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.plain.module;
+
+import java.util.ArrayList;
+import java.util.Collection;
+import java.util.Collections;
+import java.util.List;
 
 import net.sourceforge.waters.model.base.ProxyVisitor;
 import net.sourceforge.waters.model.base.VisitorException;
@@ -15,6 +20,7 @@ import net.sourceforge.waters.model.module.GraphProxy;
 import net.sourceforge.waters.model.module.IdentifierProxy;
 import net.sourceforge.waters.model.module.ModuleProxyVisitor;
 import net.sourceforge.waters.model.module.SimpleComponentProxy;
+import net.sourceforge.waters.model.module.VariableProxy;
 
 import net.sourceforge.waters.xsd.base.ComponentKind;
 
@@ -37,14 +43,38 @@ public final class SimpleComponentElement
    * @param identifier The identifier defining the name of the new simple component.
    * @param kind The kind of the new simple component.
    * @param graph The graph of the new simple component.
+   * @param variables The variables of the new simple component.
+   */
+  public SimpleComponentElement(final IdentifierProxy identifier,
+                                final ComponentKind kind,
+                                final GraphProxy graph,
+                                final Collection<? extends VariableProxy> variables)
+  {
+    super(identifier);
+    mKind = kind;
+    mGraph = graph;
+    final List<VariableProxy> variablesModifiable =
+      new ArrayList<VariableProxy>(variables);
+    mVariables =
+      Collections.unmodifiableList(variablesModifiable);
+  }
+
+  /**
+   * Creates a new simple component using default values.
+   * This constructor creates a simple component with
+   * an empty variables.
+   * @param identifier The identifier defining the name of the new simple component.
+   * @param kind The kind of the new simple component.
+   * @param graph The graph of the new simple component.
    */
   public SimpleComponentElement(final IdentifierProxy identifier,
                                 final ComponentKind kind,
                                 final GraphProxy graph)
   {
-    super(identifier);
[…]

### 2006-03-02 src/net/sourceforge/waters/plain/module/UnaryExpressionElement.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.plain.module
 //# CLASS:   UnaryExpressionElement
 //###########################################################################
-//# $Id: UnaryExpressionElement.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: UnaryExpressionElement.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.plain.module;

### 2006-03-02 src/net/sourceforge/waters/plain/module/VariableElement.java

@@ -1,0 +1,107 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.plain.module
+//# CLASS:   VariableElement
+//###########################################################################
+//# $Id: VariableElement.java,v 1.2 2006-03-02 12:12:49 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.plain.module;
+
+import net.sourceforge.waters.model.base.ProxyVisitor;
+import net.sourceforge.waters.model.base.VisitorException;
+import net.sourceforge.waters.model.module.ModuleProxyVisitor;
+import net.sourceforge.waters.model.module.SimpleExpressionProxy;
+import net.sourceforge.waters.model.module.VariableProxy;
+import net.sourceforge.waters.plain.base.Element;
+
+
+/**
+ * An immutable implementation of the {@link VariableProxy} interface.
+ *
+ * @author Robi Malik
+ */
+
+public final class VariableElement
+  extends Element
+  implements VariableProxy
+{
+
+  //#########################################################################
+  //# Constructors
+  /**
+   * Creates a new variable.
+   * @param name The name of the new variable.
+   * @param type The type of the new variable.
+   * @param initialValue The initial value of the new variable.
+   */
+  public VariableElement(final String name,
+                         final SimpleExpressionProxy type,
+                         final SimpleExpressionProxy initialValue)
+  {
+    mName = name;
+    mType = type;
+    mInitialValue = initialValue;
+  }
+
+
+  //#########################################################################
+  //# Cloning
+  public VariableElement clone()
+  {
+    return (VariableElement) super.clone();
+  }
+
+
+  //#########################################################################
+  //# Equality
+  public boolean equals(final Object partner)
+  {
+    if (super.equals(partner)) {
+      final VariableElement downcast = (VariableElement) partner;
+      return
+        mName.equals(downcast.mName) &&
+        mType.equals(downcast.mType) &&
+        mInitialValue.equals(downcast.mInitialValue);
+    } else {
+      return false;
+    }
+  }
+
+
+  //#########################################################################
+  //# Interface net.sourceforge.waters.model.base.Proxy
+  public Object acceptVisitor(final ProxyVisitor visitor)
+    throws VisitorException
+  {
+    final ModuleProxyVisitor downcast = (ModuleProxyVisitor) visitor;
[…]

### 2006-03-02 src/net/sourceforge/waters/subject/module/BooleanConstantSubject.java

@@ -1,0 +1,104 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.subject.module
+//# CLASS:   BooleanConstantSubject
+//###########################################################################
+//# $Id: BooleanConstantSubject.java,v 1.2 2006-03-02 12:12:49 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.subject.module;
+
+import net.sourceforge.waters.model.base.ProxyVisitor;
+import net.sourceforge.waters.model.base.VisitorException;
+import net.sourceforge.waters.model.module.BooleanConstantProxy;
+import net.sourceforge.waters.model.module.ModuleProxyVisitor;
+import net.sourceforge.waters.subject.base.ModelChangeEvent;
+
+
+/**
+ * The subject implementation of the {@link BooleanConstantProxy} interface.
+ *
+ * @author Robi Malik
+ */
+
+public final class BooleanConstantSubject
+  extends SimpleExpressionSubject
+  implements BooleanConstantProxy
+{
+
+  //#########################################################################
+  //# Constructors
+  /**
+   * Creates a new boolean constant.
+   * @param value The boolean value of the new boolean constant.
+   */
+  public BooleanConstantSubject(final boolean value)
+  {
+    mIsValue = value;
+  }
+
+
+  //#########################################################################
+  //# Cloning
+  public BooleanConstantSubject clone()
+  {
+    final BooleanConstantSubject cloned = (BooleanConstantSubject) super.clone();
+    return cloned;
+  }
+
+
+  //#########################################################################
+  //# Equality
+  public boolean equals(final Object partner)
+  {
+    if (super.equals(partner)) {
+      final BooleanConstantSubject downcast = (BooleanConstantSubject) partner;
+      return
+        (mIsValue == downcast.mIsValue);
+    } else {
+      return false;
+    }
+  }
+
+
+  //#########################################################################
+  //# Interface net.sourceforge.waters.model.base.Proxy
+  public Object acceptVisitor(final ProxyVisitor visitor)
+    throws VisitorException
+  {
+    final ModuleProxyVisitor downcast = (ModuleProxyVisitor) visitor;
+    return downcast.visitBooleanConstantProxy(this);
+  }
+
+
+  //#########################################################################
+  //# Interface net.sourceforge.waters.model.module.BooleanConstantProxy
[…]

### 2006-03-02 src/net/sourceforge/waters/subject/module/BoxGeometrySubject.java

@@ -4,7 +4,8 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   BoxGeometrySubject
 //###########################################################################
-//# $Id: BoxGeometrySubject.java,v 1.4 2006-02-16 04:06:18 robi Exp $
+
+//# $Id: BoxGeometrySubject.java,v 1.5 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;

### 2006-03-02 src/net/sourceforge/waters/subject/module/ComponentSubject.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   ComponentSubject
 //###########################################################################
-//# $Id: ComponentSubject.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: ComponentSubject.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;

### 2006-03-02 src/net/sourceforge/waters/subject/module/EdgeSubject.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   EdgeSubject
 //###########################################################################
-//# $Id: EdgeSubject.java,v 1.4 2006-02-16 04:06:18 robi Exp $
+//# $Id: EdgeSubject.java,v 1.5 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;
@@ -14,6 +14,7 @@ import net.sourceforge.waters.model.base.IndexedSet;
 import net.sourceforge.waters.model.base.ProxyVisitor;
 import net.sourceforge.waters.model.base.VisitorException;
 import net.sourceforge.waters.model.module.EdgeProxy;
+import net.sourceforge.waters.model.module.GuardActionBlockProxy;
 import net.sourceforge.waters.model.module.LabelBlockProxy;
 import net.sourceforge.waters.model.module.ModuleProxyVisitor;
 import net.sourceforge.waters.model.module.NodeProxy;
@@ -41,6 +42,7 @@ public final class EdgeSubject
    * @param source The source node of the new edge.
    * @param target The target node of the new edge.
    * @param labelBlock The label block of the new edge.
+   * @param guardActionBlock The guard action block of the new edge, or <CODE>null</CODE>.
    * @param geometry The rendering information of the new edge, or <CODE>null</CODE>.
    * @param startPoint The rendering information for the start point of the new edge, or <CODE>null</CODE>.
    * @param endPoint The rendering information for the end point of the new edge, or <CODE>null</CODE>.
@@ -48,6 +50,7 @@ public final class EdgeSubject
   public EdgeSubject(final NodeProxy source,
                      final NodeProxy target,
                      final LabelBlockProxy labelBlock,
+                     final GuardActionBlockProxy guardActionBlock,
                      final SplineGeometryProxy geometry,
                      final PointGeometryProxy startPoint,
                      final PointGeometryProxy endPoint)
@@ -56,6 +59,10 @@ public final class EdgeSubject
     mTarget = (NodeSubject) target;
     mLabelBlock = (LabelBlockSubject) labelBlock;
     mLabelBlock.setParent(this);
[…]

### 2006-03-02 src/net/sourceforge/waters/subject/module/ExpressionSubject.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   ExpressionSubject
 //###########################################################################
-//# $Id: ExpressionSubject.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: ExpressionSubject.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;

### 2006-03-02 src/net/sourceforge/waters/subject/module/GuardActionBlockSubject.java

@@ -1,0 +1,215 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.subject.module
+//# CLASS:   GuardActionBlockSubject
+//###########################################################################
+//# $Id: GuardActionBlockSubject.java,v 1.2 2006-03-02 12:12:49 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.subject.module;
+
+import java.util.Collection;
+import java.util.Collections;
+import java.util.List;
+
+import net.sourceforge.waters.model.base.Geometry;
+import net.sourceforge.waters.model.base.ProxyVisitor;
+import net.sourceforge.waters.model.base.VisitorException;
+import net.sourceforge.waters.model.module.BinaryExpressionProxy;
+import net.sourceforge.waters.model.module.GuardActionBlockProxy;
+import net.sourceforge.waters.model.module.LabelGeometryProxy;
+import net.sourceforge.waters.model.module.ModuleProxyVisitor;
+import net.sourceforge.waters.model.module.SimpleExpressionProxy;
+import net.sourceforge.waters.model.unchecked.Casting;
+import net.sourceforge.waters.subject.base.ArrayListSubject;
+import net.sourceforge.waters.subject.base.ListSubject;
+import net.sourceforge.waters.subject.base.ModelChangeEvent;
+import net.sourceforge.waters.subject.base.MutableSubject;
+
+
+/**
+ * The subject implementation of the {@link GuardActionBlockProxy} interface.
+ *
+ * @author Robi Malik
+ */
+
+public final class GuardActionBlockSubject
+  extends MutableSubject
+  implements GuardActionBlockProxy
+{
+
+  //#########################################################################
+  //# Constructors
+  /**
+   * Creates a new guard action block.
+   * @param guard The guard of the new guard action block, or <CODE>null</CODE>.
+   * @param actionList The action list of the new guard action block.
+   * @param geometry The geometry of the new guard action block, or <CODE>null</CODE>.
+   */
+  public GuardActionBlockSubject(final String guard,
+                                 final Collection<? extends BinaryExpressionProxy> actionList,
+                                 final LabelGeometryProxy geometry)
+  {
+    mGuard = guard;
+    mActionList = new ArrayListSubject<BinaryExpressionSubject>
[…]

### 2006-03-02 src/net/sourceforge/waters/subject/module/IntParameterSubject.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   IntParameterSubject
 //###########################################################################
-//# $Id: IntParameterSubject.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: IntParameterSubject.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;

### 2006-03-02 src/net/sourceforge/waters/subject/module/ModuleSubjectFactory.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   ModuleSubjectFactory
 //###########################################################################
-//# $Id: ModuleSubjectFactory.java,v 1.5 2006-02-22 03:35:07 robi Exp $
+//# $Id: ModuleSubjectFactory.java,v 1.6 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;
@@ -19,6 +19,7 @@ import net.sourceforge.waters.model.base.Proxy;
 import net.sourceforge.waters.model.expr.BinaryOperator;
 import net.sourceforge.waters.model.expr.UnaryOperator;
 import net.sourceforge.waters.model.module.AliasProxy;
+import net.sourceforge.waters.model.module.BinaryExpressionProxy;
 import net.sourceforge.waters.model.module.BoxGeometryProxy;
 import net.sourceforge.waters.model.module.ColorGeometryProxy;
 import net.sourceforge.waters.model.module.EdgeProxy;
@@ -26,6 +27,7 @@ import net.sourceforge.waters.model.module.EventDeclProxy;
 import net.sourceforge.waters.model.module.EventListExpressionProxy;
 import net.sourceforge.waters.model.module.ExpressionProxy;
 import net.sourceforge.waters.model.module.GraphProxy;
+import net.sourceforge.waters.model.module.GuardActionBlockProxy;
 import net.sourceforge.waters.model.module.IdentifierProxy;
 import net.sourceforge.waters.model.module.LabelBlockProxy;
 import net.sourceforge.waters.model.module.LabelGeometryProxy;
@@ -37,6 +39,7 @@ import net.sourceforge.waters.model.module.PointGeometryProxy;
 import net.sourceforge.waters.model.module.SimpleExpressionProxy;
 import net.sourceforge.waters.model.module.SimpleIdentifierProxy;
 import net.sourceforge.waters.model.module.SplineGeometryProxy;
+import net.sourceforge.waters.model.module.VariableProxy;
 
 import net.sourceforge.waters.xsd.base.ComponentKind;
 import net.sourceforge.waters.xsd.base.EventKind;
@@ -88,6 +91,16 @@ public class ModuleSubjectFactory
   }
 
   /**
+   * Creates a new boolean constant.
+   * @param value The boolean value of the new boolean constant.
+   */
+  public BooleanConstantSubject createBooleanConstantProxy
[…]

### 2006-03-02 src/net/sourceforge/waters/subject/module/RangeParameterSubject.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   RangeParameterSubject
 //###########################################################################
-//# $Id: RangeParameterSubject.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: RangeParameterSubject.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;

### 2006-03-02 src/net/sourceforge/waters/subject/module/SimpleComponentSubject.java

@@ -4,10 +4,14 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   SimpleComponentSubject
 //###########################################################################
-//# $Id: SimpleComponentSubject.java,v 1.3 2005-12-03 21:30:42 robi Exp $
+//# $Id: SimpleComponentSubject.java,v 1.4 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;
+
+import java.util.Collection;
+import java.util.Collections;
+import java.util.List;
 
 import net.sourceforge.waters.model.base.ProxyVisitor;
 import net.sourceforge.waters.model.base.VisitorException;
@@ -15,6 +19,10 @@ import net.sourceforge.waters.model.module.GraphProxy;
 import net.sourceforge.waters.model.module.IdentifierProxy;
 import net.sourceforge.waters.model.module.ModuleProxyVisitor;
 import net.sourceforge.waters.model.module.SimpleComponentProxy;
+import net.sourceforge.waters.model.module.VariableProxy;
+import net.sourceforge.waters.model.unchecked.Casting;
+import net.sourceforge.waters.subject.base.ArrayListSubject;
+import net.sourceforge.waters.subject.base.ListSubject;
 import net.sourceforge.waters.subject.base.ModelChangeEvent;
 
 import net.sourceforge.waters.xsd.base.ComponentKind;
@@ -38,15 +46,38 @@ public final class SimpleComponentSubject
    * @param identifier The identifier defining the name of the new simple component.
    * @param kind The kind of the new simple component.
    * @param graph The graph of the new simple component.
+   * @param variables The variables of the new simple component.
    */
   public SimpleComponentSubject(final IdentifierProxy identifier,
                                 final ComponentKind kind,
-                                final GraphProxy graph)
+                                final GraphProxy graph,
+                                final Collection<? extends VariableProxy> variables)
   {
     super(identifier);
     mKind = kind;
     mGraph = (GraphSubject) graph;
     mGraph.setParent(this);
+    mVariables = new ArrayListSubject<VariableSubject>
+      (variables, VariableSubject.class);
+    mVariables.setParent(this);
+  }
+
+  /**
+   * Creates a new simple component using default values.
+   * This constructor creates a simple component with
+   * an empty variables.
[…]

### 2006-03-02 src/net/sourceforge/waters/subject/module/SimpleExpressionSubject.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   SimpleExpressionSubject
 //###########################################################################
-//# $Id: SimpleExpressionSubject.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: SimpleExpressionSubject.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;

### 2006-03-02 src/net/sourceforge/waters/subject/module/SimpleIdentifierSubject.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   SimpleIdentifierSubject
 //###########################################################################
-//# $Id: SimpleIdentifierSubject.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: SimpleIdentifierSubject.java,v 1.3 2006-03-02 12:12:49 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.subject.module;

### 2006-03-02 src/net/sourceforge/waters/subject/module/SimpleNodeSubject.java

@@ -4,8 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.subject.module
 //# CLASS:   SimpleNodeSubject
 //###########################################################################
-//# $Id: SimpleNodeSubject.java,v 1.5 2006-02-16 04:06:18 robi Exp $
-//###########################################################################
+//# $Id: SimpleNodeSubject.java,v 1.6 2006-03-02 12:12:49 martin Exp $############################################################
 
 package net.sourceforge.waters.subject.module;
 

### 2006-03-02 src/net/sourceforge/waters/subject/module/VariableSubject.java

@@ -1,0 +1,164 @@
+//# -*- indent-tabs-mode: nil  c-basic-offset: 2 -*-
+//###########################################################################
+//# PROJECT: Waters
+//# PACKAGE: net.sourceforge.waters.subject.module
+//# CLASS:   VariableSubject
+//###########################################################################
+//# $Id: VariableSubject.java,v 1.2 2006-03-02 12:12:49 martin Exp $
+//###########################################################################
+
+package net.sourceforge.waters.subject.module;
+
+import net.sourceforge.waters.model.base.ProxyVisitor;
+import net.sourceforge.waters.model.base.VisitorException;
+import net.sourceforge.waters.model.module.ModuleProxyVisitor;
+import net.sourceforge.waters.model.module.SimpleExpressionProxy;
+import net.sourceforge.waters.model.module.VariableProxy;
+import net.sourceforge.waters.subject.base.ModelChangeEvent;
+import net.sourceforge.waters.subject.base.MutableSubject;
+
+
+/**
+ * The subject implementation of the {@link VariableProxy} interface.
+ *
+ * @author Robi Malik
+ */
+
+public final class VariableSubject
+  extends MutableSubject
+  implements VariableProxy
+{
+
+  //#########################################################################
+  //# Constructors
+  /**
+   * Creates a new variable.
+   * @param name The name of the new variable.
+   * @param type The type of the new variable.
+   * @param initialValue The initial value of the new variable.
+   */
+  public VariableSubject(final String name,
+                         final SimpleExpressionProxy type,
+                         final SimpleExpressionProxy initialValue)
+  {
+    mName = name;
+    mType = (SimpleExpressionSubject) type;
+    mType.setParent(this);
+    mInitialValue = (SimpleExpressionSubject) initialValue;
+    mInitialValue.setParent(this);
+  }
+
+
+  //#########################################################################
+  //# Cloning
+  public VariableSubject clone()
+  {
+    final VariableSubject cloned = (VariableSubject) super.clone();
+    cloned.mType = mType.clone();
+    cloned.mType.setParent(cloned);
+    cloned.mInitialValue = mInitialValue.clone();
+    cloned.mInitialValue.setParent(cloned);
+    return cloned;
+  }
+
+
+  //#########################################################################
+  //# Equality
+  public boolean equals(final Object partner)
+  {
+    if (super.equals(partner)) {
[…]

### 2006-03-02 src/org/supremica/automata/IO/ProjectBuildFromWaters.java

@@ -100,6 +100,7 @@ public class ProjectBuildFromWaters
 		}
 		catch (Exception ex)
 		{
+			ex.printStackTrace();
 			logger.error("Exception - Failed to compile: " + ex.getMessage(), ex);
 		}
 

### 2006-03-02 tests/src/net/sourceforge/waters/junit/WatersTestCase.java

@@ -4,14 +4,13 @@
 //# PACKAGE: net.sourceforge.waters.junit
 //# CLASS:   JAXBTestCase
 //###########################################################################
-//# $Id: WatersTestCase.java,v 1.2 2005-11-03 01:24:16 robi Exp $
+//# $Id: WatersTestCase.java,v 1.3 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 package net.sourceforge.waters.junit;
 
 import java.io.File;
 import junit.framework.TestCase;
-
 
 public abstract class WatersTestCase
   extends TestCase

### 2006-03-02 tests/src/net/sourceforge/waters/model/module/ModuleTest.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.module
 //# CLASS:   ModuleTest
 //###########################################################################
-//# $Id: ModuleTest.java,v 1.5 2006-02-22 03:35:07 robi Exp $
+//# $Id: ModuleTest.java,v 1.6 2006-03-02 12:12:50 martin Exp $
 //###########################################################################
 
 
@@ -304,7 +304,7 @@ public abstract class ModuleTest extends JAXBTestCase<ModuleProxy>
     final Point2D end = new Point(50, 25);
     final PointGeometryProxy endgeo = factory.createPointGeometryProxy(end);
     final EdgeProxy edge = factory.createEdgeProxy
-      (group, group, labelblock, null, startgeo, endgeo);
+      (group, group, labelblock, null, null, startgeo, endgeo);
     final LabelBlockProxy blocked = factory.createLabelBlockProxy();
     final List<NodeProxy> nodes = new LinkedList<NodeProxy>();
     nodes.add(node);

### 2006-03-10 src/org/supremica/automata/algorithms/scheduling/AbstractAstar.java
*** empty log message ***

@@ -65,13 +65,13 @@ public abstract class AbstractAstar
 	 * @see #ModifiedAstar.updateClosedTree(int[] node).
 	 */
 	protected TreeMap<Integer, int[]> closedTree;
-    
+
     /** 
-	 * The estimated cost for each state of each robot.
+	 * The remaining cost for each state of each robot (if run independently of other robots).
 	 * int[] = [robot_index, state_index].
 	 */
-    protected int[][] oneProdRelax;
-
+    protected int[][] remainingCosts;
+    
     /** Hashtable containing the estimated cost for each combination of two robots **/
     protected Hashtable[] twoProdRelax;
 
@@ -153,7 +153,7 @@ public abstract class AbstractAstar
 	private volatile boolean isRunning = false;
 
 	/** The dialog box that launched this scheduler */
-	private ScheduleDialog gui;
+	protected ScheduleDialog gui;
 
 	/** Decides if the schedule should be built */
 	private boolean buildSchedule;
@@ -164,14 +164,36 @@ public abstract class AbstractAstar
 	 */
 	private boolean manualExpansion;
 
+	protected Thread astarThread;
+
 
 	/****************************************************************************************/
 	/*                                 ABSTRACT METHODS                                     */
 	/****************************************************************************************/
 
+	/**
+	 * Converts the representation of the initial state to the int[]-representation of a node. 
+	 */
     protected abstract int[] makeInitialNode();
 
+	/**
+	 * Updates the closed tree if necessary, i.e. if the currently examined
+	 * node with better estimate value already is present on that tree.
+	 * If the node is added to the closed tree, its descendants are put on the 
+	 * open tree, according to the estimated remaining cost value.
+	 */
 	protected abstract void branch(int[] currNode);
+
+	/**
+	 * Calculates the total estimated cost for the node, i.e. the sum of the accumulated cost
+	 * and the estimate of the remaining cost (f(n) = g(n) + h(n)) is returned.
+	 *
+	 * @param int[] currNode - the node, whose estimated cost we want to know
+	 * @return int totalEstimatedCost - f(n) = g(n) + h(n)
+	 */
+
[…]

### 2006-03-10 src/org/supremica/automata/algorithms/scheduling/Milp.java

@@ -331,7 +331,24 @@ public class Milp
 		logger.info("OPTIMAL MAKESPAN: " + makespan + ".............................");
 		logger.info("Time to build the schedule: " + timer.elapsedTime() + "ms ");
 
-		ActionMan.getGui().addAutomaton(schedule);
+		synchronized (this)
+		{
+			org.supremica.gui.Gui theGui = null;
+			try 
+			{
+				theGui = ActionMan.getGui(); 
+				theGui.addAutomaton(schedule);
+// 				if (theGui != null)
+// 					theGui.addAutomaton(schedule);
+// 				else
+// 					logger.info("THE GUI is NULL");
+			}
+			catch (Exception ex)
+			{
+				logger.warn("EXceptiON, gui = " + theGui);
+				throw ex;
+			}
+		}
 	}
 
 

### 2006-03-10 src/org/supremica/automata/algorithms/scheduling/ModifiedAstar.java

@@ -46,6 +46,28 @@ public class ModifiedAstar
 		super(theAutomata, heuristic, manualExpansion, iterativeSearch, buildSchedule, gui);
     }
 
+	public ModifiedAstar(Automata theAutomata, boolean manualExpansion, boolean buildSchedule, ScheduleDialog gui)
+		throws Exception
+	{
+		super(theAutomata, manualExpansion, buildSchedule, gui);
+	}
+
+	/**
+	 * The defalut estimation of the remaining cost. This method is overriden 
+	 * in the subclasses. Is used if no relaxation is chosen (i.e. brute force search).
+	 */
+	public int calcEstimatedCost(int[] node)
+		throws Exception
+	{
+		return node[ACCUMULATED_COST_INDEX];
+	}
+
+	/**
+	 * Updates the closed tree if necessary, i.e. if the currently examined
+	 * node with better estimate value already is present on that tree.
+	 * If the node is added to the closed tree, its descendants are put on the 
+	 * open tree, according to the estimated remaining cost value.
+	 */
 	protected void branch(int[] currNode) 
 	{
 		try 

### 2006-03-10 src/org/supremica/automata/algorithms/scheduling/ModifiedAstarUsingOneProdRelaxation.java

@@ -1,0 +1,59 @@
+package org.supremica.automata.algorithms.scheduling;
+
+import java.util.*;
+
+import org.supremica.automata.*;
+import org.supremica.gui.ScheduleDialog;
+import org.supremica.log.*;
+import org.supremica.util.ActionTimer;
+
+public class ModifiedAstarUsingOneProdRelaxation
+	extends ModifiedAstar
+{
+	/**
+	 * The output stream
+	 */
+	private static Logger logger = LoggerFactory.createLogger(ModifiedAstarUsingOneProdRelaxation.class);
+
+	public ModifiedAstarUsingOneProdRelaxation(Automata theAutomata, boolean manualExpansion, boolean buildSchedule, ScheduleDialog gui) 
+		throws Exception 
+	{
+			super(theAutomata, manualExpansion, buildSchedule, gui);
+    }
+
+	public int calcEstimatedCost(int[] node)
+			throws Exception
+	{
+		int[] parent = getParent(node);
+	    
+		int fNode = node[ACCUMULATED_COST_INDEX] + getOneProdRelaxation(node);
+		int fParent = parent[ACCUMULATED_COST_INDEX] + getOneProdRelaxation(parent);
+
+		// The following code inside the if-loop is needed to ensure consistency of the heuristic
+		if (fParent > fNode)
+			return fParent;
+		else
+			return fNode;
+	}
+
+	/**
+	 * This method calculates the remaining cost for each robot/plant. The maximum remaining
+	 * cost is returned to be used as an estimate of the total remaining cost of the system. 
+	 * 
+	 * @param int[] node - the current node
+	 * @return int - the heuristic function, h(n), that guides the search, in this case it is the "1-product relaxation"
+	 */
+    private int getOneProdRelaxation(int[] node) {
+		int estimate = 0;
+		int[] currCosts = expander.getCosts(node);
+	
+		for (int i=0; i<activeAutomataIndex.length; i++) {
+			int altEstimate = currCosts[i] + remainingCosts[i][node[activeAutomataIndex[i]]]; 
+	    
+			if (altEstimate > estimate)
+				estimate = altEstimate;
+		}
+	
+		return estimate;
+    }
+}

### 2006-03-10 src/org/supremica/automata/algorithms/scheduling/ModifiedAstarUsingVisGraphRelaxation.java

@@ -1,0 +1,128 @@
+package org.supremica.automata.algorithms.scheduling;
+
+import java.util.*;
+
+import org.supremica.automata.*;
+import org.supremica.gui.ScheduleDialog;
+import org.supremica.log.*;
+import org.supremica.util.ActionTimer;
+
+public class ModifiedAstarUsingVisGraphRelaxation
+	extends ModifiedAstar
+{
+	/** 
+	 * Hashtable containing the estimated cost for a given time configuration 
+	 * (as returned by the visibility graph)
+	 */
+	private Hashtable<String, Integer> visGraphRelax = null;
+
+	/**
+	 * The output stream
+	 */
+	private static Logger logger = LoggerFactory.createLogger(ModifiedAstarUsingVisGraphRelaxation.class);
+
+
+	public ModifiedAstarUsingVisGraphRelaxation(Automata theAutomata, boolean manualExpansion, boolean buildSchedule, ScheduleDialog gui) 
+		throws Exception 
+	{
+			super(theAutomata, manualExpansion, buildSchedule, gui);
+    }
+
+	public void init (boolean manualExpansion)
+		throws Exception
+	{
+		super.init(manualExpansion);
+
+		visGraphRelax = new Hashtable<String, Integer>();
+
+		timer.restart();
+		preprocessVisibilityGraphs();
+		infoStr += "\tvisibility graphs calculated in " + timer.elapsedTime() + "ms\n";
+	}
+
+	public int calcEstimatedCost(int[] node)
+		throws Exception
+	{
+		int[] effTimePoint = new int[plantAutomata.size()];
+		String timePointKey = "";
+
+		// Calculate the visibility graph time coordinate, corresponding to the current node
+		for (int i=0; i<plantAutomata.size(); i++)
+		{
+			State initState = plantAutomata.getAutomatonAt(i).getInitialState();
+			int totalCycleTime = remainingCosts[i][initState.getIndex()] + initState.getCost();
+			int remainingCycleTime = remainingCosts[i][node[activeAutomataIndex[i]]] + node[CURRENT_COSTS_INDEX + i];
+
+			effTimePoint[i] = totalCycleTime - remainingCycleTime;
+			timePointKey += effTimePoint[i] + "_";
+		}
+
+		Integer previousVisibilityRelaxation = visGraphRelax.get(timePointKey);
+		if (previousVisibilityRelaxation == null)
+		{
[…]

### 2006-03-10 src/org/supremica/automata/algorithms/scheduling/VisGraphScheduler.java

@@ -42,6 +42,9 @@ public class VisGraphScheduler
 	/** To draw or not to draw - this is the question... */
 	private boolean toDrawVisibilityGraph;
 
+	/** Determines if this visibility graph is only used for relaxation (and is thus called by another scheduler) */
+	private boolean isRelaxationProvider = false;
+
 	private ActionTimer timer = new ActionTimer();
 
 	/** The index mapping handler */
@@ -53,12 +56,14 @@ public class VisGraphScheduler
 	/** The dialog box that launched this scheduler */
 	private ScheduleDialog gui;
 
-	public VisGraphScheduler (Automata robots, Automata zones, boolean toDrawVisibilityGraph)
-		throws Exception
-	{
-		this(robots, zones, toDrawVisibilityGraph, null);
-	}
+	/** 
+	 * Assures that the mutex zones of the graph have been initialized 
+	 * (is needed for correct functionning when this thread is started from
+	 * another thread for relaxation purposes).
+	 */
+	private boolean isInitialized = false;
 
+	
 	public VisGraphScheduler (Automata theAutomata, boolean toDrawVisibilityGraph, ScheduleDialog gui)
 		throws Exception
 	{
@@ -68,11 +73,18 @@ public class VisGraphScheduler
 	public VisGraphScheduler (Automata robots, Automata zones, boolean toDrawVisibilityGraph, ScheduleDialog gui)
 		throws Exception
 	{
+		this(robots, zones, toDrawVisibilityGraph, false, gui);
+	}
+	
+	public VisGraphScheduler (Automata robots, Automata zones, boolean toDrawVisibilityGraph, boolean isRelaxationProvider, ScheduleDialog gui)
+		throws Exception
+	{
 		this.toDrawVisibilityGraph = toDrawVisibilityGraph;
-
+		
 		this.robots = robots;
 		this.zones = zones;
 		this.gui = gui;
+		this.isRelaxationProvider = isRelaxationProvider;
 
 		Thread vgThread = new Thread(this);
 		isRunning = true;
@@ -85,48 +97,39 @@ public class VisGraphScheduler
 		{
 			if (isRunning)
 			{
-				timer.restart();
-				extractGraphTimes();
-
-				// If the call came from GUI, print the preprocessing time
-				if (gui != null)
-				{
-					logger.info("Preprocessing in " + timer.elapsedTime() + "ms");
-				}
-			}
-
-			if (isRunning)
-			{
 				init();
 			}
 
[…]

### 2006-03-10 src/org/supremica/gui/ScheduleDialog.java

@@ -19,9 +19,12 @@ public class ScheduleDialog
 	private static final String MILP = "MILP";
 	private static final String VIS_GRAPH = "Visibility Graph";
 
+	private static final String ONE_PRODUCT_RELAXATION = "1-product relax";
+	private static final String VIS_GRAPH_RELAXATION = "visibility graph";
+
     private static final long serialVersionUID = 1L;
     private static final String[] optiMethodNames = new String[]{MODIFIED_A_STAR, MODIFIED_VGA_STAR, MILP, VIS_GRAPH}; //, "Modified IDA*", "Modified SMA*"};
-    private static final String[] heuristicsNames = new String[]{"1-product relax", "visibility graph", "brute force"}; //"2-product relax",
+    private static final String[] heuristicsNames = new String[]{ONE_PRODUCT_RELAXATION, VIS_GRAPH_RELAXATION, "brute force"}; //"2-product relax",
     private static Logger logger = LoggerFactory.createLogger(ScheduleDialog.class);
     private JComboBox optiMethodsBox, heuristicsBox;
     private JCheckBox nodeExpander, buildAutomaton, vgDrawer;
@@ -171,11 +174,30 @@ public class ScheduleDialog
 			
 			if (optiMethodsBox.getSelectedItem().equals(MODIFIED_A_STAR))
 			{
-				sched = new ModifiedAstar(ActionMan.getGui().getSelectedAutomata(), (String) heuristicsBox.getSelectedItem(), nodeExpander.isSelected(), false, buildAutomaton.isSelected(), this);
+				String selectedHeuristic = (String) heuristicsBox.getSelectedItem();
+				if (selectedHeuristic.equals(VIS_GRAPH_RELAXATION))
+				{
+					if (ActionMan.getGui().getSelectedAutomata().getPlantAutomata().size() == 2)
+					{
+						sched = new VisGraphScheduler(ActionMan.getGui().getSelectedAutomata(), vgDrawer.isSelected(), this);
+					}
+					else
+					{
+						sched = new ModifiedAstarUsingVisGraphRelaxation(ActionMan.getGui().getSelectedAutomata(), nodeExpander.isSelected(), buildAutomaton.isSelected(), this);
+					}
+				}
+				else if (selectedHeuristic.equals(ONE_PRODUCT_RELAXATION))
+				{
[…]

### 2006-03-10 doc/EFA/efa.tex
*** empty log message ***

@@
  a specification in EFA form. To analyze the specification against the plant and at the same time
-use the modular structure in the translation algorithm we must translate them simultaneously.
+use the modular structure in step 2 in the translation algorithm we must translate the
+specification and the plant simultaneously.
+
+Before we calculate a supervisor from the modular model obtained in step 2 we divide the alphabet of $A'$ (the
+relabelled alphabet)
+into as many subsets as the number of events (events that occur in transitions) in $A_e$. Each subset
+$\Sigma_\sigma=\{\text{events in } A' \text{ obtained by relabelling } \sigma \}$
+is an equivalence class representing the event $\sigma$ in the original alphabet. The supervisor
+calculated from the modular model with the relabelled alphabet have one important restriction,
+ it must either accept all events in an equivalence class or forbid all events in an equivalence class. Of course
+  we need only to consider the controllable events in $A_e$.
+ The key is to treat all the relabelled events as uncontrollable and control them
+ simultaneously with the original controllable event
+ of $A_e$ that defines the their equivalence class. This
+ can easily be obtained using regular algorithms by adding a simple automata shown in fig.
 
@@
 To avoid such situations we require that action functions of common
-variables are identical.
+variables are identical?
 

### 2006-03-16 src/net/sourceforge/waters/gui/EditorSurface.java
*** empty log message ***

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.gui
 //# CLASS:   EditorSurface
 //###########################################################################
-//# $Id: EditorSurface.java,v 1.45 2006-03-13 12:46:24 martin Exp $
+//# $Id: EditorSurface.java,v 1.46 2006-03-16 12:09:16 flordal Exp $
 //###########################################################################
 
 
@@ -185,7 +185,6 @@ public class EditorSurface
 		for (int i = 0; i < nodeGroups.size(); i++)
 		{
 			EditorNodeGroup n = (EditorNodeGroup) nodeGroups.get(i);
-
 			n.drawObject(g, isSelected(n));
 
 			ArrayList groups = new ArrayList();
@@ -265,14 +264,14 @@ public class EditorSurface
 			}
 		}
 		
-		if (newGroup != null) {
-			    newGroup.drawObject(g, isSelected(newGroup));
-			}
-
+		if (newGroup != null) 
+		{
+			newGroup.drawObject(g, isSelected(newGroup));
+		}
+		
 		for (int i = 0; i < edges.size(); i++)
 		{
 			EditorEdge edge = (EditorEdge) edges.get(i);
-
 			edge.drawObject(g, isSelected(edge));
 
 			/*
@@ -295,21 +294,18 @@ public class EditorSurface
 		for (int i = 0; i < nodes.size(); i++)
 		{
 			EditorNode n = (EditorNode) nodes.get(i);
-
 			n.drawObject(g, isSelected(n));
 		}
 
 		for (int i = 0; i < labels.size(); i++)
 		{
 			EditorLabel l = (EditorLabel) labels.get(i);
-
 			l.drawObject(g, isSelected(l) || isSelected(l.getParent()));
 		}
 
 		for (int i = 0; i < events.size(); i++)
 		{
 			EditorLabelGroup l = (EditorLabelGroup) events.get(i);
-
 			l.setPanelLocation(isSelected(l) || isSelected(l.getParent()));
 
 			// Why is this done here? Why are labelgroups treated differently?
@@ -327,15 +323,13 @@ public class EditorSurface
 		
 		for (int i = 0; i < mGuardActionBlocks.size(); i++)
 		{
+			EditorGuardActionBlock block = (EditorGuardActionBlock) mGuardActionBlocks.get(i);
+			block.setPanelLocation();
 
[…]

### 2006-03-16 src/net/sourceforge/waters/model/compiler/CompilerOperatorTable.java

@@ -4,7 +4,7 @@
 //# PACKAGE: net.sourceforge.waters.model.compiler
 //# CLASS:   CompilerOperatorTable
 //###########################################################################
-//# $Id: CompilerOperatorTable.java,v 1.3 2006-03-02 12:12:50 martin Exp $
+//# $Id: CompilerOperatorTable.java,v 1.4 2006-03-16 12:09:16 flordal Exp $
 //###########################################################################
 
 package net.sourceforge.waters.model.compiler;
@@ -43,6 +43,7 @@ public class CompilerOperatorTable extends AbstractOperatorTable {
     store(new BinaryPlusOperator());
     store(new BinaryMinusOperator());
     store(new BinaryEqualsOperator());
+    store(new BinaryNotEqualsOperator());
     store(new BinaryRangeOperator());
     store(new UnaryMinusOperator());
 
@@ -367,6 +368,21 @@ public class CompilerOperatorTable extends AbstractOperatorTable {
 
   }
 
+  //#########################################################################
+  //# Inner Class BinaryNotEqualsOperator, inherits from BinaryEqualsOperator
+  private static class BinaryNotEqualsOperator extends BinaryEqualsOperator
+  {
+    public String getName()
+    {
+      return OPNAME_NOTEQUALS;
+    }
+
+    public CompiledIntValue eval(final Value lhsValue, final Value rhsValue)
+    {
+      final boolean result = lhsValue.equals(rhsValue);
+      return new CompiledIntValue(!result);     
+    }
+  }
 
   //#########################################################################
   //# Inner Class BinaryRangeOperator
@@ -461,6 +477,7 @@ public class CompilerOperatorTable extends AbstractOperatorTable {
     new CompilerOperatorTable();
 
   private static final String OPNAME_EQUALS = "==";
+  private static final String OPNAME_NOTEQUALS = "!=";
   private static final String OPNAME_MINUS = "-";
   private static final String OPNAME_PLUS = "+";
   private static final String OPNAME_RANGE = "..";

### 2006-03-16 src/org/supremica/automata/IO/ProjectBuildFromWaters.java

@@ -1,4 +1,3 @@
-//# -*- tab-width: 4  indent-tabs-mode: t  c-basic-offset: 4 -*-
 /*
  *  Supremica Software License Agreement
  *
@@ -64,139 +63,143 @@ import org.supremica.log.*;
 
 public class ProjectBuildFromWaters
 {
-	private static Logger logger = LoggerFactory.createLogger(ProjectBuildFromWaters.class);
-	private ProjectFactory theProjectFactory = null;
-	private Project currProject = null;
+    private static Logger logger = LoggerFactory.createLogger(ProjectBuildFromWaters.class);
+    private ProjectFactory theProjectFactory = null;
+    private Project currProject = null;
 
-	public ProjectBuildFromWaters()
+    public ProjectBuildFromWaters()
+    {
+	this.theProjectFactory = new DefaultProjectFactory();
+    }
+
+    public ProjectBuildFromWaters(ProjectFactory theProjectFactory)
+    {
+	this.theProjectFactory = theProjectFactory;
+    }
+
+    public Project build(ModuleProxy module)
+    {
+	if (module == null)
 	{
-		this.theProjectFactory = new DefaultProjectFactory();
+	    throw new IllegalArgumentException("argument must be non null");
+	}
+	Project currProject = theProjectFactory.getProject();
+	currProject.setName(module.getName());
+
+	final ProductDESProxyFactory factory =
+	    ProductDESElementFactory.getInstance();
+	DocumentManager mDocumentManager = new DocumentManager();
+	ModuleCompiler compiler =
+	    new ModuleCompiler(mDocumentManager, factory, module);
+
+	ProductDESProxy des = null;
+	try
+	{
+	    des = compiler.compile();
+	}
+	catch (Exception ex)
+	{
+	    ex.printStackTrace();
+	    logger.error("Exception - Failed to compile: " + ex.getMessage(), ex);
 	}
 
-	public ProjectBuildFromWaters(ProjectFactory theProjectFactory)
+	if (des == null)
 	{
-		this.theProjectFactory = theProjectFactory;
+	    return currProject;
 	}
 
-	public Project build(ModuleProxy module)
+	Collection theWatersAutomata = des.getAutomata();
+	for (Iterator autIt = theWatersAutomata.iterator(); autIt.hasNext(); )
 	{
-		if (module == null)
+	    AutomatonProxy currWatersAutomaton = (AutomatonProxy)autIt.next();
+	    Automaton currSupremicaAutomaton = new Automaton();
[…]

### 2006-03-16 src/org/supremica/automata/algorithms/minimization/AutomataMinimizer.java

@@ -47,747 +47,679 @@
  * Supremica is owned and represented by KA.
  */
 package org.supremica.automata.algorithms.minimization;
-
 import org.supremica.log.*;
 import org.supremica.automata.*;
 import org.supremica.automata.algorithms.*;
 import org.supremica.gui.*;
 import org.supremica.util.ActionTimer;
 import org.supremica.properties.SupremicaProperties;
-
 import java.util.*;
-
 public class AutomataMinimizer
-	implements Stoppable
+    implements Stoppable
 {
-	private static Logger logger = LoggerFactory.createLogger(AutomataMinimizer.class);
-
-	// Stoppable stuff
-	private ExecutionDialog executionDialog;
-	private Stoppable threadToStop = null;
-	private boolean stopRequested = false;
-
-	/** The automata being minimized (may be a copy of the original). */
-	private Automata theAutomata;
-
-	/** The supplied options. */
-	private MinimizationOptions options;
-
-	/** The strategy to use, in order of priority */
-	private LinkedList<MinimizationStrategy> strategyList;
-
-	/** The heuristic to use, in order of priority */
-	private LinkedList<MinimizationHeuristic> heuristicList;
-
-	////////////////
-	// Statistics //
-	////////////////
-
-	/** Largest number of states come across. */
-	private int mostStates = 0;
-	/** Largest number of transitions come across. */
-	private int mostTransitions = 0;
-
-	/** Number of automata in the beginning. */
-	private int initialNbrOfAutomata;
-
-	/** Event to automata map, for choosing the next task in compositional minimization. */
-	private EventToAutomataMap eventToAutomataMap;
-
-	/**
-	 * Basic constructor.
-	 */
-	public AutomataMinimizer(Automata theAutomata)
-	{
-		initialNbrOfAutomata = theAutomata.nbrOfAutomata();
-		this.theAutomata = theAutomata;
-	}
-
-	/**
-	 * Sets the executionDialog of this AutomataMinimizer. If executionDialog is null,
-	 * the dialog is not updated.
-	 */
-	public void setExecutionDialog(ExecutionDialog executionDialog)
-	{
-		this.executionDialog = executionDialog;
-	}
-
-	/**
-	 * Returns minimized automaton, minimized with respect to the supplied options.
-	 */
-	public Automaton getCompositionalMinimization(MinimizationOptions options)
-		throws Exception
-	{
-		this.options = options;
[…]

### 2006-03-16 src/org/supremica/external/robotCoordination/CellExaminer.java

@@ -22,17 +22,17 @@ import java.awt.Color;
 public class CellExaminer
     extends JDialog
 {
-	// private static final long serialVersionUID = 1L;
-	private static Logger logger = LoggerFactory.createLogger(CellExaminer.class);
-	private Frame owner = null;
-	private JPanel contentPane = null;
+    // private static final long serialVersionUID = 1L;
+    private static Logger logger = LoggerFactory.createLogger(CellExaminer.class);
+    private Frame owner = null;
+    private JPanel contentPane = null;
 
-	// Robot stuff
-	private Cell cell;
+    // Robot stuff
+    private Cell cell;
 
-	// Supremica stuff
-	private Automata zoneAutomata;
-	private Automata robotAutomata;
+    // Supremica stuff
+    private Automata zoneAutomata;
+    private Automata robotAutomata;
 
     // Automata constants
     public final static String STARTSTATE_NAME = "start";
@@ -42,906 +42,906 @@ public class CellExaminer
     public final static String BOOKEDSTATE_NAME = "Booked";
     public final static String UNDERSCORE = "_";
 
-	/** The prefix of a zone's name. */
-	static final String ZONE_PREFIX = "Zone_";
+    /** The prefix of a zone's name. */
+    static final String ZONE_PREFIX = "Zone_";
 
-	// Colors
-	static final Color RED = Color.RED;
-	static final Color ORANGE = Color.ORANGE;
-	static final Color GREEN = Color.GREEN;
-	static final Color BLUE = Color.BLUE;
-	static final Color BLACK = Color.BLACK;
-	// Transparency
-	static final double TRANSPARENCY = 0.9;
+    // Colors
+    static final Color RED = Color.RED;
+    static final Color ORANGE = Color.ORANGE;
+    static final Color GREEN = Color.GREEN;
+    static final Color BLUE = Color.BLUE;
+    static final Color BLACK = Color.BLACK;
+    // Transparency
+    static final double TRANSPARENCY = 0.9;
 
-	// Demo file
-	// final String DEMOSTATION_FILENAME = "C:/temp/RobSuprTestStation/RobSuprTest.stn";
-	// final String DEMOSTATION_FILENAME = "C:/temp/DomStations/DemoSafe.stn";
-	String DEMOSTATION_FILENAME = "DemoSafe.stn";
+    // Demo file
+    // final String DEMOSTATION_FILENAME = "C:/temp/RobSuprTestStation/RobSuprTest.stn";
[…]

### 2006-03-16 src/org/supremica/external/robotCoordination/RobotStudio/RSCell.java

@@ -82,7 +82,7 @@ public class RSCell
     /** The name of the IPart containing the mutex zones. */
     final static String ZONEPART_NAME = "Mutex_Zones";
     //final static String BOXPART_NAME = "Box_Set";
-	final static String BOXPART_NAME = ZONEPART_NAME;
+    final static String BOXPART_NAME = ZONEPART_NAME;
     final static String VOLUMEPART_NAME = "Volumes_Set";
 
     final static String ZONEENTITY_BASENAME = "Zone";
@@ -386,178 +386,178 @@ public class RSCell
     public static Automata buildRobotAutomata(List<Robot> robots)
 		throws Exception
     {
-		Automata robotAut = new Automata();
-
-		// Iterate over the robots...
-		for (Iterator robotIt = robots.iterator(); robotIt.hasNext();)
+	Automata robotAut = new Automata();
+	
+	// Iterate over the robots...
+	for (Iterator robotIt = robots.iterator(); robotIt.hasNext();)
+	{
+	    Robot robot = (Robot) robotIt.next();
+	    
+	    // ONE AUTOMATON FOR THE ROBOT ITSELF //
+	    Automaton aut = new Automaton(robot.getName());
+	    aut.setType(AutomatonType.Plant);
+	    // Build the states...
+	    State initial = new State(CellExaminer.STARTSTATE_NAME);
+	    initial.setInitial(true);
+	    aut.addState(initial);
+	    initial.setCost(0);
+	    State marked = new State(CellExaminer.FINISHSTATE_NAME);
+	    marked.setAccepting(true);
+	    aut.addState(marked);
+	    marked.setCost(0);
+	    List<Configuration> posList = robot.getConfigurations();
+	    for (int i=0; i < posList.size(); i++)
 	    {
-			Robot robot = (Robot) robotIt.next();
-
-			// ONE AUTOMATON FOR THE ROBOT ITSELF //
-			Automaton aut = new Automaton(robot.getName());
-			aut.setType(AutomatonType.Plant);
-			// Build the states...
-			State initial = new State(CellExaminer.STARTSTATE_NAME);
-			initial.setInitial(true);
-			aut.addState(initial);
-			initial.setCost(0);
-			State marked = new State(CellExaminer.FINISHSTATE_NAME);
-			marked.setAccepting(true);
-			aut.addState(marked);
-			marked.setCost(0);
-			List<Configuration> posList = robot.getConfigurations();
[…]

### 2006-03-16 src/org/supremica/gui/ActionMan.java

@@ -74,6 +74,14 @@ import org.supremica.gui.useractions.*;
 import org.supremica.gui.texteditor.TextFrame;
 import org.swixml.SwingEngine;
 
+// Waters importing
+import net.sourceforge.waters.subject.module.*;
+import net.sourceforge.waters.model.module.*;
+import net.sourceforge.waters.model.marshaller.*;
+import net.sourceforge.waters.model.expr.*;
+import net.sourceforge.waters.model.compiler.CompilerOperatorTable;
+import java.net.URI;
+
 // Robot coordination junk
 //import org.supremica.external.robotCoordinationABB.*;
 //import org.supremica.automata.algorithms.RobotStudioLink;
@@ -2152,6 +2160,17 @@ public class ActionMan
 		};
 	}
 
+	public static void fileImportWaters(Gui gui)
+	{
+		new FileImporter(FileDialogs.getWatersFileImporter(), gui)    // anonymous class
+		{
+			void openFile(Gui g, File f)
+			{
+				importWatersFile(g, f);
+			}
+		};
+	}
+
 	public static void fileImportHYB(Gui gui)
 	{
 		new FileImporter(FileDialogs.getHYBFileImporter(), gui)    // anonymous class
@@ -2461,6 +2480,36 @@ public class ActionMan
 			int nbrOfAddedAutomata = gui.addAutomata(currAutomata);
 
 			gui.info("Successfully imported " + nbrOfAddedAutomata + " automata.");
+		}
+		catch (Exception ex)
+		{
+			logger.error("Error while importing " + file.getAbsolutePath() + ". " + ex);
+			logger.debug(ex.getStackTrace());
+
+			return;
+		}
+	}
+
+	public static void importWatersFile(Gui gui, File file)
+	{
+		gui.info("Importing " + file.getAbsolutePath() + " ...");
+
+		try
+		{
+		    // Build Waters ModuleProxy
+		    final ModuleProxyFactory factory =
+			ModuleSubjectFactory.getInstance();
+		    final OperatorTable optable = CompilerOperatorTable.getInstance();
+		    final ProxyUnmarshaller<ModuleProxy> unMarshaller =
+			new JAXBModuleMarshaller(factory, optable);
+		    final URI uri = file.toURI();
+		    ModuleProxy module = (ModuleSubject) unMarshaller.unmarshal(uri);
+
[…]

### 2006-03-16 src/org/supremica/gui/ExampleTemplates.java

@@ -179,6 +179,7 @@ public class ExampleTemplates
 		thisGroup.addItem(new TemplateItem("Ball Process - Gatekeeper", prefix + "ballProcessGatekeeper.xml"));
 		thisGroup.addItem(new TemplateItem("Car Window Control System", prefix + "big_bmw.xml"));
 		thisGroup.addItem(new TemplateItem("Observation Equivalence Execise", prefix + "ObservationEquivalenceExercise.xml"));
+		thisGroup.addItem(new TemplateItem("Passenger Land-Transportation System", prefix + "PLanTS.xml"));
 		thisGroup.addItem(new TemplateItem("Professors, Pen and Paper", prefix + "ProfessorsPenPaper.xml"));
 		thisGroup.addItem(new TemplateItem("Telecommunications Network (incomplete)", prefix + "telecommunicationsNetwork.xml"));
     }
@@ -293,6 +294,7 @@ public class ExampleTemplates
 		String prefix = extraPrefix + "/Games/";
 
 		thisGroup.addItem(new TemplateItem("Wine Merchant's Problem", prefix + "WineMerchant.xml"));
+		thisGroup.addItem(new TemplateItem("Man, Wolf, Goat, Cabbage", extraPrefix + "/CCSBookExercises/" + "Ex2_6.xml"));
     }
 
     /*

### 2006-03-16 src/org/supremica/gui/FileDialogs.java

@@ -1,5 +1,5 @@
 //# -*- tab-width: 4  indent-tabs-mode: t  c-basic-offset: 4 -*-
-//# $Id: FileDialogs.java,v 1.27 2005-11-28 16:52:26 knut Exp $
+//# $Id: FileDialogs.java,v 1.28 2006-03-16 12:09:16 flordal Exp $
 
 /*
  * Supremica Software License Agreement
@@ -218,6 +218,16 @@ public class FileDialogs
 		fileImporter.addChoosableFileFilter(fd.getDGRFFilter());
 		fileImporter.addChoosableFileFilter(fd.getVMODFilter());
 		fileImporter.setFileFilter(fd.getVPRJFilter());
+
+		return fileImporter;
+	}
+
+	public static JFileChooser getWatersFileImporter()
+	{
+		JFileChooser fileImporter = fd.getFileImporter();
+
+		fileImporter.resetChoosableFileFilters();
+		fileImporter.setFileFilter(fd.getWMODFilter());
 
 		return fileImporter;
 	}
@@ -520,6 +530,16 @@ public class FileDialogs
 		return mainvmodFilter;
 	}
 
+	private FileFilter getDGRFFilter()
+	{
+		if (dgrfFilter == null)
+		{
+			dgrfFilter = makeFileFilter(".dgrf", "VALID Graph files (*.dgrf)");
+		}
+
+		return dgrfFilter;
+	}
+
 	private FileFilter getWMODFilter()
 	{
 		if (wmodFilter == null)
@@ -529,16 +549,6 @@ public class FileDialogs
 		}
 
 		return wmodFilter;
-	}
-
-	private FileFilter getDGRFFilter()
-	{
-		if (dgrfFilter == null)
-		{
-			dgrfFilter = makeFileFilter(".dgrf", "VALID Graph files (*.dgrf)");
-		}
-
-		return dgrfFilter;
 	}
 
 	private FileFilter getHYBFilter()

### 2006-03-16 src/org/supremica/gui/MainMenuBar.java

@@ -230,6 +230,16 @@ public class MainMenuBar
 		menuFileImport.setToolTipText("Import file");
 	    menuFile.add(menuFileImport);
 
+	    // File.Import.Waters
+	    JMenuItem menuFileImportWaters = new JMenuItem("From Waters...");
+	    menuFileImport.add(menuFileImportWaters);
+	    menuFileImportWaters.addActionListener(new ActionListener()
+			{
+				public void actionPerformed(ActionEvent e)
+				{
+					ActionMan.fileImportWaters(ActionMan.getGui());
+				}
+			});
 	    // File.Import.UMDES
 	    JMenuItem menuFileImportUMDES = new JMenuItem("From UMDES...");
 	    menuFileImport.add(menuFileImportUMDES);

### 2006-03-16 doc/EFA/efa.tex
*** empty log message ***

@@
 that are triggered by common events. To avoid contradictions when synchronizing
-automata, we do not allow this for the updating actions of variables
+automata, we do not allow this for the updating actions of global variables
 that are shared between automata.\\
@@
 $X_k$, $k=1,...,n$, are finite nonempty sets. For instance, $X_k$
-could be any finite subset of $\mathbb{Z}$ or the set of states in  some
-other finite deterministic automaton. The finite set $X$ is the
+could be any finite subset of $\mathbb{Z}$, the set of states an automaton or
+an input signal. The finite set $X$ is the
 domain of definition of an $n$-tuple of variables
-$x=(x_1,...,x_n)$ with initial value $x_0=(x_{10},...,x_{n0})$. If
-a variable corresponds to the set of states in an automaton, we
+$x=(x_1,...,x_n)$ with initial value $x_0=(x_{10},...,x_{n0})$. If a variable is
+an input signal to the system that $A$ describes or the value
+of the variable represents the state that an automaton occupies, we
 shall call it a state variable. The initial value of a state
@@
 since it is the main reason why EFA's enable efficient modular algorithms to perform
-verifications of supervisors.
+verifications and to find supervisors.
 \subsection{How to use the variables}
 All logical systems can easily be represented by
-EFA models. When using EFA's as a modelling tool it important to understand how to choose the variables.
+EFA models. When using EFA's as a modelling tool it important to choose variables in a structured way.
 A variable $x$ can represent many things, some examples  are: the states that an object can be in,
@@
 takes for a process to execute different operations. In many applications the variables can be represented by
- integers or enumeration objects.
+ integers or enumeration objects. One way to standardize model building with EFA is to each EFA divide the
+ variables in input variables and output variables.
 
-\subsubsection{Integer variables}
+\subsubsection{Input variables}
 
-\subsubsection{Enumeration variables}
+\subsubsection{Output variables}
 
@@
 that when $A_e$ only has ordinary variables $X=X_v$. If $A_e$ also has
-state variables the guard conditions on these variables are evaluated. When a variable
-corresponds to the set of states
-in an other automata this is not possible. Therefore its guard conditions are set to
+state variables the guard conditions on these variables can not be evaluated. Therefore its guard conditions are set to
 TRUE when the translation is performed. To simplify the
